### 2012-02-07 paper/abstract.tex
The paper directory added. It's easier to write it here than on kruso.

% This is LLNCS.DEM the demonstration file of
% the LaTeX macro package from Springer-Verlag
% for Lecture Notes in Computer Science, version 1.1
\documentclass[10pt, a4paper]{article}
\oddsidemargin 4mm
\evensidemargin 4mm
\usepackage{lrec2006}
\usepackage[colorlinks,urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}
%\usepackage[hungarian]{babel}
%\selectlanguage{hungarian}
%
\begin{document}

\title{Rapid creation of large-scale corpora and frequency dictionaries}

\name{Attila Zs\'eder$^1$, G\'abor Recski$^{1}$, D\'aniel Varga$^2$, Andr\'as Kornai$^1$}
\address{$^1$Computer and Automation Research Institute\\Hungarian Academy of Sciences\\{\tt\{recski,zseder,kornai\}@sztaki.hu}\\
$^2$Media Education and Research Center\\Budapest University of Technology and Economics\\{\tt daniel@mokk.bme.hu}
}

%beigerunk ket tucatot es a szoftvert
%sztaki szotar osszes nyelve is minusz amiben nem nyomjuk le Baronit 
%TODO es textcat is, hivatkozassal
%Baroniekrol ne seggnyalast, hanem angol nemet olaszra oket "for the largest languages"
%TODO miert mas ez mint a wac pipeline, Daniel?
%TODO ha minden említett korpuszhoz lenne hivatkozás, az plusz féloldalnyi irodalomjegyzék

\abstract{We describe, and make public, large-scale language resources and the
  toolchain used in their creation, for several medium density European languages}

\maketitleabstract

Using the web as a source of linguistic data is by no means a new idea: the
first efforts in this direction were made over a decade ago (Resnik 1999,
Varantola 2000, Davies 2001)\nocite{Resnik:1999}, and in 2003 {\it
  Computational Linguistics} devoted an entire special issue to the subject
(Kilgarriff and Grefenstette 2003).\nocite{Kilgarriff:2003} It is all the more
surprising that a decade later language resources based on web corpora are,
aside from a handful of major languages, still largely missing. Perhaps
unsurprisingly, the largest European languages (English, German, and French)
fared best, as gigaword corpora and frequency dictionaries have already been
created for these as part of the \texttt{Wacky} project (Baroni and Kilgarriff
2006, Baroni et al. 2009)\nocite{baroni2009wacky}\nocite{baroni2006large}. The
respective corpora \texttt{ukWac, deWac}, and \texttt{frWac} contain billions
of tokens and the derived frequency dictionaries (though not the corpora
themselves) are freely available for download. Another discernible tendency of
the decade is the rise of national corpora (Tadi\'c
2002,\nocite{tadic2002building} Przepi{\'o}rkowski et
al. 2008,\nocite{przepiorkowski2008towards} Kucera
2002).\nocite{kucera2002czech} Remarkably, the fruits of such state- and
EU-funded efforts tend to remain behind query interface walls: individual
results pertaining to individual words may be accessible, but the data as a
whole is not available for download (see Table \ref{table:access}).

Since the fundamental idea is quite sound for medium- and small-density
languages as well, we would have expected there to be a significant selection
of linguistic data accessed by means of crawling. But for most medium European
languages frequency dictionaries derived from gigaword corpora are still
unavailable. Our goal in this paper is to show that a fairly simple pipeline
of entirely generic crawling and only minimally language-specific data
cleaning steps can produce remarkable basic language resources, in particular
corpora and frequency dictionaries, that are considerably larger than those
currently available.
%We also present a toolchain which uses exclusively open-source technology and which will be made available to the public in its entirety.
For seven languages the frequency dictionaries are already publicly accessible
and over two dozen will be available by the time of the final paper. By that
time we shall also release the open-source toolchain we created.

First we describe the process of creating corpora using a combination of
open-source NLP tools. Since the well-known
\texttt{Heritrix}\footnote{\href{http://crawler.archive.org}{http://crawler.archive.org}}
crawler tends to slow down and even halt after several days of operation,
which makes significant human supervision essential to its operation, we use
the less well known but considerably faster and autonomous {\tt wire} crawler
(Castillo et al 2005).\nocite{castillo2005wire} In all fairness, what makes
{\tt heretrix} slow is precisely its excellence in the task it was designed
for, building complete snapshots, and what makes {\tt wire} fast is the
radical pruning of sites (e.g. that no timed-out site is tried a second time),
a strategy only made possible by the fact that all we need are large samples,
completeness is not a goal. In our experience, {\tt wire} will easily sustain
10-20 GB/day throughputs, while {\tt heretrix} will slow down by an order of
magnitude after the first day.

Subsequent stages of the pipeline used in creating the \textit{*Wac} corpora
are described in detail in the papers of Baroni and his co-workers cited
above, but only a few components of the toolchain are publicly
accessible. Therefore, we created our own tools (Hal\'acsy et
al. 2008)\nocite{halacsy2008parallel}, and made sure their runtime is
negligible compared to the time it takes to crawl. By now, our tools are
capable of processing a week's worth of crawl data in a matter of hours, while
the filtering process used in the \texttt{Wacky} project is reported to have
taken several days for each language.  All components of our toolchain are
open source (LGPL) and the packaged pipeline will be freely downloadable by
the time the final paper is submitted.

For seven European languages (Catalan, Czech, Croatian, Dutch, Italian,
Polish, Spanish) we have already gathered enough data to create large-scale
corpora. For these languages, table \ref{table:existing} shows the size of the
largest existing corpora that we know of. Table \ref{table:access} shows the
availability of these resources. Remarkably, only {\tt itWaC} is actually 
available without limitations.

\begin{table}[h]
\centering
\begin{tabular}{|l|c|c|}
\hline
corpus & download & search \\%& URL\\
\hline
CUCWeb & no & yes \\
Czech National Corpus& no & yes \\%& \href{http://ucnk.ff.cuni.cz/}{http://ucnk.ff.cuni.cz/} \\
Croatian National Corpus& no & yes\\% & \href{http://www.hnk.ffzg.hr/}{http://www.hnk.ffzg.hr/} \\
Dutch Parallel Corpus & no & no \\%&\href{http://www.kuleuven-kortrijk.be/DPC}{http://www.kuleuven-kortrijk.be/DPC} \\
itWaC & yes & no \\%& \href{http://wacky.sslmit.unibo.it/}{http://wacky.sslmit.unibo.it/} \\
Polish National Corpus & no & yes\\% &\href{http://nkjp.pl}{http://nkjp.pl} \\
Corpus del Espanol & no & yes \\%&\href{http://www.corpusdelespanol.org/}{http://www.corpusdelespanol.org/} \\
\hline
\end{tabular}
\caption{Corpus access}
\label{table:access}
\end{table}

\begin{table*}[ht]
\centering
\begin{tabular}{|l|l|r|l|l|}
\hline
Language & Largest corpus & tokens & Reference & URL\\
\hline
Catalan & CUCWeb & 166m & Boleda et al. 2006\nocite{boleda2006cucweb} & \href{http://ramsesii.upf.es/cucweb/}{http://ramsesii.upf.es/cucweb/}\\
Czech & Czech National Corpus& 1300m & Kucera 2002\nocite{kucera2002czech} & \href{http://ucnk.ff.cuni.cz/}{http://ucnk.ff.cuni.cz/}\\
Croatian & Croatian National Corpus& 100m & Tadic 2002\nocite{tadic2002building}& \href{http://www.hnk.ffzg.hr/}{http://www.hnk.ffzg.hr/} \\
Dutch & Dutch Parallel Corpus & 10m & Paulussen et al. 2006\nocite{paulussen2006dutch} & \href{http://www.kuleuven-kortrijk.be/DPC}{http://www.kuleuven-kortrijk.be/DPC} \\
Italian & itWaC & 1900m & Baroni et al. 2009& \href{http://wacky.sslmit.unibo.it/}{http://wacky.sslmit.unibo.it/} \\
Polish & Polish National Corpus & 1200m &Przepi{\'o}rkowski et al. 2008\nocite{przepiorkowski2008towards}& \href{http://nkjp.pl}{http://nkjp.pl} \\
Spanish & Corpus del Espanol & 100m & Davies 2001\nocite{davies2001creating}& \href{http://www.corpusdelespanol.org/}{http://www.corpusdelespanol.org/}\\
\hline
\end{tabular}
\caption{Existing corpora}
\label{table:existing}
\end{table*}

In our own work, first we collected HTML documents from the relevant top-level domains
(\texttt{.cat}, \texttt{.cz}, \texttt{.hr}, etc.). The data obtained was then
converted to raw text, split into sentences and tokenized. For language
detection we used Gertjan van Noord’s n-gram language classifier
\texttt{textcat}\footnote{\href{http://www.let.rug.nl/\~vannoord/TextCat}{http://www.let.rug.nl/$\sim$vannoord/TextCat}}
To further clean our data, we passed each set of documents through the
\texttt{hunspell} (Németh et al. 2004)\nocite{nemeth2004leveraging}
spellchecker and discarded those in which the ratio of misspelled words were
above some language-specific threshold. This kind of filtering is in itself an
efficient method for language identification and also reduces all kinds of
noise in raw text obtained by crawling the web. This produced corpora with
sizes in the hundred millions (see table \ref{table:size}). Since further
crawling of the web is currently in progress, we shall most probably provide
gigaword corpora in all of these, as well as another 12-15 languages.

%The number of tokens and lemmas for each language can be seen in table \ref{table:size}.

\begin{table}
\centering
\begin{tabular}{|l|r|r|}
\hline
Language & tokens & lemmas\\
\hline
Catalan & 492m & 62k\\
Czech & 655m & 189k\\
Croatian & 382m & 168k\\
Dutch & 463m & 95k\\
Italian & 944m & 79k\\
Polish & 366m & 152k\\
Spanish & 528m & 56k\\
\hline
\end{tabular}
\caption{Corpus sizes by language}
\label{table:size}
\end{table}

The next step involved counting the number of occurences of each token and
thus creating a frequency dictionary of all word forms of a language that are
present in the data gathered. We then passed the lists through
\texttt{hunspell} for the purpose of stemming and summed the frequency of
stems in order to create a frequency dictionary of lemmas.

In average, \texttt{hunspell} found some 10 percent of all word forms
ambiguous and returned multiple stems. The frequency dictionaries we created
contain two figures for each stem: one was obtained by choosing the shortest
possible stem for each ambiguous form, while the other is the result of
summing the frequencies of all tokens for which the given stem was among the
options returned by \texttt{hunspell}. Both the frequency dictionaries and the
tokenized corpora are available for download at
\href{http://hlt.sztaki.hu/resources}{http://hlt.sztaki.hu/resources}. The
ratio of tokens unrecognized by \texttt{hunspell} can be seen in table
\ref{table:recognized}.

\begin{table}
\centering
\begin{tabular}{|l|r|r|r|}
\hline
Language & tokens & unknown & ratio\\
\hline
Catalan & 492m & 84m & $17.04\%$\\
Czech & 655m & 79k & $12.00\%$ \\
Croatian & 383m & 74m & $19.24\%$\\
Dutch & 463m & 58m & $12.50\%$\\
Italian & 944m & 131k & $13.83\%$\\
Polish & 366m & 55m & $14.93\%$\\
Spanish & 528m & 72m & $13.61\%$\\
\hline
\end{tabular}
\caption{Ratio of unrecognized word forms}
\label{table:recognized}
\end{table}

The pipeline we have created enables us to efficiently create good quality
corpora and frequency dictionaries as newer data is crawled. For Catalan, 
Croatian, Dutch, and Spanish we have already created the largest freely
downloadable corpus, and in the final paper we hope to present figures that 
improve on the current state of the art for several more languages.

%In order to create good quality corpora from the available text we passed
%each set of documents through the \texttt{hunspell} spellchecker and
%discarded those in which the ratio of misspelled words were above some
%language-specific threshold. This kind of filtering is an efficient method
%for language identification and reduces all kinds of noise in raw text
%obtained by crawling the web. The size of the corpora thus obtained can be
%seen in table \ref{table:cleanedSize}.

%These figures show that the corpora obtained are larger than the Croatian

%Ezért lesz különösen kurvajó a gyakorisági szótár, ezért lemmatizálunk! Olykor morfológiailag elemzünk! Végül konklúzió!

\bibliographystyle{lrec2006}
\bibliography{webcorp}{}

\end{document}

%Catalan & 490901211 & \\
%Czech & & \\
%Croatian & 382167628 & \\
%Dutch & 463162799 & \\
%Italian & & \\
%Polish & 365751955 & \\
%Spanish & 528249527 & \\

### 2012-02-08 paper/engNERwiki.tex
Started writing my part.

%
% File acl2012.tex
%
% Contact: Maggie Li (cswjli@comp.polyu.edu.hk), Michael White (mwhite@ling.osu.edu)
%%
%% Based on the style files for ACL2008 by Joakim Nivre and Noah Smith
%% and that of ACL2010 by Jing-Shin Chang and Philipp Koehn

\documentclass[11pt]{article}
\usepackage{acl2012}
\usepackage{times}
\usepackage{latexsym}
\usepackage{amsmath}
\usepackage{multirow}
\usepackage{url}
\usepackage[utf8]{inputenc}
\DeclareMathOperator*{\argmax}{arg\,max}
\setlength\titlebox{6.5cm}    % Expanding the titlebox

\title{XXX}

%\author{Dávid Márk Nemeskey \\
%  Research Institute for Computer Science and Automation \\
%  Hungarian Academy of Sciences \\
%  H-1111 Lágymányosi út 11, Budapest \\
%  {\tt nemeskey.david@sztaki.hu} \\\And
%  Eszter Simon \\
%  Research Institute for Linguistics \\
%  Hungarian Academy of Sciences \\
%  H-1068 Benczúr utca 33, Budapest \\
%  {\tt eszter@nytud.hu} \\}

\date{}

\begin{document}
\maketitle
\begin{abstract}
  This document contains the instructions for preparing a camera-ready manuscript for the proceedings of ACL2012. The document itself conforms to its own specifications, and is therefore an example of what your manuscript should look like. These instructions should be used for both papers submitted for review and for final versions of accepted papers. Authors are asked to conform to all the directions reported in this document.
\end{abstract}

\section{Introduction}

One of the main trends in NLP is the need for reducing the annotation cost. After examining the calls for papers of nowadays conferences and workshops there seems to be three major directions to achieve this aim.

The first approach is the application of unsupervised and semi-supervised learning techniques. Researchers following this approach try to not only reduce, but avoid the annotation cost with minimizing the need for annotated data. These methods can be applied to any language or genre for which adequate raw text resources are available.

The second approach is using collaborative annotation and/or collaboratively constructed resources, such as Wikipedia, Wiktionary or Linked Open Data. Researchers apply the Web-as-corpus approach, often dealing with such datasets as the Google 5G 1T collection or DBPedia. Thanks for the huge amount of Wikipedia articles, the possibility of building large enough corpora is given for less resourced languages, as well.

The third approach to reduce the manual intervention required in language resource production, and thus ultimately production costs is automatically generating resources or merging existing ones. However merging language resources faces NLP researchers with the problem of combining different annotation schemas.

Named Entity Recognition (NER) is one of the important subtasks of Information Extraction. NER contains two main substeps: recognizing the Named Entities (NEs) in unstructured text and classifying them into preselected classes

mostly newspaper articles, domain-specific

In this article we present a Named Entity tagged corpus for English -- fully automatically generated from Wikipedia. 
The applied annotation schema is the CoNLL tagset, which is one of the major standards applied in the field of Named Entity Recognition (NER).

The paper is structured as follows.

\section{Related Work}
% Ide kell? Vagy Conclusion elé?

\section{Creation of the Corpus}  % Dzsííízösz
XXX: past or present tense?
% Introduction
We decided to use Wikipedia because

The categories of Wikipedia are too numerous and ; not a tree;

Wikipedia proves to be a rich source of information about its entities; aside from the article text, a huge amount of data is encompassed in infoboxes, templates and the category structure. Since our goal is to create a NER training corpus, our interest is limited to the article text. Templates are removed from the articles; however, the names of the templates have been retained for , but all data (including infobox)

Beside regular articles, Wikipedia also contains redirect and disambiguation pages, which require special treatment. Redirect pages are empty and can be safely ignored altogether. Disambiguation pages contain text, usually in the form of an enumeration of page links with a short description. Full sentences are scarce and the structure of these pages makes it hard to extract valid training sentences from them. Therefore, disambiguation pages have been dropped from the corpus as well.

The corpus is based on the Wikipedia snapshot as of XXX. The xml files were parsed with the XXX (Attila) parser. The raw text was tokenized with a modified version of the XXX sentence tokenizer in NLTK (CITE). The data was then converted into the tab separated format used by ConLL (XXX, CITE).

Though not included in the corpus, redirect pages can improve recall of the gold standard (XXX). 
% 1. Parsing Wikipedia (hunpos)
% TODO: redirect redirect pages

\subsection{The DBpedia -- ConLL mapping}

The entities in DBpedia are organized into a type hierarchy, available as an OWL (CITE) ontology. The taxonomy 
% description of the DBpedia ontology
On the surface, this hierarchy seems the parallel the four categories (\texttt{PER, LOC, ORG and MISC}) XXX. However, closer inspection reveals that the similarity is superficial. The ConLL NER annotation guide (CITE?: http://www.cnts.ua.ac.be/conll2003/ner/annotation.txt)
The full mapping can be found in LINK Appendix A (XXX or not).

% 3. Training sentence extraction
%   a. only known links
%   b. title parts
%   c. parts of links mention earlier
%   d. date regexp
% 4. Test sentence extraction?
%   a. unknown links, no unlinked names
%   b. might be good for quality evaluation / recursive tagging (next paper?)
% 5. teaching the original DBpedia categories?

\section{Creation of the Hungarian Corpus}
% Inconsistent template names, esp. across languages
% Comparison of the Hungarian DBpedia vs English + mapping?

\section{Evaluation}

\section{Conclusions and Future Work}

\section{General Instructions}

Manuscripts must be in two-column format. Exceptions to the two-column format include the title, 
authors' names and complete addresses, which must be centered at the top of the first page, 
and any full-width figures or tables (see the guidelines in Subsection~\ref{ssec:first}). {\bf Type single-spaced}. 
Start all pages directly under the top margin. See the guide-lines later regarding formatting the first page. Do not number the pages.

\subsection{Electronically-available resources}

ACL2012 provides this description in \LaTeX2e (acl2012.tex) and PDF format (acl2012.pdf), along with the LATEX2e style file used to format it (acl2012.sty) and an ACL bibliography style (acl.bst). These files are all available at \url{http://www.acl2012.org}.  A Microsoft Word template file (acl2012.dot) is also available at the same URL. We strongly recommend the use of these style files, which have been appropriately tailored for the ACL2012 proceedings. If you have an option, we recommend that you use the \LaTeX2e version. \textbf{If you will be using the Microsoft Word template, we suggest that you anonymize your source file so that the pdf produced does not retain your identity.} This can be done by removing any personal information from your source
document properties.

\subsection{Format of Electronic Manuscript}
\label{sect:pdf}

For the production of the electronic manuscript you must use Adobe's
Portable Document Format (PDF). This format can be generated from
postscript files: on Linux/Unix systems, you can use {\tt ps2pdf} for this
purpose; under Microsoft Windows, you can use Adobe's Distiller, or
if you have {\tt cygwin} installed, you can use {\tt dvipdf} or
{\tt ps2pdf}.  Note
that some word processing programs generate PDF which may not include
all the necessary fonts (esp. tree diagrams, symbols). When you print
or create the PDF file, there is usually an option in your printer
setup to include none, all or just non-standard fonts.  Please make
sure that you select the option of including ALL the fonts.  {\em Before sending it, test your PDF by printing it from a computer different from the one where it was created}. Moreover,
some word processor may generate very large postscript/PDF files,
where each page is rendered as an image. Such images may reproduce
poorly.  In this case, try alternative ways to obtain the postscript
and/or PDF.  One way on some systems is to install a driver for a
postscript printer, send your document to the printer specifying
``Output to a file'', then convert the file to PDF.

Additionally, it is of utmost importance to specify the {\bf US-Letter format} (8.5in $\times$ 11in) when formatting the paper. When working with {\tt dvips}, for instance, one should specify {\tt -t letter}.

Print-outs of the PDF file on US-Letter paper should be identical to the
hardcopy version.  If you cannot meet the above requirements about the
production of your electronic submission, please contact the
publication chair above as soon as possible.

\subsection{Layout}
\label{ssec:layout}

Format manuscripts two columns to a page, in the manner these
instructions are formatted. The exact dimensions for a page on US-letter
paper are:

\begin{itemize}
\item Left and right margins: 1in
\item Top margin:1in
\item Bottom margin: 1in
\item Column width: 3.15in
\item Column height: 9in
\item Gap between columns: 0.2in
\end{itemize}

\noindent Papers should not be submitted on any other paper size. If you cannot meet the above requirements about the production of your electronic submission, please contact the publication chair above as soon as possible.

\subsection{Fonts}

For reasons of uniformity, Adobe's {\bf Times Roman} font should be
used. In \LaTeX2e{} this is accomplished by putting

\begin{quote}
\begin{verbatim}
\usepackage{times}
\usepackage{latexsym}
\end{verbatim}
\end{quote}
in the preamble. If Times Roman is unavailable, use {\bf Computer
  Modern Roman} (\LaTeX2e{}'s default).  Note that the latter is about
  10\% less dense than Adobe's Times Roman font.

\begin{table}[h]
\begin{center}
\begin{tabular}{|l|rl|}
\hline \bf Type of Text & \bf Font Size & \bf Style \\ \hline
paper title & 15 pt & bold \\
author names & 12 pt & bold \\
author affiliation & 12 pt & \\
the word ``Abstract'' & 12 pt & bold \\
section titles & 12 pt & bold \\
document text & 11 pt  &\\
captions & 11 pt & \\
abstract text & 10 pt & \\
bibliography & 10 pt & \\
footnotes & 9 pt & \\
\hline
\end{tabular}
\end{center}
\caption{\label{font-table} Font guide. }
\end{table}

\subsection{The First Page}
\label{ssec:first}

Center the title, author's name(s) and affiliation(s) across both
columns. Do not use footnotes for affiliations.  Do not include the
paper ID number assigned during the submission process.
Use the two-column format only when you begin the abstract.

{\bf Title}: Place the title centered at the top of the first page, in
a 15 point bold font.  (For a complete guide to font sizes and styles, see Table~\ref{font-table}.)
Long title should be typed on two lines without
a blank line intervening. Approximately, put the title at 1in from the
top of the page, followed by a blank line, then the author's names(s),
and the affiliation on the following line.  Do not use only initials
for given names (middle initials are allowed). Do not format surnames
in all capitals (e.g., ``Zhou,'' not ``ZHOU'').  The affiliation should
contain the author's complete address, and if possible an electronic
mail address. Leave about 0.75in between the affiliation and the body
of the first page. The title, author names and addresses should be completely identical to those entered to the electronic paper submission website in order to maintain the consistency of author information among all publications of the conference.

{\bf Abstract}: Type the abstract at the beginning of the first
column.  The width of the abstract text should be smaller than the
width of the columns for the text in the body of the paper by about
0.25in on each side.  Center the word {\bf Abstract} in a 12 point
bold font above the body of the abstract. The abstract should be a
concise summary of the general thesis and conclusions of the paper.
It should be no longer than 200 words. The abstract text should be in 10 point font.

{\bf Text}: Begin typing the main body of the text immediately after
the abstract, observing the two-column format as shown in
the present document. Do not include page numbers.

{\bf Indent} when starting a new paragraph. For reasons of uniformity,
use Adobe's {\bf Times Roman} fonts, with 11 points for text and
subsection headings, 12 points for section headings and 15 points for
the title.  If Times Roman is unavailable, use {\bf Computer Modern
  Roman} (\LaTeX2e's default; see section \ref{sect:pdf} above).
Note that the latter is about 10\% less dense than Adobe's Times Roman
font.

\subsection{Sections}

{\bf Headings}: Type and label section and subsection headings in the
style shown on the present document.  Use numbered sections (Arabic
numerals) in order to facilitate cross references. Number subsections
with the section number and the subsection number separated by a dot,
in Arabic numerals. Do not number subsubsections.

{\bf Citations}: Citations within the text appear
in parentheses as~\cite{Gusfield:97} or, if the author's name appears in
the text itself, as Gusfield~\shortcite{Gusfield:97}. Append lowercase letters to the year in cases of ambiguities. Treat double authors as in~\cite{Aho:72}, but write as in~\cite{Chandra:81} when more than two authors are involved. Collapse multiple citations as in~\cite{Gusfield:97,Aho:72}. Also refrain from using full citations as sentence constituents. We suggest that instead of
\begin{quote}
  ``\cite{Gusfield:97} showed that ...''
\end{quote}
you use
\begin{quote}
``Gusfield \shortcite{Gusfield:97}   showed that ...''
\end{quote}

If you are using the provided \LaTeX{} and Bib\TeX{} style files, you
can use the command \verb|\newcite| to get ``author (year)'' citations.

As reviewing will be double-blind, the submitted version of the papers should not include the
authors' names and affiliations. Furthermore, self-references that
reveal the author's identity, e.g.,
\begin{quote}
``We previously showed \cite{Gusfield:97} ...''
\end{quote}
should be avoided. Instead, use citations such as
\begin{quote}
``Gusfield \shortcite{Gusfield:97}
previously showed ... ''
\end{quote}

Please do not  use anonymous
citations and  do not include acknowledgements when submitting your papers. Papers that do not conform
to these requirements may be rejected without review.

\textbf{References}: Gather the full set of references together under
the heading {\bf References}; place the section before any Appendices,
unless they contain references. Arrange the references alphabetically
by first author, rather than by order of occurrence in the text.
Provide as complete a citation as possible, using a consistent format,
such as the one for {\em Computational Linguistics\/} or the one in the
{\em Publication Manual of the American
Psychological Association\/}~\cite{APA:83}.  Use of full names for
authors rather than initials is preferred.  A list of abbreviations
for common computer science journals can be found in the ACM
{\em Computing Reviews\/}~\cite{ACM:83}.

The \LaTeX{} and Bib\TeX{} style files provided roughly fit the
American Psychological Association format, allowing regular citations,
short citations and multiple citations as described above.

{\bf Appendices}: Appendices, if any, directly follow the text and the
references (but see above).  Letter them in sequence and provide an
informative title: {\bf Appendix A. Title of Appendix}.

\textbf{Acknowledgment} sections should go as a last section immediately
before the references. Do not number the acknowledgement section.

\subsection{Footnotes}

{\bf Footnotes}: Put footnotes at the bottom of the page. They may
be numbered or referred to by asterisks or other
symbols.\footnote{This is how a footnote should appear.} Footnotes
should be separated from the text by a line.\footnote{Note the
line separating the footnotes from the text.}  Footnotes should be in 9 point font.

\subsection{Graphics}

{\bf Illustrations}: Place figures, tables, and photographs in the
paper near where they are first discussed, rather than at the end, if
possible.  Wide illustrations may run across both columns and should be placed at
the top of a page. Color illustrations are discouraged, unless you have verified that
they will be understandable when printed in black ink.

{\bf Captions}: Provide a caption for every illustration; number each one
sequentially in the form:  ``Figure 1. Caption of the Figure.'' ``Table 1.
Caption of the Table.''  Type the captions of the figures and
tables below the body, using 10 point text.

\section{Translation of non-English Terms}

It is also advised to supplement non-English characters and terms
with appropriate transliterations and/or translations
since not all readers understand all such characters and terms.

Inline transliteration or translation can be represented in
the order of: original-form transliteration ``translation''.

\section{Length of Submission}
\label{sec:length}

Long papers may consist of up to eight (8) pages of content and an
additional two (2) pages of references, and short papers may consist
of up to four (4) pages of content and two (2) additional pages of
references.  Papers that do not conform to the specified length and
formatting requirements are subject to re-submission.

\section{Other Issues}

Those papers that had software and/or dataset submitted for the review process, should also submit it 
with the camera-ready paper. Besides, the software and/or dataset should not be anonymous.

Please note that the publications of ACL2012 will be publicly available at ACL Anthology 
(http://aclweb.org/anthology-new/) on July 2nd, 2012, one week before the start of the conference. 
Since some of the authors may have plans to file patents related to their papers in the conference, 
we are sending this reminder that July 2nd, 2012 may be considered to be the official publication date, 
instead of the opening day of the conference.

%\section*{Acknowledgments}

%The authors are grateful to Attila Zséder and Gábor Recski for their respective work on Wikipedia parsing and {\tt hunner}.
%Do not number the acknowledgment section. Do not include this section when submitting your paper for review.

\begin{thebibliography}{}

\bibitem[\protect\citename{Aho and Ullman}1972]{Aho:72}
Alfred~V. Aho and Jeffrey~D. Ullman.
\newblock 1972.
\newblock {\em The Theory of Parsing, Translation and Compiling}, volume~1.
\newblock Prentice-{Hall}, Englewood Cliffs, NJ.

\bibitem[\protect\citename{{American Psychological Association}}1983]{APA:83}
{American Psychological Association}.
\newblock 1983.
\newblock {\em Publications Manual}.
\newblock American Psychological Association, Washington, DC.

\bibitem[\protect\citename{{Association for Computing Machinery}}1983]{ACM:83}
{Association for Computing Machinery}.
\newblock 1983.
\newblock {\em Computing Reviews}, 24(11):503--512.

\bibitem[\protect\citename{Chandra \bgroup et al.\egroup }1981]{Chandra:81}
Ashok~K. Chandra, Dexter~C. Kozen, and Larry~J. Stockmeyer.
\newblock 1981.
\newblock Alternation.
\newblock {\em Journal of the Association for Computing Machinery},
  28(1):114--133.

\bibitem[\protect\citename{Gusfield}1997]{Gusfield:97}
Dan Gusfield.
\newblock 1997.
\newblock {\em Algorithms on Strings, Trees and Sequences}.
\newblock Cambridge University Press, Cambridge, UK.

\end{thebibliography}

\end{document}

### 2012-02-09 paper/engNERwiki.tex
A few more lines (chapter 2 introduction + parsing).

@@
 \section{Related Work}
+% Wikify! 
 % Ide kell? Vagy Conclusion elé?
@@
 \section{Creation of the Corpus}  % Dzsííízösz
+
+% Introduction: Wikipedia + DBpedia
+The corpora are available in the tab separated format used by ConLL (XXX, CITE).
+In this format, one line contains one word, its linguistic features and the named entity tags.
+Sentence boundaries are marked by empty lines.
+The linguistic features include the lemmatized form of the word, part-of-speech tag and chunk tag.
+Two named entity tags are included with each word: one is the most specific DBpedia category it belongs to, if any; the other is the ConLL named entity tag.
+While the named entity tags can be considered "gold", the linguistic features are provided on a "best-effort" basis (XXX: check the LREC paper to find a better wording for this).
 XXX: past or present tense?
-% Introduction
+
 We decided to use Wikipedia because
@@
 Since our goal is to create a NER training corpus, our interest is limited to the article text.
-Templates are removed from the articles; however, the names of the templates have been retained for , but all data (including infobox)
+Templates are removed from the articles; however, the names of the templates have been retained for XXX, but all data (including infobox)
 
 Beside regular articles, Wikipedia also contains redirect and disambiguation pages, which require special treatment.
-Redirect pages are empty and can be safely ignored altogether.
+Redirect pages are empty and are safe to ignore altogether.
 Disambiguation pages contain text, usually in the form of an enumeration of page links with a short description.
@@
 The raw text was tokenized with a modified version of the XXX sentence tokenizer in NLTK (CITE).
-The data was then converted into the tab separated format used by ConLL (XXX, CITE).
+The words were lemmatized with XXX and POS tagged by HunPos (CITE).
+Chunking was performed by HunChunk (CITE).
+Experiments have shown that HunPos obtains STAT f-score on the Brown corpus; HunChunk performs at around STAT, CITE?.
 
 Though not included in the corpus, redirect pages can improve recall of the gold standard (XXX). 
-% 1. Parsing Wikipedia (hunpos)
 % TODO: redirect redirect pages

### 2012-02-09 paper/engNERwiki.tex
A few more lines.

@@
 
-The entities in DBpedia are organized into a type hierarchy, available as an OWL (CITE) ontology. The taxonomy 
+The classes in DBpedia are organized into a type hierarchy, available as an OWL (CITE) ontology.
+The ontology contains the 350 most frequent categories of Wikipedia, arranged into a taxonomy tree under the base class \texttt{owl:Thing}.
+Most of the classes belong into the 5 largest sub-hierarchies: \texttt{Agent} (which is further divided into \texttt{Person} and \texttt{Organisation}), \texttt{Event}, \texttt{Place}, \texttt{Species} and \texttt{Work}.
+The taxonomy is rather flat: the top level contains 44 classes and there are several nodes with a branching factor of 20.
 % description of the DBpedia ontology
-On the surface, this hierarchy seems the parallel the four categories (\texttt{PER, LOC, ORG and MISC}) XXX.
+On the surface, this hierarchy seems to parallel the four categories (\texttt{PER, LOC, ORG and MISC}) XXX.
 However, closer inspection reveals that the similarity is superficial.
@@
 The full mapping can be found in LINK Appendix A (XXX or not).
+
+\subsection{Extraction of the Training Sentences}
 

### 2012-02-12 scripts/dbpedia_class.py
Fixed dbpedia_class to use only DBpedia classes.

@@ -32,7 +32,7 @@ class OwlClassHierarchy(xml.sax.handler.ContentHandler):
                 self.cls = OwlClassHierarchy._strip_uri(uri)
         elif name == 'rdfs:subClassOf':
             uri = attrs.get('rdf:resource')
-            if self.cls is not None and uri is not None:
+            if self.cls is not None and uri is not None and uri.startswith('http://dbpedia.org/ontology'):
                 self.super_cls.append(OwlClassHierarchy._strip_uri(uri))
 
     def endElement(self, name):
@@ -117,7 +117,7 @@ if __name__ == '__main__':
         sys.exit()
 
     handler = OwlClassHierarchy(sys.argv[1])
-#    print handler.get_sorted()
     handler.create_isa_map()
     print handler.is_a('University', 'Organisation')
+#    print handler.get_sorted()
 

### 2012-02-12 scripts/dbpedia_class.py
Added a method that draws a graph of the ontology.

@@ -5,6 +5,8 @@ import sys
 import re
 import copy
 import xml.sax.handler
+import networkx as nx
+import matplotlib.pyplot as plt
 
 class OwlClassHierarchy(xml.sax.handler.ContentHandler):
     URI_PATTERN = re.compile(r'http:.*/(.+)')
@@ -102,6 +104,20 @@ class OwlClassHierarchy(xml.sax.handler.ContentHandler):
             self.sorted = dict((t[1], t[0]) for t in enumerate(self.sorted))
         return self.sorted
 
+    def draw_graph(self):
+        """Draws a graph of the ontology classes. Again, use only for small
+        ontologies."""
+        G = nx.DiGraph()
+        G.add_nodes_from(self.tree.keys())
+        for cls, parents in self.tree.iteritems():
+            for parent in parents:
+                G.add_edge(cls, parent)
+        pos = nx.spring_layout(G, scale=0.1)
+        nx.draw_networkx_nodes(G, pos)
+        nx.draw_networkx_edges(G, pos, width=1.0, alpha=0.5)
+        nx.draw_networkx_labels(G, pos, font_weight='bold')
+        plt.show()
+
     @staticmethod
     def _strip_uri(uri):
         m = OwlClassHierarchy.URI_PATTERN.match(uri)
@@ -120,4 +136,9 @@ if __name__ == '__main__':
     handler.create_isa_map()
     print handler.is_a('University', 'Organisation')
 #    print handler.get_sorted()
+#    handler.draw_graph()
+#    for child, parents in handler.tree.iteritems():
+#        if len(parents) > 1:
+#            print child, parents
 
+#    print 'Library', handler.isa_map['Library']

### 2012-02-12 paper/engNERwiki.tex
Wrote quite a few lines to the article.

@@
 
+% Introduction + entity -> class -> category.
+
 The classes in DBpedia are organized into a type hierarchy, available as an OWL (CITE) ontology.
-The ontology contains the 350 most frequent categories of Wikipedia, arranged into a taxonomy tree under the base class \texttt{owl:Thing}.
-Most of the classes belong into the 5 largest sub-hierarchies: \texttt{Agent} (which is further divided into \texttt{Person} and \texttt{Organisation}), \texttt{Event}, \texttt{Place}, \texttt{Species} and \texttt{Work}.
+The ontology contains the 320 most frequent categories of Wikipedia, arranged into a taxonomy under the base class \texttt{owl:Thing}.
+Most of the classes belong into the 5 largest sub-hierarchies (here named after their topmost classes XXX): \texttt{Agent} (which is further divided into \texttt{Person} and \texttt{Organisation}), \texttt{Event}, \texttt{Place}, \texttt{Species} and \texttt{Work}.
 The taxonomy is rather flat: the top level contains 44 classes and there are several nodes with a branching factor of 20.
-% description of the DBpedia ontology
-On the surface, this hierarchy seems to parallel the four categories (\texttt{PER, LOC, ORG and MISC}) XXX.
-However, closer inspection reveals that the similarity is superficial.
-The ConLL NER annotation guide (CITE?: http://www.cnts.ua.ac.be/conll2003/ner/annotation.txt)
+
+It is not difficult to see the parallels XXX between the DBpedia sub-hierarchies \texttt{Person}, \texttt{Organisation} and \texttt{Place} and the ConLL NER categories \texttt{PER}, \texttt{ORG} and \texttt{LOC}.
+The fourth category, MISC is more elusive; according to the ConLL NER annotation guide (CITE?: http://www.cnts.ua.ac.be/conll2003/ner/annotation.txt), the sub-hierarchies \texttt{Event} and \texttt{Work} belong to this category, as well as various other class outside the main sub-hierarchies.
+
+While the correspondence described above holds for most classes in the XXX sub-hierarchies, there are some exceptions.
+For instance, the class \texttt{SportsLeague} is part of the \texttt{Organisation} sub-hierarchy, but in the ConLL corpus, sport leagues are tagged as MISC.
+Similarly, the class \texttt{SportsTeamSeason} is under \texttt{Event}, by which account it would be classified as MISC; however, ConLL does not recognize such entities.
+To account for these misclassification, as well as the miscellaneous classes outside the main hierarchies, we have introduced a mapping mechanism.
+We created a file of DBpedia class -- NER category mappings.
+Whenever an entity is evaluated, we look up its class, as well as the ancestors of its class, in the mapping, and assign to the entity the category of the class that matches it most closely -- i.e.
+the category of the entity's class, if it is in the list, or that of its closest ancestor.
+XXX!!!
+If no match is found, the entity is tagged with \texttt{0}.
+Since we take advantage of the inheritence hierarchy, the mapping list remains short: it contains only the root classes of the main sub-hierarchies, exceptions of the like XXX mentioned above and the various classes that belong to the MISC category according to the ConLL annotation guideline.
+
+As of version 3.7, the DBpedia ontology allows multiple superclasses, making the hierarchy a directed acyclic graph.
+This introduces the problem of selecting the right superclass, and hence, ConLL mapping for classes with more than one parent.
+In version 3.7, the only such class is \textit{Library}, which can be traced back to both \textit{Place} and \textit{Organisation}.
+There are two ways to tackle this problem.
+The first is to follow the DBpedia guidelines and only retain the first parent.
+The second method is manually deciding the right NER mapping for the ambiguous XXX class and adding it to the mapping file.
+The first approach, unfortunately, already fails to assign the correct ConLL NER tag to our only example.
+The reason behind this failure is that multiple parents in an ontology show an inherent ambiguity in the class in question: it is entirely natural to think of a library as a location as well as a public organization.
+DBpedia's guidelines prefers Organisation; ConLL's LOC.
+Therefore, only the second, manual approach ensures correct tagging.
+XXX
+
 The full mapping can be found in LINK Appendix A (XXX or not).
@@
 \subsection{Extraction of the Training Sentences}
+
+The basic unit of a NER training corpus is the sentence. In this section, we describe how training sentences are extracted from the preprocessed Wikipedia pages.
 
@@
 \section{Creation of the Hungarian Corpus}
+
+% Good for middle-sized languages
 % Inconsistent template names, esp. across languages
-% Comparison of the Hungarian DBpedia vs English + mapping?
+% Comparison of the Hungarian DBpedia vs English + mapping
 
 \section{Evaluation}
+
+% vs. ConLL / Szeged
+% Corpus sizes, performance, tagging conventions
 

### 2012-02-13 paper/engNERwiki.tex
Swapped the order of two subsections.
Just ****ing around...

@@
 The raw text was tokenized with a modified version of the XXX sentence tokenizer in NLTK (CITE).
-The words were lemmatized with XXX and POS tagged by HunPos (CITE).
-Chunking was performed by HunChunk (CITE).
+The words were lemmatized with XXX and POS tagged with HunPos (CITE).
+Chunking was performed with HunChunk (CITE).
 Experiments have shown that HunPos obtains STAT f-score on the Brown corpus; HunChunk performs at around STAT, CITE?.
 
+\subsection{Extraction of the Training Sentences}
+
+The basic unit of a NER training corpus is the sentence. In this section, we describe how training sentences are extracted from the preprocessed Wikipedia pages.
+
+\subsubsection{Sentence selection}
+
+A NER training corpus consists of sentences that contain "gold" quality information about the named entities (or lack thereof) occuring in them: XXX In order to achieve
+
+The text in Wikipedia is 
+Unknown entities might or might not be
+
+whose validity is
+
+Some  Resolving these links results in a STAT increase in recall.
 Though not included in the corpus, redirect pages can improve recall of the gold standard (XXX). 
 % TODO: redirect redirect pages
+
+% 3. Training sentence extraction
+%   a. only known links
+%   b. title parts
+%   c. parts of links mention earlier
+%   d. date regexp
+% recall+precision?
+
+\subsubsection*{Filtering}
+
+Some sentences that were selected by the process above
+
+% 4. Test sentence extraction?
+%   a. unknown links, no unlinked names
+%   b. might be good for quality evaluation / recursive tagging (next paper?)
+% 5. teaching the original DBpedia categories?
 
@@
 The full mapping can be found in LINK Appendix A (XXX or not).
-
-\subsection{Extraction of the Training Sentences}
-
-The basic unit of a NER training corpus is the sentence. In this section, we describe how training sentences are extracted from the preprocessed Wikipedia pages.
-
-% 3. Training sentence extraction
-%   a. only known links
-%   b. title parts
-%   c. parts of links mention earlier
-%   d. date regexp
-% 4. Test sentence extraction?
-%   a. unknown links, no unlinked names
-%   b. might be good for quality evaluation / recursive tagging (next paper?)
-% 5. teaching the original DBpedia categories?
 

### 2012-02-14 paper/engNERwiki.tex
A few more sections added.

--- 2012-02-12 paper/engNERwiki.tex
@@
 The raw text was tokenized with a modified version of the XXX sentence tokenizer in NLTK (CITE).
-The words were lemmatized with XXX and POS tagged by HunPos (CITE).
-Chunking was performed by HunChunk (CITE).
+The words were lemmatized with XXX and POS tagged with HunPos (CITE).
+Chunking was performed with HunChunk (CITE).
 Experiments have shown that HunPos obtains STAT f-score on the Brown corpus; HunChunk performs at around STAT, CITE?.
 
-Though not included in the corpus, redirect pages can improve recall of the gold standard (XXX). 
-% TODO: redirect redirect pages
+\subsection{Extraction of the Training Sentences}
+
+A NER training corpus consists of sentences that accurately tag named entities occurring in them.
+In order to automatically prepare such sentences, two tasks need to be performed: identification of entities in the sentence and tagging them with the correct tag.
+Sentences, for which accurate tagging could not be accomplished, must be removed from the corpus.
+In this section, we describe how these steps were implemented, as well as the post-filtering we employed to dispose of correctly tagged but unusable training sentences.
+
+% TODO: we want to be on the strict side!
+
+\subsubsection{Entity identification}
+
+Our approach to entity identification relies on Wikipedia cross-references found in the article text.
+We assume that individual Wikipedia articles describe named entities.
+A link to an article can then be perceived as a mapping that identifies its anchor text with a particular entity.
+There are exceptions to this rule, however.
+Disambiguation pages are used when the page title is ambiguous, and can refer to several entities.
+If a link points to a disambiguation page, we acknowledge the anchor text as an entity, but mark it unknown.
+A similar measure is taken if the link target XXX does not exist.
+Links to redirect pages are resolved to point instead to the redirect target, after which they are handled as regular cross-references.
+
+Not all entities are marked with links, however.
+Authors may neglect to add XXX references to their text.
+Furthermore, in a Wikipedia article XXX, typically only the first mention of a particular entity is linked to the corresponding page.
+Subsequent mentions are unmarked and often incomplete -- e.g.
+family names are used instead of full names.
+
+% Numbers, years.
+
+\subsubsection{Tagging}
+
+We use DBpedia as
+
+Only sentences for whose entities the correct category could be obtained from DBpedia are added to the corpus; those that contain unknown entities are discarded.
+Sentences without any entities are retained as well, so that the corpus mirrors (XXX the distribution of the entities in the corpus mirrors their distribution in Wikipedia?) the distribution of entities in Wikipedia as closely as possible; this is also in line with how the ConLL corpus is assembled.
+(XXX or is it)
+
+Unknown entities might or might not be
+
+% TODO: monetary, etc. units?
+
+% 3. Training sentence extraction
+%   a. only known links
+%   b. title parts
+%   c. parts of links mention earlier
+%   d. date regexp
+% recall+precision?
+
+\subsubsection{Filtering}
+
+Some sentences that were selected by the process above
+
+% 4. Test sentence extraction?
+%   a. unknown links, no unlinked names
+%   b. might be good for quality evaluation / recursive tagging (next paper?)
+% 5. teaching the original DBpedia categories?
 
@@
 If no match is found, the entity is tagged with \texttt{0}.
-Since we take advantage of the inheritence hierarchy, the mapping list remains short: it contains only the root classes of the main sub-hierarchies, exceptions of the like XXX mentioned above and the various classes that belong to the MISC category according to the ConLL annotation guideline.
+Since we take advantage of the inheritance hierarchy, the mapping list remains short: it contains only the root classes of the main sub-hierarchies, exceptions of the like XXX mentioned above and the various classes that belong to the MISC category according to the ConLL annotation guideline.
 
@@
 
-\subsection{Extraction of the Training Sentences}
+\subsection{Remarks}
 
-The basic unit of a NER training corpus is the sentence. In this section, we describe how training sentences are extracted from the preprocessed Wikipedia pages.
+Strictly speaking, our original assumption of equating Wikipedia articles with named entities is not valid: many pages describe common nouns (Book, Aircraft), currencies (Euro, Dollar), calendrical units XXX (January, the year 1999) or other concepts that fall outside the scope of NER.
+However, such entities are either missing from DBpedia altogether, or belong to classes mapped to the NER category \texttt{0} (not an entity).
+This "ontological filtering" ensures that
+% TODO: this is not true, we need filters!
 
-% 3. Training sentence extraction
-%   a. only known links
-%   b. title parts
-%   c. parts of links mention earlier
-%   d. date regexp
-% 4. Test sentence extraction?
-%   a. unknown links, no unlinked names
-%   b. might be good for quality evaluation / recursive tagging (next paper?)
-% 5. teaching the original DBpedia categories?
+In this paper, we describe % not only ConLL
 

### 2012-02-19 paper/engNERwiki.tex
Wrote a lot of stuff.

@@
 
-% TODO: we want to be on the strict side!
-
 \subsubsection{Entity identification}
+% TODO: identification and tagging happens in parallel, else the recall oriented approach cannot be explained.
 
@@
 Authors may neglect to add XXX references to their text.
-Furthermore, in a Wikipedia article XXX, typically only the first mention of a particular entity is linked to the corresponding page.
+Furthermore, in a Wikipedia article XXX, typically only the first occurrence of a particular entity is linked to the corresponding page.
 Subsequent mentions are unmarked and often incomplete -- e.g.
 family names are used instead of full names.
+We devised XXX two methods to account for unmarked entities: one that emphasizes XXX precision, and the other, recall.
+XXX
 
-% Numbers, years.
+% Or NPP
+
+The precision-oriented approach seeks to minimize the risk of a word getting incorrectly tagged XXX in the corpus.
+A short study of the ConLL annotation guidelines reveals that the words in the entities of interest to us all start with capital letter.
+According to this method XXX, all sentences that contain words that start with capital letter, but are not part of the anchor text of a link, XXX are discarded.
+The rationale behind this approach is that the sheer size of Wikipedia guarantees that we end up with a sizeable corpus even with such a strict policy.
+XXX
+
+The recall-oriented method tries to associate unmarked entities with entities already identified on the same page.
+Only if the attempt fails is the sentence that contains the words discarded.
+This approach aims to generate more training sentences with as little loss of precision as possible.
+
+The algorithm maintains a list of entities already indentified on the page. The list includes
+\begin{itemize}
+\item The page title; % if it is in DBpedia
+\item Title of redirect pages that target the current page; % TODO
+\item Entities marked with links in the text thus far.
+\end{itemize}
+
+When an unmarked entity is encountered, the list is consulted to find an item that contains it in its entirety; i.e.
+all of its words and in the same order.
+If either exactly one such item is found, or the matching entities in the list all have the same NER tag, the unmarked entity is tagged accordingly.
+Otherwise the sentence is discarded.
+
+Both approaches are evaluated in the next chapter.
 
@@
 
+%TODO: merge with Entity Identification
 We use DBpedia as
@@
 Only sentences for whose entities the correct category could be obtained from DBpedia are added to the corpus; those that contain unknown entities are discarded.
-Sentences without any entities are retained as well, so that the corpus mirrors (XXX the distribution of the entities in the corpus mirrors their distribution in Wikipedia?) the distribution of entities in Wikipedia as closely as possible; this is also in line with how the ConLL corpus is assembled.
+Sentences without any entities are retained as well, so that the corpus mirrors (XXX the distribution of entities in the corpus mirrors their distribution in Wikipedia?) the distribution of entities in Wikipedia as closely as possible; this is also in line with how the ConLL corpus is assembled.
 (XXX or is it)
 
-Unknown entities might or might not be
+\subsubsection{Exceptions}
+
+Strictly speaking, our original assumption of equating Wikipedia articles with named entities is not valid: many pages describe common nouns (Book, Aircraft), numbers (1, 42), currencies (Euro, Dollar) or other concepts that fall outside the scope of NER.
+In this section, we describe the measures taken to avoid errors introduced by these false entities.
+
+Links to number pages are filtered by the surface form of the page title. To Common nouns are filtered by their surface form and POS category: links that contain no proper nouns are discarded
+
+Proper entities However, such entities are either missing from DBpedia altogether, or we can map the classes they belong to to the NER category \texttt{0} (not an entity).
+This "ontological filtering" ensures that
+
+Dates require special attention because
+
+Wikipedia 
+% Numbers, years.
 
 % TODO: monetary, etc. units?
-
-% 3. Training sentence extraction
-%   a. only known links
-%   b. title parts
-%   c. parts of links mention earlier
-%   d. date regexp
-% recall+precision?
 
@@
 
-Strictly speaking, our original assumption of equating Wikipedia articles with named entities is not valid: many pages describe common nouns (Book, Aircraft), currencies (Euro, Dollar), calendrical units XXX (January, the year 1999) or other concepts that fall outside the scope of NER.
-However, such entities are either missing from DBpedia altogether, or belong to classes mapped to the NER category \texttt{0} (not an entity).
-This "ontological filtering" ensures that
-% TODO: this is not true, we need filters!
-
 In this paper, we describe % not only ConLL
@@
 % Comparison of the Hungarian DBpedia vs English + mapping
+% Not all entities are written w/ capital letters => we need lists.
 

### 2012-02-20 paper/engNERwiki.tex
Filtering next...

@@
 In this section, we describe the measures taken to avoid errors introduced by these false entities.
+We shall modify the two approaches -- precision and recall-oriented -- accordingly.
 
-Links to number pages are filtered by the surface form of the page title. To Common nouns are filtered by their surface form and POS category: links that contain no proper nouns are discarded
+Proper entities that are not deemed NE by ConLL cause XXX little concern: they are either missing from DBpedia altogether, or we can map the classes they belong to to the NER category \texttt{0} (not an entity).
+This "ontological filtering" ensures that we need no adjustment XXX to our original method.
 
-Proper entities However, such entities are either missing from DBpedia altogether, or we can map the classes they belong to to the NER category \texttt{0} (not an entity).
-This "ontological filtering" ensures that
+Links to number and common noun pages are handled differently by the two approaches.
+The precision-oriented method treats them similarly to links to unknown entities and discards the sentences in which they occur.
+The recall-oriented approach tries to distinguish them from links to entities so that otherwise valid training sentences can be retained XXX.
+The method is as follows XXX.
+\begin{itemize}
+\item Common nouns are filtered by POS category: links that contain no proper nouns are ignored.
+\item Numbers behave more like proper nouns in that they can refer to entities, such as 737 to the Boeing 737. Therefore, links to number pages are filtered by the surface form of the page title.
+\end{itemize}
 
-Dates require special attention because
-
-Wikipedia 
-% Numbers, years.
+Time expressions require special attention due to two reasons.
+First, Wikipedia has pages for years, months and concrete dates, therefore our algorithm might mistake them for entities.
+Second, as dates are made up of numbers and proper nouns, they can be mistaken for unmarked entities.
+To avoid these errors, we have create a set of regular expressions that recognize dates and time expressions and changed the recall-oriented method to not treat them as entities.
 

### 2012-02-21 scripts/extract_dbpedia_types.py
dbpedia_map is now tab-separated.

@@ -108,7 +108,7 @@ def __read_map(map_file):
     with FileReader(map_file, encoding='utf-8').open() as mappings:
         for mapping in mappings:
             try:
-                key, value = mapping.strip().split(' ')
+                key, value = mapping.strip().split("\t")
                 type_map[key] = value
             except (ValueError):
                 continue

### 2012-02-21 paper/engNERwiki.tex

--- 2012-02-14 paper/engNERwiki.tex
@@
 %\author{Dávid Márk Nemeskey \\
-%  Research Institute for Computer Science and Automation \\
+%  XXX Computer and Automation Research Institute\\
 %  Hungarian Academy of Sciences \\
@@
 \end{abstract}
+
+\section{Credits}
+
+This document has been adapted from the instructions for earlier ACL proceedings, including those for ACL-2008 by Johanna D.
+Moore, Simone Teufel, James Allan, and Sadaoki Furui, those for ACL-2005 by Hwee Tou Ng and Kemal Oflazer, those for ACL-2002 by Eugene Charniak and Dekang Lin, and earlier ACL and EACL formats.
+Those versions were written by several people, including John Chen, Henry S.
+Thompson and Donald Walker.
+Additional elements were taken from the formatting instructions of the {\em International Joint Conference on Artificial Intelligence}.
 
@@
 \section{Related Work}
-% Wikify! 
 % Ide kell? Vagy Conclusion elé?
@@
 \section{Creation of the Corpus}  % Dzsííízösz
-
-% Introduction: Wikipedia + DBpedia
-The corpora are available in the tab separated format used by ConLL (XXX, CITE).
-In this format, one line contains one word, its linguistic features and the named entity tags.
-Sentence boundaries are marked by empty lines.
-The linguistic features include the lemmatized form of the word, part-of-speech tag and chunk tag.
-Two named entity tags are included with each word: one is the most specific DBpedia category it belongs to, if any; the other is the ConLL named entity tag.
-While the named entity tags can be considered "gold", the linguistic features are provided on a "best-effort" basis (XXX: check the LREC paper to find a better wording for this).
-XXX: past or present tense?
-
-We decided to use Wikipedia because
-
-The categories of Wikipedia are too numerous and ; not a tree;
-
-Wikipedia proves to be a rich source of information about its entities; aside from the article text, a huge amount of data is encompassed in infoboxes, templates and the category structure.
-Since our goal is to create a NER training corpus, our interest is limited to the article text.
-Templates are removed from the articles; however, the names of the templates have been retained for XXX, but all data (including infobox)
-
-Beside regular articles, Wikipedia also contains redirect and disambiguation pages, which require special treatment.
-Redirect pages are empty and are safe to ignore altogether.
-Disambiguation pages contain text, usually in the form of an enumeration of page links with a short description.
-Full sentences are scarce and the structure of these pages makes it hard to extract valid training sentences from them.
-Therefore, disambiguation pages have been dropped from the corpus as well.
-
-The corpus is based on the Wikipedia snapshot as of XXX.
-The xml files were parsed with the XXX (Attila) parser.
-The raw text was tokenized with a modified version of the XXX sentence tokenizer in NLTK (CITE).
-The words were lemmatized with XXX and POS tagged with HunPos (CITE).
-Chunking was performed with HunChunk (CITE).
-Experiments have shown that HunPos obtains STAT f-score on the Brown corpus; HunChunk performs at around STAT, CITE?.
-
-\subsection{Extraction of the Training Sentences}
-
-A NER training corpus consists of sentences that accurately tag named entities occurring in them.
-In order to automatically prepare such sentences, two tasks need to be performed: identification of entities in the sentence and tagging them with the correct tag.
-Sentences, for which accurate tagging could not be accomplished, must be removed from the corpus.
-In this section, we describe how these steps were implemented, as well as the post-filtering we employed to dispose of correctly tagged but unusable training sentences.
-
-% TODO: we want to be on the strict side!
-
-\subsubsection{Entity identification}
-
-Our approach to entity identification relies on Wikipedia cross-references found in the article text.
-We assume that individual Wikipedia articles describe named entities.
-A link to an article can then be perceived as a mapping that identifies its anchor text with a particular entity.
-There are exceptions to this rule, however.
-Disambiguation pages are used when the page title is ambiguous, and can refer to several entities.
-If a link points to a disambiguation page, we acknowledge the anchor text as an entity, but mark it unknown.
-A similar measure is taken if the link target XXX does not exist.
-Links to redirect pages are resolved to point instead to the redirect target, after which they are handled as regular cross-references.
-
-Not all entities are marked with links, however.
-Authors may neglect to add XXX references to their text.
-Furthermore, in a Wikipedia article XXX, typically only the first mention of a particular entity is linked to the corresponding page.
-Subsequent mentions are unmarked and often incomplete -- e.g.
-family names are used instead of full names.
-
-% Numbers, years.
-
-\subsubsection{Tagging}
-
-We use DBpedia as
-
-Only sentences for whose entities the correct category could be obtained from DBpedia are added to the corpus; those that contain unknown entities are discarded.
-Sentences without any entities are retained as well, so that the corpus mirrors (XXX the distribution of the entities in the corpus mirrors their distribution in Wikipedia?) the distribution of entities in Wikipedia as closely as possible; this is also in line with how the ConLL corpus is assembled.
-(XXX or is it)
-
-Unknown entities might or might not be
-
-% TODO: monetary, etc. units?
-
+% 1. Parsing Wikipedia (hunpos)
+% 2. The DBpedia -> ConLL mapping
 % 3. Training sentence extraction
@@
 %   d. date regexp
-% recall+precision?
-
-\subsubsection{Filtering}
-
-Some sentences that were selected by the process above
-
 % 4. Test sentence extraction?
@@
 
-\subsection{The DBpedia -- ConLL mapping}
-
-% Introduction + entity -> class -> category.
-
-The classes in DBpedia are organized into a type hierarchy, available as an OWL (CITE) ontology.
-The ontology contains the 320 most frequent categories of Wikipedia, arranged into a taxonomy under the base class \texttt{owl:Thing}.
-Most of the classes belong into the 5 largest sub-hierarchies (here named after their topmost classes XXX): \texttt{Agent} (which is further divided into \texttt{Person} and \texttt{Organisation}), \texttt{Event}, \texttt{Place}, \texttt{Species} and \texttt{Work}.
-The taxonomy is rather flat: the top level contains 44 classes and there are several nodes with a branching factor of 20.
-
-It is not difficult to see the parallels XXX between the DBpedia sub-hierarchies \texttt{Person}, \texttt{Organisation} and \texttt{Place} and the ConLL NER categories \texttt{PER}, \texttt{ORG} and \texttt{LOC}.
-The fourth category, MISC is more elusive; according to the ConLL NER annotation guide (CITE?: http://www.cnts.ua.ac.be/conll2003/ner/annotation.txt), the sub-hierarchies \texttt{Event} and \texttt{Work} belong to this category, as well as various other class outside the main sub-hierarchies.
-
-While the correspondence described above holds for most classes in the XXX sub-hierarchies, there are some exceptions.
-For instance, the class \texttt{SportsLeague} is part of the \texttt{Organisation} sub-hierarchy, but in the ConLL corpus, sport leagues are tagged as MISC.
-Similarly, the class \texttt{SportsTeamSeason} is under \texttt{Event}, by which account it would be classified as MISC; however, ConLL does not recognize such entities.
-To account for these misclassification, as well as the miscellaneous classes outside the main hierarchies, we have introduced a mapping mechanism.
-We created a file of DBpedia class -- NER category mappings.
-Whenever an entity is evaluated, we look up its class, as well as the ancestors of its class, in the mapping, and assign to the entity the category of the class that matches it most closely -- i.e.
-the category of the entity's class, if it is in the list, or that of its closest ancestor.
-XXX!!!
-If no match is found, the entity is tagged with \texttt{0}.
-Since we take advantage of the inheritance hierarchy, the mapping list remains short: it contains only the root classes of the main sub-hierarchies, exceptions of the like XXX mentioned above and the various classes that belong to the MISC category according to the ConLL annotation guideline.
-
-As of version 3.7, the DBpedia ontology allows multiple superclasses, making the hierarchy a directed acyclic graph.
-This introduces the problem of selecting the right superclass, and hence, ConLL mapping for classes with more than one parent.
-In version 3.7, the only such class is \textit{Library}, which can be traced back to both \textit{Place} and \textit{Organisation}.
-There are two ways to tackle this problem.
-The first is to follow the DBpedia guidelines and only retain the first parent.
-The second method is manually deciding the right NER mapping for the ambiguous XXX class and adding it to the mapping file.
-The first approach, unfortunately, already fails to assign the correct ConLL NER tag to our only example.
-The reason behind this failure is that multiple parents in an ontology show an inherent ambiguity in the class in question: it is entirely natural to think of a library as a location as well as a public organization.
-DBpedia's guidelines prefers Organisation; ConLL's LOC.
-Therefore, only the second, manual approach ensures correct tagging.
-XXX
-
-The full mapping can be found in LINK Appendix A (XXX or not).
-
-\subsection{Remarks}
-
-Strictly speaking, our original assumption of equating Wikipedia articles with named entities is not valid: many pages describe common nouns (Book, Aircraft), currencies (Euro, Dollar), calendrical units XXX (January, the year 1999) or other concepts that fall outside the scope of NER.
-However, such entities are either missing from DBpedia altogether, or belong to classes mapped to the NER category \texttt{0} (not an entity).
-This "ontological filtering" ensures that
-% TODO: this is not true, we need filters!
-
-In this paper, we describe % not only ConLL
-
-\section{Creation of the Hungarian Corpus}
-
-% Good for middle-sized languages
-% Inconsistent template names, esp. across languages
-% Comparison of the Hungarian DBpedia vs English + mapping
-
 \section{Evaluation}
-
-% vs. ConLL / Szeged
-% Corpus sizes, performance, tagging conventions
 

### 2012-02-21 paper/engNERwiki.tex

--- 2012-02-20 paper/engNERwiki.tex
@@
 The paper is structured as follows.
-
-\section{Related Work}
-% Wikify! 
-% Ide kell? Vagy Conclusion elé?
 
@@
 
-Some sentences that were selected by the process above
+There are many incomplete sentences in the Wikipedia text: image captions, enumerations items, contents of table cells, etc.
+On the one hand, these sentence fragments may prove to be of too low quality to be of any use in the traditional NER task.
+On the other hand, they could prove to be invaluable when training a named entity recognizer for user generated content.
+We have added a post-filtering step to recognize and label these fragments, so that users of the corpus can decide what to do with them.
+XXX
+
+Manual inspection has found two conditions that can be used to detect fragments.
+A sentence is labelled as such if it either lacks a punctuation mark at the end or it contains no finite verbs.
+Our final corpus contains STAT full sentences and STAT fragments.
+% TODO: Last sentence to evaluation? But then this whole subsection is too short.
 
@@
 
-In this paper, we describe % not only ConLL
+In this paper, we describe % not only ConLL? Do we need this?
 
@@
 % Corpus sizes, performance, tagging conventions
+
+\section{Related Work}
+% Wikify!
+% stuff from Mining Meaning from Wikipedia
 

### 2012-02-21 scripts/ner_filter_incomplete_sentences.py
Added ner_filter_incomplete_sentences.py.

@@ -1,0 +1,89 @@
+## Filters incomplete sentences from a conll file.
+
+from langtools.io.conll2.conll_reader import ConllReader, ConllCallback
+from langtools.utils.file_utils import FileWriter
+
+class SentenceFilterCallback(ConllCallback):
+    """Filters all incomplete sentences, i.e. those that don't end in a period,
+    question mark, etc. and those that don't have a verb in them."""
+    def __init__(self):
+        self.sentence = []
+        self.filters = []
+
+    def addFilter(self, filter):
+        """Adds a SentenceFilter to the filter list."""
+        if filter is not None and filter not in self.filters:
+            self.filters.append(filter)
+
+    def fileStart(self, file_name):
+        """Opens output files for the sentences kept, as well as one for each
+        filter, where the sentences filtered by that particular filter are
+        written."""
+        self.kept_file = FileWriter(file_name + '.kept').open()
+        self.filtered_files = [FileWriter(file_name + '.f' + str(i)).open()
+                               for i in xrange(len(self.filters))]
+
+    def sentenceStart(self):
+        self.sentence = []
+
+    def word(self, attributes):
+        self.sentence.append(attributes)
+
+    def sentenceEnd(self):
+        if len(self.sentence) > 0:
+            for i, filter in enumerate(self.filters):
+                if not filter.filter(self.sentence):
+                    self.filtered_files[i].write(u"\n".join(
+                        u"\t".join(word) for word in self.sentence))
+                    self.filtered_files[i].write(u"\n\n")
+                    return
+            self.kept_file.write(u"\n".join(
+                u"\t".join(word) for word in self.sentence))
+            self.kept_file.write(u"\n\n")
+
+    def fileEnd(self):
+        """Closes all the files."""
+        self.kept_file.close()
+        for ff in self.filtered_files:
+            ff.close()
+
+class SentenceFilter(object):
+    """Base class for sentence filters."""
+    def filter(self, sentence):
+        """
+        @param sentence is a list of word attributes.
+        @return @c True, if the sentence is to be kept; @c False otherwise.
+        """
+        pass
+
+class TerminalPunctuationFilter(SentenceFilter):
+    """Filters sentences that lack terminal punctuation."""
+    def __init__(self, lemma_pos=0):
+        """@param lemma_pos the position of the lemma field."""
+        self.lemma_pos = lemma_pos
[…]

### 2012-02-21 scripts/resolve_redirects.py
Added search_disambig_redirect_paths.py from langtools as resolve_redirects.py.

@@ -1,0 +1,160 @@
+import sys
+from collections import defaultdict
+import logging
+
+"""
+This scripts collects all incoming paths that are built of only redirect and
+disambiguation pages
+Circles are detected and skipped, only a stderr message warns the user.
+
+Formats (description of what a line should contain)(TAB separated inputs):
+  page_ids:
+    id namespace title is_redirect
+  links:
+    src_id src_namespace tgt_title
+  redirect_pages:
+  disambig_pages:
+  normal_pages:
+    title
+"""
+
+
+def read_to_set(f):
+    s = set()
+    for l in f:
+        l = l.strip()
+        s.add(l)
+    return s
+
+def read_links(f, title_to_id, id_to_title, reverse=False):
+    links = defaultdict(set)
+    c = 0
+    for l in f:
+        c += 1
+        if c % 1000000 == 0:
+            logging.info("%f read." % (c / 1000000.0))
+        le = l.strip().split("\t")
+        src_id = int(le[0])
+        # skip pages that we don't know like useless namespaces
+        if not src_id in id_to_title:
+            continue
+
+        if le[1] != "0":
+            continue
+        try:
+            tgt_title = le[2]
+        except IndexError:
+            continue
+        try:
+            tgt_id = title_to_id[tgt_title]
+        except KeyError:
+            continue
+        if not reverse:
+            links[tgt_id].add(src_id)
+        else:
+            links[src_id].add(tgt_id)
+    return links
+
+def read_ids(f):
+    title_to_id = {}
+    id_to_title = {}
+    for l in f:
+        le = l.strip().split("\t")
+        title_to_id[le[2]] = int(le[0])
+        id_to_title[int(le[0])] = le[2]
+    return title_to_id, id_to_title
+
+def read_all(redirect_pages_file, disambig_pages_file, normal_pages_file, page_ids_file, links_file, is_reverse):
+    redirect_pages = read_to_set(redirect_pages_file)
+    disambig_pages = read_to_set(disambig_pages_file)
+    normal_pages = read_to_set(normal_pages_file)
+
+    import gc
+    gc.disable()
+    title_to_id, id_to_title = read_ids(page_ids_file)
+    logging.info("ids read.")
+    dr_pages = set(title_to_id[page] for page in (redirect_pages | disambig_pages) if page in title_to_id)
+    del redirect_pages
+    del disambig_pages
+    links = read_links(links_file, title_to_id, id_to_title, is_reverse)
+    gc.enable()
[…]

### 2012-02-21 scripts/read_wiki_files.py
File reader functions refactored to read_wiki_files.py.

@@ -1,0 +1,56 @@
+from collections import defaultdict
+import logging
+
+"""
+Contains functions that read the various files that make up a Wikipedia dump."""
+
+def read_to_set(f):
+    """Reads the lines of a file into a set. Used for reading the redirect and
+    disambiguation page list files."""
+    s = set()
+    for l in f:
+        l = l.strip()
+        s.add(l)
+    return s
+
+def read_ids(f):
+    """Reads the page id file, and returns two maps: title to id and
+    id to title."""
+    title_to_id = {}
+    id_to_title = {}
+    for line in f:
+        fields = line.strip().split("\t")
+        title_to_id[fields[2]] = int(fields[0])
+        id_to_title[int(fields[0])] = fields[2]
+    return title_to_id, id_to_title
+
+def read_links(f, title_to_id, id_to_title, reverse=False):
+    """Reads the links file and returns it as a map."""
+    links = defaultdict(set)
+    count = 0
+    for line in f:
+        count += 1
+        if count % 1000000 == 0:
+            logging.info("%f read." % (count / 1000000.0))
+        fields = line.strip().split("\t")
+        src_id = int(fields[0])
+        # skip pages that we don't know like useless namespaces
+        if not src_id in id_to_title:
+            continue
+
+        if fields[1] != "0":
+            continue
+        try:
+            tgt_title = fields[2]
+        except IndexError:
+            continue
+        try:
+            tgt_id = title_to_id[tgt_title]
+        except KeyError:
+            continue
+        if not reverse:
+            links[tgt_id].add(src_id)
+        else:
+            links[src_id].add(tgt_id)
+    return links
+

### 2012-02-21 scripts/resolve_redirects.py

@@ -1,144 +1,48 @@
 import sys
-from collections import defaultdict
 import logging
+from read_wiki_files import read_ids, read_to_set, read_links
+from optparse import OptionParser
 
-"""
-This scripts collects all incoming paths that are built of only redirect and
-disambiguation pages
-Circles are detected and skipped, only a stderr message warns the user.
+"""Resolves redirect pages and tags them with the same NER categories as the
+target pages."""
 
-Formats (description of what a line should contain)(TAB separated inputs):
-  page_ids:
-    id namespace title is_redirect
-  links:
-    src_id src_namespace tgt_title
-  redirect_pages:
-  disambig_pages:
-  normal_pages:
-    title
-"""
-
-
-def read_to_set(f):
-    s = set()
-    for l in f:
-        l = l.strip()
-        s.add(l)
-    return s
-
-def read_links(f, title_to_id, id_to_title, reverse=False):
-    links = defaultdict(set)
-    c = 0
-    for l in f:
-        c += 1
-        if c % 1000000 == 0:
-            logging.info("%f read." % (c / 1000000.0))
-        le = l.strip().split("\t")
-        src_id = int(le[0])
-        # skip pages that we don't know like useless namespaces
-        if not src_id in id_to_title:
-            continue
-
-        if le[1] != "0":
-            continue
-        try:
-            tgt_title = le[2]
-        except IndexError:
-            continue
-        try:
-            tgt_id = title_to_id[tgt_title]
-        except KeyError:
-            continue
-        if not reverse:
-            links[tgt_id].add(src_id)
-        else:
-            links[src_id].add(tgt_id)
-    return links
-
-def read_ids(f):
-    title_to_id = {}
-    id_to_title = {}
-    for l in f:
-        le = l.strip().split("\t")
-        title_to_id[le[2]] = int(le[0])
-        id_to_title[int(le[0])] = le[2]
-    return title_to_id, id_to_title
-
-def read_all(redirect_pages_file, disambig_pages_file, normal_pages_file, page_ids_file, links_file, is_reverse):
+def read_all(redirect_pages_file, page_ids_file, links_file):
+    """Reads the files necessary for the resolution of redirects."""
     redirect_pages = read_to_set(redirect_pages_file)
-    disambig_pages = read_to_set(disambig_pages_file)
-    normal_pages = read_to_set(normal_pages_file)
+    logging.info("Redirect pages read.")
 
     import gc
     gc.disable()
[…]

### 2012-02-23 scripts/resolve_redirects.py
I don't know.

@@ -7,7 +7,11 @@ from optparse import OptionParser
 target pages."""
 
 def read_all(redirect_pages_file, page_ids_file, links_file):
-    """Reads the files necessary for the resolution of redirects."""
+    """Reads the files necessary for the resolution of redirects.
+    Returns: - the ids of the redirect pages;
+             - the links map, reversed;
+             - the title to id map;
+             - the id to title map."""
     redirect_pages = read_to_set(redirect_pages_file)
     logging.info("Redirect pages read.")
 
@@ -20,7 +24,7 @@ def read_all(redirect_pages_file, page_ids_file, links_file):
     del redirect_pages
     links = read_links(links_file, title_to_id, id_to_title, True)
     gc.enable()
-    return redirect_ids, links, (title_to_id, id_to_title)
+    return redirect_ids, links, title_to_id, id_to_title
 
 def run(map_file, redirect_ids, links, title_to_id, id_to_title):
     for line in map_file:
@@ -44,24 +48,27 @@ def run(map_file, redirect_ids, links, title_to_id, id_to_title):
                 continue
             print "{0}\t{1}".format(redirect_title, ner)
 
-def main():
-    logging.basicConfig(level=logging.DEBUG, format="%(asctime)s : %(module)s - %(levelname)s - %(message)s")
-    page_ids_file = file(sys.argv[1])
-    links_file = file(sys.argv[2])
-    redirect_pages_file = file(sys.argv[3])
-    disambig_pages_file = file(sys.argv[4])
-    normal_pages_file = file(sys.argv[5])
-    is_reverse = (bool(int(sys.argv[6])) if len(sys.argv) > 6 else False)
+def main(redirects_file, links_file, pages_file, original_mappings):
+    redirects_file = file(redirects_file)
+    links_file = file(links_file)
+    pages_file = file(pages_file)
+    original_mappings = file(original_mappings)
 
-    normal_pages, dr_pages, links, (title_to_id, id_to_title) = read_all(redirect_pages_file, disambig_pages_file, normal_pages_file, page_ids_file, links_file, is_reverse)
+    redirects, links, title_to_id, id_to_title = read_all(redirect_pages_file, page_ids_file, links_file)
     #import cProfile
     #cProfile.run("run(normal_pages, dr_pages, links, title_to_id, id_to_title, is_reverse)")
-    run(normal_pages, dr_pages, links, title_to_id, id_to_title, is_reverse)
-
-if __name__ == "__main__":
[…]

### 2012-02-23 paper/engNERwiki.tex
A tiny little change.

@@
 
-Proper entities that are not deemed NE by ConLL cause XXX little concern: they are either missing from DBpedia altogether, or we can map the classes they belong to to the NER category \texttt{0} (not an entity).
+Proper entities that are not deemed NE under the ConLL scheme cause XXX little concern: they are either missing from DBpedia altogether, or we can map the classes they belong to to the NER category \texttt{0} (not an entity).
 This "ontological filtering" ensures that we need no adjustment XXX to our original method.

### 2012-02-23 scripts/resolve_redirects.py
resolve_redirects.py finished (?)

@@ -1,7 +1,6 @@
 import sys
 import logging
 from read_wiki_files import read_ids, read_to_set, read_links
-from optparse import OptionParser
 
 """Resolves redirect pages and tags them with the same NER categories as the
 target pages."""
@@ -54,7 +53,7 @@ def main(redirects_file, links_file, pages_file, original_mappings):
     pages_file = file(pages_file)
     original_mappings = file(original_mappings)
 
-    redirects, links, title_to_id, id_to_title = read_all(redirect_pages_file, page_ids_file, links_file)
+    redirects, links, title_to_id, id_to_title = read_all(redirects_file, pages_file, links_file)
     #import cProfile
     #cProfile.run("run(normal_pages, dr_pages, links, title_to_id, id_to_title, is_reverse)")
     run(original_mappings, redirects, links, title_to_id, id_to_title)

### 2012-03-05 paper/engNERwiki.tex
Introduction, Related work

@@
 \begin{abstract}
-  This document contains the instructions for preparing a camera-ready manuscript for the proceedings of ACL2012.
-The document itself conforms to its own specifications, and is therefore an example of what your manuscript should look like.
-These instructions should be used for both papers submitted for review and for final versions of accepted papers.
-Authors are asked to conform to all the directions reported in this document.
+
 \end{abstract}
@@
 
-One of the main trends in NLP is the need for reducing the annotation cost.
+Named Entity Recognition (NER), the task of identifying Named Entities (NEs) in unstructured texts and classifying them into pre-selected classes, is one of the most important subtasks in many NLP tasks, such as Information Retrieval, Information Extraction or Machine Translation.
+The NER task was introduced with the 6th Message Understanding Conference (MUC) in 1995.
+In MUC shared tasks the NER consists of three subtasks: entity names, temporal expressions and number expressions.
+However there is a general agreement in the NER community about the inclusion of temporal expressions and some numerical expressions such as amounts of money and other types of units, the most studied types are names of persons, locations and organizations.
+The fourth type, called "miscellaneous", is used in the CoNLL NER tasks in 2002 and 2003, and includes proper names falling outside the three classic types.
+Since then, MUC and CoNLL datasets and annotation schemes have been the major standards applied in the field of NER.
+
+The standard datasets are highly domain-specific (mostly newspaper articles about company activities and defense related activities), and are restricted in size, as well.
+Thus they are not large enough to train and test robust NE-taggers on them.
+Moreover, merging these datasets to get a bigger training corpus faces NLP researchers with the problem of combining different tagsets and annotation schemes.
+However, to build a robust supervised machine learning system for NER, large enough, accurately annotated corpora are essential.
+Annotating large amount of texts with linguistic information is a time-consuming, highly skilled and delicate work.
+Mainly due to this fact, one of the main challenges in NLP is reducing the annotation cost.
+
+One approach to reduce the manual intervention required in language resource production, and thus ultimately production costs is automatically generating resources.
+An other approach to achieve this aim is using collaborative annotation and/or collaboratively constructed resources, such as Wikipedia, Wiktionary or Linked Open Data.
+Researchers apply the Web-as-corpus approach, often dealing with such datasets as the Google 5G 1T collection or DBPedia.
+Thanks for the huge amount of Wikipedia articles, the possibility of building large enough corpora is given for less resourced languages, as well.
+
+%One of the main trends in NLP is the need for reducing the annotation cost.
 After examining the calls for papers of nowadays conferences and workshops there seems to be three major directions to achieve this aim.
 
-The first approach is the application of unsupervised and semi-supervised learning techniques.
+%The first approach is the application of unsupervised and semi-supervised learning techniques.
 Researchers following this approach try to not only reduce, but avoid the annotation cost with minimizing the need for annotated data.
@@
 
-The second approach is using collaborative annotation and/or collaboratively constructed resources, such as Wikipedia, Wiktionary or Linked Open Data.
-Researchers apply the Web-as-corpus approach, often dealing with such datasets as the Google 5G 1T collection or DBPedia.
-Thanks for the huge amount of Wikipedia articles, the possibility of building large enough corpora is given for less resourced languages, as well.
+In this paper we present a kind of combination of the approaches mentioned above.
+We built NE-tagged corpora for English and Hungarian -- fully automatically generated from Wikipedia.
+Our goal was to create gold standard corpora which can serve as train, devel and test datasets for automatic NE-taggers.
 
-The third approach to reduce the manual intervention required in language resource production, and thus ultimately production costs is automatically generating resources or merging existing ones.
-However merging language resources faces NLP researchers with the problem of combining different annotation schemas.
+The paper is structured as follows. XXX
 
-Named Entity Recognition (NER) is one of the important subtasks of Information Extraction.
-NER contains two main substeps: recognizing the Named Entities (NEs) in unstructured text and classifying them into preselected classes
+\section{Related Work}
+% Wikify!
+% stuff from Mining Meaning from Wikipedia
+%NER and Wikipedia
 
-mostly newspaper articles, domain-specific
+Wikipedia is a goldmine of information; it is applied for several NLP tasks, e.g.~word sense disambiguation, ontology and thesaurus building, multilingual Information Retrieval, and Question Answering, as Medelyan et al.~\shortcite{Medelyan:09} point out in their fully detailed survey about mining meaning from Wikipedia.
 
-In this article we present a Named Entity tagged corpus for English -- fully automatically generated from Wikipedia. 
-The applied annotation schema is the CoNLL tagset, which is one of the major standards applied in the field of Named Entity Recognition (NER).
+Wikipedia is recognized as one of the largest available collection of entities, and also as an external knowledge to improve the accuracy of NER.
+The most obvious utilization of Wikipedia for NER is extracting gazetteers containing person names, locations or organizations \cite{Toral:06}.
+Creating dictionaries of entities is also a common step of named entity disambiguation \cite{Bunescu:06,Cucerzan:07}.
+Nadeau et al.
+\shortcite{Nadeau:06} use similar lists in an unsupervised NER system, while Kazama and Torisawa \shortcite{Kazama:07} showed that automatic extraction of category labels from Wikipedia improves the accuracy of a supervised NE-tagger.
 
-The paper is structured as follows.
+%Exploiting Wikipedia as external knowledge for NER:
+%%-creating entity lists  
+%%-disambiguating named entities
+%%-automatic extraction of category labels from Wikipedia to improve the accuracy of a CRF-based NE-tagger \cite{Kazama:07}
+%-entity ranking
+%-typing named entities
+
+Automatic creation of training data from Wikipedia for NER:
+-classification of articles into CoNLL entity classes \cite{Nothman:08}
 
@@
 % Corpus sizes, performance, tagging conventions
-
-\section{Related Work}
-% Wikify!
-% stuff from Mining Meaning from Wikipedia
 
@@
 
+\bibitem[\protect\citename{Nothman \bgroup et al.\egroup}2008]{Nothman:08}
+Joel Nothman, James R. Curran, and Tara Murphy.
+\newblock 2008.
+\newblock Transforming Wikipedia into Named Entity Training Data.
+\newblock In: {\em Proceedings of the Australasian Language Technology Workshop}, Vol 6., pages 124--132.
+
+\bibitem[\protect\citename{Kazama and Torisawa}2007]{Kazama:07}
+Jun'ichi Kazama and Kentaro Torisawa.
+\newblock 2007.
+\newblock Exploiting Wikipedia as External Knowledge for Named Entity Recognition.
+\newblock In: {\em Proceedings of the 2007 Joint Conference on Empirical Methods in Natural Language Processing and Computational Natural Language Learning}, pages 698--707.
+
+\bibitem[\protect\citename{Medelyan  \bgroup et al.\egroup}2009]{Medelyan:09}
+Olena Medelyan, David Milne, Catherine Legg, and Ian H. Witten.
+\newblock 2009.
+\newblock Mining meaning from Wikipedia.
+\newblock {\em International Journal of Human-Computer Studies}, 67: 716--754.
+
+\bibitem[\protect\citename{Toral and Munoz}2006]{Toral:06}
+A. Toral and R. Munoz.
+\newblock 2006.
+\newblock A proposal to automatically build and maintain gazetteers for Named Entity Recognition by using Wikipedia.
+\newblock In: {\em EACL 2006}.
+
+\bibitem[\protect\citename{Bunescu and Pasca}2006]{Bunescu:06}
+B. Bunescu and M. Pasca.
+\newblock 2006.
+\newblock Using encyclopedic knowledge for named entity disambiguation.
+\newblock In: {\em Proceedings of the 11th Conference of the European Chapter of the Association for Computational Linguistics}, pages 9--16.
+
+\bibitem[\protect\citename{Cucerzan}2007]{Cucerzan:07}
+S. Cucerzan.
+\newblock 2007.
+\newblock Large-scale named entity disambiguation based on Wikipedia data.
+\newblock In: {\em Proceedings of the 2007 Joint Conference on Empirical Methods in Natural Language Processing and Computational Natural Language Learning}.
+\newblock Prague, Czech Republic, June 2007. pages 708--716.
+
+\bibitem[\protect\citename{Nadeau  \bgroup et al.\egroup}2006]{Nadeau:06} 
+David Nadeau, Peter D. Turney and Stan Matwin.
+\newblock 2006.
+\newblock Unsupervised named entity recognition: Generating gazetteers and resolving ambiguity. 
+\newblock In: {\em Proceedings of the 19th Canadian Conference on Artificial Intelligence}, volume 4013 of {\em LNCS}, pages 266--277.
+
 \bibitem[\protect\citename{Aho and Ullman}1972]{Aho:72}

### 2012-03-08 scripts/extract_dbpedia_types.py
Bugs in extract_dbpedia_types.py fixed.

@@ -8,6 +8,7 @@ from langtools.utils.cmd_utils import get_params_sing
 from dbpedia_class import OwlClassHierarchy
 
 _type_re = re.compile(r'^\s*<http://dbpedia.org/resource/(.+?)>\s+<http://(.+?)>\s+<http://(.+?)>')
+_dbpedia_prefix = 'dbpedia'
 
 def extract_dbpedia_type(type_stream):
     """A generator that reads type_stream line-by-line and outputs page title - 
@@ -17,7 +18,7 @@ def extract_dbpedia_type(type_stream):
     for line in type_stream:
         match = _type_re.match(line)
         if match:
-            if match.group(2).endswith('#type'):
+            if match.group(2).endswith('#type') and match.group(3).startswith(_dbpedia_prefix):
                 title = match.group(1)
                 slash = match.group(3).rfind('/')
                 if 0 <= slash < len(match.group(3)) - 1:
@@ -29,17 +30,17 @@ def merge_pairs(pairs):
     """Merges subsequent pairs whose key (the first field) is the same. The
     resulting pair will have the same key as the original and a list of the
     values (seconds fields) of the merged pairs as value."""
-    value = []
+    value = set()
     key = ''
     for pair in pairs:
         if key != pair[0]:
             if len(value) > 0:
                 yield (key, value)
             key = pair[0]
-            value = []
-        value.append(pair[1])
+            value = set()
+        value.add(pair[1])
     if len(value) > 0:
-        yield (key, value)
+        yield (key, list(value))
     return
 
 def filter_general(pairs, hierarchy, what_to_keep=None):
@@ -114,14 +115,20 @@ def __read_map(map_file):
                 continue
     return type_map
 
+def print_usage_and_exit():
+    sys.stderr.write('Usage: {0} dbpedia_type_file [-c classes_OWL_file] [-m NE_mappings]\n'.format(__file__))
+    sys.exit()
+
 if __name__ == '__main__':
     import sys
     try:
         params, args = get_params_sing(sys.argv[1:], 'c:m:', '', 1)
     except ValueError as ve:
         sys.stderr.write(ve + "\n")
-        sys.stderr.write('Usage: {0} dbpedia_type_file [-c classes_OWL_file] [-m NE_mappings]\n'.format(__file__))
-        sys.exit()
+        print_usage_and_exit()
+    
+    if len(args) != 1:
+        print_usage_and_exit()
 
[…]

### 2012-03-08 scripts/ner_list_to_hu.py
ner_list_to_hu.py added.

@@ -1,0 +1,38 @@
+"""Converts the NER list to Hungarian (DBpedia uses English URIs)."""
+
+import sys
+
+def read_interlang_map(interlang_file):
+    """Reads the English -> Hungarian article mapping."""
+    import gc
+    gc.disable()
+    ret = {}
+    with open(interlang_file, 'r') as infile:
+        for line in infile:
+            try:
+                en, hu = line.strip().split("\t")
+                ret[en] = hu
+            except Exception:
+                sys.stderr.write("Invalid line: " + line)
+    gc.enable()
+    return ret
+
+def convert_ner_lang(ner_file, lang_mapping):
+    """Converts the NER list according to @p lang_mapping."""
+    with open(ner_file, 'r') as infile:
+        for line in infile:
+            try:
+                page, cat = line.strip().split("\t")
+                hu_page = lang_mapping[page]
+                print "{0}\t{1}".format(hu_page, cat)
+            except KeyError:
+                sys.stderr.write("Unknown page " + page)
+
+if __name__ == '__main__':
+    if len(sys.argv) != 2:
+        sys.stderr.write("Usage: {0} ner_list language_mapping\n\n".format(__file__))
+        sys.exit()
+
+    lang_mapping = read_interlang_map(sys.argv[2])
+    convert_ner_lang(sys.argv[1], lang_mapping)
+

### 2012-03-08 scripts/ner_list_to_hu.py
Keep option added to ner_list_to_hu.py

@@ -1,6 +1,7 @@
 """Converts the NER list to Hungarian (DBpedia uses English URIs)."""
 
 import sys
+from optparse import OptionParser
 
 def read_interlang_map(interlang_file):
     """Reads the English -> Hungarian article mapping."""
@@ -17,22 +18,30 @@ def read_interlang_map(interlang_file):
     gc.enable()
     return ret
 
-def convert_ner_lang(ner_file, lang_mapping):
+def convert_ner_lang(ner_file, lang_mapping, keep=False):
     """Converts the NER list according to @p lang_mapping."""
     with open(ner_file, 'r') as infile:
         for line in infile:
             try:
                 page, cat = line.strip().split("\t")
+                if keep:
+                    print "{0}\t{1}".format(page, cat)
                 hu_page = lang_mapping[page]
                 print "{0}\t{1}".format(hu_page, cat)
             except KeyError:
-                sys.stderr.write("Unknown page " + page)
+                sys.stderr.write("Unknown page " + page + "\n")
 
 if __name__ == '__main__':
-    if len(sys.argv) != 2:
-        sys.stderr.write("Usage: {0} ner_list language_mapping\n\n".format(__file__))
+    option_parser = OptionParser()
+    option_parser.add_option("-k", "--keep", dest="keep",
+            action="store_true", default=False,
+            help="keeps the English entries")
+    options, args = option_parser.parse_args()
+
+    if len(args) != 2:
+        sys.stderr.write("Usage: {0} [options] ner_list language_mapping\n\n".format(__file__))
         sys.exit()
 
-    lang_mapping = read_interlang_map(sys.argv[2])
-    convert_ner_lang(sys.argv[1], lang_mapping)
+    lang_mapping = read_interlang_map(args[1])
+    convert_ner_lang(args[0], lang_mapping, options.keep)
 

### 2012-03-12 paper/engNERwiki.tex
mai megbeszeltek es otletek Nothman-bol beleirva

@@
 Creating dictionaries of entities is also a common step of named entity disambiguation \cite{Bunescu:06,Cucerzan:07}.
-Nadeau et al.
-\shortcite{Nadeau:06} use similar lists in an unsupervised NER system, while Kazama and Torisawa \shortcite{Kazama:07} showed that automatic extraction of category labels from Wikipedia improves the accuracy of a supervised NE-tagger.
+Not only supervised systems use lists generated from Wikipedia: Nadeau et al.
+\shortcite{Nadeau:06} use similar lists in an unsupervised NER system.
+
+Wikipedia's knowledge may be incorporated as features in NER learning, e.g.
+Kazama and Torisawa \shortcite{Kazama:07} showed that automatic extraction of category labels from Wikipedia improves the accuracy of a supervised NE-tagger.
+
+A third approach to improve NER with Wikipedia is the automatic creation of training data.
+Richman and Schone \shortcite{Richman:08} built corpora for less commonly taught languages annotated with NE-tags.
+They used the inherent category structure of Wikipedia to determine the NE-type of a proposed entity.
+Nothman et al.~\shortcite{Nothman:08} used a similar method to create a NE-annotated text in English.
+They transformed the Wikipedia links into NE-annotations by classifying the target articles into standard entity classes.
+Their approach to classification is based primarily on category head nouns and definitional opening sentences of articles.
+
+not open access XXX
+
+más nyelvekre is csinálták XXX
+
+domain-specificity: not really
+
+Szeged NER korpusz kicsi, ahhoz képest ez nagy
+
+DBPediát előrébb
 
@@
 
-Automatic creation of training data from Wikipedia for NER:
--classification of articles into CoNLL entity classes \cite{Nothman:08}
+%Automatic creation of training data from Wikipedia for NER:
+%-classification of articles into CoNLL entity classes \cite{Nothman:08}
 
@@
 
+from \cite{Nothman:09}: Exceptions are made for common titlecase words, e.g. I, Mr., June, and sentence-initial words.
+
 % TODO: monetary, etc. units?
@@
 
+In Wikipedia the nationalities (e.g.
+American and Australian) links to the corresponding places (America and Australia, respectively).
+However, in CoNLL scheme these types of entities take parts of MISC category.
+XXX Postprocessing: list look-up (nationalities, adjective forms of person who live in the corresponding country); if the name is POS-tagged as adjective, then it is not a NE
+
+%nem biztos, hogy ide kell
+Link boundaries need to be adjusted, e.g. to remove excess punctuation.
+
 \subsection{Remarks}
@@
 \section{Creation of the Hungarian Corpus}
+
+tokenizálás: built in-house, tailored for Hungarian, because of the specialities of Hungarian (e.g. period after years in date expressions)
+
+ocamorph + hundisambig, a lemmatizálás miatt
+
+a KR-ban nincs jelölve a proper noun külön
+
+DBPedia--CoNLL mapping for Hungarian: az angol oldalak vannak megcímkézve --> csak azok a magyar oldalak vannak letöltve és feldolgozva, amik megvannak angolul is
+	--> future work: kibővíteni azokkal a magyar cikkekkel is, amik nincsenek angolul
+
+filtering: az igés szűrő itt nem lesz
+
+most van: 102 millió szó a disambiguation pages-zel együtt
+az angol: 1,7 milliárd
+
+according to the Szeged NER guideline
+
+a nyelv és népnevek magyarul nem NE-k, vagyis nem kell az annotálásukkal foglalkozni
+
+a tulajdonnevekhez kötőjellel kapcsolódó köznevek kérdésesek (pl. SAP-szoftvercég, Matáv-csoport)
+
+a tulajdonnevekből képzett szavak sem NE-k
 
@@
 
+measuring Wikipedia's inconsistency using n-gram tag variation (\cite{Nothman:09})
+ez jó mérőszám lehet a gold standard korpuszok összehasonlítására
+használhatjuk fordítva is: ha egy adott nucleus ugyanolyan n-gram környezetben nagyon sokszor van X-ként taggelve, és kevésszer Y-ként, akkor az Y-okat is átcsinálhatnánk X-szé
+
 \section{Conclusions and Future Work}
+
+az eredeti DBPedia osztályokat is megtarthatnánk, hiszen nem mindig célravezető a CoNLL osztályok használata, mivel azok elég tág kategóriák, amikbe olyan dolgok is tartoznak, amik nagyon eltérőek is lehetnek
+
+később majd kirakjuk a scripteket is, és akár újra és újra frissíthető lesz
 
@@
 
+\bibitem[\protect\citename{Nothman \bgroup et al.\egroup}2009]{Nothman:09}
+\newblock 2009.
+\newblock Analysing Wikipedia and Gold-Standard Corpora for NER Training.
+\newblock In: {\em Proceedings of the 12th Conference of the European Chapter of the ACL}, pages 612--620.
+
+\bibitem[\protect\citename{Richman and Schone}2008]{Richman:08}
+Alexander E. Richman and Patrick Schone.
+\newblock 2008.
+\newblock Mining Wiki Resources for Multilingual Named Entity Recognition.
+\newblock In: {\em Proceedings of ACL-08: HLT}, pages 1--9.
+
 \bibitem[\protect\citename{Nothman \bgroup et al.\egroup}2008]{Nothman:08}

### 2012-03-12 scripts/extract_dbpedia_types.py
Stuff.

@@ -43,13 +43,13 @@ def merge_pairs(pairs):
         yield (key, list(value))
     return
 
-def filter_general(pairs, hierarchy, what_to_keep=None):
-    """Filters categories from pair's value list whose descendant categories
+def filter_general(lines, hierarchy, what_to_keep=None):
+    """Filters categories from lines' value list whose descendant categories
     are in the list as well. If the set @p what_to_keep is specified, all
     categories not present in in it are discarded."""
     hierarchy.create_isa_map()
-    for pair in pairs:
-        key, values = pair
+    for line in lines:
+        key, values = line
         if what_to_keep is not None:
             values = [v for v in values if v in what_to_keep]
         values.sort(key=hierarchy.get_sorted().get, reverse=True)
@@ -61,21 +61,25 @@ def filter_general(pairs, hierarchy, what_to_keep=None):
                     break
             else:
                 filtered_values.append(curr)
-        yield (key, filtered_values)
+        yield [key, filtered_values]
     return
 
-def filter_type(pairs, type_map):
-    """Filters pairs with certain types. Only pairs whose keys are in the keys
-    of C{type_map} are retained and their keys are renamed according to the
-    mapping in C{type_map}."""
-    for pair in pairs:
+def filter_type(lines, type_map, keep=False):
+    """Filters lines with certain types. Only lines whose types are in the keys
+    of C{type_map} are retained and their types are renamed according to the
+    mapping in C{type_map}. If @p keep is @c True, the types are not renamed,
+    but the new types are appended to the line."""
+    for line in lines:
         new_types = []
-        for type in pair[1]:
+        for type in line[1]:
             new_type = type_map.get(type)
             if new_type is not None:
                 new_types.append(new_type)
         if len(new_types) > 0:
-            yield (pair[0], new_types)
+            if keep:
+                yield line + [new_types]
+            else:
+                yield [line[0], new_types]
     return
 
 def decode_title(title):
@@ -122,7 +126,7 @@ def print_usage_and_exit():
 if __name__ == '__main__':
     import sys
     try:
-        params, args = get_params_sing(sys.argv[1:], 'c:m:', '', 1)
+        params, args = get_params_sing(sys.argv[1:], 'c:m:k', '', 1)
     except ValueError as ve:
         sys.stderr.write(ve + "\n")
         print_usage_and_exit()
[…]

### 2012-03-12 scripts/ner_list_to_hu.py

@@ -3,31 +3,32 @@
 import sys
 from optparse import OptionParser
 
-def read_interlang_map(interlang_file):
+def read_correspondence_map(correspondence_file):
     """Reads the English -> Hungarian article mapping."""
     import gc
     gc.disable()
     ret = {}
-    with open(interlang_file, 'r') as infile:
+    with open(correspondence_file, 'r') as infile:
         for line in infile:
             try:
-                en, hu = line.strip().split("\t")
-                ret[en] = hu
+                orig, others = line.strip().split("\t", 1)
+                ret[orig] = others.split("\t")
             except Exception:
                 sys.stderr.write("Invalid line: " + line)
     gc.enable()
     return ret
 
-def convert_ner_lang(ner_file, lang_mapping, keep=False):
-    """Converts the NER list according to @p lang_mapping."""
+def expand_ner_list(ner_file, correspondence_mapping, keep=False):
+    """Converts the NER list according to @p correspondence_mapping."""
     with open(ner_file, 'r') as infile:
         for line in infile:
             try:
-                page, cat = line.strip().split("\t")
+                page, cat = line.strip().split("\t", 1)
                 if keep:
                     print "{0}\t{1}".format(page, cat)
-                hu_page = lang_mapping[page]
-                print "{0}\t{1}".format(hu_page, cat)
+                new_pages = correspondence_mapping[page]
+                for new_page in new_pages:
+                    print "{0}\t{1}".format(new_page, cat)
             except KeyError:
                 sys.stderr.write("Unknown page " + page + "\n")
 
@@ -39,7 +40,13 @@ if __name__ == '__main__':
     options, args = option_parser.parse_args()
 
     if len(args) != 2:
-        sys.stderr.write("Usage: {0} [options] ner_list language_mapping\n\n".format(__file__))
+        sys.stderr.write("Usage: {0} [options] ner_list correspondence_mapping\n\n".format(__file__))
+        sys.stderr.write("Expands the NER list with pages from a page correspondence mapping ")
+        sys.stderr.write("(e.g. redirects, interlanguage links, etc.). The mapping file is ")
+        sys.stderr.write("tab-separated, the first field being a page already in the list, ")
+        sys.stderr.write("and the others the new pages to be added with the same categories ")
+        sys.stderr.write("as the original.\n\n")
+        sys.stderr.write("To see the options, call the script with the -h flag.\n\n")
         sys.exit()
 
     lang_mapping = read_interlang_map(args[1])

### 2012-03-20 scripts/select_ner_training2.py
Fixed all kinds of problems.

@@ -66,7 +66,7 @@ class SentenceData(object):
     def write(self, out):
         """Writes the sentence to the output file."""
         for word in self.sentence:
-            out.write(word + '\n')
+            out.write(word.encode('iso-8859-2', 'xmlcharrefreplace') + '\n')
         out.write('\n')
         self.num_words += len(self.sentence)
         self.num_train += 1
@@ -85,7 +85,8 @@ class Tries(object):
         """
         self.lt = lt
         self.paths = {}
-        self.words = set()
+        self.words = set()     # The words that occur in paths
+        self.prefixes = set()  # Prefixes for the entries in paths
 
     def add_title(self, title, category):
         """
@@ -94,7 +95,7 @@ class Tries(object):
         """
         par = title.rfind('(')
         if par != -1 and par != 0:
-            title = title[par - 1:]
+            title = title[0 : par]
         words = self.lt.word_tokenize(title)
         self.add_anchor(words, category)
 
@@ -110,10 +111,11 @@ class Tries(object):
 
     def add_path(self, words, category):
         """Adds a path (words of an entity)."""
-        if type(words) == list:
-            words = tuple(words)
+        words = tuple(word.lower() for word in words)
         for word in words:
             self.words.add(word)
+        for i in xrange(len(words)):
+            self.prefixes.add(tuple(words[0 : i + 1]))
         self.paths[words] = category
 
     def get_category(self, words):
@@ -130,6 +132,14 @@ class Tries(object):
         else:
             tpl = tuple(words)
             return tpl, self.paths.get(tpl, None)  # lower() ?
+
+    def is_prefix(self, words, word):
+        """Checks if words + word is a prefix to an entity."""
+        # Last words: raw or lemma
+        words_raw = [token[NERTrainingCallback.RAW].lower() for token in words]
+        tpl1 = tuple(words_raw + [word[NERTrainingCallback.RAW].lower()])
+        tpl2 = tuple(words_raw + [word[NERTrainingCallback.LEMMA].lower()])
+        return tpl1 in self.prefixes or tpl2 in self.prefixes
 
     def clear(self):
         """Clears the trie."""
@@ -212,11 +222,11 @@ class NERTrainingCallback(DefaultConllCallback):
     def fileStart(self, file_name):
         """Opens the output file."""
         DefaultConllCallback.fileStart(self, file_name)
-        self._out = FileWriter(file_name + '.ner').open()
+        self._out = open(file_name + '.ner', 'w')
         if self._keep_discarded:
[…]

### 2012-03-20 scripts/select_ner_training2.py
Comment added.

@@ -337,6 +337,13 @@ class NERTrainingCallback(DefaultConllCallback):
         self.tmp.append(attributes)
 
     def __is_NNP_prefix(self, attributes):
+        """
+        Trie functionality... maybe I should have implemented it correctly.
+        Anyway, this method is here to handle not titlecase words in
+        a NE (Harry Potter and the Deathly Hollows). If such words are met,
+        but the chunk read so far is a prefix to a known entity, we do
+        not throw it away just then.
+        """
         for i in xrange(len(self.tmp)):
             if self._trie.is_prefix(self.tmp[i:], attributes):
                 return True

### 2012-03-21 scripts/select_ner_training2.py
Lots of bug fixes: - punctuation marks stripped from the end of the links (.!?:) - some other stuff I forgot

@@ -31,6 +31,12 @@ from langtools.utils import misc
 # - lowercase link to uppercase LOC (Hun)
 # - redirect mapping
 
+__zero_ner_types = set([0, '0', 'O'])
+def zero_ner_type():
+    return 'O'
+def is_ner_type_zero(ner):
+    return ner in __zero_ner_types
+
 class SentenceData(object):
     def __init__(self):
         self.clear_sentence()
@@ -61,7 +67,7 @@ class SentenceData(object):
 
     def ner_bi(self):
         """Prints the ner type and also converts 0 to O."""
-        return '{0}-{1}'.format(self.bi, self.ner_type) if str(self.ner_type) != '0' else 'O'
+        return '{0}-{1}'.format(self.bi, self.ner_type) if not is_ner_type_zero(self.ner_type) else zero_ner_type()
     
     def write(self, out):
         """Writes the sentence to the output file."""
@@ -213,8 +219,10 @@ class NERTrainingCallback(DefaultConllCallback):
             for line in redirects:
                 kv = line.strip().split("\t")
                 if len(kv) >= 2 and kv[0].lower() in self._gold_map:
-                    redirs = self._redirect_map.get(kv[0].lower(), [])
-                    redirs.extend(l.lower() for l in kv[1:])
+                    redirs = self._redirect_map.get(kv[0].lower(), set())
+                    for l in kv[1:]:
+                        redirs.add(l.lower())
+#                    redirs.add(l.lower() for l in kv[1:])
                     self._redirect_map[kv[0].lower()] = redirs
             gc.enable()
 
@@ -301,7 +309,7 @@ class NERTrainingCallback(DefaultConllCallback):
             else:
 #                            print "NER -> NO/NNP"
                 self.__add_tmp()
-                self._sent.ner_type = 0
+                self._sent.ner_type = zero_ner_type()
                 if self.__is_NNP(attributes):
                     self._mode = NERTrainingCallback.NNP_LINK
                 else:
@@ -316,7 +324,7 @@ class NERTrainingCallback(DefaultConllCallback):
 #                            print "NNP -> NO"
                     self.__add_tmp()
                     self._mode = NERTrainingCallback.NO_LINK
-                    self._sent.ner_type = 0
+                    self._sent.ner_type = zero_ner_type()
 #            elif len(self.tmp) == 8:  # Too long
 #                self.__add_tmp()
[…]

### 2012-03-21 scripts/select_ner_training2.py
Whatever did I commit the last time?

@@ -189,6 +189,7 @@ class NERTrainingCallback(DefaultConllCallback):
         self._trie = trie        # the NERs mentioned on the page
         self._mode = NERTrainingCallback.NO_LINK  # state machine state
         self._punct = set(u'.?!:')  # Sentences should end with one of these
+        self._link_punct = set(u'.?!:,;')  # To strip from the end of links
         self._keep_discarded = keep_discarded
         self._keep_filtered = keep_filtered
 
@@ -368,7 +369,7 @@ class NERTrainingCallback(DefaultConllCallback):
             # The link may end with punctuation marks -- let's remove them!
             puncts = 0
             for token in reversed(self.tmp):
-                if token[NERTrainingCallback.LEMMA] in self._punct:
+                if token[NERTrainingCallback.LEMMA] in self._link_punct:
                     puncts += 1
                 else:
                     break

### 2012-03-22 scripts/extract_date_pages_from_templates.py
Faster date extractor method added.

@@ -1,23 +1,34 @@
-from langtools.io.conll2.conll_reader import DefaultConllCallback, ConllReader
+"""Extracts the date pages from the .templates file. Should be faster than
+from the content files."""
+
+from langtools.utils.file_utils import FileReader
 from langtools.utils.cascading_config import CascadingConfigParser
 import sys
+import re
 
-class DateTemplateCallback(DefaultConllCallback):
+class DateTemplateFinder(object):
+    PAGE_PATTERN = re.compile("^%%#PAGE\s(.+)$")
+
     def __init__(self, date_templates):
         self.date_templates = set(t.strip() for t in date_templates.decode('utf-8').split(','))
-        self.docs = 0
+        self.page = None
 
-    def documentStart(self, title):
-        DefaultConllCallback.documentStart(self, title)
-        self.docs += 1
-        if self.docs % 1000 == 0:
-            sys.stderr.write("{0}\n".format(self.docs))
-    
-    def templates(self, templates):
-        DefaultConllCallback.templates(self, templates)
-
-        if len(self.date_templates & set(templates)) > 0:
-            print "{0}\t{1}".format(self.cc_title.encode('utf-8'), 0)
+    def read(self, template_file):
+        with FileReader(template_file).open() as infile:
+            for line in infile:
+                line = line.strip()
+                m = DateTemplateFinder.PAGE_PATTERN.match(line)
+                if m is not None:
+                    self.page = m.group(1)
+                elif self.page is not None:
+                    if line.startswith("Template\t"):
+                        template = line[9:]
+                        if template in self.date_templates:
+                            print "{0}\t{1}".format(self.page.encode('utf-8'), 0)
+                            self.page = None
+        sys.stderr.write("File {0} parsed.\n".format(template_file))
+        sys.stderr.flush()
+        sys.stdout.flush()
 
 if __name__ == '__main__':
     from optparse import OptionParser
@@ -40,8 +51,7 @@ if __name__ == '__main__':
     config_parser = CascadingConfigParser(args[0])
     config = dict(config_parser.items(options.language + '-wikimedia'))
 
-    dtc = DateTemplateCallback(config['date-templates'])
-    cr = ConllReader([dtc])
+    dtc = DateTemplateFinder(config['date-templates'])
     for wiki_file in args[1:]:
-        cr.read(wiki_file)
+        dtc.read(wiki_file)
 

### 2012-03-22 scripts/select_ner_training_en.py
select_ner_training_en.py added; half of the personal title problem solved.

@@ -3,26 +3,40 @@ considered a training sentence, all links in it are leading to wikipedia pages
 whose NE category is known."""
 
 import sys
+import re
 import os
 import os.path
 if sys.version_info[0] >= 3:
-    from conll_reader3 import DefaultConllCallback, ConllReader
-    from file_utils3 import *
+    from langtools.io.conll2.conll_reader3 import DefaultConllCallback, ConllReader
+    from langtools.utils.file_utils3 import *
 else:
-    from conll_reader import DefaultConllCallback, ConllReader
-    from file_utils import *
-import cmd_utils
+    from langtools.io.conll2.conll_reader import DefaultConllCallback, ConllReader
+    from langtools.utils.file_utils import *
+import langtools.utils.cmd_utils
+from langtools.utils.cascading_config import CascadingConfigParser
+from langtools.utils.useful import all_partitions
+from langtools.utils.language_config import LanguageTools
+from langtools.utils import misc
+from trie import *
 
 # Functionality needed:
-# - language selection (English is different from Hungarian)
 # - from Nothman et al, 2008 (and our ideas, which they blantantly stole in
 #   their paper :)):
-#   .link inference (article title, redirects, words in links)
-#   .capitalization (first words, dates, personal titles) - last two English only
-#   .adjectival forms (nationalities) - English only
+#   .capitalization (personal titles)
 #   .lowercase links: drop / analyse incoming link capitalization
-# We need:
+# Done:
+# - from Nothman et al, 2008 (and our ideas, which they blantantly stole in
+#   their paper :)):
+#   .capitalization (first words, dates)
+#   .link inference (article title, redirects, words in links)
+#   .adjectival forms (nationalities) - English only
 # - redirect mapping
+
+__zero_ner_types = set([0, '0', 'O'])
+def zero_ner_type():
+    return 'O'
+def is_ner_type_zero(ner):
+    return ner in __zero_ner_types
 
 class SentenceData(object):
     def __init__(self):
@@ -53,12 +67,13 @@ class SentenceData(object):
 
     def ner_bi(self):
         """Prints the ner type and also converts 0 to O."""
-        return '{0}-{1}'.format(self.bi, self.ner_type) if self.ner_type != 0 else 'O'
+        return '{0}-{1}'.format(self.bi, self.ner_type) if not is_ner_type_zero(self.ner_type) else zero_ner_type()
     
     def write(self, out):
         """Writes the sentence to the output file."""
[…]

### 2012-03-22 scripts/select_ner_training_hu.py

@@ -261,7 +261,7 @@ class NERTrainingCallback(DefaultConllCallback):
             self._mode = NERTrainingCallback.NO_LINK
             self.tmp = []
             self.sentence_words = []
-    
+
     def word(self, attributes):
         """
         Puts all tokens to self.sentence_words, so that we can do sentence-
@@ -509,7 +509,6 @@ class NERTrainingCallback(DefaultConllCallback):
         """If the sentence contains links only to entities whose types are
         known, it is written to the output file."""
 #        print "NERTrainingCallback:sentenceEnd"
-        sys.stdout.flush()
         if self.cc_field.lower() == 'body':
             if len(self.sentence_words) > 1:
                 if (self.__ends_in_punct() and self.__starts_with_uppercase() and

### 2012-03-22 scripts/select_ner_training_en.py
JJ -> MISC tested.

@@ -320,6 +320,7 @@ class NERTrainingCallback(DefaultConllCallback):
         #print "ATTR", attributes
         if self._mode == NERTrainingCallback.NER_LINK:
             if attributes[1] == 'B-link':
+                print "B-link1", attributes
 #                            print "NER -> NER"
                 self.__add_tmp()
                 self.__start_link(attributes)
@@ -337,6 +338,7 @@ class NERTrainingCallback(DefaultConllCallback):
         elif self._mode == NERTrainingCallback.NNP_LINK:
             if attributes[1] == 'B-link':
 #                            print "NNP -> NER"
+                print "B-link2", attributes
                 self.__add_tmp()
                 self.__start_link(attributes)
             elif not self.__is_NNP(attributes):
@@ -352,6 +354,7 @@ class NERTrainingCallback(DefaultConllCallback):
 #                            print "NNP -> NNP"
         elif self._mode == NERTrainingCallback.NO_LINK:
             if attributes[1] == 'B-link':
+                print "B-link3", attributes
 #                            print "NO -> NER"
                 self.__add_tmp()
                 self.__start_link(attributes)
@@ -518,10 +521,11 @@ class NERTrainingCallback(DefaultConllCallback):
 
     def __start_link(self, attributes):
         """Temporary, to refactor."""
-#        print "NERTrainingCallback:__start_link"
+        print "NERTrainingCallback:__start_link", attributes
         self._sent.ner_type = self.__get_gold(attributes[NERTrainingCallback.LINK])
         self._mode = NERTrainingCallback.NER_LINK
         self._sent.bi = 'B'
+        print "NER", self._sent.ner_type
 
         # Lowercase links point to common nouns, or are non-NER pointers, such
         # as "azonos ci'mu""

### 2012-03-23 scripts/select_ner_training_en.py
Personal titles handled correctly in all cases.

@@ -22,7 +22,6 @@ from trie import *
 # Functionality needed:
 # - from Nothman et al, 2008 (and our ideas, which they blantantly stole in
 #   their paper :)):
-#   .capitalization (personal titles)
 #   .lowercase links: drop / analyse incoming link capitalization
 # Done:
 # - from Nothman et al, 2008 (and our ideas, which they blantantly stole in
@@ -30,6 +29,7 @@ from trie import *
 #   .capitalization (first words, dates)
 #   .link inference (article title, redirects, words in links)
 #   .adjectival forms (nationalities) - English only
+#   .capitalization (personal titles)
 # - redirect mapping
 
 __zero_ner_types = set([0, '0', 'O'])
@@ -107,7 +107,7 @@ class Tries(object):
         self.add_anchor(words, category)
 
     def add_anchor(self, words, category):
-        tpl, old_category = self.get_category(words)
+        tpl, old_category = self.get_category_or_none(words)
         if old_category is None:
             self.add_path(tpl, category)
             if category == 'PER':
@@ -125,7 +125,7 @@ class Tries(object):
             self.prefixes.add(tuple(words[0 : i + 1]))
         self.paths[words] = category
 
-    def get_category(self, words):
+    def get_category_or_none(self, words):
         """Returns the word tuple and its category."""
         if type(words[0]) == list:
             # Last words: raw or lemma
@@ -139,6 +139,12 @@ class Tries(object):
         else:
             tpl = tuple(words)
             return tpl, self.paths.get(tpl, None)  # lower() ?
+
+    def get_category(self, words):
+        tpl, cat = self.get_category_or_none(words)
+        if cat is None:
+            cat = 'UNK'
+        return tpl, cat
 
     def is_prefix(self, words, word):
         """Checks if words + word is a prefix to an entity."""
@@ -320,7 +326,6 @@ class NERTrainingCallback(DefaultConllCallback):
         #print "ATTR", attributes
         if self._mode == NERTrainingCallback.NER_LINK:
             if attributes[1] == 'B-link':
-                print "B-link1", attributes
 #                            print "NER -> NER"
                 self.__add_tmp()
                 self.__start_link(attributes)
@@ -338,7 +343,6 @@ class NERTrainingCallback(DefaultConllCallback):
         elif self._mode == NERTrainingCallback.NNP_LINK:
             if attributes[1] == 'B-link':
 #                            print "NNP -> NER"
-                print "B-link2", attributes
[…]

### 2012-03-23 paper/engNERwiki.tex
Deleted the original content (style guide).

@@
 
-\section{General Instructions}
-
-Manuscripts must be in two-column format. Exceptions to the two-column format include the title, 
-authors' names and complete addresses, which must be centered at the top of the first page, 
-and any full-width figures or tables (see the guidelines in Subsection~\ref{ssec:first}). {\bf Type single-spaced}. 
-Start all pages directly under the top margin. See the guide-lines later regarding formatting the first page. Do not number the pages.
-
-\subsection{Electronically-available resources}
-
-ACL2012 provides this description in \LaTeX2e (acl2012.tex) and PDF format (acl2012.pdf), along with the LATEX2e style file used to format it (acl2012.sty) and an ACL bibliography style (acl.bst).
-These files are all available at \url{http://www.acl2012.org}.
-A Microsoft Word template file (acl2012.dot) is also available at the same URL.
-We strongly recommend the use of these style files, which have been appropriately tailored for the ACL2012 proceedings.
-If you have an option, we recommend that you use the \LaTeX2e version.
-\textbf{If you will be using the Microsoft Word template, we suggest that you anonymize your source file so that the pdf produced does not retain your identity.} This can be done by removing any personal information from your source
-document properties.
-
-\subsection{Format of Electronic Manuscript}
-\label{sect:pdf}
-
-For the production of the electronic manuscript you must use Adobe's
-Portable Document Format (PDF). This format can be generated from
-postscript files: on Linux/Unix systems, you can use {\tt ps2pdf} for this
-purpose; under Microsoft Windows, you can use Adobe's Distiller, or
-if you have {\tt cygwin} installed, you can use {\tt dvipdf} or
-{\tt ps2pdf}.  Note
-that some word processing programs generate PDF which may not include
-all the necessary fonts (esp. tree diagrams, symbols). When you print
-or create the PDF file, there is usually an option in your printer
-setup to include none, all or just non-standard fonts.  Please make
-sure that you select the option of including ALL the fonts.  {\em Before sending it, test your PDF by printing it from a computer different from the one where it was created}. Moreover,
-some word processor may generate very large postscript/PDF files,
-where each page is rendered as an image. Such images may reproduce
-poorly.  In this case, try alternative ways to obtain the postscript
-and/or PDF.  One way on some systems is to install a driver for a
-postscript printer, send your document to the printer specifying
-``Output to a file'', then convert the file to PDF.
-
-Additionally, it is of utmost importance to specify the {\bf US-Letter format} (8.5in $\times$ 11in) when formatting the paper.
-When working with {\tt dvips}, for instance, one should specify {\tt -t letter}.
-
-Print-outs of the PDF file on US-Letter paper should be identical to the
-hardcopy version.  If you cannot meet the above requirements about the
-production of your electronic submission, please contact the
-publication chair above as soon as possible.
-
-\subsection{Layout}
-\label{ssec:layout}
-
-Format manuscripts two columns to a page, in the manner these
-instructions are formatted. The exact dimensions for a page on US-letter
-paper are:
-
-\begin{itemize}
-\item Left and right margins: 1in
-\item Top margin:1in
-\item Bottom margin: 1in
-\item Column width: 3.15in
-\item Column height: 9in
-\item Gap between columns: 0.2in
-\end{itemize}
-
-\noindent Papers should not be submitted on any other paper size.
-If you cannot meet the above requirements about the production of your electronic submission, please contact the publication chair above as soon as possible.
-
-\subsection{Fonts}
-
-For reasons of uniformity, Adobe's {\bf Times Roman} font should be
-used. In \LaTeX2e{} this is accomplished by putting
-
-\begin{quote}
-\begin{verbatim}
-\usepackage{times}
-\usepackage{latexsym}
-\end{verbatim}
-\end{quote}
-in the preamble. If Times Roman is unavailable, use {\bf Computer
-  Modern Roman} (\LaTeX2e{}'s default).  Note that the latter is about
-  10\% less dense than Adobe's Times Roman font.
-
-\begin{table}[h]
-\begin{center}
-\begin{tabular}{|l|rl|}
-\hline \bf Type of Text & \bf Font Size & \bf Style \\ \hline
-paper title & 15 pt & bold \\
-author names & 12 pt & bold \\
-author affiliation & 12 pt & \\
-the word ``Abstract'' & 12 pt & bold \\
-section titles & 12 pt & bold \\
-document text & 11 pt  &\\
-captions & 11 pt & \\
-abstract text & 10 pt & \\
-bibliography & 10 pt & \\
-footnotes & 9 pt & \\
-\hline
-\end{tabular}
-\end{center}
-\caption{\label{font-table} Font guide. }
-\end{table}
-
-\subsection{The First Page}
-\label{ssec:first}
-
-Center the title, author's name(s) and affiliation(s) across both
-columns. Do not use footnotes for affiliations.  Do not include the
-paper ID number assigned during the submission process.
-Use the two-column format only when you begin the abstract.
-
-{\bf Title}: Place the title centered at the top of the first page, in
-a 15 point bold font.  (For a complete guide to font sizes and styles, see Table~\ref{font-table}.)
-Long title should be typed on two lines without
-a blank line intervening. Approximately, put the title at 1in from the
-top of the page, followed by a blank line, then the author's names(s),
-and the affiliation on the following line.  Do not use only initials
-for given names (middle initials are allowed). Do not format surnames
-in all capitals (e.g., ``Zhou,'' not ``ZHOU'').  The affiliation should
-contain the author's complete address, and if possible an electronic
-mail address. Leave about 0.75in between the affiliation and the body
-of the first page.
-The title, author names and addresses should be completely identical to those entered to the electronic paper submission website in order to maintain the consistency of author information among all publications of the conference.
-
-{\bf Abstract}: Type the abstract at the beginning of the first
-column.  The width of the abstract text should be smaller than the
-width of the columns for the text in the body of the paper by about
-0.25in on each side.  Center the word {\bf Abstract} in a 12 point
-bold font above the body of the abstract. The abstract should be a
-concise summary of the general thesis and conclusions of the paper.
-It should be no longer than 200 words. The abstract text should be in 10 point font.
-
-{\bf Text}: Begin typing the main body of the text immediately after
-the abstract, observing the two-column format as shown in
-the present document. Do not include page numbers.
-
-{\bf Indent} when starting a new paragraph. For reasons of uniformity,
-use Adobe's {\bf Times Roman} fonts, with 11 points for text and
-subsection headings, 12 points for section headings and 15 points for
-the title.  If Times Roman is unavailable, use {\bf Computer Modern
-  Roman} (\LaTeX2e's default; see section \ref{sect:pdf} above).
-Note that the latter is about 10\% less dense than Adobe's Times Roman
-font.
-
-\subsection{Sections}
-
-{\bf Headings}: Type and label section and subsection headings in the
-style shown on the present document.  Use numbered sections (Arabic
-numerals) in order to facilitate cross references. Number subsections
-with the section number and the subsection number separated by a dot,
-in Arabic numerals. Do not number subsubsections.
-
-{\bf Citations}: Citations within the text appear
-in parentheses as~\cite{Gusfield:97} or, if the author's name appears in
-the text itself, as Gusfield~\shortcite{Gusfield:97}.
-Append lowercase letters to the year in cases of ambiguities.
-Treat double authors as in~\cite{Aho:72}, but write as in~\cite{Chandra:81} when more than two authors are involved.
-Collapse multiple citations as in~\cite{Gusfield:97,Aho:72}.
-Also refrain from using full citations as sentence constituents.
-We suggest that instead of
-\begin{quote}
-  ``\cite{Gusfield:97} showed that ...''
-\end{quote}
-you use
-\begin{quote}
-``Gusfield \shortcite{Gusfield:97}   showed that ...''
-\end{quote}
-
-If you are using the provided \LaTeX{} and Bib\TeX{} style files, you
-can use the command \verb|\newcite| to get ``author (year)'' citations.
-
-As reviewing will be double-blind, the submitted version of the papers should not include the
-authors' names and affiliations. Furthermore, self-references that
-reveal the author's identity, e.g.,
-\begin{quote}
-``We previously showed \cite{Gusfield:97} ...''
-\end{quote}
-should be avoided. Instead, use citations such as
-\begin{quote}
-``Gusfield \shortcite{Gusfield:97}
-previously showed ... ''
-\end{quote}
-
-Please do not  use anonymous
-citations and  do not include acknowledgements when submitting your papers. Papers that do not conform
-to these requirements may be rejected without review.
-
-\textbf{References}: Gather the full set of references together under
-the heading {\bf References}; place the section before any Appendices,
-unless they contain references. Arrange the references alphabetically
-by first author, rather than by order of occurrence in the text.
-Provide as complete a citation as possible, using a consistent format,
-such as the one for {\em Computational Linguistics\/} or the one in the
-{\em Publication Manual of the American
-Psychological Association\/}~\cite{APA:83}.  Use of full names for
-authors rather than initials is preferred.  A list of abbreviations
-for common computer science journals can be found in the ACM
-{\em Computing Reviews\/}~\cite{ACM:83}.
-
-The \LaTeX{} and Bib\TeX{} style files provided roughly fit the
-American Psychological Association format, allowing regular citations,
-short citations and multiple citations as described above.
-
-{\bf Appendices}: Appendices, if any, directly follow the text and the
-references (but see above).  Letter them in sequence and provide an
-informative title: {\bf Appendix A. Title of Appendix}.
-
-\textbf{Acknowledgment} sections should go as a last section immediately
-before the references. Do not number the acknowledgement section.
-
-\subsection{Footnotes}
-
-{\bf Footnotes}: Put footnotes at the bottom of the page. They may
-be numbered or referred to by asterisks or other
-symbols.\footnote{This is how a footnote should appear.} Footnotes
-should be separated from the text by a line.\footnote{Note the
-line separating the footnotes from the text.}  Footnotes should be in 9 point font.
-
-\subsection{Graphics}
-
-{\bf Illustrations}: Place figures, tables, and photographs in the
-paper near where they are first discussed, rather than at the end, if
-possible.  Wide illustrations may run across both columns and should be placed at
-the top of a page. Color illustrations are discouraged, unless you have verified that
-they will be understandable when printed in black ink.
-
-{\bf Captions}: Provide a caption for every illustration; number each one
-sequentially in the form:  ``Figure 1. Caption of the Figure.'' ``Table 1.
-Caption of the Table.''  Type the captions of the figures and
-tables below the body, using 10 point text.
-
-\section{Translation of non-English Terms}
-
-It is also advised to supplement non-English characters and terms
-with appropriate transliterations and/or translations
-since not all readers understand all such characters and terms.
-
-Inline transliteration or translation can be represented in
-the order of: original-form transliteration ``translation''.
-
-\section{Length of Submission}
-\label{sec:length}
-
-Long papers may consist of up to eight (8) pages of content and an
-additional two (2) pages of references, and short papers may consist
-of up to four (4) pages of content and two (2) additional pages of
-references.  Papers that do not conform to the specified length and
-formatting requirements are subject to re-submission.
-
-\section{Other Issues}
-
-Those papers that had software and/or dataset submitted for the review process, should also submit it 
-with the camera-ready paper. Besides, the software and/or dataset should not be anonymous.
-
-Please note that the publications of ACL2012 will be publicly available at ACL Anthology 
-(http://aclweb.org/anthology-new/) on July 2nd, 2012, one week before the start of the conference. 
-Since some of the authors may have plans to file patents related to their papers in the conference, 
-we are sending this reminder that July 2nd, 2012 may be considered to be the official publication date, 
-instead of the opening day of the conference.
-
 %\section*{Acknowledgments}
@@
 %The authors are grateful to Attila Zséder and Gábor Recski for their respective work on Wikipedia parsing and {\tt hunner}.
-%Do not number the acknowledgment section. Do not include this section when submitting your paper for review.
+% + CE(A?)SAR
 

### 2012-03-23 paper/engNERwiki.tex
A few lines added.

@@
 Named Entity Recognition (NER), the task of identifying Named Entities (NEs) in unstructured texts and classifying them into pre-selected classes, is one of the most important subtasks in many NLP tasks, such as Information Retrieval, Information Extraction or Machine Translation.
-The NER task was introduced with the 6th Message Understanding Conference (MUC) in 1995.
+The NER task was introduced with the 6th Message Understanding Conference (MUC) in 1995 (CITE).
 In MUC shared tasks the NER consists of three subtasks: entity names, temporal expressions and number expressions.
 However there is a general agreement in the NER community about the inclusion of temporal expressions and some numerical expressions such as amounts of money and other types of units, the most studied types are names of persons, locations and organizations.
-The fourth type, called "miscellaneous", is used in the CoNLL NER tasks in 2002 and 2003, and includes proper names falling outside the three classic types.
+The fourth type, called "miscellaneous", is used in the CoNLL NER tasks in 2002 and 2003 (CITE), and includes proper names falling outside the three classic types.
 Since then, MUC and CoNLL datasets and annotation schemes have been the major standards applied in the field of NER.
@@
 
+\section{Corpus Description}
+
+% TODO: pls írd meg az első pár mondatot és nevezd át a sectiont!
+
+The corpora are available in the tab separated format used by ConLL (XXX, CITE).
+In this format, one line contains one word, its linguistic features and the named entity tags.
+Sentence boundaries are marked by empty lines.
+The linguistic features include the lemmatized form of the word, its part-of-speech and chunk tags.
+Two named entity tags are included with each word: one is the most specific DBpedia category it belongs to, if any; the other is the ConLL named entity tag.
+While the named entity tags can be considered "silver", the linguistic features are provided on a "best-effort" basis.
+
+The corpus is available under the Creative Commons Attribution-Sharealike 3.0 Unported License (CC-BY-SA), the same license under which the text of Wikipedia is released.
+The data files can be freely downloaded from ...\footnote{To be filled in in the camera-ready version.}.
+
 \section{Creation of the Corpus}  % Dzsííízösz
@@
 % Introduction: Wikipedia + DBpedia
-The corpora are available in the tab separated format used by ConLL (XXX, CITE).
-In this format, one line contains one word, its linguistic features and the named entity tags.
-Sentence boundaries are marked by empty lines.
-The linguistic features include the lemmatized form of the word, part-of-speech tag and chunk tag.
-Two named entity tags are included with each word: one is the most specific DBpedia category it belongs to, if any; the other is the ConLL named entity tag.
-While the named entity tags can be considered "gold", the linguistic features are provided on a "best-effort" basis (XXX: check the LREC paper to find a better wording for this).
-XXX: past or present tense?
-
-We decided to use Wikipedia because
-
-The categories of Wikipedia are too numerous and ; not a tree;
 
@@
 XXX!!!
-If no match is found, the entity is tagged with \texttt{0}.
+If no match is found, the entity is tagged with \texttt{O}.
 Since we take advantage of the inheritance hierarchy, the mapping list remains short: it contains only the root classes of the main sub-hierarchies, exceptions of the like XXX mentioned above and the various classes that belong to the MISC category according to the ConLL annotation guideline.
 
-As of version 3.7, the DBpedia ontology allows multiple superclasses, making the hierarchy a directed acyclic graph.
+As of version 3.7, the DBpedia ontology allows multiple superclasses, making the hierarchy a directed acyclic graph (CITE).
 This introduces the problem of selecting the right superclass, and hence, ConLL mapping for classes with more than one parent.
@@
 
-\subsection{Remarks}
+\section{Creation of the Hungarian Corpus}
 
-In this paper, we describe % not only ConLL? Do we need this?
+In addition to differences in grammar,
 
-\section{Creation of the Hungarian Corpus}
+The Szeged NER guideline relies heavily on the rules of capitalization to decide which words should be marked as named entities.
+In Hungarian, apart from the first word in the sentence, only proper nouns are capitalized.
+Thus, the following concepts are not considered named entities in Hungarian XXX:
+\begin{itemize}
+\item Names of languages and ethnic groups
+\item Words created from proper nouns through derivation, most prominently
+\item Adjectival form of location names
+\item Months % angolban sem
+\end{itemize}
 
@@
 
-a nyelv és népnevek magyarul nem NE-k, vagyis nem kell az annotálásukkal foglalkozni
-
 a tulajdonnevekhez kötőjellel kapcsolódó köznevek kérdésesek (pl. SAP-szoftvercég, Matáv-csoport)
-
-a tulajdonnevekből képzett szavak sem NE-k
 
@@
 % Corpus sizes, performance, tagging conventions
+
+% Hun -> not utf-8, loss of precision?
 
@@
 
-\bibitem[\protect\citename{Aho and Ullman}1972]{Aho:72}
-Alfred~V. Aho and Jeffrey~D. Ullman.
-\newblock 1972.
-\newblock {\em The Theory of Parsing, Translation and Compiling}, volume~1.
-\newblock Prentice-{Hall}, Englewood Cliffs, NJ.
-
-\bibitem[\protect\citename{{American Psychological Association}}1983]{APA:83}
-{American Psychological Association}.
-\newblock 1983.
-\newblock {\em Publications Manual}.
-\newblock American Psychological Association, Washington, DC.
-
-\bibitem[\protect\citename{{Association for Computing Machinery}}1983]{ACM:83}
-{Association for Computing Machinery}.
-\newblock 1983.
-\newblock {\em Computing Reviews}, 24(11):503--512.
-
-\bibitem[\protect\citename{Chandra \bgroup et al.\egroup }1981]{Chandra:81}
-Ashok~K. Chandra, Dexter~C. Kozen, and Larry~J. Stockmeyer.
-\newblock 1981.
-\newblock Alternation.
-\newblock {\em Journal of the Association for Computing Machinery},
-  28(1):114--133.
-
-\bibitem[\protect\citename{Gusfield}1997]{Gusfield:97}
-Dan Gusfield.
-\newblock 1997.
-\newblock {\em Algorithms on Strings, Trees and Sequences}.
-\newblock Cambridge University Press, Cambridge, UK.
+\bibitem[\protect\citename{Bizer \bgroup et al.\egroup}2009]{Bizer:09} 
+Christian Bizer, Jens Lehmann, Georgi Kobilarov, Sören Auer, Christian Becker, Richard Cyganiak, Sebastian Hellmann.
+\newblock 2009.
+\newblock DBpedia -- A Crystallization Point for the Web of Data. 
+\newblock In: {\em Journal of Web Semantics: Science, Services and Agents on the World Wide Web}, Issue 7, pages 154--165.
 

### 2012-03-23 scripts/select_ner_training_en.py
- Fixed bugs in select_ner_training_en.py - Updated README with instructions for English

@@ -442,7 +442,8 @@ class NERTrainingCallback(DefaultConllCallback):
                     if (self.tmp[last - 1][NERTrainingCallback.POS].startswith(u'J')
                         and not self.tmp[last - 1][NERTrainingCallback.RAW].lower()
                             in self.tmp[last - 1][NERTrainingCallback.LINK].lower()
-                        and self._sent.ner_type != 'UNK'):
+                        and self._sent.ner_type != 'UNK'
+                        and self._sent.ner_type != zero_ner_type()):
                         sys.stderr.write("Adj entity: {0}\n".format(self.tmp[0 : last]))
                         self._sent.ner_type = 'MISC'
                     # Add the link stripped of punctuation marks
@@ -454,14 +455,15 @@ class NERTrainingCallback(DefaultConllCallback):
                         self._sent.append(attributes)
                     self._trie.add_anchor(self.tmp[begin : last], self._sent.ner_type)
 
+                # Unknown link: we must throw the sentence away
+                if self._sent.ner_type == 'UNK':
+                    self._sent.links_lost += 1
+
             # And the rest as regular text
             self._sent.ner_type = zero_ner_type()
             for attributes in self.tmp[last:]:
                 self._sent.append(attributes)
 
-            # Unknown link: we must throw the sentence away
-            if self._sent.ner_type == 'UNK':
-                self.links_lost += 1
         elif self._mode == NERTrainingCallback.NO_LINK:
             for attributes in self.tmp:
                 self._sent.append(attributes)
@@ -531,7 +533,7 @@ class NERTrainingCallback(DefaultConllCallback):
                                         self._sent.append(attributes)
 
                                     if self._sent.ner_type == 'UNK':
-                                        self.links_lost += 1
+                                        self._sent.links_lost += 1
 
                                 # And the rest as regular text
                                 self._sent.ner_type = zero_ner_type()
@@ -581,11 +583,13 @@ class NERTrainingCallback(DefaultConllCallback):
         # Lowercase links point to common nouns, or are non-NER pointers, such
         # as "azonos ci'mu""
         if attributes[NERTrainingCallback.RAW][0].islower():
+            print "ISLOWER", attributes, self._sent.ner_type
             # All lower-case links are of type 0
             self._sent.ner_type = zero_ner_type()
-        else:
-            if not is_ner_type_zero(self._sent.ner_type):
-    #            print "LINK", attributes, self._sent.ner_type
[…]

### 2012-03-24 paper/engNERwiki.tex
Added Hungarian DBpedia subsection.

@@
 
-The paper is structured as follows. XXX
+The paper is structured as follows. XXX In section~\ref{sec:hun}, we show how the algorithm was adapted to Hungarian.
 
@@
 \section{Creation of the Hungarian Corpus}
+\label{sec:hun}
+
+The procedure described in the last section was used to generate the Hungarian corpus as well.
+However, the differences between English and Hungarian prompted XXX several modifications in its steps.
+In this section we describe how the algorithm was adapted to Hungarian.
+XXX!!!
+
+in tool a
+
+\subsection{Data availability}
+
+For some of the bigger languages, DBpedia has internationalized "chapters" -- data dumps based on the Wikipedia of that particular language.
+Hungarian, being a relatively small language XXX, does not have a separate chapter.
+Instead, the Hungarian entity list comprises of the pages in the English list that have their equivalents in the Hungarian Wikipedia.
+Two consequences follow.
+First, in order to identify which pages denote entities in the Hungarian Wikipedia, an additional step is required, in which the Hungarian equivalents of the English pages are added to the entity list.
+The English titles are retained, as due to the relatively small size of the Hungarian Wikipedia, in-article links sometimes point to English articles.
+Second, entities without a page in the English Wikipedia are absent from the entity list.
+This gives rise to two potential problems.
+One is that compared to English, the list is relatively shorter: the entity/page ratio is 12.12\%, as opposed to the 37.66\% of the English Wikipedia.
+The other, since mostly Hungarian people, places and organizations are missing, a NER tagger that takes the surface forms of words into account might be mislead as to the language model of entity names.
+To overcome these problems, the list has to be extended with Hungarian entity pages that do not have a corresponding English page.
+We leave this for future work.
 

### 2012-03-24 paper/engNERwiki.tex
Another subsection.

@@
 
-The corpus is based on the Wikipedia snapshot as of XXX.
-The xml files were parsed with the XXX (Attila) parser.
+The English corpus is based on the Wikipedia snapshot as of XXX.
+The xml files were parsed with the mwlib (CITE) parser.
 The raw text was tokenized with a modified version of the XXX sentence tokenizer in NLTK (CITE).
@@
 
+\subsection{Tools}
+
+While NLTK has excellent tools for English, Hungarian is not supported.
+Therefore, we worked with the tools developed by the Hungarian computational linguist community.
+For sentence tokenization we used an in-house statistical chunker tailored for Hungarian.
+It has been trained on the Szeged corpus and it handles the peculiarities of Hungarian ortography, such as the periods placed after numbers in date expressions.
+Although HunPos, as its name implies, is fully capable of POS tagging Hungarian text, we opted for Hunmorph (CITE), a morphological analyzer that also XXX lemmatizes the words.
+Hundisambig (CITE) was used to select the right analysis based on the word context\footnote{Incidentally, Hundisambig relies on HunPos for its HMM model.}.
+
+The Hungarian part-of-speech taggers output KR codes (CITE).
+In addition to the POS category, the KR code also includes inflection information, such as the number, person and case XXX, which makes it much better suited to agglutinative languages than Penn Treebank POS tags.
+One shortcoming of the KR code is that it does not differenciate between common and proper nouns.
+Since in Hungarian only proper nouns are capitalized, we can usually decide whether a noun is proper or not based on its ortography.
+However, this rule cannot be used if the noun is at the beginning of a sentence, as words at the beginning of sentences are always capitalized.
+Hence, all sentences that begin with nouns have been removed from the corpus.
+
 \subsection{Data availability}
@@
 \end{itemize}
-
-tokenizálás: built in-house, tailored for Hungarian, because of the specialities of Hungarian (e.g. period after years in date expressions)
-
-ocamorph + hundisambig, a lemmatizálás miatt
-
-a KR-ban nincs jelölve a proper noun külön
-
-DBPedia--CoNLL mapping for Hungarian: az angol oldalak vannak megcímkézve --> csak azok a magyar oldalak vannak letöltve és feldolgozva, amik megvannak angolul is
-	--> future work: kibővíteni azokkal a magyar cikkekkel is, amik nincsenek angolul
 

### 2012-03-24 paper/engNERwiki.tex
Related work --> Wikipedia and NER Evaluation intro

--- 2012-03-24 paper/engNERwiki.tex
@@
 Named Entity Recognition (NER), the task of identifying Named Entities (NEs) in unstructured texts and classifying them into pre-selected classes, is one of the most important subtasks in many NLP tasks, such as Information Retrieval, Information Extraction or Machine Translation.
-The NER task was introduced with the 6th Message Understanding Conference (MUC) in 1995 (CITE).
+The NER task was introduced with the 6th Message Understanding Conference (MUC) in 1995 \cite{Grishman:96}.
 In MUC shared tasks the NER consists of three subtasks: entity names, temporal expressions and number expressions.
 However there is a general agreement in the NER community about the inclusion of temporal expressions and some numerical expressions such as amounts of money and other types of units, the most studied types are names of persons, locations and organizations.
-The fourth type, called "miscellaneous", is used in the CoNLL NER tasks in 2002 and 2003 (CITE), and includes proper names falling outside the three classic types.
+The fourth type, called "miscellaneous'', is used in the CoNLL NER tasks in 2002 \cite{Tjong:02} and 2003 \cite{Tjong:03}, and includes proper names falling outside the three classic types.
 Since then, MUC and CoNLL datasets and annotation schemes have been the major standards applied in the field of NER.
@@
 We built NE-tagged corpora for English and Hungarian -- fully automatically generated from Wikipedia.
-Our goal was to create gold standard corpora which can serve as train, devel and test datasets for automatic NE-taggers.
+%Our goal was to create gold standard corpora which can serve as train, devel and test datasets for automatic NE-taggers.
 
@@
 
-\section{Related Work}
+\section{Wikipedia and NER}
 % Wikify!
@@
 Wikipedia is recognized as one of the largest available collection of entities, and also as an external knowledge to improve the accuracy of NER.
-The most obvious utilization of Wikipedia for NER is extracting gazetteers containing person names, locations or organizations \cite{Toral:06}.
+The most obvious utilization of Wikipedia for NER is extracting gazetteers containing person names, locations or organizations (e.g.
+Toral and Munoz \shortcite{Toral:06}).
 Creating dictionaries of entities is also a common step of named entity disambiguation \cite{Bunescu:06,Cucerzan:07}.
@@
 
-not open access XXX
+To our knowledge, such type of automatically built corpora are not freely available, with the only one exception of the Semantically Annotated Snapshot of the English Wikipedia \cite{Zaragoza:07}.
+It provides a wide range of linguistic information: POS tags, dependency labels, WordNet super senses and NE annotation according to WSJ and CoNLL tagsets.
+However, automatically NE-annotated corpus for Hungarian does not exist.
+Now we present freely available, automatically NE-tagged corpora for English and Hungarian.
 
-más nyelvekre is csinálták XXX
+%Szeged NER korpusz kicsi, ahhoz képest ez nagy
 
-domain-specificity: not really
-
-Szeged NER korpusz kicsi, ahhoz képest ez nagy
-
-DBPediát előrébb
-
-%Exploiting Wikipedia as external knowledge for NER:
-%%-creating entity lists  
-%%-disambiguating named entities
-%%-automatic extraction of category labels from Wikipedia to improve the accuracy of a CRF-based NE-tagger \cite{Kazama:07}
-%-entity ranking
-%-typing named entities
-
-%Automatic creation of training data from Wikipedia for NER:
-%-classification of articles into CoNLL entity classes \cite{Nothman:08}
+%DBPediát előrébb
 
@@
 \subsection{The DBpedia -- ConLL mapping}
+\label{dbpedia}
+% Introduction + entity -> class -> category.
 
-% Introduction + entity -> class -> category.
+%Despite the aforementioned methods, we applied a new approach to recognize and classify NEs in the corpora generated from Wikipedia.
+We used the DBpedia ontology classes to map Wikipedia categories to standard NE tags (see more details in section \ref{dbpedia}).
 
@@
 
+Having the obvious advantages, an automatically generated corpus can not serve as a gold standard dataset.
+Then what can we do with these so called "silver standard corpora" (SSC)?
+They can be very useful for improving NER in more ways: (a) for less commonly used languages, they can serve as training corpora in absence of gold standard datasets; (b) they can serve as (supplementary or independent) training sets for domains differing from newswire; (c) they can be sources of huge entity lists, and (d) feature extraction.
+
 measuring Wikipedia's inconsistency using n-gram tag variation (\cite{Nothman:09})
@@
 \begin{thebibliography}{}
+
+\bibitem[\protect\citename{Zaragoza \bgroup et al.\egroup}2007]{Zaragoza:07}
+H. Zaragoza and J. Atserias and M. Ciaramita and G. Attardi.
+\newblock 2007.
+\newblock Semantically Annotated Snapshot of the English Wikipedia v.1 (SW1).
+\newblock \url{http://www.yr-bcn.es/semanticWikipedia}
+
+\bibitem[\protect\citename{Tjong Kim Sang}2002]{Tjong:02}
+Erik F. Tjong Kim Sang.
+\newblock 2002.
+\newblock Introduction to the CoNLL-2002 shared task: Language-independent named entity recognition.
+\newblock In: {\em Proceedings of the 6th Conference on Natural Language Learning}, pages 1--4, Taipei, Taiwan.
+
+\bibitem[\protect\citename{Tjong Kim Sang and De Meulder}2003]{Tjong:03}
+Erik F. Tjong Kim Sang and Fien De Meulder.
+\newblock 2003.
+\newblock Introduction to the CoNLL-2003 shared task: Language-independent named entity recognition. 
+\newblock In: {\em Proceedings of the 7th Conference on Natural Language Learning}, pages 142--147, Edmonton, Canada.
+
+\bibitem[\protect\citename{Grishman and Sundheim}1996]{Grishman:96}
+Ralph Grishman and B. Sundheim.
+\newblock 1996.
+\newblock Message Understanding Conference -- 6. 
+\newblock In: {\em Proc. International Conference on Computational Linguistics}.
 

### 2012-03-24 paper/engNERwiki.tex
Changes.

--- 2012-03-24 paper/engNERwiki.tex
@@
 \section{Creation of the Corpus}  % Dzsííízösz
+\label{sec:create}
 
@@
 It has been trained on the Szeged corpus and it handles the peculiarities of Hungarian ortography, such as the periods placed after numbers in date expressions.
-Although HunPos, as its name implies, is fully capable of POS tagging Hungarian text, we opted for Hunmorph (CITE), a morphological analyzer that also XXX lemmatizes the words.
+Although HunPos, as its name implies, is fully capable of POS tagging Hungarian text, we opted for Hunmorph (CITE), a morphological analyzer that, apart from PAS tagging, also lemmatizes the words.
 Hundisambig (CITE) was used to select the right analysis based on the word context\footnote{Incidentally, Hundisambig relies on HunPos for its HMM model.}.
@@
 One shortcoming of the KR code is that it does not differenciate between common and proper nouns.
-Since in Hungarian only proper nouns are capitalized, we can usually decide whether a noun is proper or not based on its ortography.
+Since in Hungarian only proper nouns are capitalized, we can usually decide whether a noun is proper or not based on the initial letter.
 However, this rule cannot be used if the noun is at the beginning of a sentence, as words at the beginning of sentences are always capitalized.
@@
 
-\subsection{Data availability}
+\subsection{Data Availability}
 

### 2012-03-24 paper/engNERwiki.tex
Wikipedia and NER fejezet befejezve plusz egyed aprosagok

--- 2012-03-24 paper/engNERwiki.tex
@@
 Researchers apply the Web-as-corpus approach, often dealing with such datasets as the Google 5G 1T collection or DBPedia.
-Thanks for the huge amount of Wikipedia articles, the possibility of building large enough corpora is given for less resourced languages, as well.
 
@@
 
+Despite the aforementioned methods, we applied a new approach to recognize and classify NEs in the corpora generated from Wikipedia.
+We used the DBpedia ontology classes to map Wikipedia categories to standard NE tags (see more details in section \ref{dbpedia}).
+
 To our knowledge, such type of automatically built corpora are not freely available, with the only one exception of the Semantically Annotated Snapshot of the English Wikipedia \cite{Zaragoza:07}.
 It provides a wide range of linguistic information: POS tags, dependency labels, WordNet super senses and NE annotation according to WSJ and CoNLL tagsets.
-However, automatically NE-annotated corpus for Hungarian does not exist.
-Now we present freely available, automatically NE-tagged corpora for English and Hungarian.
+However, they followed a different way: the NEs were tagged by open source softwares.
+Our corpora will be available under the same license as Wikipedia's one (Creative Commons Attribution-Sharealike 3.0 Unported License).
 
-%Szeged NER korpusz kicsi, ahhoz képest ez nagy
+Thanks for the huge amount of Wikipedia articles, the possibility of building large enough corpora is given for less resourced languages, as well.
+Since our method is mainly language-independent, it can be easily applied for languages which are structurally differing from English.
+Hungarian is a highly agglutinative language, with free word order and other particularities detailed later in section \ref{sec:hun}.
+There are human annotated CoNLL-style datasets, but this presented here is the first large enough automatically NE-annotated corpus for Hungarian.
 
-%DBPediát előrébb
-
-\section{Corpus Description}
-
-% TODO: pls írd meg az első pár mondatot és nevezd át a sectiont!
-
-The corpora are available in the tab separated format used by ConLL (XXX, CITE).
-In this format, one line contains one word, its linguistic features and the named entity tags.
-Sentence boundaries are marked by empty lines.
-The linguistic features include the lemmatized form of the word, its part-of-speech and chunk tags.
-Two named entity tags are included with each word: one is the most specific DBpedia category it belongs to, if any; the other is the ConLL named entity tag.
-While the named entity tags can be considered "silver", the linguistic features are provided on a "best-effort" basis.
-
-The corpus is available under the Creative Commons Attribution-Sharealike 3.0 Unported License (CC-BY-SA), the same license under which the text of Wikipedia is released.
-The data files can be freely downloaded from ...\footnote{To be filled in in the camera-ready version.}.
-
-\section{Creation of the Corpus}  % Dzsííízösz
+\section{Creating the English Corpus}  % Dzsííízösz
 
@@
 Wikipedia proves to be a rich source of information about its entities; aside from the article text, a huge amount of data is encompassed in infoboxes, templates and the category structure.
-Since our goal is to create a NER training corpus, our interest is limited to the article text.
+Since our goal is to create a NE-annotated corpus containing large amount of texts, our interest is limited to the article text.
 Templates are removed from the articles; however, the names of the templates have been retained for XXX, but all data (including infobox)
@@
 Beside regular articles, Wikipedia also contains redirect and disambiguation pages, which require special treatment.
-Redirect pages are empty and are safe to ignore altogether.
+Redirect pages are empty, and are safe to ignore altogether.
 Disambiguation pages contain text, usually in the form of an enumeration of page links with a short description.
-Full sentences are scarce and the structure of these pages makes it hard to extract valid training sentences from them.
-Therefore, disambiguation pages have been dropped from the corpus as well.
+Full sentences are scarce, and the structure of these pages makes it hard to extract valid training sentences from them.
+Therefore, disambiguation pages are dropped from the corpus, as well.
 
 The English corpus is based on the Wikipedia snapshot as of XXX.
-The xml files were parsed with the mwlib (CITE) parser.
+The XML files were parsed with the mwlib (CITE) parser.
 The raw text was tokenized with a modified version of the XXX sentence tokenizer in NLTK (CITE).
-The words were lemmatized with XXX and POS tagged with HunPos (CITE).
-Chunking was performed with HunChunk (CITE).
-Experiments have shown that HunPos obtains STAT f-score on the Brown corpus; HunChunk performs at around STAT, CITE?.
+The words were lemmatized with XXX and POS tagged with HunPos \cite{Halacsy:07}.
 
-\subsection{Extraction of the Training Sentences}
+%Chunking was performed with HunChunk (CITE). Experiments have shown that HunPos obtains STAT f-score on the Brown corpus; HunChunk performs at around STAT, CITE?.
 
-A NER training corpus consists of sentences that accurately tag named entities occurring in them.
-In order to automatically prepare such sentences, two tasks need to be performed: identification of entities in the sentence and tagging them with the correct tag.
+\subsection{Extracting and Filtering Sentences}
+
+A NER corpus consists of sentences that accurately tag NEs occurring in them.
+In order to automatically prepare such sentences, two tasks need to be performed: identifying the entities in the sentence and tagging them with the correct tag.
 Sentences, for which accurate tagging could not be accomplished, must be removed from the corpus.
-In this section, we describe how these steps were implemented, as well as the post-filtering we employed to dispose of correctly tagged but unusable training sentences.
+In this section, we describe how these steps were implemented, as well as the post-filtering we employed to dispose of correctly tagged, but unusable training sentences.
 
-\subsubsection{Entity identification}
+\subsubsection{Identifying the entities}
 % TODO: identification and tagging happens in parallel, else the recall oriented approach cannot be explained.
@@
 A link to an article can then be perceived as a mapping that identifies its anchor text with a particular entity.
-There are exceptions to this rule, however.
+However, there are exceptions to this rule.
 Disambiguation pages are used when the page title is ambiguous, and can refer to several entities.
@@
 
-Not all entities are marked with links, however.
+However, not all entities are marked with links.
 Authors may neglect to add XXX references to their text.
 Furthermore, in a Wikipedia article XXX, typically only the first occurrence of a particular entity is linked to the corresponding page.
-Subsequent mentions are unmarked and often incomplete -- e.g.
-family names are used instead of full names.
+Subsequent mentions are unmarked and often incomplete -- e.g.~family names are used instead of full names.
 We devised XXX two methods to account for unmarked entities: one that emphasizes XXX precision, and the other, recall.
@@
 
-The precision-oriented approach seeks to minimize the risk of a word getting incorrectly tagged XXX in the corpus.
-A short study of the ConLL annotation guidelines reveals that the words in the entities of interest to us all start with capital letter.
-According to this method XXX, all sentences that contain words that start with capital letter, but are not part of the anchor text of a link, XXX are discarded.
+The precision-oriented approach seeks to minimize the risk of words getting incorrectly tagged XXX in the corpus. Names and their parts mostly start with capital letter, thus 
+%A short study of the CoNLL annotation guidelines reveals that the words in the entities of interest to us all start with capital letter. According to this method XXX, 
+all sentences that contain capitalized words, but are not part of the anchor text of a link, XXX were discarded.
 The rationale behind this approach is that the sheer size of Wikipedia guarantees that we end up with a sizeable corpus even with such a strict policy.
@@
 
-The recall-oriented method tries to associate unmarked entities with entities already identified on the same page.
-Only if the attempt fails is the sentence that contains the words discarded.
-This approach aims to generate more training sentences with as little loss of precision as possible.
+The recall-oriented method tries to associate unmarked entities with entities already identified on the same page. Only if the attempt fails is the sentence that contains the words discarded. 
+%Ez utóbbi biztosan jó mondat?
+This approach aims to generate more sentences with as little loss of precision as possible.
 
-The algorithm maintains a list of entities already indentified on the page. The list includes
+The algorithm maintains a list of entities already indentified on the page. The list includes:
 \begin{itemize}
@@
 
-When an unmarked entity is encountered, the list is consulted to find an item that contains it in its entirety; i.e.
-all of its words and in the same order.
-If either exactly one such item is found, or the matching entities in the list all have the same NER tag, the unmarked entity is tagged accordingly.
+When an unmarked entity is encountered, the list is consulted to find an item that contains it in its entirety; i.e.~all of its words, in the same order.
+If either exactly one such item is found, or the matching entities in the list all have the same NE tag, the unmarked entity is tagged accordingly.
 Otherwise the sentence is discarded.
 
-Both approaches are evaluated in the next chapter.
+%Both approaches are evaluated in the next chapter.
 
-\subsubsection{Tagging}
-
-%TODO: merge with Entity Identification
+%Szerintem ide kéne a DBpedia--CoNLL mapping
 We use DBpedia as
@@
 Only sentences for whose entities the correct category could be obtained from DBpedia are added to the corpus; those that contain unknown entities are discarded.
-Sentences without any entities are retained as well, so that the corpus mirrors (XXX the distribution of entities in the corpus mirrors their distribution in Wikipedia?) the distribution of entities in Wikipedia as closely as possible; this is also in line with how the ConLL corpus is assembled.
-(XXX or is it)
+
+%\subsubsection{Tagging}
+
+%TODO: merge with Entity Identification
+%megtörtént
 
@@
 % Introduction + entity -> class -> category.
-
-%Despite the aforementioned methods, we applied a new approach to recognize and classify NEs in the corpora generated from Wikipedia.
-We used the DBpedia ontology classes to map Wikipedia categories to standard NE tags (see more details in section \ref{dbpedia}).
 
@@
 
+\section{Corpus Description}
+
+% TODO: pls írd meg az első pár mondatot és nevezd át a sectiont!
+
+The corpora are available in the tab separated format used by ConLL (XXX, CITE).
+In this format, one line contains one word, its linguistic features and the named entity tags.
+Sentence boundaries are marked by empty lines.
+The linguistic features include the lemmatized form of the word, its part-of-speech and chunk tags.
+Two named entity tags are included with each word: one is the most specific DBpedia category it belongs to, if any; the other is the ConLL named entity tag.
+While the named entity tags can be considered "silver", the linguistic features are provided on a "best-effort" basis.
+
+The corpus is available under the Creative Commons Attribution-Sharealike 3.0 Unported License (CC-BY-SA), the same license under which the text of Wikipedia is released.
+The data files can be freely downloaded from ...\footnote{To be filled in in the camera-ready version.}.
+
 \section{Evaluation}
@@
 They can be very useful for improving NER in more ways: (a) for less commonly used languages, they can serve as training corpora in absence of gold standard datasets; (b) they can serve as (supplementary or independent) training sets for domains differing from newswire; (c) they can be sources of huge entity lists, and (d) feature extraction.
+
+Sentences without any entities are retained as well, so that the corpus mirrors (XXX the distribution of entities in the corpus mirrors their distribution in Wikipedia?) the distribution of entities in Wikipedia as closely as possible; this is also in line with how the ConLL corpus is assembled.
+(XXX or is it)
 
@@
 \begin{thebibliography}{}
+
+\bibitem[\protect\citename{Halácsy \bgroup et al.\egroup}2007]{Halacsy:07}
+P. Halácsy, A. Kornai and Cs. Oravecz.
+\newblock 2007.
+\newblock Hunpos -- an open source trigram tagger. 
+\newblock In: {\em Proceedings of the 45th Annual Meeting of the Association for Computational Linguistics}, pages 209--212.
 

### 2012-03-24 paper/engNERwiki.tex

@@
 \section{Creating the English Corpus}  % Dzsííízösz
+\label{sec:create}
 

### 2012-03-24 paper/engNERwiki.tex
I don't know.

--- 2012-03-24 paper/engNERwiki.tex
@@
 
-The procedure described in the last section was used to generate the Hungarian corpus as well.
+Hungarian differs from English not only as a language, but in data and tool coverage and NER practices as well.
+The procedure described in the previous section was used to generate the Hungarian corpus as well.
 However, the differences between English and Hungarian prompted XXX several modifications in its steps.
@@
 
+\subsection{Training Corpus}
+
 In addition to differences in grammar,
@@
 az angol: 1,7 milliárd
-
-according to the Szeged NER guideline
 

### 2012-03-24 paper/engNERwiki.tex
Last part of the Hungarian section.

@@
 
+% Good for middle-sized languages
+% Inconsistent template names, esp. across languages
+% Not all entities are written w/ capital letters => we need lists.
+
 \section{Creation of the Hungarian Corpus}
@@
 
-Hungarian differs from English not only as a language, but in data and tool coverage and NER practices as well.
 The procedure described in the previous section was used to generate the Hungarian corpus as well.
-However, the differences between English and Hungarian prompted XXX several modifications in its steps.
-In this section we describe how the algorithm was adapted to Hungarian.
-XXX!!!
-
-in tool a
+However, Hungarian differs from English in several aspects: as a language, in data and tool coverage and NER practices.
+In this section we describe these differences and the changes they prompted in the algorithm.
 
@@
 Two consequences follow.
+
 First, in order to identify which pages denote entities in the Hungarian Wikipedia, an additional step is required, in which the Hungarian equivalents of the English pages are added to the entity list.
 The English titles are retained, as due to the relatively small size of the Hungarian Wikipedia, in-article links sometimes point to English articles.
+
 Second, entities without a page in the English Wikipedia are absent from the entity list.
@@
 
-\subsection{Training Corpus}
+\subsection{NER Conventions}
 
-In addition to differences in grammar,
+% Eszter
+The largest (?, the only?, most cited?) NER corpus in Hungarian is the Szeged NER corpus, which will be described by Eszter.
+:) The Szeged corpus uses the same named entity categories as ConLL; however, the convention of what constitutes an entity is slightly different.
 
 The Szeged NER guideline relies heavily on the rules of capitalization to decide which words should be marked as named entities.
-In Hungarian, apart from the first word in the sentence, only proper nouns are capitalized.
-Thus, the following concepts are not considered named entities in Hungarian XXX:
+As mentioned earlier, in Hungarian, apart from the first word in the sentence, only proper nouns are capitalized.
+The following concepts are not proper nouns in Hungarian and thus are not considered named entities according to the Szeged guideline:
 \begin{itemize}
 \item Names of languages and ethnic groups
-\item Words created from proper nouns through derivation, most prominently
+\item Words created from proper nouns through derivation; most prominently
 \item Adjectival form of location names
-\item Months % angolban sem
+\item Months
 \end{itemize}
 
-filtering: az igés szűrő itt nem lesz
+The strong correlation between capitalization and named entities makes recognition of entities in text relatively simple: a word is part of an entity if it is capitalized.
+% Eszter
+a tulajdonnevekhez kötőjellel kapcsolódó köznevek kérdésesek (pl. SAP-szoftvercég, Matáv-csoport)
 
+% Ezt hová?
 most van: 102 millió szó a disambiguation pages-zel együtt
 az angol: 1,7 milliárd
-
-a tulajdonnevekhez kötőjellel kapcsolódó köznevek kérdésesek (pl. SAP-szoftvercég, Matáv-csoport)
-
-% Good for middle-sized languages
-% Inconsistent template names, esp. across languages
-% Comparison of the Hungarian DBpedia vs English + mapping
-% Not all entities are written w/ capital letters => we need lists.
 

### 2012-03-24 paper/engNERwiki.tex
Evaluation elso resze megirva

--- 2012-03-24 paper/engNERwiki.tex
@@
 
-\section{Creation of the Hungarian Corpus}
+\section{Creating the Hungarian Corpus}
 \label{sec:hun}
 
-Hungarian differs from English not only as a language, but in data and tool coverage and NER practices as well.
-The procedure described in the previous section was used to generate the Hungarian corpus as well.
+The procedure described in the last section was used to generate the Hungarian corpus as well.
 However, the differences between English and Hungarian prompted XXX several modifications in its steps.
@@
 
-\subsection{Training Corpus}
-
 In addition to differences in grammar,
@@
 az angol: 1,7 milliárd
+
+according to the Szeged NER guideline
 
@@
 % Corpus sizes, performance, tagging conventions
-
 % Hun -> not utf-8, loss of precision?
+%measuring Wikipedia's inconsistency using n-gram tag variation (\cite{Nothman:09})
+%ez jó mérőszám lehet a gold standard korpuszok összehasonlítására
+%használhatjuk fordítva is: ha egy adott nucleus ugyanolyan n-gram környezetben nagyon sokszor van X-ként taggelve, és kevésszer Y-ként, akkor az Y-okat is átcsinálhatnánk X-szé
 
@@
 Then what can we do with these so called "silver standard corpora" (SSC)?
-They can be very useful for improving NER in more ways: (a) for less commonly used languages, they can serve as training corpora in absence of gold standard datasets; (b) they can serve as (supplementary or independent) training sets for domains differing from newswire; (c) they can be sources of huge entity lists, and (d) feature extraction.
+They can be very useful for improving NER in more ways: (a) for less resourced languages, they can serve as training corpora in absence of gold standard datasets; (b) they can serve as (supplementary or independent) training sets for domains differing from newswire; (c) they can be sources of huge entity lists, and (d) feature extraction.
 
-Sentences without any entities are retained as well, so that the corpus mirrors (XXX the distribution of entities in the corpus mirrors their distribution in Wikipedia?) the distribution of entities in Wikipedia as closely as possible; this is also in line with how the ConLL corpus is assembled.
+To evaluate our corpora we used a maximum entropy NE-tagger \cite{Varga:07}, which was originally developed for labeling NEs in Hungarian texts, but it can be tuned for different languages, as well.
+The feature set was as follows:
+
+\begin{itemize}
+\item gazetteer features;
+\item sentence start and end position;
+\item Boolean-valued orthographic properties of the word form;
+\item string-valued surface properties of the word form;
+\item morphological information.
+\end{itemize}
+
+We used the CoNLL standard method for evaluation.
+According to this, an automatic labeling is correct if it gives the same start and end position, and the same NE class, as the gold standard.
+Based on this, precision and recall can be calculated, and the F-measure, as usual, the harmonic mean of these two values.
+
+\subsection{Wikipedia data}
+
+Our automatic annotation process retain all of the Wikipedia sentences, which remained after our two-fold filtering method.
+Thus sentences without NEs are also parts of the whole corpus.
+The rationale behind this is that we wanted to reserve the original distribution of names in Wikipedia as much as possible.
+However, after further investigation of NE density in our Wikipedia corpora and gold standard corpora (see table \ref{size}), we decided the sentences without NEs not to include in evaluation datasets.
+
+\begin{table*}[ht]
+\begin{center}
+\begin{tabular}{lllllll}
+\hline  & \bf enwiki & \bf enwiki filtered & \bf CoNLL & \bf huwiki & \bf huwiki filtered  & \bf Szeged NER \\ \hline
+token & 60,520,819 & 21,718,854 & 302,811 &  19,108,027 & 3,512,249  & 225,963\\
+NE & 3,169,863 & 3,169,863 & 50,758 & 456,281 & 456,281  & 25,896\\
+NE density & 5.23\% & 14.59\% & 16.76\% & 2.38\% & 12.99\%  & 11.46\%\\
+\hline
+\end{tabular}
+\end{center}
+\caption{\label{size} Corpus size and NE-density. }
+\end{table*}
+
+%Sentences without any entities are retained as well, so that the corpus mirrors (XXX the distribution of entities in the corpus mirrors their distribution in Wikipedia?) the distribution of entities in Wikipedia as closely as possible; this is also in line with how the ConLL corpus is assembled.
 (XXX or is it)
 
-measuring Wikipedia's inconsistency using n-gram tag variation (\cite{Nothman:09})
-ez jó mérőszám lehet a gold standard korpuszok összehasonlítására
-használhatjuk fordítva is: ha egy adott nucleus ugyanolyan n-gram környezetben nagyon sokszor van X-ként taggelve, és kevésszer Y-ként, akkor az Y-okat is átcsinálhatnánk X-szé
+\subsection{Experiments}
+
+We evaluated our corpora against (a) itself (splitted into train and test sets); (b) manually annotated English and Hungarian corpora.
+
+\subsection{Results}
 
@@
 \begin{thebibliography}{}
+
+\bibitem[\protect\citename{Varga and Simon}2007]{Varga:07}
+Dániel Varga and Eszter Simon.
+\newblock 2007.
+\newblock Hungarian named entity recognition with a maximum entropy approach.
+\newblock {\em Acta Cybernetica}, 18: 293--301.
 

### 2012-03-24 paper/engNERwiki.tex

@@
 
-The procedure described in the last section was used to generate the Hungarian corpus as well.
-However, the differences between English and Hungarian prompted XXX several modifications in its steps.
-In this section we describe how the algorithm was adapted to Hungarian.
-XXX!!!
-
-in tool a
+The procedure described in the previous section was used to generate the Hungarian corpus as well.
+However, Hungarian differs from English in several aspects: as a language, in data and tool coverage and NER practices.
+In this section we describe these differences and the changes they prompted in the algorithm.
 
@@
 
-In addition to differences in grammar,
+\subsection{NER Conventions}
+
+% Eszter
+The largest (?, the only?, most cited?) NER corpus in Hungarian is the Szeged NER corpus, which will be described by Eszter.
+:) The Szeged corpus uses the same named entity categories as ConLL; however, the convention of what constitutes an entity is slightly different.
 
 The Szeged NER guideline relies heavily on the rules of capitalization to decide which words should be marked as named entities.
-In Hungarian, apart from the first word in the sentence, only proper nouns are capitalized.
-Thus, the following concepts are not considered named entities in Hungarian XXX:
+As mentioned earlier, in Hungarian, apart from the first word in the sentence, only proper nouns are capitalized.
+The following concepts are not proper nouns in Hungarian and thus are not considered named entities according to the Szeged guideline:
 \begin{itemize}
 \item Names of languages and ethnic groups
-\item Words created from proper nouns through derivation, most prominently
+\item Words created from proper nouns through derivation; most prominently
 \item Adjectival form of location names
-\item Months % angolban sem
+\item Months
 \end{itemize}
 
-filtering: az igés szűrő itt nem lesz
+The strong correlation between capitalization and named entities makes recognition of entities in text relatively simple: a word is part of an entity if it is capitalized.
+% Eszter
+a tulajdonnevekhez kötőjellel kapcsolódó köznevek kérdésesek (pl. SAP-szoftvercég, Matáv-csoport)
 
+% Ezt hová?
 most van: 102 millió szó a disambiguation pages-zel együtt
 az angol: 1,7 milliárd
-
-according to the Szeged NER guideline
-
-a tulajdonnevekhez kötőjellel kapcsolódó köznevek kérdésesek (pl. SAP-szoftvercég, Matáv-csoport)
-
-% Good for middle-sized languages
-% Inconsistent template names, esp. across languages
-% Comparison of the Hungarian DBpedia vs English + mapping
-% Not all entities are written w/ capital letters => we need lists.
 

### 2012-03-24 paper/engNERwiki.tex
Stuff.

@@
 \section{Introduction}
+\label{sec:intro}
 
@@
 In MUC shared tasks the NER consists of three subtasks: entity names, temporal expressions and number expressions.
-However there is a general agreement in the NER community about the inclusion of temporal expressions and some numerical expressions such as amounts of money and other types of units, the most studied types are names of persons, locations and organizations.
-The fourth type, called "miscellaneous'', is used in the CoNLL NER tasks in 2002 \cite{Tjong:02} and 2003 \cite{Tjong:03}, and includes proper names falling outside the three classic types.
+Although there is a general agreement in the NER community about the inclusion of temporal expressions and some numerical expressions such as amounts of money and other types of units, the most studied types are names of persons, locations and organizations.
+The fourth type, called "miscellaneous'', was introduced in the CoNLL NER tasks in 2002 \cite{Tjong:02} and 2003 \cite{Tjong:03}, and includes proper names falling outside the three classic types.
 Since then, MUC and CoNLL datasets and annotation schemes have been the major standards applied in the field of NER.
@@
 An other approach to achieve this aim is using collaborative annotation and/or collaboratively constructed resources, such as Wikipedia, Wiktionary or Linked Open Data.
-Researchers apply the Web-as-corpus approach, often dealing with such datasets as the Google 5G 1T collection or DBPedia.
+Researchers apply the Web-as-corpus approach, often dealing with such datasets as Google's Web 1T 5-gram collection or DBPedia.
 
@@
 \section{Wikipedia and NER}
+\label{sec:related}
 % Wikify!
@@
 
-Wikipedia is a goldmine of information; it is applied for several NLP tasks, e.g.~word sense disambiguation, ontology and thesaurus building, multilingual Information Retrieval, and Question Answering, as Medelyan et al.~\shortcite{Medelyan:09} point out in their fully detailed survey about mining meaning from Wikipedia.
+Wikipedia is a free, multilingual Internet encyclopedia, written collaboratively by volunteers.
+(CITE http://en.wikipedia.org/wiki/Wikipedia).
+It is a goldmine of information; at the time of writing, Wikipedia contains about 21 million interlinked articles.
+Of these, 3,903,467 are English and 212,120 are Hungarian.
+XXX
+
+Wikipedia has been applied for several NLP tasks, e.g.~word sense disambiguation, ontology and thesaurus building, multilingual Information Retrieval, and Question Answering, as Medelyan et al.~\shortcite{Medelyan:09} point out in their fully detailed survey about mining meaning from Wikipedia.
 
@@
 
-To our knowledge, such type of automatically built corpora are not freely available, with the only one exception of the Semantically Annotated Snapshot of the English Wikipedia \cite{Zaragoza:07}.
+To our knowledge, such type of automatically built corpora are not freely available, with the one exception of the Semantically Annotated Snapshot of the English Wikipedia \cite{Zaragoza:07}.
 It provides a wide range of linguistic information: POS tags, dependency labels, WordNet super senses and NE annotation according to WSJ and CoNLL tagsets.
 However, they followed a different way: the NEs were tagged by open source softwares.
-Our corpora will be available under the same license as Wikipedia's one (Creative Commons Attribution-Sharealike 3.0 Unported License).
+Our corpora is available under the same license as Wikipedia's one (Creative Commons Attribution-Sharealike 3.0 Unported License).
 
-Thanks for the huge amount of Wikipedia articles, the possibility of building large enough corpora is given for less resourced languages, as well.
-Since our method is mainly language-independent, it can be easily applied for languages which are structurally differing from English.
-Hungarian is a highly agglutinative language, with free word order and other particularities detailed later in section \ref{sec:hun}.
-There are human annotated CoNLL-style datasets, but this presented here is the first large enough automatically NE-annotated corpus for Hungarian.
+Thanks for the huge amount of Wikipedia articles, the possibility of building large enough (XXX ez mit jelent?) corpora is given for less resourced languages, as well.
+Since our method is mainly XXX language-independent, it can be easily applied for languages which are structurally differing from English.
+Hungarian is a highly agglutinative language, with free word order and other particularities XXX detailed later in section \ref{sec:hun}.
+There are human annotated CoNLL-style datasets, but the one presented here is the first large enough automatically NE-annotated corpus for Hungarian.
 
@@
 
-% Introduction: Wikipedia + DBpedia
+Our goal is to create a large NE-annotated corpus, automatically generated from Wikipedia articles. We follow a similar path XXX to \cite{Nothman:08}, who broke up XXX the process into four steps:
+\begin{enumerate}
+\item Classify all articles into entity classes
+\item Split Wikipedia articles into sentences
+\item Label \texttt{NE}s according to link targets
+\item Select sentences for inclusion in a corpus
+\end{enumerate}
 
-Wikipedia proves to be a rich source of information about its entities; aside from the article text, a huge amount of data is encompassed in infoboxes, templates and the category structure.
-Since our goal is to create a NE-annotated corpus containing large amount of texts, our interest is limited to the article text.
-Templates are removed from the articles; however, the names of the templates have been retained for XXX, but all data (including infobox)
+In this section, we describe how these steps were implemented.
+While the approach in essence is language-agnostic, the exact details XXX have to be accomodated to the target language.
+This section explains the general idea and its execution for English; the next section describes how the idea can be adapted to Hungarian.
+
+\subsection{Parsing Wikipedia}
+
+Wikipedia is very rich source of information; aside from the article text, a huge amount of data is encompassed in infoboxes, templates and the category structure.
+Our task requires the use of only the links between the articles and the article text.
+In addition to in-article links, our method takes advantage of the redirect and interlanguage links, available as SQL dumps.
 
@@
 Full sentences are scarce, and the structure of these pages makes it hard to extract valid training sentences from them.
-Therefore, disambiguation pages are dropped from the corpus, as well.
+Therefore, disambiguation pages are dropped from the corpus as well.
 
-The English corpus is based on the Wikipedia snapshot as of XXX.
-The XML files were parsed with the mwlib (CITE) parser.
-The raw text was tokenized with a modified version of the XXX sentence tokenizer in NLTK (CITE).
-The words were lemmatized with XXX and POS tagged with HunPos \cite{Halacsy:07}.
+The English corpus is based on the Wikipedia snapshot as of January 15, 2011.
+The XML files were parsed with the \texttt{mwlib} (CITE http://code.pediapress.com) parser.
+The raw text was tokenized with a modified version of the Punkt sentence and word tokenizers \cite{Kiss:06} in NLTK \cite{Loper:02}.
+The words were lemmatized with the Wordnet lemmatizer, also in NLTK, and POS tagged with HunPos \cite{Halacsy:07}.
 
 %Chunking was performed with HunChunk (CITE). Experiments have shown that HunPos obtains STAT f-score on the Brown corpus; HunChunk performs at around STAT, CITE?.
+
+\subsection{Articles as Entities}
+\label{dbpedia}
+
+Most authors
+
+The classes in DBpedia are organized into a type hierarchy, available as an OWL (CITE) ontology.
+The ontology contains the 320 most frequent categories of Wikipedia, arranged into a taxonomy under the base class \texttt{owl:Thing}.
+Most of the classes belong into the 5 largest sub-hierarchies (here named after their topmost classes XXX): \texttt{Agent} (which is further divided into \texttt{Person} and \texttt{Organisation}), \texttt{Event}, \texttt{Place}, \texttt{Species} and \texttt{Work}.
+The taxonomy is rather flat: the top level contains 44 classes and there are several nodes with a branching factor of 20.
+
+It is not difficult to see the parallels XXX between the DBpedia sub-hierarchies \texttt{Person}, \texttt{Organisation} and \texttt{Place} and the ConLL NER categories \texttt{PER}, \texttt{ORG} and \texttt{LOC}.
+The fourth category, MISC is more elusive; according to the ConLL NER annotation guide (CITE?: http://www.cnts.ua.ac.be/conll2003/ner/annotation.txt), the sub-hierarchies \texttt{Event} and \texttt{Work} belong to this category, as well as various other class outside the main sub-hierarchies.
+
+While the correspondence described above holds for most classes in the XXX sub-hierarchies, there are some exceptions.
+For instance, the class \texttt{SportsLeague} is part of the \texttt{Organisation} sub-hierarchy, but in the ConLL corpus, sport leagues are tagged as MISC.
+Similarly, the class \texttt{SportsTeamSeason} is under \texttt{Event}, by which account it would be classified as MISC; however, ConLL does not recognize such entities.
+To account for these misclassification, as well as the miscellaneous classes outside the main hierarchies, we have introduced a mapping mechanism.
+We created a file of DBpedia class -- NER category mappings.
+Whenever an entity is evaluated, we look up its class, as well as the ancestors of its class, in the mapping, and assign to the entity the category of the class that matches it most closely -- i.e.
+the category of the entity's class, if it is in the list, or that of its closest ancestor.
+XXX!!!
+If no match is found, the entity is tagged with \texttt{O}.
+Since we take advantage of the inheritance hierarchy, the mapping list remains short: it contains only the root classes of the main sub-hierarchies, exceptions of the like XXX mentioned above and the various classes that belong to the MISC category according to the ConLL annotation guideline.
+
+As of version 3.7, the DBpedia ontology allows multiple superclasses, making the hierarchy a directed acyclic graph (CITE).
+This introduces the problem of selecting the right superclass, and hence, ConLL mapping for classes with more than one parent.
+In version 3.7, the only such class is \textit{Library}, which can be traced back to both \textit{Place} and \textit{Organisation}.
+There are two ways to tackle this problem.
+The first is to follow the DBpedia guidelines and only retain the first parent.
+The second method is manually deciding the right NER mapping for the ambiguous XXX class and adding it to the mapping file.
+The first approach, unfortunately, already fails to assign the correct ConLL NER tag to our only example.
+The reason behind this failure is that multiple parents in an ontology show an inherent ambiguity in the class in question: it is entirely natural to think of a library as a location as well as a public organization.
+DBpedia's guidelines prefers Organisation; ConLL's LOC.
+Therefore, only the second, manual approach ensures correct tagging.
+XXX
+
+The full mapping can be found in LINK Appendix A (XXX or not).
 
@@
 % 5. teaching the original DBpedia categories?
-
-\subsection{The DBpedia -- ConLL mapping}
-\label{dbpedia}
-% Introduction + entity -> class -> category.
-
-The classes in DBpedia are organized into a type hierarchy, available as an OWL (CITE) ontology.
-The ontology contains the 320 most frequent categories of Wikipedia, arranged into a taxonomy under the base class \texttt{owl:Thing}.
-Most of the classes belong into the 5 largest sub-hierarchies (here named after their topmost classes XXX): \texttt{Agent} (which is further divided into \texttt{Person} and \texttt{Organisation}), \texttt{Event}, \texttt{Place}, \texttt{Species} and \texttt{Work}.
-The taxonomy is rather flat: the top level contains 44 classes and there are several nodes with a branching factor of 20.
-
-It is not difficult to see the parallels XXX between the DBpedia sub-hierarchies \texttt{Person}, \texttt{Organisation} and \texttt{Place} and the ConLL NER categories \texttt{PER}, \texttt{ORG} and \texttt{LOC}.
-The fourth category, MISC is more elusive; according to the ConLL NER annotation guide (CITE?: http://www.cnts.ua.ac.be/conll2003/ner/annotation.txt), the sub-hierarchies \texttt{Event} and \texttt{Work} belong to this category, as well as various other class outside the main sub-hierarchies.
-
-While the correspondence described above holds for most classes in the XXX sub-hierarchies, there are some exceptions.
-For instance, the class \texttt{SportsLeague} is part of the \texttt{Organisation} sub-hierarchy, but in the ConLL corpus, sport leagues are tagged as MISC.
-Similarly, the class \texttt{SportsTeamSeason} is under \texttt{Event}, by which account it would be classified as MISC; however, ConLL does not recognize such entities.
-To account for these misclassification, as well as the miscellaneous classes outside the main hierarchies, we have introduced a mapping mechanism.
-We created a file of DBpedia class -- NER category mappings.
-Whenever an entity is evaluated, we look up its class, as well as the ancestors of its class, in the mapping, and assign to the entity the category of the class that matches it most closely -- i.e.
-the category of the entity's class, if it is in the list, or that of its closest ancestor.
-XXX!!!
-If no match is found, the entity is tagged with \texttt{O}.
-Since we take advantage of the inheritance hierarchy, the mapping list remains short: it contains only the root classes of the main sub-hierarchies, exceptions of the like XXX mentioned above and the various classes that belong to the MISC category according to the ConLL annotation guideline.
-
-As of version 3.7, the DBpedia ontology allows multiple superclasses, making the hierarchy a directed acyclic graph (CITE).
-This introduces the problem of selecting the right superclass, and hence, ConLL mapping for classes with more than one parent.
-In version 3.7, the only such class is \textit{Library}, which can be traced back to both \textit{Place} and \textit{Organisation}.
-There are two ways to tackle this problem.
-The first is to follow the DBpedia guidelines and only retain the first parent.
-The second method is manually deciding the right NER mapping for the ambiguous XXX class and adding it to the mapping file.
-The first approach, unfortunately, already fails to assign the correct ConLL NER tag to our only example.
-The reason behind this failure is that multiple parents in an ontology show an inherent ambiguity in the class in question: it is entirely natural to think of a library as a location as well as a public organization.
-DBpedia's guidelines prefers Organisation; ConLL's LOC.
-Therefore, only the second, manual approach ensures correct tagging.
-XXX
-
-The full mapping can be found in LINK Appendix A (XXX or not).
 
@@
 
+\bibitem[\protect\citename{Loper \bgroup et al.\egroup}2002]{Loper:02}
+Edward Loper, Steven Bird.
+\newblock 2002.
+\newblock NLTK: the Natural Language Toolkit. 
+\newblock In: {\em Proceedings of the ACL-02 Workshop on Effective tools and methodologies for teaching natural language processing and computational linguistics}, Volume 1, pages 63--70.
+
+\bibitem[\protect\citename{Loper \bgroup et al.\egroup}2002]{Kiss:06}
+Tibor Kiss, Jan Strunk.
+\newblock 2006.
+\newblock Unsupervised Multilingual Sentence Boundary Detection. 
+\newblock In: {\em Computational Linguistics}, 32 (4): pages 485--525.
+
 \end{thebibliography}
@@
 \end{document}
+
+% Jobban ki kell hangsúlyozni az introductionben, és a related worksben, hogy szabadon elérhető training korpuszt építettünk
+% És azt, hogy Hungarian middle, English nagy.

### 2012-03-24 paper/engNERwiki.tex
Evaluation nagy resze megirva

--- 2012-03-24 paper/engNERwiki.tex
@@
 In this section, we describe how these steps were implemented, as well as the post-filtering we employed to dispose of correctly tagged, but unusable training sentences.
+
+%We elected to use the CoNLL NE types. stb.
 
@@
 Our automatic annotation process retain all of the Wikipedia sentences, which remained after our two-fold filtering method.
-Thus sentences without NEs are also parts of the whole corpus.
+Thus sentences without NEs take also part of the whole corpus.
 The rationale behind this is that we wanted to reserve the original distribution of names in Wikipedia as much as possible.
@@
 \end{center}
-\caption{\label{size} Corpus size and NE-density. }
+\caption{\label{size} Corpus size and NE-density.}
 \end{table*}
 
-%Sentences without any entities are retained as well, so that the corpus mirrors (XXX the distribution of entities in the corpus mirrors their distribution in Wikipedia?) the distribution of entities in Wikipedia as closely as possible; this is also in line with how the ConLL corpus is assembled.
-(XXX or is it)
+\subsection{Experiments and results}
 
-\subsection{Experiments}
+\begin{table*}[ht]
+\begin{center}
+\begin{tabular}{lllll}
+\hline \bf train & \bf test & \bf precision & \bf recall & \bf F-measure \\ \hline
+CoNLL & CoNLL & 85.13 & 85.13 & 85.13 \\
+enwiki & enwiki & 72.46 & 73.33 &  72.89 \\
+enwiki & CoNLL & 56.55 & 49.77 & 52.94 \\
+CoNLL with wikilists & CoNLL & 86.33 & 86.35 & 86.34 \\
+CoNLL with wikitags & CoNLL & 85.88 & 85.94 & 85.91 \\
+\hline
+\end{tabular}
+\end{center}
+\caption{\label{enresults} English results.}
+\end{table*}
 
-We evaluated our corpora against (a) itself (splitted into train and test sets); (b) manually annotated English and Hungarian corpora.
+\begin{table*}[ht]
+\begin{center}
+\begin{tabular}{lllll}
+\hline \bf train & \bf test & \bf precision & \bf recall & \bf F-measure \\ \hline
+Szeged NER & Szeged NER & 94.50 & 94.35 & 94.43 \\
+huwiki & huwiki & 90.64 & 88.91 &  89.76 \\
+huwiki & Szeged NER & 63.08 & 70.46 & 66.57 \\
+Szeged NER with wikilists & Szeged NER & 95.48 & 95.48 & 95.48 \\
+Szeged NER with wikitags & Szeged NER & 95.38 & 94.92 & 95.15 \\
+\hline
+\end{tabular}
+\end{center}
+\caption{\label{huresults} Hungarian results.}
+\end{table*}
 
-\subsection{Results}
+We evaluated our English corpus against (a) itself (divided into train and test sets); and (b) a manually annotated English corpus.
+Since the filtered English Wikipedia corpus -- containing only the sentences with NEs -- is still very huge, and we are limited by time and memory, our experiments were performed with a portion of 3.5 million tokens (the size of our filtered Hungarian corpus is similar).
+It was divided into train and test sets (90\%-10\%).
+
+For the English cross-corpus evaluation the CoNLL-2003 dataset was chosen, mostly because of the similarity in tagset and format.
+It is well-studied (e.g.
+XXX) that cross-checking between different corpora does not lead to F-measure improvement.
+Stylistic and domain differences certainly affect NER performance, and the different annotation schemes pose several compatibility problems.
+Nothman et al.~\shortcite{Nothman:08} showed that each set of gold standard training data performs much better on corresponding test sets than on test sets from other sources.
+The situation here is similar: the NE-tagger trained on Wikipedia does not achieve as high performance tested against CoNLL test set as one trained on the own training set (see table \ref{enresults} and \ref{huresults} for results).
+
+However, Wikipedia-derived corpora can also be used for improving NER performance in other ways.
+First, we collected huge gazetteer lists from the corpus for each NE category, which improved the overall F-measure given to the NE-tagger training and testing on CoNLL dataset.
+A second trial was labeling the CoNLL datasets by the model trained on Wikipedia corpus, and giving these labels as extra features to the next CoNLL train.
+Both methods result in improved overall F-measure on CoNLL test set.
+
+Results of experiments done on the Hungarian corpus look similar.
 

### 2012-03-24 paper/engNERwiki.tex
DBpedia.

--- 2012-03-24 paper/engNERwiki.tex
@@
 An other approach to achieve this aim is using collaborative annotation and/or collaboratively constructed resources, such as Wikipedia, Wiktionary or Linked Open Data.
-Researchers apply the Web-as-corpus approach, often dealing with such datasets as Google's Web 1T 5-gram collection or DBPedia.
+Researchers apply the Web-as-corpus approach, often dealing with such datasets as Google's Web 1T 5-gram collection or DBpedia.
 
@@
 
-Despite the aforementioned methods, we applied a new approach to recognize and classify NEs in the corpora generated from Wikipedia.
-We used the DBpedia ontology classes to map Wikipedia categories to standard NE tags (see more details in section \ref{dbpedia}).
+In contrast to the aforementioned methods, we applied a new approach to recognize and classify NEs in the corpora generated from Wikipedia.
+We mapped the DBpedia ontology classes to standard NE tags and assigned them to Wikipedia entities (see more details in section \ref{dbpedia}).
 
@@
 
-Most authors
+Many authors, such as Kazama and Torisawa \shortcite{Kazama:07} and Nothman et al \shortcite{Nothman:09} used unsupervised methods based on Wikipedia categories and text to classify articles into entity categories.
+To avoid the inevitable classification errors, we obtain entity type information from the DBpedia knowledge base \cite{Bizer:09}.
+This enables us to have high precision information about entity types at the expense of recall: out of the 3,903,467 English Wikipedia pages, 1,470,293 are covered by DBpedia.
 
-The classes in DBpedia are organized into a type hierarchy, available as an OWL (CITE) ontology.
-The ontology contains the 320 most frequent categories of Wikipedia, arranged into a taxonomy under the base class \texttt{owl:Thing}.
+DBpedia is an open database that presents information about entities in Wikipedia in structured form.
+Each entity corresponds to a Wikipedia article.
+DBpedia provides various information about entities including type, properties, home pages, etc.
+Here we limit our interest to the type system.
+
+The types in DBpedia are organized into a class hierarchy, available as an OWL (CITE) ontology.
+The ontology contains the 320 most frequent entity categories, arranged into a taxonomy under the base class \texttt{owl:Thing}.
 Most of the classes belong into the 5 largest sub-hierarchies (here named after their topmost classes XXX): \texttt{Agent} (which is further divided into \texttt{Person} and \texttt{Organisation}), \texttt{Event}, \texttt{Place}, \texttt{Species} and \texttt{Work}.
 The taxonomy is rather flat: the top level contains 44 classes and there are several nodes with a branching factor of 20.
+
+The type of entities is extracted automatically from Wikipedia categories.
+However, the mapping between Wikipedia categories and classes in the DBpedia ontology is manually defined.
+This, together with the fact that the existance of the reference ontology prevents the proliferation of categories observable in Wikipedia \cite{Bizer:09}, ensures that type information in DBpedia can be considered gold quality.
 
@@
 
-The full mapping can be found in LINK Appendix A (XXX or not).
+The full mapping can be found in Appendix~\ref{sec:appa}.
 
@@
 
+\pagebreak
+% HEEEEELP!
+\appendix
+\section{DBpedia to ConLL Mapping}
+\label{sec:appa}
+
+\begin{table*}[ht]
+\begin{center}
+\begin{tabular}{llll}
+\hline DBpedia & \bf CoNLL & \bf DBpedia & \bf CoNLL \\ \hline
+Person & PER & Library & LOC \\
+Place & LOC & MeanOfTransportation & MISC \\
+Organization & ORG & PeriodicalLiterature & ORG \\
+Organisation & ORG & ProgrammingLanguage & MISC \\
+Award & MISC & Project & MISC \\
+EthnicGroup & MISC & SportsLeague & MISC \\
+Event & MISC & SportsTeamSeason & O \\
+Holiday & MISC & Weapon & MISC \\
+Ideology & MISC & Work & MISC \\
+Language & MISC & & \\
+\hline
+\end{tabular}
+\end{center}
+\end{table*}
+
 \end{document}

### 2012-03-25 paper/engNERwiki.tex
Ha!

@@
 Similarly, the class \texttt{SportsTeamSeason} is under \texttt{Event}, by which account it would be classified as MISC; however, ConLL does not recognize such entities.
+
 To account for these misclassification, as well as the miscellaneous classes outside the main hierarchies, we have introduced a mapping mechanism.
 We created a file of DBpedia class -- NER category mappings.
-Whenever an entity is evaluated, we look up its class, as well as the ancestors of its class, in the mapping, and assign to the entity the category of the class that matches it most closely -- i.e.
-the category of the entity's class, if it is in the list, or that of its closest ancestor.
-XXX!!!
+Whenever an entity is evaluated, we look up its class, as well as the ancestors of its class, in the mapping and assign to the entity the category of the class that matches it most closely.
 If no match is found, the entity is tagged with \texttt{O}.
-Since we take advantage of the inheritance hierarchy, the mapping list remains short: it contains only the root classes of the main sub-hierarchies, exceptions of the like XXX mentioned above and the various classes that belong to the MISC category according to the ConLL annotation guideline.
+Since we take advantage of the inheritance hierarchy, the mapping list remains short: it contains only the root classes of the main sub-hierarchies, exceptions like those mentioned above and the various classes that belong to the MISC category according to the ConLL annotation guideline.
 
@@
 The second method is manually deciding the right NER mapping for the ambiguous XXX class and adding it to the mapping file.
+
 The first approach, unfortunately, already fails to assign the correct ConLL NER tag to our only example.
@@
 
-\subsection{Extracting and Filtering Sentences}
+Using the mapping described above, we compile an entity list that contains all entities in DBpedia tagged with the appropriate XXX CoNLL named entity category.
 
-A NER corpus consists of sentences that accurately tag NEs occurring in them.
+\subsection{Named Entity Labeling}
+
+A NER corpus consists of sentences where NEs are accurately tagged.
 In order to automatically prepare such sentences, two tasks need to be performed: identifying the entities in the sentence and tagging them with the correct tag.
 Sentences, for which accurate tagging could not be accomplished, must be removed from the corpus.
-In this section, we describe how these steps were implemented, as well as the post-filtering we employed to dispose of correctly tagged, but unusable training sentences.
-
-%We elected to use the CoNLL NE types. stb.
-
-\subsubsection{Identifying the entities}
-% TODO: identification and tagging happens in parallel, else the recall oriented approach cannot be explained.
 
@@
 A link to an article can then be perceived as a mapping that identifies its anchor text with a particular entity.
-However, there are exceptions to this rule.
-Disambiguation pages are used when the page title is ambiguous, and can refer to several entities.
-If a link points to a disambiguation page, we acknowledge the anchor text as an entity, but mark it unknown.
-A similar measure is taken if the link target XXX does not exist.
+
+Discovered entities are tagged with the CoNLL category assigned to them in the entity list extracted from DBpedia.
+If the link target is not in the entity list, or the link points to a disambiguation page, we cannot determine the type of the entity and tag it as \texttt{UNK}.
 Links to redirect pages are resolved to point instead to the redirect target, after which they are handled as regular cross-references.
 
-However, not all entities are marked with links.
-Authors may neglect to add XXX references to their text.
-Furthermore, in a Wikipedia article XXX, typically only the first occurrence of a particular entity is linked to the corresponding page.
+There are several
+
+\subsubsection{Non-entity Links}
+
+Strictly speaking, our original assumption of equating Wikipedia articles with named entities is not valid: many pages describe common nouns (Book, Aircraft), numbers (1, 42), currencies (Euro, Dollar) or other concepts that fall outside the scope of NER.
+
+\begin{description}
+\item[Common nouns] Links to common nouns are filtered by POS tags; if a link contains no \texttt{NNP}s, it is ignored.
+\item[Dates] Wikipedia has pages for years, months and concrete dates, therefore our algorithm might mistake them for entities.
+To prevent this problem, we compiled a list of calendar-related entities and added them to the main entity list tagged with the CoNLL category \texttt{O}.
+\item[Lowercase links] 
+\end{description}
+
+\subsubsection{Unmarked Entities}
+
+\subsubsection{Special Cases}
+
+Not all entities are marked with links.
+Authors may neglect to add references to their text.
+Furthermore, in a Wikipedia article, typically only the first occurrence of a particular entity is linked to the corresponding page.
 Subsequent mentions are unmarked and often incomplete -- e.g.~family names are used instead of full names.
-We devised XXX two methods to account for unmarked entities: one that emphasizes XXX precision, and the other, recall.
-XXX
-
-% Or NPP
-
-The precision-oriented approach seeks to minimize the risk of words getting incorrectly tagged XXX in the corpus. Names and their parts mostly start with capital letter, thus 
-%A short study of the CoNLL annotation guidelines reveals that the words in the entities of interest to us all start with capital letter. According to this method XXX, 
-all sentences that contain capitalized words, but are not part of the anchor text of a link, XXX were discarded.
-The rationale behind this approach is that the sheer size of Wikipedia guarantees that we end up with a sizeable corpus even with such a strict policy.
-XXX
+To account for such entities,
 
 The recall-oriented method tries to associate unmarked entities with entities already identified on the same page. Only if the attempt fails is the sentence that contains the words discarded. 
-%Ez utóbbi biztosan jó mondat?
 This approach aims to generate more sentences with as little loss of precision as possible.
@@
 \begin{itemize}
-\item The page title; % if it is in DBpedia
+\item The page title, if we know its category;
 \item Title of redirect pages that target the current page; % TODO
@@
 \subsubsection{Exceptions}
-
-Strictly speaking, our original assumption of equating Wikipedia articles with named entities is not valid: many pages describe common nouns (Book, Aircraft), numbers (1, 42), currencies (Euro, Dollar) or other concepts that fall outside the scope of NER.
-In this section, we describe the measures taken to avoid errors introduced by these false entities.
-We shall modify the two approaches -- precision and recall-oriented -- accordingly.
 
@@
 % TODO: Last sentence to evaluation? But then this whole subsection is too short.
-
-% 4. Test sentence extraction?
-%   a. unknown links, no unlinked names
-%   b. might be good for quality evaluation / recursive tagging (next paper?)
-% 5. teaching the original DBpedia categories?
 
@@
 
-Our automatic annotation process retain all of the Wikipedia sentences, which remained after our two-fold filtering method.
-Thus sentences without NEs take also part of the whole corpus.
+Our automatic annotation process retains all of the Wikipedia sentences, which remained after our two-fold filtering method.
+Thus sentences without NEs are also included in the corpus.
 The rationale behind this is that we wanted to reserve the original distribution of names in Wikipedia as much as possible.
-However, after further investigation of NE density in our Wikipedia corpora and gold standard corpora (see table \ref{size}), we decided the sentences without NEs not to include in evaluation datasets.
+However, after further investigation of the NE density in our Wikipedia corpora and gold standard corpora (see table \ref{size}), we decided not to include the sentences without NEs in evaluation datasets.
 
@@
 We evaluated our English corpus against (a) itself (divided into train and test sets); and (b) a manually annotated English corpus.
-Since the filtered English Wikipedia corpus -- containing only the sentences with NEs -- is still very huge, and we are limited by time and memory, our experiments were performed with a portion of 3.5 million tokens (the size of our filtered Hungarian corpus is similar).
+Since the filtered English Wikipedia corpus -- containing only the sentences with NEs -- is still very huge, and we are limited by time and memory, our experiments were performed with a portion of 3.5 million tokens (the size of our filtered Hungarian corpus).
 It was divided into train and test sets (90\%-10\%).
@@
 
-\bibitem[\protect\citename{Loper \bgroup et al.\egroup}2002]{Kiss:06}
+\bibitem[\protect\citename{Kiss and Strunk}2006]{Kiss:06}
 Tibor Kiss, Jan Strunk.
@@
 
-\pagebreak
 % HEEEEELP!

### 2012-03-25 paper/engNERwiki.tex
references rendberakva nagyjabol, plusz dolgok a magyar reszhez

--- 2012-03-24 paper/engNERwiki.tex
@@
 In this paper we present a kind of combination of the approaches mentioned above.
-We built NE-tagged corpora for English and Hungarian -- fully automatically generated from Wikipedia.
+We built freely available NE-tagged corpora for English and Hungarian -- fully automatically generated from Wikipedia.
 %Our goal was to create gold standard corpora which can serve as train, devel and test datasets for automatic NE-taggers.
@@
 % Wikify!
-% stuff from Mining Meaning from Wikipedia
-%NER and Wikipedia
 
-Wikipedia is a free, multilingual Internet encyclopedia, written collaboratively by volunteers.
-(CITE http://en.wikipedia.org/wiki/Wikipedia).
+Wikipedia is a free, multilingual Internet encyclopedia, written collaboratively by volunteers\footnote{http://en.wikipedia.org/wiki/Wikipedia}.
 It is a goldmine of information; at the time of writing, Wikipedia contains about 21 million interlinked articles.
 Of these, 3,903,467 are English and 212,120 are Hungarian.
-XXX
+XXX %kell ez ide?
 
-Wikipedia has been applied for several NLP tasks, e.g.~word sense disambiguation, ontology and thesaurus building, multilingual Information Retrieval, and Question Answering, as Medelyan et al.~\shortcite{Medelyan:09} point out in their fully detailed survey about mining meaning from Wikipedia.
+Wikipedia has been applied for several NLP tasks, e.g.~word sense disambiguation, ontology and thesaurus building, multilingual information retrieval, and question answering, as Medelyan et al.~\shortcite{Medelyan:09} point out in their fully detailed survey about mining meaning from Wikipedia.
 
 Wikipedia is recognized as one of the largest available collection of entities, and also as an external knowledge to improve the accuracy of NER.
-The most obvious utilization of Wikipedia for NER is extracting gazetteers containing person names, locations or organizations (e.g.
-Toral and Munoz \shortcite{Toral:06}).
+The most obvious utilization of Wikipedia for NER is extracting gazetteers containing person names, locations or organizations (e.g.~Toral and Mu\~noz \shortcite{Toral:06}).
 Creating dictionaries of entities is also a common step of named entity disambiguation \cite{Bunescu:06,Cucerzan:07}.
-Not only supervised systems use lists generated from Wikipedia: Nadeau et al.
-\shortcite{Nadeau:06} use similar lists in an unsupervised NER system.
+Not only supervised systems use lists generated from Wikipedia: Nadeau et al.~\shortcite{Nadeau:06} use similar lists in an unsupervised NER system.
 
-Wikipedia's knowledge may be incorporated as features in NER learning, e.g.
-Kazama and Torisawa \shortcite{Kazama:07} showed that automatic extraction of category labels from Wikipedia improves the accuracy of a supervised NE-tagger.
+Wikipedia's knowledge may be incorporated as features in NER learning, e.g.~Kazama and Torisawa \shortcite{Kazama:07} showed that automatic extraction of category labels from Wikipedia improves the accuracy of a supervised NE-tagger.
 
@@
 
-In contrast to the aforementioned methods, we applied a new approach to recognize and classify NEs in the corpora generated from Wikipedia.
+In contrast to the aforementioned methods, we applied a new approach to recognize and classify NEs in corpora generated from Wikipedia.
 We mapped the DBpedia ontology classes to standard NE tags and assigned them to Wikipedia entities (see more details in section \ref{dbpedia}).
@@
 It provides a wide range of linguistic information: POS tags, dependency labels, WordNet super senses and NE annotation according to WSJ and CoNLL tagsets.
-However, they followed a different way: the NEs were tagged by open source softwares.
+However, they followed a different way: the NEs were tagged by high quality open source softwares.
 Our corpora is available under the same license as Wikipedia's one (Creative Commons Attribution-Sharealike 3.0 Unported License).
 
-Thanks for the huge amount of Wikipedia articles, the possibility of building large enough (XXX ez mit jelent?) corpora is given for less resourced languages, as well.
+Thanks for the huge amount of Wikipedia articles, the possibility of building large enough 
+%(XXX ez mit jelent?) elég nagy :)
+corpora is given for less resourced languages, as well.
 Since our method is mainly XXX language-independent, it can be easily applied for languages which are structurally differing from English.
 Hungarian is a highly agglutinative language, with free word order and other particularities XXX detailed later in section \ref{sec:hun}.
-There are human annotated CoNLL-style datasets, but the one presented here is the first large enough automatically NE-annotated corpus for Hungarian.
+There are human annotated CoNLL-style datasets, but the one presented here is the first automatically NE-annotated corpus for Hungarian.
 
-\section{Creating the English Corpus}  % Dzsííízösz
+\section{Creating the English Corpus} 
 \label{sec:create}
@@
 Our goal is to create a large NE-annotated corpus, automatically generated from Wikipedia articles. We follow a similar path XXX to \cite{Nothman:08}, who broke up XXX the process into four steps:
+
 \begin{enumerate}
-\item Classify all articles into entity classes
-\item Split Wikipedia articles into sentences
-\item Label \texttt{NE}s according to link targets
-\item Select sentences for inclusion in a corpus
+\item Classify all articles into entity classes.
+\item Split Wikipedia articles into sentences.
+\item Label NEs according to link targets.
+\item Select sentences for inclusion in the corpus.
 \end{enumerate}
@@
 In this section, we describe how these steps were implemented.
-While the approach in essence is language-agnostic, the exact details XXX have to be accomodated to the target language.
-This section explains the general idea and its execution for English; the next section describes how the idea can be adapted to Hungarian.
+While the approach in essence is language-independent, the exact details XXX have to be accomodated to the target language.
+This section explains the general idea and its execution for English; the next section describes how the idea is adapted to Hungarian.
 
@@
 
-Wikipedia is very rich source of information; aside from the article text, a huge amount of data is encompassed in infoboxes, templates and the category structure.
+Wikipedia is a very rich source of information; aside from the article text, a huge amount of data is encompassed in infoboxes, templates and the category structure.
 Our task requires the use of only the links between the articles and the article text.
@@
 The English corpus is based on the Wikipedia snapshot as of January 15, 2011.
-The XML files were parsed with the \texttt{mwlib} (CITE http://code.pediapress.com) parser.
+The XML files were parsed with the \texttt{mwlib} parser\footnote{http://code.pediapress.com}.
 The raw text was tokenized with a modified version of the Punkt sentence and word tokenizers \cite{Kiss:06} in NLTK \cite{Loper:02}.
-The words were lemmatized with the Wordnet lemmatizer, also in NLTK, and POS tagged with HunPos \cite{Halacsy:07}.
+The words were lemmatized with the Wordnet Lemmatizer, also in NLTK, and POS tagged with HunPos \cite{Halacsy:07}.
 
@@
 
-Many authors, such as Kazama and Torisawa \shortcite{Kazama:07} and Nothman et al \shortcite{Nothman:09} used unsupervised methods based on Wikipedia categories and text to classify articles into entity categories.
+Many authors, such as Kazama and Torisawa \shortcite{Kazama:07} and Nothman et al.~\shortcite{Nothman:08} used semi-supervised methods based on Wikipedia categories and text to classify articles into entity categories.
 To avoid the inevitable classification errors, we obtain entity type information from the DBpedia knowledge base \cite{Bizer:09}.
@@
 
-The types in DBpedia are organized into a class hierarchy, available as an OWL (CITE) ontology.
+The types in DBpedia are organized into a class hierarchy, available as an OWL\footnote{http://www.w3.org/TR/owl-ref/} ontology.
 The ontology contains the 320 most frequent entity categories, arranged into a taxonomy under the base class \texttt{owl:Thing}.
@@
 However, the mapping between Wikipedia categories and classes in the DBpedia ontology is manually defined.
-This, together with the fact that the existance of the reference ontology prevents the proliferation of categories observable in Wikipedia \cite{Bizer:09}, ensures that type information in DBpedia can be considered gold quality.
+This, together with the fact that the existence of the reference ontology prevents the proliferation of categories observable in Wikipedia \cite{Bizer:09}, ensures that type information in DBpedia can be considered gold quality.
 
-It is not difficult to see the parallels XXX between the DBpedia sub-hierarchies \texttt{Person}, \texttt{Organisation} and \texttt{Place} and the ConLL NER categories \texttt{PER}, \texttt{ORG} and \texttt{LOC}.
-The fourth category, MISC is more elusive; according to the ConLL NER annotation guide (CITE?: http://www.cnts.ua.ac.be/conll2003/ner/annotation.txt), the sub-hierarchies \texttt{Event} and \texttt{Work} belong to this category, as well as various other class outside the main sub-hierarchies.
+From the available NER annotation standards we elected to use the CoNLL \cite{Tjong:03} NE types.
+It is not difficult to see the parallels XXX between the DBpedia sub-hierarchies \texttt{Person}, \texttt{Organisation} and \texttt{Place} and the CoNLL NE categories PERSON, ORGANIZATION and LOCATION.
+The fourth category, MISCELLANEOUS is more elusive; according to the ConLL NER annotation guide\footnote{http://www.cnts.ua.ac.be/conll2003/ner/annotation.txt}, the sub-hierarchies \texttt{Event} and \texttt{Work} belong to this category, as well as various other classes outside the main hierarchies.
 
 While the correspondence described above holds for most classes in the XXX sub-hierarchies, there are some exceptions.
-For instance, the class \texttt{SportsLeague} is part of the \texttt{Organisation} sub-hierarchy, but in the ConLL corpus, sport leagues are tagged as MISC.
+For instance, the class \texttt{SportsLeague} is part of the \texttt{Organisation} sub-hierarchy, but according to the CoNLL annotation scheme, they are tagged as MISC.
 Similarly, the class \texttt{SportsTeamSeason} is under \texttt{Event}, by which account it would be classified as MISC; however, ConLL does not recognize such entities.
 To account for these misclassification, as well as the miscellaneous classes outside the main hierarchies, we have introduced a mapping mechanism.
-We created a file of DBpedia class -- NER category mappings.
-Whenever an entity is evaluated, we look up its class, as well as the ancestors of its class, in the mapping, and assign to the entity the category of the class that matches it most closely -- i.e.
-the category of the entity's class, if it is in the list, or that of its closest ancestor.
+We created a file of DBpedia class--NE category mappings.
+Whenever an entity is evaluated, we look up its class, as well as the ancestors of its class, in the mapping, and assign to the entity the category of the class that matches it most closely -- i.e.~the category of the entity's class, if it is in the list, or that of its closest ancestor.
 XXX!!!
 If no match is found, the entity is tagged with \texttt{O}.
-Since we take advantage of the inheritance hierarchy, the mapping list remains short: it contains only the root classes of the main sub-hierarchies, exceptions of the like XXX mentioned above and the various classes that belong to the MISC category according to the ConLL annotation guideline.
+Since we take advantage of the inheritance hierarchy, the mapping list remains short: it contains only the root classes of the main sub-hierarchies, exceptions of the like XXX mentioned above and the various classes that belong to the MISC category according to the CoNLL annotation guideline.
 
 As of version 3.7, the DBpedia ontology allows multiple superclasses, making the hierarchy a directed acyclic graph (CITE).
-This introduces the problem of selecting the right superclass, and hence, ConLL mapping for classes with more than one parent.
+This introduces the problem of selecting the right superclass, and hence, CoNLL mapping for classes with more than one parent.
 In version 3.7, the only such class is \textit{Library}, which can be traced back to both \textit{Place} and \textit{Organisation}.
@@
 The first is to follow the DBpedia guidelines and only retain the first parent.
-The second method is manually deciding the right NER mapping for the ambiguous XXX class and adding it to the mapping file.
-The first approach, unfortunately, already fails to assign the correct ConLL NER tag to our only example.
+The second method is manually deciding the right NE mapping for the ambiguous XXX class and adding it to the mapping file.
+The first approach, unfortunately, already fails to assign the correct CoNLL NE tag to our only example.
 The reason behind this failure is that multiple parents in an ontology show an inherent ambiguity in the class in question: it is entirely natural to think of a library as a location as well as a public organization.
-DBpedia's guidelines prefers Organisation; ConLL's LOC.
+DBpedia's guidelines prefers Organisation; while ConLL's one refers it as a Location.
 Therefore, only the second, manual approach ensures correct tagging.
@@
 
-%We elected to use the CoNLL NE types. stb.
-
 \subsubsection{Identifying the entities}
@@
 Our approach to entity identification relies on Wikipedia cross-references found in the article text.
-We assume that individual Wikipedia articles describe named entities.
+We assume that individual Wikipedia articles describe NEs.
 A link to an article can then be perceived as a mapping that identifies its anchor text with a particular entity.
@@
 
-%Both approaches are evaluated in the next chapter.
-
-%Szerintem ide kéne a DBpedia--CoNLL mapping
-We use DBpedia as
-
 Only sentences for whose entities the correct category could be obtained from DBpedia are added to the corpus; those that contain unknown entities are discarded.
@@
 
-Strictly speaking, our original assumption of equating Wikipedia articles with named entities is not valid: many pages describe common nouns (Book, Aircraft), numbers (1, 42), currencies (Euro, Dollar) or other concepts that fall outside the scope of NER.
+Strictly speaking, our original assumption of equating Wikipedia articles with NEs is not valid: many pages describe common nouns (Book, Aircraft), numbers (1, 42), currencies (Euro, Dollar) or other concepts that fall outside the scope of CoNLL tagset or of NER in general.
 In this section, we describe the measures taken to avoid errors introduced by these false entities.
@@
 
-Proper entities that are not deemed NE under the ConLL scheme cause XXX little concern: they are either missing from DBpedia altogether, or we can map the classes they belong to to the NER category \texttt{0} (not an entity).
+Proper names that are not deemed NE under the ConLL scheme
+%ilyeneknek nem szabad lenniük a MISC miatt
+ cause XXX little concern: they are either missing from DBpedia altogether, or we can map the classes they belong to to the NER category \texttt{0} (not an entity).
 This "ontological filtering" ensures that we need no adjustment XXX to our original method.
@@
 The precision-oriented method treats them similarly to links to unknown entities and discards the sentences in which they occur.
-The recall-oriented approach tries to distinguish them from links to entities so that otherwise valid training sentences can be retained XXX.
+The recall-oriented approach tries to distinguish them from links to entities so that otherwise valid sentences can be retained XXX.
 The method is as follows XXX.
@@
 
-from \cite{Nothman:09}: Exceptions are made for common titlecase words, e.g. I, Mr., June, and sentence-initial words.
+XXX from \cite{Nothman:09}: Exceptions are made for common titlecase words, e.g. I, Mr., June, and sentence-initial words.
 
@@
 On the one hand, these sentence fragments may prove to be of too low quality to be of any use in the traditional NER task.
-On the other hand, they could prove to be invaluable when training a named entity recognizer for user generated content.
+On the other hand, they could prove to be invaluable when training a NER tagger for User Generated Content, which is well known to be noisy and fragmented.
 We have added a post-filtering step to recognize and label these fragments, so that users of the corpus can decide what to do with them.
@@
 Manual inspection has found two conditions that can be used to detect fragments.
-A sentence is labelled as such if it either lacks a punctuation mark at the end or it contains no finite verbs.
-Our final corpus contains STAT full sentences and STAT fragments.
+A sentence is labelled as such if it either lacks a punctuation mark at the end, or it contains no finite verbs.
+Our final corpus contains XXXSTAT full sentences and STAT fragments.
+%Ezeket leméred, vagy mérjem én? Ha utóbbi, akkor melyikben kell mérni?
 % TODO: Last sentence to evaluation? But then this whole subsection is too short.
@@
 
-In Wikipedia the nationalities (e.g.
-American and Australian) links to the corresponding places (America and Australia, respectively).
+In Wikipedia the nationalities (e.g.~American and Australian) links to the corresponding places (America and Australia, respectively).
 However, in CoNLL scheme these types of entities take parts of MISC category.
@@
 %nem biztos, hogy ide kell
-Link boundaries need to be adjusted, e.g. to remove excess punctuation.
+Link boundaries need to be adjusted, e.g.~to remove excess punctuation.
 
@@
 
-The procedure described in the previous section was used to generate the Hungarian corpus as well.
-However, Hungarian differs from English in several aspects: as a language, in data and tool coverage and NER practices.
-In this section we describe these differences and the changes they prompted in the algorithm.
+The procedure described in the previous section was used to generate the Hungarian corpus, as well.
+However, linguistic diversity pose several problems even in the NER task.
+In this section we describe the differences between the two languages related to labeling NEs, and the changes they prompted in the algorithm.
 
@@
 
-While NLTK has excellent tools for English, Hungarian is not supported.
-Therefore, we worked with the tools developed by the Hungarian computational linguist community.
-For sentence tokenization we used an in-house statistical chunker tailored for Hungarian.
-It has been trained on the Szeged corpus and it handles the peculiarities of Hungarian ortography, such as the periods placed after numbers in date expressions.
-Although HunPos, as its name implies, is fully capable of POS tagging Hungarian text, we opted for Hunmorph (CITE), a morphological analyzer that, apart from PAS tagging, also lemmatizes the words.
+Although Hungarian is reckoned to be a less resourced language, and it is not supperted in NLTK, several high quality language processing tools have been developed for Hungarian in recent years.
+For tokenization and sentence segmentation we used an in-house statistical tool tailored for Hungarian.
+It has been trained on the largest manually annotated Hungarian corpus \cite{Csendes:04}, and it handles the peculiarities of Hungarian ortography, such as the periods placed after numbers in date expressions.
+Lemmatization was performed by \texttt{hunmorph} \cite{Tron:05} and \texttt{hundisambig}, latter being an in-house disambiguator to select the right analysis based on the word context.
+
+%Although HunPos, as its name implies, is fully capable of POS tagging Hungarian text, we opted for Hunmorph (CITE), a morphological analyzer that, apart from PAS tagging, also lemmatizes the words.
 Hundisambig (CITE) was used to select the right analysis based on the word context\footnote{Incidentally, Hundisambig relies on HunPos for its HMM model.}.
 
-The Hungarian part-of-speech taggers output KR codes (CITE).
-In addition to the POS category, the KR code also includes inflection information, such as the number, person and case XXX, which makes it much better suited to agglutinative languages than Penn Treebank POS tags.
+The Hungarian morphological analyzer outputs KR-codes \cite{Kornai:04}.
+In addition to the POS category, the KR-code also includes inflection information, such as the number, person and case XXX, which makes it much better suited to agglutinative languages than Penn Treebank POS tags.
 One shortcoming of the KR code is that it does not differenciate between common and proper nouns.
@@
 
-% vs. ConLL / Szeged
 % Corpus sizes, performance, tagging conventions
@@
 To evaluate our corpora we used a maximum entropy NE-tagger \cite{Varga:07}, which was originally developed for labeling NEs in Hungarian texts, but it can be tuned for different languages, as well.
-The feature set was as follows:
+Corpus-specific features (e.g.~NP chunks, Wikipedia links) were removed to get better comparability, so the feature set was as follows:
 
@@
 However, Wikipedia-derived corpora can also be used for improving NER performance in other ways.
-First, we collected huge gazetteer lists from the corpus for each NE category, which improved the overall F-measure given to the NE-tagger training and testing on CoNLL dataset.
+First, like e.g.~\cite{Toral:06} we collected huge gazetteer lists from the corpus for each NE category, which improved the overall F-measure given to the NE-tagger training and testing on CoNLL dataset.
 A second trial was labeling the CoNLL datasets by the model trained on Wikipedia corpus, and giving these labels as extra features to the next CoNLL train.
@@
 
-az eredeti DBPedia osztályokat is megtarthatnánk, hiszen nem mindig célravezető a CoNLL osztályok használata, mivel azok elég tág kategóriák, amikbe olyan dolgok is tartoznak, amik nagyon eltérőek is lehetnek
+az eredeti DBpedia osztályokat is megtarthatnánk, hiszen nem mindig célravezető a CoNLL osztályok használata, mivel azok elég tág kategóriák, amikbe olyan dolgok is tartoznak, amik nagyon eltérőek is lehetnek
 
@@
 
-\bibitem[\protect\citename{Varga and Simon}2007]{Varga:07}
-Dániel Varga and Eszter Simon.
+\bibitem[\protect\citename{Balasuriya \bgroup et al.\egroup}2009]{Balasuriya:09}
+Dominic Balasuriya, Nicky Ringland, Joel Nothman, Tara Murphy, James R. Curran
+\newblock 2009.
+\newblock Named Entity Recognition in Wikipedia.
+\newblock In: {\em Proceedings of the 2009 Workshop on the People's Web Meets NLP, ACL-IJCNLP}, pages 10--18.
+
+\bibitem[\protect\citename{Bizer \bgroup et al.\egroup}2009]{Bizer:09} 
+Christian Bizer, Jens Lehmann, Georgi Kobilarov, Sören Auer, Christian Becker, Richard Cyganiak, Sebastian Hellmann.
+\newblock 2009.
+\newblock DBpedia -- A Crystallization Point for the Web of Data. 
+\newblock In: {\em Journal of Web Semantics: Science, Services and Agents on the World Wide Web}, Issue 7, pages 154--165.
+
+\bibitem[\protect\citename{Bunescu and Pasca}2006]{Bunescu:06}
+B. Bunescu and M. Pasca.
+\newblock 2006.
+\newblock Using encyclopedic knowledge for named entity disambiguation.
+\newblock In: {\em Proceedings of the 11th Conference of the European Chapter of the Association for Computational Linguistics}, pages 9--16.
+
+\bibitem[\protect\citename{Csendes \bgroup et al.\egroup}2004]{Csendes:04} 
+Dóra Csendes, János Csirik, Tibor Gyimóthy.
+\newblock 2004.
+\newblock The Szeged Corpus: A POS tagged and Syntactically Annotated Hungarian Natural Language Corpus. 
+\newblock In: {\em Proceedings of TSD 2004}, vol. 3206, pages 41--49.
+
+\bibitem[\protect\citename{Cucerzan}2007]{Cucerzan:07}
+S. Cucerzan.
 \newblock 2007.
-\newblock Hungarian named entity recognition with a maximum entropy approach.
-\newblock {\em Acta Cybernetica}, 18: 293--301.
+\newblock Large-scale named entity disambiguation based on Wikipedia data.
+\newblock In: {\em Proceedings of the 2007 Joint Conference on Empirical Methods in Natural Language Processing and Computational Natural Language Learning}.
+\newblock Prague, Czech Republic, June 2007. pages 708--716.
+
+\bibitem[\protect\citename{Grishman and Sundheim}1996]{Grishman:96}
+Ralph Grishman and B. Sundheim.
+\newblock 1996.
+\newblock Message Understanding Conference -- 6. 
+\newblock In: {\em Proc. International Conference on Computational Linguistics}.
 
@@
 
-\bibitem[\protect\citename{Zaragoza \bgroup et al.\egroup}2007]{Zaragoza:07}
-H. Zaragoza and J. Atserias and M. Ciaramita and G. Attardi.
+\bibitem[\protect\citename{Kazama and Torisawa}2007]{Kazama:07}
+Jun'ichi Kazama and Kentaro Torisawa.
 \newblock 2007.
-\newblock Semantically Annotated Snapshot of the English Wikipedia v.1 (SW1).
-\newblock \url{http://www.yr-bcn.es/semanticWikipedia}
+\newblock Exploiting Wikipedia as External Knowledge for Named Entity Recognition.
+\newblock In: {\em Proceedings of the 2007 Joint Conference on Empirical Methods in Natural Language Processing and Computational Natural Language Learning}, pages 698--707.
+
+\bibitem[\protect\citename{Kiss and Strunk}2002]{Kiss:06}
+Tibor Kiss, Jan Strunk.
+\newblock 2006.
+\newblock Unsupervised Multilingual Sentence Boundary Detection. 
+\newblock In: {\em Computational Linguistics}, 32 (4): pages 485--525.
+
+\bibitem[\protect\citename{Kornai \bgroup et al.\egroup}2004]{Kornai:04}
+András Kornai, Péter Rebrus, Péter Vajda, Péter Halácsy, András Rung, Viktor Trón.
+\newblock 2004.
+\newblock Általános célú morfológiai elemző kimeneti formalizmusa (The output formalism of a general-purpose morphological analyzer).
+\newblock In: {\em Proceedings of the 2nd Hungarian Computational Linguistics Conference}.
+
+\bibitem[\protect\citename{Loper and Bird}2002]{Loper:02}
+Edward Loper, Steven Bird.
+\newblock 2002.
+\newblock NLTK: the Natural Language Toolkit. 
+\newblock In: {\em Proceedings of the ACL-02 Workshop on Effective tools and methodologies for teaching natural language processing and computational linguistics}, Volume 1, pages 63--70.
+
+\bibitem[\protect\citename{Medelyan  \bgroup et al.\egroup}2009]{Medelyan:09}
+Olena Medelyan, David Milne, Catherine Legg, and Ian H. Witten.
+\newblock 2009.
+\newblock Mining meaning from Wikipedia.
+\newblock {\em International Journal of Human-Computer Studies}, 67: 716--754.
+
+\bibitem[\protect\citename{Nadeau  \bgroup et al.\egroup}2006]{Nadeau:06} 
+David Nadeau, Peter D. Turney and Stan Matwin.
+\newblock 2006.
+\newblock Unsupervised named entity recognition: Generating gazetteers and resolving ambiguity. 
+\newblock In: {\em Proceedings of the 19th Canadian Conference on Artificial Intelligence}, volume 4013 of {\em LNCS}, pages 266--277.
+
+\bibitem[\protect\citename{Nothman \bgroup et al.\egroup}2008]{Nothman:08}
+Joel Nothman, James R. Curran, and Tara Murphy.
+\newblock 2008.
+\newblock Transforming Wikipedia into Named Entity Training Data.
+\newblock In: {\em Proceedings of the Australasian Language Technology Workshop}, Vol 6., pages 124--132.
+
+\bibitem[\protect\citename{Nothman \bgroup et al.\egroup}2009]{Nothman:09}
+Joel Nothman and Tara Murphy and James R. Curran.
+\newblock 2009.
+\newblock Analysing Wikipedia and Gold-Standard Corpora for NER Training.
+\newblock In: {\em Proceedings of the 12th Conference of the European Chapter of the ACL}, pages 612--620.
+
+\bibitem[\protect\citename{Richman and Schone}2008]{Richman:08}
+Alexander E. Richman and Patrick Schone.
+\newblock 2008.
+\newblock Mining Wiki Resources for Multilingual Named Entity Recognition.
+\newblock In: {\em Proceedings of ACL-08: HLT}, pages 1--9.
 
@@
 
-\bibitem[\protect\citename{Grishman and Sundheim}1996]{Grishman:96}
-Ralph Grishman and B. Sundheim.
-\newblock 1996.
-\newblock Message Understanding Conference -- 6. 
-\newblock In: {\em Proc. International Conference on Computational Linguistics}.
-
-\bibitem[\protect\citename{Nothman \bgroup et al.\egroup}2009]{Nothman:09}
-\newblock 2009.
-\newblock Analysing Wikipedia and Gold-Standard Corpora for NER Training.
-\newblock In: {\em Proceedings of the 12th Conference of the European Chapter of the ACL}, pages 612--620.
-
-\bibitem[\protect\citename{Richman and Schone}2008]{Richman:08}
-Alexander E. Richman and Patrick Schone.
-\newblock 2008.
-\newblock Mining Wiki Resources for Multilingual Named Entity Recognition.
-\newblock In: {\em Proceedings of ACL-08: HLT}, pages 1--9.
-
-\bibitem[\protect\citename{Nothman \bgroup et al.\egroup}2008]{Nothman:08}
-Joel Nothman, James R. Curran, and Tara Murphy.
-\newblock 2008.
-\newblock Transforming Wikipedia into Named Entity Training Data.
-\newblock In: {\em Proceedings of the Australasian Language Technology Workshop}, Vol 6., pages 124--132.
-
-\bibitem[\protect\citename{Kazama and Torisawa}2007]{Kazama:07}
-Jun'ichi Kazama and Kentaro Torisawa.
-\newblock 2007.
-\newblock Exploiting Wikipedia as External Knowledge for Named Entity Recognition.
-\newblock In: {\em Proceedings of the 2007 Joint Conference on Empirical Methods in Natural Language Processing and Computational Natural Language Learning}, pages 698--707.
-
-\bibitem[\protect\citename{Medelyan  \bgroup et al.\egroup}2009]{Medelyan:09}
-Olena Medelyan, David Milne, Catherine Legg, and Ian H. Witten.
-\newblock 2009.
-\newblock Mining meaning from Wikipedia.
-\newblock {\em International Journal of Human-Computer Studies}, 67: 716--754.
-
-\bibitem[\protect\citename{Toral and Munoz}2006]{Toral:06}
-A. Toral and R. Munoz.
+\bibitem[\protect\citename{Toral and Mu\~noz}2006]{Toral:06}
+A. Toral and R. Mu\~noz.
 \newblock 2006.
@@
 
-\bibitem[\protect\citename{Bunescu and Pasca}2006]{Bunescu:06}
-B. Bunescu and M. Pasca.
-\newblock 2006.
-\newblock Using encyclopedic knowledge for named entity disambiguation.
-\newblock In: {\em Proceedings of the 11th Conference of the European Chapter of the Association for Computational Linguistics}, pages 9--16.
+\bibitem[\protect\citename{Trón \bgroup et al.\egroup}2005]{Tron:05}
+Viktor Trón, György Gyepesi, Péter Halácsy, András Kornai, László Németh, Dániel Varga.
+\newblock 2005.
+\newblock Hunmorph: open source word analysis.
+\newblock In: {\em Proceedings of the ACL 2005 Workshop on Software}.
 
-\bibitem[\protect\citename{Cucerzan}2007]{Cucerzan:07}
-S. Cucerzan.
+\bibitem[\protect\citename{Varga and Simon}2007]{Varga:07}
+Dániel Varga and Eszter Simon.
 \newblock 2007.
-\newblock Large-scale named entity disambiguation based on Wikipedia data.
-\newblock In: {\em Proceedings of the 2007 Joint Conference on Empirical Methods in Natural Language Processing and Computational Natural Language Learning}.
-\newblock Prague, Czech Republic, June 2007. pages 708--716.
+\newblock Hungarian named entity recognition with a maximum entropy approach.
+\newblock {\em Acta Cybernetica}, 18: 293--301.
 
-\bibitem[\protect\citename{Nadeau  \bgroup et al.\egroup}2006]{Nadeau:06} 
-David Nadeau, Peter D. Turney and Stan Matwin.
-\newblock 2006.
-\newblock Unsupervised named entity recognition: Generating gazetteers and resolving ambiguity. 
-\newblock In: {\em Proceedings of the 19th Canadian Conference on Artificial Intelligence}, volume 4013 of {\em LNCS}, pages 266--277.
-
-\bibitem[\protect\citename{Bizer \bgroup et al.\egroup}2009]{Bizer:09} 
-Christian Bizer, Jens Lehmann, Georgi Kobilarov, Sören Auer, Christian Becker, Richard Cyganiak, Sebastian Hellmann.
-\newblock 2009.
-\newblock DBpedia -- A Crystallization Point for the Web of Data. 
-\newblock In: {\em Journal of Web Semantics: Science, Services and Agents on the World Wide Web}, Issue 7, pages 154--165.
-
-\bibitem[\protect\citename{Loper \bgroup et al.\egroup}2002]{Loper:02}
-Edward Loper, Steven Bird.
-\newblock 2002.
-\newblock NLTK: the Natural Language Toolkit. 
-\newblock In: {\em Proceedings of the ACL-02 Workshop on Effective tools and methodologies for teaching natural language processing and computational linguistics}, Volume 1, pages 63--70.
-
-\bibitem[\protect\citename{Loper \bgroup et al.\egroup}2002]{Kiss:06}
-Tibor Kiss, Jan Strunk.
-\newblock 2006.
-\newblock Unsupervised Multilingual Sentence Boundary Detection. 
-\newblock In: {\em Computational Linguistics}, 32 (4): pages 485--525.
+\bibitem[\protect\citename{Zaragoza \bgroup et al.\egroup}2007]{Zaragoza:07}
+H. Zaragoza and J. Atserias and M. Ciaramita and G. Attardi.
+\newblock 2007.
+\newblock Semantically Annotated Snapshot of the English Wikipedia v.1 (SW1).
+\newblock \url{http://www.yr-bcn.es/semanticWikipedia}
 

### 2012-03-25 paper/engNERwiki.tex
Almost done?

--- 2012-03-25 paper/engNERwiki.tex
@@
 
-Many authors, such as Kazama and Torisawa \shortcite{Kazama:07} and Nothman et al \shortcite{Nothman:09} used unsupervised methods based on Wikipedia categories and text to classify articles into entity categories.
+Many authors, such as Kazama and Torisawa \shortcite{Kazama:07} and Nothman et al \shortcite{Nothman:08} used unsupervised methods based on Wikipedia categories and text to classify articles into entity categories.
 To avoid the inevitable classification errors, we obtain entity type information from the DBpedia knowledge base \cite{Bizer:09}.
@@
 
-Using the mapping described above, we compile an entity list that contains all entities in DBpedia tagged with the appropriate XXX CoNLL named entity category.
+Using the mapping described above, we compile an entity list that contains all entities in DBpedia tagged with the appropriate CoNLL named entity category.
+
+We would like to note that even though we use CoNLL categories, our method can be modified to work with any tagset compatible with the DBpedia ontology.
+Indeed, the DBpedia classes define a named entity tagset themselves.
+We leave the exploration of these possibilities for future work.
 
@@
 Links to redirect pages are resolved to point instead to the redirect target, after which they are handled as regular cross-references.
+Finally, sentences with UNK links in them are removed from the corpus.
 
-There are several
+There are several ways to improve the method described above to increase precision, sentence coverage and to account for peculiarities in the CoNLL guidelines.
+These are detailed in the next sub-sections.
 
@@
 
-Strictly speaking, our original assumption of equating Wikipedia articles with named entities is not valid: many pages describe common nouns (Book, Aircraft), numbers (1, 42), currencies (Euro, Dollar) or other concepts that fall outside the scope of NER.
+Strictly speaking, our original assumption of equating Wikipedia articles with named entities is not valid: many pages describe common nouns (Book, Aircraft), calendar-related concepts (March 15, 2007), or other concepts that fall outside the scope of NER.
+To increase sentence coverage, we modified the algorithm to prevent it from misclassifying links to these pages as unknown entities and discarding the sentence.
 
@@
 \item[Common nouns] Links to common nouns are filtered by POS tags; if a link contains no \texttt{NNP}s, it is ignored.
-\item[Dates] Wikipedia has pages for years, months and concrete dates, therefore our algorithm might mistake them for entities.
-To prevent this problem, we compiled a list of calendar-related entities and added them to the main entity list tagged with the CoNLL category \texttt{O}.
-\item[Lowercase links] 
+\item[Dates] Time expressions require special attention, because dates and months are often linked to the respective Wikipedia pages.
+We circumvented this problem by compiling a list of calendar-related pages and adding them to the main entity list tagged with the CoNLL category \texttt{O}.
+\item[Lowercase links] Entities are sometimes referred to by common nouns, such as \textit{republic} to \textit{Roman Republic}, \texttt{LOC}, which results in a lowercase link.
+These references are not considered named entities and are ignored.
 \end{description}
@@
 \subsubsection{Unmarked Entities}
-
-\subsubsection{Special Cases}
 
@@
 Subsequent mentions are unmarked and often incomplete -- e.g.~family names are used instead of full names.
-To account for such entities,
+To account for such mentions, we try to associate capitalized words in the article text with entities we have tagged thus far.
 
-The recall-oriented method tries to associate unmarked entities with entities already identified on the same page. Only if the attempt fails is the sentence that contains the words discarded. 
-This approach aims to generate more sentences with as little loss of precision as possible.
-
-The algorithm maintains a list of entities already indentified on the page. The list includes:
+For each article, we build a set that contains all entities discovered so far, as well as their aliases:
 \begin{itemize}
-\item The page title, if we know its category;
-\item Title of redirect pages that target the current page; % TODO
-\item Entities marked with links in the text thus far.
+\item If the current page is a NE:
+  \begin{itemize}
+  \item The page title, with Wikipedia antics, such as a category name in parentheses removed
+  \item Title of redirect pages that target the current page
+  \item Title of the current page in other languages, extracted from the interlanguage link file, as Wikipedia usually mentions the original name of foreign entities
+  \end{itemize}
+\item Entities marked with links in the text thus far
+\item Certain subsets of the above:
+  \begin{itemize}
+  \item First and last names of \texttt{PER} entities
+  \item Numbers, such as \textit{737} in \textit{Boeing 737}
+  \end{itemize}
 \end{itemize}
 
-When an unmarked entity is encountered, the list is consulted to find an item that contains it in its entirety; i.e.~all of its words, in the same order.
-If either exactly one such item is found, or the matching entities in the list all have the same NE tag, the unmarked entity is tagged accordingly.
-Otherwise the sentence is discarded.
+When an unmarked entity is encountered, the set is consulted to find an item that contains it in its entirety; i.e.~all of its words, in the same order.
+If either exactly one such item is found, or the matching entities in the list all have the same NE tag, the unmarked entity is tagged accordingly; otherwise it is tagged as \texttt{UNK}.
 
-%Both approaches are evaluated in the next chapter.
+% Eszter: ehhez a felső listához hivatkozzam Nothmant? A fele tőle jön, viszont mi máshogy csináltuk kicsit.
+Our solution is similar to Nothman's \shortcite{Nothman:09}.
+Instead of a trie, however, we use a set, albeit with more alias types.
+We expect more precise tagging from our slightly more rigorous solution.
 
-%Szerintem ide kéne a DBpedia--CoNLL mapping
-We use DBpedia as
+\subsubsection{Special Cases}
 
-Only sentences for whose entities the correct category could be obtained from DBpedia are added to the corpus; those that contain unknown entities are discarded.
+There are some special cases to the algorithm described thus far. These are detailed below.
 
-%\subsubsection{Tagging}
+\begin{description}
+\item[Derived words] According to the CoNLL guidelines, word derived from name entities must be tagged as \texttt{MISC}.
+We complied with this rule by tagging each entity whose head is not a noun, as well as when the link's anchor text is not contained in the entity's name, as \texttt{MISC}.
+The most prominent example for such entities are ethnic groups, which can be linked to their home country, a \texttt{LOC}; e.g.
+\textit{Turkish} to \textit{Turkey}.
+Our solution assigns the correct tag to these entities.
+\item[First word in a sentence] As first words are always capitalized, labelling them is difficult if they are unlinked and not contained in the entity alias set.
+We base the decision on the POS tag of the first word: if it is \texttt{NNP}, we tag it as \texttt{UNK}; otherwise, \texttt{O}.
+\item[Reference cleansing] Page titles and anchor texts may contain more than just the entity name.
+Personal titles are part of the entity name in Wikipedia, but not in CoNLL, and punctuation marks around the entity may become part of the link by mistake.
+We tag all punctuation marks after the entity name as \texttt{O}.
 
-%TODO: merge with Entity Identification
-%megtörtént
-
-\subsubsection{Exceptions}
-
-Proper entities that are not deemed NE under the ConLL scheme cause XXX little concern: they are either missing from DBpedia altogether, or we can map the classes they belong to to the NER category \texttt{0} (not an entity).
-This "ontological filtering" ensures that we need no adjustment XXX to our original method.
-
-Links to number and common noun pages are handled differently by the two approaches.
-The precision-oriented method treats them similarly to links to unknown entities and discards the sentences in which they occur.
-The recall-oriented approach tries to distinguish them from links to entities so that otherwise valid training sentences can be retained XXX.
-The method is as follows XXX.
-\begin{itemize}
-\item Common nouns are filtered by POS category: links that contain no proper nouns are ignored.
-\item Numbers behave more like proper nouns in that they can refer to entities, such as 737 to the Boeing 737. Therefore, links to number pages are filtered by the surface form of the page title.
-\end{itemize}
-
-Time expressions require special attention due to two reasons.
-First, Wikipedia has pages for years, months and concrete dates, therefore our algorithm might mistake them for entities.
-Second, as dates are made up of numbers and proper nouns, they can be mistaken for unmarked entities.
-To avoid these errors, we have create a set of regular expressions that recognize dates and time expressions and changed the recall-oriented method to not treat them as entities.
-
-from \cite{Nothman:09}: Exceptions are made for common titlecase words, e.g. I, Mr., June, and sentence-initial words.
-
-% TODO: monetary, etc. units?
+To tackle personal titles, we extracted a list from the Wikipedia page \textit{List of titles}, which contains titles in many languages.
+We manually removed all titles that also function as given names, such as \textit{Regina}.
+If a link to a \texttt{PER} or \texttt{UNK} entity, or an unlinked entity starts with, or consists solely of a title in the list, we tag the words that make up the title as \texttt{O}.
+\item[Incidental capitalization] Various non-\texttt{NNP} words in English are capitalized: names of months, the \textit{I} pronoun and non-entity acronyms, such as \textit{RSVP}.
+While the latter two group is unlikely to appear in Wikipedia text, we assembled a list of these words and tag them as \texttt{O} unless they are part of the alias set.
+\end{description}
 

### 2012-03-25 paper/engNERwiki.tex
Hungarian resz megirva

--- 2012-03-25 paper/engNERwiki.tex
@@
 For instance, the class \texttt{SportsLeague} is part of the \texttt{Organisation} sub-hierarchy, but according to the CoNLL annotation scheme, they are tagged as MISC.
-Similarly, the class \texttt{SportsTeamSeason} is under \texttt{Event}, by which account it would be classified as MISC; however, ConLL does not recognize such entities.
+Similarly, the class \texttt{SportsTeamSeason} is under \texttt{Event}, by which account it would be classified as MISC; however, CoNLL does not recognize such entities.
+
 To account for these misclassification, as well as the miscellaneous classes outside the main hierarchies, we have introduced a mapping mechanism.
 We created a file of DBpedia class--NE category mappings.
-Whenever an entity is evaluated, we look up its class, as well as the ancestors of its class, in the mapping, and assign to the entity the category of the class that matches it most closely -- i.e.~the category of the entity's class, if it is in the list, or that of its closest ancestor.
-XXX!!!
+Whenever an entity is evaluated, we look up its class, as well as the ancestors of its class, in the mapping, and assign the category of the class that matches it most closely to the entity.
 If no match is found, the entity is tagged with \texttt{O}.
-Since we take advantage of the inheritance hierarchy, the mapping list remains short: it contains only the root classes of the main sub-hierarchies, exceptions of the like XXX mentioned above and the various classes that belong to the MISC category according to the CoNLL annotation guideline.
+Since we take advantage of the inheritance hierarchy, the mapping list remains short: it contains only the root classes of the main hierarchies, exceptions like those mentioned above, and the various classes that belong to the MISC category according to the CoNLL annotation guideline.
 
@@
 The second method is manually deciding the right NE mapping for the ambiguous XXX class and adding it to the mapping file.
-The first approach, unfortunately, already fails to assign the correct CoNLL NE tag to our only example.
+
+The first approach, unfortunately, already fails to assign the correct ConLL NE tag to our only example.
 The reason behind this failure is that multiple parents in an ontology show an inherent ambiguity in the class in question: it is entirely natural to think of a library as a location as well as a public organization.
-DBpedia's guidelines prefers Organisation; while ConLL's one refers it as a Location.
+DBpedia's guidelines prefers Organisation, while ConLL's one refers it as a Location.
 Therefore, only the second, manual approach ensures correct tagging.
@@
 
-\subsection{Extracting and Filtering Sentences}
+Using the mapping described above, we compile an entity list that contains all entities in DBpedia tagged with the appropriate XXX CoNLL NE category.
 
-A NER corpus consists of sentences that accurately tag NEs occurring in them.
+\subsection{Named Entity Labeling}
+
+A NER corpus consists of sentences where NEs are accurately tagged.
 In order to automatically prepare such sentences, two tasks need to be performed: identifying the entities in the sentence and tagging them with the correct tag.
 Sentences, for which accurate tagging could not be accomplished, must be removed from the corpus.
-In this section, we describe how these steps were implemented, as well as the post-filtering we employed to dispose of correctly tagged, but unusable training sentences.
 
-\subsubsection{Identifying the entities}
-% TODO: identification and tagging happens in parallel, else the recall oriented approach cannot be explained.
+Our approach to entity identification relies on Wikipedia cross-references found in the article text.
+We assume that individual Wikipedia articles describe named entities.
+A link to an article can then be perceived as a mapping that identifies its anchor text with a particular entity.
+
+Discovered entities are tagged with the CoNLL category assigned to them in the entity list extracted from DBpedia.
+If the link target is not in the entity list, or the link points to a disambiguation page, we cannot determine the type of the entity and tag it as \texttt{UNK}.
+Links to redirect pages are resolved to point instead to the redirect target, after which they are handled as regular cross-references.
+
+There are several
+
+\subsubsection{Non-entity Links}
 
@@
 
-However, not all entities are marked with links.
-Authors may neglect to add XXX references to their text.
-Furthermore, in a Wikipedia article XXX, typically only the first occurrence of a particular entity is linked to the corresponding page.
+%Strictly speaking, our original assumption of equating Wikipedia articles with named entities is not valid: many pages describe common nouns (Book, Aircraft), numbers (1, 42), currencies (Euro, Dollar) or other concepts that fall outside the scope of NER.
+
+\begin{description}
+\item[Common nouns] Links to common nouns are filtered by POS tags; if a link contains no \texttt{NNP}s, it is ignored.
+\item[Dates] Wikipedia has pages for years, months and concrete dates, therefore our algorithm might mistake them for entities.
+To prevent this problem, we compiled a list of calendar-related entities and added them to the main entity list tagged with the CoNLL category \texttt{O}.
+\item[Lowercase links] 
+\end{description}
+
+\subsubsection{Unmarked Entities}
+
+\subsubsection{Special Cases}
+
+Not all entities are marked with links.
+Authors may neglect to add references to their text.
+Furthermore, in a Wikipedia article, typically only the first occurrence of a particular entity is linked to the corresponding page.
 Subsequent mentions are unmarked and often incomplete -- e.g.~family names are used instead of full names.
-We devised XXX two methods to account for unmarked entities: one that emphasizes XXX precision, and the other, recall.
-XXX
-
-% Or NPP
-
-The precision-oriented approach seeks to minimize the risk of words getting incorrectly tagged XXX in the corpus. Names and their parts mostly start with capital letter, thus 
-%A short study of the CoNLL annotation guidelines reveals that the words in the entities of interest to us all start with capital letter. According to this method XXX, 
-all sentences that contain capitalized words, but are not part of the anchor text of a link, XXX were discarded.
-The rationale behind this approach is that the sheer size of Wikipedia guarantees that we end up with a sizeable corpus even with such a strict policy.
-XXX
+To account for such entities,
 
 The recall-oriented method tries to associate unmarked entities with entities already identified on the same page. Only if the attempt fails is the sentence that contains the words discarded. 
-%Ez utóbbi biztosan jó mondat?
 This approach aims to generate more sentences with as little loss of precision as possible.
@@
 
-%\subsubsection{Tagging}
-
-%TODO: merge with Entity Identification
-%megtörtént
-
 \subsubsection{Exceptions}
 
-Strictly speaking, our original assumption of equating Wikipedia articles with NEs is not valid: many pages describe common nouns (Book, Aircraft), numbers (1, 42), currencies (Euro, Dollar) or other concepts that fall outside the scope of CoNLL tagset or of NER in general.
+%Strictly speaking, our original assumption of equating Wikipedia articles with NEs is not valid: many pages describe common nouns (Book, Aircraft), numbers (1, 42), currencies (Euro, Dollar) or other concepts that fall outside the scope of CoNLL tagset or of NER in general.
 In this section, we describe the measures taken to avoid errors introduced by these false entities.
@@
 The method is as follows XXX.
+
 \begin{itemize}
@@
 
-% 4. Test sentence extraction?
-%   a. unknown links, no unlinked names
-%   b. might be good for quality evaluation / recursive tagging (next paper?)
-% 5. teaching the original DBpedia categories?
-
-In Wikipedia the nationalities (e.g.~American and Australian) links to the corresponding places (America and Australia, respectively).
+In Wikipedia the nationalities (e.g.
+American and Australian) links to the corresponding places (America and Australia, respectively).
 However, in CoNLL scheme these types of entities take parts of MISC category.
@@
 
-\subsection{Tools}
+\subsection{Parsing the Hungarian Wikipedia}
 
@@
 It has been trained on the largest manually annotated Hungarian corpus \cite{Csendes:04}, and it handles the peculiarities of Hungarian ortography, such as the periods placed after numbers in date expressions.
-Lemmatization was performed by \texttt{hunmorph} \cite{Tron:05} and \texttt{hundisambig}, latter being an in-house disambiguator to select the right analysis based on the word context.
+Lemmatization was performed by \texttt{hunmorph} \cite{Tron:05} and \texttt{hundisambig}, the latter being an in-house disambiguator to select the right analysis based on the word context.
 
@@
 
-The Hungarian morphological analyzer outputs KR-codes \cite{Kornai:04}.
-In addition to the POS category, the KR-code also includes inflection information, such as the number, person and case XXX, which makes it much better suited to agglutinative languages than Penn Treebank POS tags.
-One shortcoming of the KR code is that it does not differenciate between common and proper nouns.
+Hungarian is called a synthetic language: for the most part, it expresses grammatical elements in a single word form using affixes, as opposed to isolating languages (as English), which tend to employ separate words, e.g.~prepositions, pronouns, auxiliaries, for expressing grammatical phenomena.
+The Hungarian morphological analyzer outputs KR-codes \cite{Kornai:04}, which, in addition to the POS category, also includes inflectional information, which makes it much better suited to agglutinative languages, than Penn Treebank POS tags.
+One shortcoming of the KR-code is that it does not differenciate between common and proper nouns.
 Since in Hungarian only proper nouns are capitalized, we can usually decide whether a noun is proper or not based on the initial letter.
@@
 
-\subsection{Data Availability}
+\subsection{Named Entity Labeling in Hungarian}
 
-For some of the bigger languages, DBpedia has internationalized "chapters" -- data dumps based on the Wikipedia of that particular language.
-Hungarian, being a relatively small language XXX, does not have a separate chapter.
+For some of the well-resourced languages, DBpedia has internationalized "chapters" -- data dumps based on the Wikipedia of those particular languages.
+Hungarian does not have a separate chapter.
 Instead, the Hungarian entity list comprises of the pages in the English list that have their equivalents in the Hungarian Wikipedia.
@@
 First, in order to identify which pages denote entities in the Hungarian Wikipedia, an additional step is required, in which the Hungarian equivalents of the English pages are added to the entity list.
-The English titles are retained, as due to the relatively small size of the Hungarian Wikipedia, in-article links sometimes point to English articles.
+The English titles are retained, as due to the medium size of the Hungarian Wikipedia, in-article links sometimes point to English articles.
 
@@
 
-\subsection{NER Conventions}
+To annotate our corpus with NE-tags, we chose to follow the annotation guidelines of the largest human-annotated NER corpus for Hungarian, Szeged NER corpus \cite{Szarvas:06}.
+It is quite similar to the CoNLL standards: it contains business newswire texts, it comprises ca.~200,000 tokens, and the text has been annotated with NE class labels in line with the CoNLL annotation scheme.
+However, the convention of what constitutes a NE is slightly different for Hungarian.
 
-% Eszter
-The largest (?, the only?, most cited?) NER corpus in Hungarian is the Szeged NER corpus, which will be described by Eszter.
-:) The Szeged corpus uses the same named entity categories as ConLL; however, the convention of what constitutes an entity is slightly different.
+\subsubsection{Special cases}
 
-The Szeged NER guideline relies heavily on the rules of capitalization to decide which words should be marked as named entities.
+The Szeged NER guideline relies heavily on the rules of capitalization to decide which words should be marked as NEs.
 As mentioned earlier, in Hungarian, apart from the first word in the sentence, only proper nouns are capitalized.
-The following concepts are not proper nouns in Hungarian and thus are not considered named entities according to the Szeged guideline:
+The following concepts are not proper nouns in Hungarian, and thus are not considered as NEs according to the Szeged NER guideline:
+
 \begin{itemize}
-\item Names of languages and ethnic groups
-\item Words created from proper nouns through derivation; most prominently
-\item Adjectival form of location names
-\item Months
+\item names of languages, nationalities, religions, poltical ideologies;
+\item adjectives derived from a NE;
+\item names of months, days, holidays;
+\item names of special events and wars.
 \end{itemize}
 
-The strong correlation between capitalization and named entities makes recognition of entities in text relatively simple: a word is part of an entity if it is capitalized.
-% Eszter
-a tulajdonnevekhez kötőjellel kapcsolódó köznevek kérdésesek (pl. SAP-szoftvercég, Matáv-csoport)
+There is an other special case in Hungarian: opposite to English, the number of compund words is quite large in Hungarian, and NEs can also be subjects of compounding.
+In this case the common noun following the NE is joint with a hyphen, so they constitute one token.
+However, the joint common noun can modify the original sense of NE, depending on the semantics of the common noun.
+For example: in the compound \textit{SAP-szoftvercég} ['SAP software company'] the common noun does not change the labeling, but in the case of the compound \textit{WorldCom-botrány} ['WorldCom scandal'] the NE tag changes from ORG to O.
+
+Inflections of acronyms and foreign names ending with non-pronounced vowel have similar surface forms to the aforementioned compounds, e.g.~\textit{MTI-t} ['MTI'+ACC], \textit{Shakespeare-rel} ['with Shakespeare'].
+It is important to distinguish these types of hyphenated NEs, because inflections do not change NE labeling in contrast to some type of compounds.
+Due to the surface similarity of these phenomena the solution of this problem is not obvious, and needs more investigation.
 
 % Ezt hová?
-most van: 102 millió szó a disambiguation pages-zel együtt
-az angol: 1,7 milliárd
+%most van: 102 millió szó a disambiguation pages-zel együtt
+%az angol: 1,7 milliárd
 
@@
 Results of experiments done on the Hungarian corpus look similar.
+
+Although two human-annotated NER corpora exist for Hungarian, we chose the Szeged NER corpus \cite{}, which follows the CoNLL annotation scheme. It contains ca.
+
+The strong correlation between capitalization and named entities makes recognition of entities in text relatively simple: a word is part of an entity if it is capitalized.
 
@@
 
+\bibitem[\protect\citename{Szarvas \bgroup et al.\egroup}2006]{Szarvas:06}
+György Szarvas, Richárd Farkas, András Kocsor.
+\newblock 2006.
+\newblock A highly accurate Named Entity corpus for Hungarian. 
+\newblock In: {\em Proceedings of International Conference on Language Resources and Evaluation}.
+
 \bibitem[\protect\citename{Tjong Kim Sang}2002]{Tjong:02}

### 2012-03-25 paper/engNERwiki.tex

--- 2012-03-25 paper/engNERwiki.tex
@@
 
-\subsection{Extracting and Filtering Sentences}
+Using the mapping described above, we compile an entity list that contains all entities in DBpedia tagged with the appropriate CoNLL named entity category.
 
-A NER corpus consists of sentences that accurately tag NEs occurring in them.
+We would like to note that even though we use CoNLL categories, our method can be modified to work with any tagset compatible with the DBpedia ontology.
+Indeed, the DBpedia classes define a named entity tagset themselves.
+We leave the exploration of these possibilities for future work.
+
+\subsection{Named Entity Labeling}
+
+A NER corpus consists of sentences where NEs are accurately tagged.
 In order to automatically prepare such sentences, two tasks need to be performed: identifying the entities in the sentence and tagging them with the correct tag.
 Sentences, for which accurate tagging could not be accomplished, must be removed from the corpus.
-In this section, we describe how these steps were implemented, as well as the post-filtering we employed to dispose of correctly tagged, but unusable training sentences.
-
-\subsubsection{Identifying the entities}
-% TODO: identification and tagging happens in parallel, else the recall oriented approach cannot be explained.
 
 Our approach to entity identification relies on Wikipedia cross-references found in the article text.
-We assume that individual Wikipedia articles describe NEs.
+We assume that individual Wikipedia articles describe named entities.
 A link to an article can then be perceived as a mapping that identifies its anchor text with a particular entity.
-However, there are exceptions to this rule.
-Disambiguation pages are used when the page title is ambiguous, and can refer to several entities.
-If a link points to a disambiguation page, we acknowledge the anchor text as an entity, but mark it unknown.
-A similar measure is taken if the link target XXX does not exist.
+
+Discovered entities are tagged with the CoNLL category assigned to them in the entity list extracted from DBpedia.
+If the link target is not in the entity list, or the link points to a disambiguation page, we cannot determine the type of the entity and tag it as \texttt{UNK}.
 Links to redirect pages are resolved to point instead to the redirect target, after which they are handled as regular cross-references.
+Finally, sentences with UNK links in them are removed from the corpus.
 
-However, not all entities are marked with links.
-Authors may neglect to add XXX references to their text.
-Furthermore, in a Wikipedia article XXX, typically only the first occurrence of a particular entity is linked to the corresponding page.
+There are several ways to improve the method described above to increase precision, sentence coverage and to account for peculiarities in the CoNLL guidelines.
+These are detailed in the next sub-sections.
+
+\subsubsection{Non-entity Links}
+
+Strictly speaking, our original assumption of equating Wikipedia articles with named entities is not valid: many pages describe common nouns (Book, Aircraft), calendar-related concepts (March 15, 2007), or other concepts that fall outside the scope of NER.
+To increase sentence coverage, we modified the algorithm to prevent it from misclassifying links to these pages as unknown entities and discarding the sentence.
+
+\begin{description}
+\item[Common nouns] Links to common nouns are filtered by POS tags; if a link contains no \texttt{NNP}s, it is ignored.
+\item[Dates] Time expressions require special attention, because dates and months are often linked to the respective Wikipedia pages.
+We circumvented this problem by compiling a list of calendar-related pages and adding them to the main entity list tagged with the CoNLL category \texttt{O}.
+\item[Lowercase links] Entities are sometimes referred to by common nouns, such as \textit{republic} to \textit{Roman Republic}, \texttt{LOC}, which results in a lowercase link.
+These references are not considered named entities and are ignored.
+\end{description}
+
+\subsubsection{Unmarked Entities}
+
+Not all entities are marked with links.
+Authors may neglect to add references to their text.
+Furthermore, in a Wikipedia article, typically only the first occurrence of a particular entity is linked to the corresponding page.
 Subsequent mentions are unmarked and often incomplete -- e.g.~family names are used instead of full names.
-We devised XXX two methods to account for unmarked entities: one that emphasizes XXX precision, and the other, recall.
-XXX
+To account for such mentions, we try to associate capitalized words in the article text with entities we have tagged thus far.
 
-% Or NPP
-
-The precision-oriented approach seeks to minimize the risk of words getting incorrectly tagged XXX in the corpus. Names and their parts mostly start with capital letter, thus 
-%A short study of the CoNLL annotation guidelines reveals that the words in the entities of interest to us all start with capital letter. According to this method XXX, 
-all sentences that contain capitalized words, but are not part of the anchor text of a link, XXX were discarded.
-The rationale behind this approach is that the sheer size of Wikipedia guarantees that we end up with a sizeable corpus even with such a strict policy.
-XXX
-
-The recall-oriented method tries to associate unmarked entities with entities already identified on the same page. Only if the attempt fails is the sentence that contains the words discarded. 
-%Ez utóbbi biztosan jó mondat?
-This approach aims to generate more sentences with as little loss of precision as possible.
-
-The algorithm maintains a list of entities already indentified on the page. The list includes:
+For each article, we build a set that contains all entities discovered so far, as well as their aliases:
 \begin{itemize}
-\item The page title; % if it is in DBpedia
-\item Title of redirect pages that target the current page; % TODO
-\item Entities marked with links in the text thus far.
+\item If the current page is a NE:
+  \begin{itemize}
+  \item The page title, with Wikipedia antics, such as a category name in parentheses removed
+  \item Title of redirect pages that target the current page
+  \item Title of the current page in other languages, extracted from the interlanguage link file, as Wikipedia usually mentions the original name of foreign entities
+  \end{itemize}
+\item Entities marked with links in the text thus far
+\item Certain subsets of the above:
+  \begin{itemize}
+  \item First and last names of \texttt{PER} entities
+  \item Numbers, such as \textit{737} in \textit{Boeing 737}
+  \end{itemize}
 \end{itemize}
 
-When an unmarked entity is encountered, the list is consulted to find an item that contains it in its entirety; i.e.~all of its words, in the same order.
-If either exactly one such item is found, or the matching entities in the list all have the same NE tag, the unmarked entity is tagged accordingly.
-Otherwise the sentence is discarded.
+When an unmarked entity is encountered, the set is consulted to find an item that contains it in its entirety; i.e.~all of its words, in the same order.
+If either exactly one such item is found, or the matching entities in the list all have the same NE tag, the unmarked entity is tagged accordingly; otherwise it is tagged as \texttt{UNK}.
 
-Only sentences for whose entities the correct category could be obtained from DBpedia are added to the corpus; those that contain unknown entities are discarded.
+% Eszter: ehhez a felső listához hivatkozzam Nothmant? A fele tőle jön, viszont mi máshogy csináltuk kicsit.
+Our solution is similar to Nothman's \shortcite{Nothman:09}.
+Instead of a trie, however, we use a set, albeit with more alias types.
+We expect more precise tagging from our slightly more rigorous solution.
 
-%\subsubsection{Tagging}
+\subsubsection{Special Cases}
 
-%TODO: merge with Entity Identification
-%megtörtént
+There are some special cases to the algorithm described thus far. These are detailed below.
 
-\subsubsection{Exceptions}
+\begin{description}
+\item[Derived words] According to the CoNLL guidelines, word derived from name entities must be tagged as \texttt{MISC}.
+We complied with this rule by tagging each entity whose head is not a noun, as well as when the link's anchor text is not contained in the entity's name, as \texttt{MISC}.
+The most prominent example for such entities are ethnic groups, which can be linked to their home country, a \texttt{LOC}; e.g.
+\textit{Turkish} to \textit{Turkey}.
+Our solution assigns the correct tag to these entities.
+\item[First word in a sentence] As first words are always capitalized, labelling them is difficult if they are unlinked and not contained in the entity alias set.
+We base the decision on the POS tag of the first word: if it is \texttt{NNP}, we tag it as \texttt{UNK}; otherwise, \texttt{O}.
+\item[Reference cleansing] Page titles and anchor texts may contain more than just the entity name.
+Personal titles are part of the entity name in Wikipedia, but not in CoNLL, and punctuation marks around the entity may become part of the link by mistake.
+We tag all punctuation marks after the entity name as \texttt{O}.
 
-Strictly speaking, our original assumption of equating Wikipedia articles with NEs is not valid: many pages describe common nouns (Book, Aircraft), numbers (1, 42), currencies (Euro, Dollar) or other concepts that fall outside the scope of CoNLL tagset or of NER in general.
-In this section, we describe the measures taken to avoid errors introduced by these false entities.
-We shall modify the two approaches -- precision and recall-oriented -- accordingly.
-
-Proper names that are not deemed NE under the ConLL scheme
-%ilyeneknek nem szabad lenniük a MISC miatt
- cause XXX little concern: they are either missing from DBpedia altogether, or we can map the classes they belong to to the NER category \texttt{0} (not an entity).
-This "ontological filtering" ensures that we need no adjustment XXX to our original method.
-
-Links to number and common noun pages are handled differently by the two approaches.
-The precision-oriented method treats them similarly to links to unknown entities and discards the sentences in which they occur.
-The recall-oriented approach tries to distinguish them from links to entities so that otherwise valid sentences can be retained XXX.
-The method is as follows XXX.
-\begin{itemize}
-\item Common nouns are filtered by POS category: links that contain no proper nouns are ignored.
-\item Numbers behave more like proper nouns in that they can refer to entities, such as 737 to the Boeing 737. Therefore, links to number pages are filtered by the surface form of the page title.
-\end{itemize}
-
-Time expressions require special attention due to two reasons.
-First, Wikipedia has pages for years, months and concrete dates, therefore our algorithm might mistake them for entities.
-Second, as dates are made up of numbers and proper nouns, they can be mistaken for unmarked entities.
-To avoid these errors, we have create a set of regular expressions that recognize dates and time expressions and changed the recall-oriented method to not treat them as entities.
-
-XXX from \cite{Nothman:09}: Exceptions are made for common titlecase words, e.g. I, Mr., June, and sentence-initial words.
-
-% TODO: monetary, etc. units?
+To tackle personal titles, we extracted a list from the Wikipedia page \textit{List of titles}, which contains titles in many languages.
+We manually removed all titles that also function as given names, such as \textit{Regina}.
+If a link to a \texttt{PER} or \texttt{UNK} entity, or an unlinked entity starts with, or consists solely of a title in the list, we tag the words that make up the title as \texttt{O}.
+\item[Incidental capitalization] Various non-\texttt{NNP} words in English are capitalized: names of months, the \textit{I} pronoun and non-entity acronyms, such as \textit{RSVP}.
+While the latter two group is unlikely to appear in Wikipedia text, we assembled a list of these words and tag them as \texttt{O} unless they are part of the alias set.
+\end{description}
 

### 2012-03-25 paper/engNERwiki.tex

@@
 
+Although two human-annotated NER corpora exist for Hungarian, we chose the Szeged NER corpus \cite{}, which follows the CoNLL annotation scheme. It contains ca.
+
+The strong correlation between capitalization and named entities makes recognition of entities in text relatively simple: a word is part of an entity if it is capitalized.
+
 \section{Conclusions and Future Work}

### 2012-03-25 paper/engNERwiki.tex
09->08.
Appendix removed.

@@
 % Eszter: ehhez a felső listához hivatkozzam Nothmant? A fele tőle jön, viszont mi máshogy csináltuk kicsit.
-Our solution is similar to Nothman's \shortcite{Nothman:09}.
+Our solution is similar to Nothman's \shortcite{Nothman:08}.
 Instead of a trie, however, we use a set, albeit with more alias types.
@@
 \begin{itemize}
-\item names of languages, nationalities, religions, poltical ideologies;
+\item names of languages, nationalities, religions, political ideologies;
 \item adjectives derived from a NE;
@@
 
-There is an other special case in Hungarian: opposite to English, the number of compund words is quite large in Hungarian, and NEs can also be subjects of compounding.
+There is another special case in Hungarian: opposite to English, the number of compund words is quite large in Hungarian, and NEs can also be subjects of compounding.
 In this case the common noun following the NE is joint with a hyphen, so they constitute one token.
@@
 
-% HEEEEELP!
-\appendix
-\section{DBpedia to ConLL Mapping}
-\label{sec:appa}
-
-\begin{table*}[ht]
-\begin{center}
-\begin{tabular}{llll}
-\hline DBpedia & \bf CoNLL & \bf DBpedia & \bf CoNLL \\ \hline
-Person & PER & Library & LOC \\
-Place & LOC & MeanOfTransportation & MISC \\
-Organization & ORG & PeriodicalLiterature & ORG \\
-Organisation & ORG & ProgrammingLanguage & MISC \\
-Award & MISC & Project & MISC \\
-EthnicGroup & MISC & SportsLeague & MISC \\
-Event & MISC & SportsTeamSeason & O \\
-Holiday & MISC & Weapon & MISC \\
-Ideology & MISC & Work & MISC \\
-Language & MISC & & \\
-\hline
-\end{tabular}
-\end{center}
-\end{table*}
+%\appendix
+%\section{DBpedia to ConLL Mapping}
+%\label{sec:appa}
+%
+%\begin{table*}[ht]
+%\begin{center}
+%\begin{tabular}{llll}
+%\hline DBpedia & \bf CoNLL & \bf DBpedia & \bf CoNLL \\ \hline
+%Person & PER & Library & LOC \\
+%Place & LOC & MeanOfTransportation & MISC \\
+%Organization & ORG & PeriodicalLiterature & ORG \\
+%Organisation & ORG & ProgrammingLanguage & MISC \\
+%Award & MISC & Project & MISC \\
+%EthnicGroup & MISC & SportsLeague & MISC \\
+%Event & MISC & SportsTeamSeason & O \\
+%Holiday & MISC & Weapon & MISC \\
+%Ideology & MISC & Work & MISC \\
+%Language & MISC & & \\
+%\hline
+%\end{tabular}
+%\end{center}
+%\end{table*}
 

### 2012-03-25 paper/engNERwiki.tex
Data description

--- 2012-03-25 paper/engNERwiki.tex
@@
 
-Using the mapping described above, we compile an entity list that contains all entities in DBpedia tagged with the appropriate CoNLL named entity category.
-
-We would like to note that even though we use CoNLL categories, our method can be modified to work with any tagset compatible with the DBpedia ontology.
-Indeed, the DBpedia classes define a named entity tagset themselves.
-We leave the exploration of these possibilities for future work.
+Using the mapping described above, we compile an entity list that contains all entities in DBpedia tagged with the appropriate XXX CoNLL NE category.
 
@@
 Links to redirect pages are resolved to point instead to the redirect target, after which they are handled as regular cross-references.
-Finally, sentences with UNK links in them are removed from the corpus.
 
-There are several ways to improve the method described above to increase precision, sentence coverage and to account for peculiarities in the CoNLL guidelines.
-These are detailed in the next sub-sections.
+There are several
 
@@
 
-Strictly speaking, our original assumption of equating Wikipedia articles with named entities is not valid: many pages describe common nouns (Book, Aircraft), calendar-related concepts (March 15, 2007), or other concepts that fall outside the scope of NER.
-To increase sentence coverage, we modified the algorithm to prevent it from misclassifying links to these pages as unknown entities and discarding the sentence.
+Our approach to entity identification relies on Wikipedia cross-references found in the article text.
+We assume that individual Wikipedia articles describe NEs.
+A link to an article can then be perceived as a mapping that identifies its anchor text with a particular entity.
+However, there are exceptions to this rule.
+Disambiguation pages are used when the page title is ambiguous, and can refer to several entities.
+If a link points to a disambiguation page, we acknowledge the anchor text as an entity, but mark it unknown.
+A similar measure is taken if the link target XXX does not exist.
+Links to redirect pages are resolved to point instead to the redirect target, after which they are handled as regular cross-references.
+
+%Strictly speaking, our original assumption of equating Wikipedia articles with named entities is not valid: many pages describe common nouns (Book, Aircraft), numbers (1, 42), currencies (Euro, Dollar) or other concepts that fall outside the scope of NER.
 
@@
 \item[Common nouns] Links to common nouns are filtered by POS tags; if a link contains no \texttt{NNP}s, it is ignored.
-\item[Dates] Time expressions require special attention, because dates and months are often linked to the respective Wikipedia pages.
-We circumvented this problem by compiling a list of calendar-related pages and adding them to the main entity list tagged with the CoNLL category \texttt{O}.
-\item[Lowercase links] Entities are sometimes referred to by common nouns, such as \textit{republic} to \textit{Roman Republic}, \texttt{LOC}, which results in a lowercase link.
-These references are not considered named entities and are ignored.
+\item[Dates] Wikipedia has pages for years, months and concrete dates, therefore our algorithm might mistake them for entities.
+To prevent this problem, we compiled a list of calendar-related entities and added them to the main entity list tagged with the CoNLL category \texttt{O}.
+\item[Lowercase links] 
 \end{description}
@@
 \subsubsection{Unmarked Entities}
+
+\subsubsection{Special Cases}
 
@@
 Subsequent mentions are unmarked and often incomplete -- e.g.~family names are used instead of full names.
-To account for such mentions, we try to associate capitalized words in the article text with entities we have tagged thus far.
+To account for such entities,
 
-For each article, we build a set that contains all entities discovered so far, as well as their aliases:
+The recall-oriented method tries to associate unmarked entities with entities already identified on the same page. Only if the attempt fails is the sentence that contains the words discarded. 
+This approach aims to generate more sentences with as little loss of precision as possible.
+
+The algorithm maintains a list of entities already indentified on the page. The list includes:
 \begin{itemize}
-\item If the current page is a NE:
-  \begin{itemize}
-  \item The page title, with Wikipedia antics, such as a category name in parentheses removed
-  \item Title of redirect pages that target the current page
-  \item Title of the current page in other languages, extracted from the interlanguage link file, as Wikipedia usually mentions the original name of foreign entities
-  \end{itemize}
-\item Entities marked with links in the text thus far
-\item Certain subsets of the above:
-  \begin{itemize}
-  \item First and last names of \texttt{PER} entities
-  \item Numbers, such as \textit{737} in \textit{Boeing 737}
-  \end{itemize}
+\item The page title, if we know its category;
+\item Title of redirect pages that target the current page; % TODO
+\item Entities marked with links in the text thus far.
 \end{itemize}
 
-When an unmarked entity is encountered, the set is consulted to find an item that contains it in its entirety; i.e.~all of its words, in the same order.
-If either exactly one such item is found, or the matching entities in the list all have the same NE tag, the unmarked entity is tagged accordingly; otherwise it is tagged as \texttt{UNK}.
+When an unmarked entity is encountered, the list is consulted to find an item that contains it in its entirety; i.e.~all of its words, in the same order.
+If either exactly one such item is found, or the matching entities in the list all have the same NE tag, the unmarked entity is tagged accordingly.
+Otherwise the sentence is discarded.
 
-% Eszter: ehhez a felső listához hivatkozzam Nothmant? A fele tőle jön, viszont mi máshogy csináltuk kicsit.
-Our solution is similar to Nothman's \shortcite{Nothman:09}.
-Instead of a trie, however, we use a set, albeit with more alias types.
-We expect more precise tagging from our slightly more rigorous solution.
+Only sentences for whose entities the correct category could be obtained from DBpedia are added to the corpus; those that contain unknown entities are discarded.
 
-\subsubsection{Special Cases}
+\subsubsection{Exceptions}
 
-There are some special cases to the algorithm described thus far. These are detailed below.
+%Strictly speaking, our original assumption of equating Wikipedia articles with NEs is not valid: many pages describe common nouns (Book, Aircraft), numbers (1, 42), currencies (Euro, Dollar) or other concepts that fall outside the scope of CoNLL tagset or of NER in general.
+In this section, we describe the measures taken to avoid errors introduced by these false entities.
+We shall modify the two approaches -- precision and recall-oriented -- accordingly.
 
-\begin{description}
-\item[Derived words] According to the CoNLL guidelines, word derived from name entities must be tagged as \texttt{MISC}.
-We complied with this rule by tagging each entity whose head is not a noun, as well as when the link's anchor text is not contained in the entity's name, as \texttt{MISC}.
-The most prominent example for such entities are ethnic groups, which can be linked to their home country, a \texttt{LOC}; e.g.
-\textit{Turkish} to \textit{Turkey}.
-Our solution assigns the correct tag to these entities.
-\item[First word in a sentence] As first words are always capitalized, labelling them is difficult if they are unlinked and not contained in the entity alias set.
-We base the decision on the POS tag of the first word: if it is \texttt{NNP}, we tag it as \texttt{UNK}; otherwise, \texttt{O}.
-\item[Reference cleansing] Page titles and anchor texts may contain more than just the entity name.
-Personal titles are part of the entity name in Wikipedia, but not in CoNLL, and punctuation marks around the entity may become part of the link by mistake.
-We tag all punctuation marks after the entity name as \texttt{O}.
+Proper names that are not deemed NE under the ConLL scheme
+%ilyeneknek nem szabad lenniük a MISC miatt
+ cause XXX little concern: they are either missing from DBpedia altogether, or we can map the classes they belong to to the NER category \texttt{0} (not an entity).
+This "ontological filtering" ensures that we need no adjustment XXX to our original method.
 
-To tackle personal titles, we extracted a list from the Wikipedia page \textit{List of titles}, which contains titles in many languages.
-We manually removed all titles that also function as given names, such as \textit{Regina}.
-If a link to a \texttt{PER} or \texttt{UNK} entity, or an unlinked entity starts with, or consists solely of a title in the list, we tag the words that make up the title as \texttt{O}.
-\item[Incidental capitalization] Various non-\texttt{NNP} words in English are capitalized: names of months, the \textit{I} pronoun and non-entity acronyms, such as \textit{RSVP}.
-While the latter two group is unlikely to appear in Wikipedia text, we assembled a list of these words and tag them as \texttt{O} unless they are part of the alias set.
-\end{description}
+Links to number and common noun pages are handled differently by the two approaches.
+The precision-oriented method treats them similarly to links to unknown entities and discards the sentences in which they occur.
+The recall-oriented approach tries to distinguish them from links to entities so that otherwise valid sentences can be retained XXX.
+The method is as follows XXX.
+
+\begin{itemize}
+\item Common nouns are filtered by POS category: links that contain no proper nouns are ignored.
+\item Numbers behave more like proper nouns in that they can refer to entities, such as 737 to the Boeing 737. Therefore, links to number pages are filtered by the surface form of the page title.
+\end{itemize}
+
+Time expressions require special attention due to two reasons.
+First, Wikipedia has pages for years, months and concrete dates, therefore our algorithm might mistake them for entities.
+Second, as dates are made up of numbers and proper nouns, they can be mistaken for unmarked entities.
+To avoid these errors, we have create a set of regular expressions that recognize dates and time expressions and changed the recall-oriented method to not treat them as entities.
+
+XXX from \cite{Nothman:09}: Exceptions are made for common titlecase words, e.g. I, Mr., June, and sentence-initial words.
+
+% TODO: monetary, etc. units?
 
@@
 
-\section{Corpus Description}
+\section{Data Description}
 
-% TODO: pls írd meg az első pár mondatot és nevezd át a sectiont!
+The corpora are available under the Creative Commons Attribution-Sharealike 3.0 Unported License (CC-BY-SA), the same license under which the text of Wikipedia is released.
+The data files can be freely downloaded from ...\footnote{To be filled in in the camera-ready version.}.
 
-The corpora are available in the tab separated format used by ConLL (XXX, CITE).
-In this format, one line contains one word, its linguistic features and the named entity tags.
+The corpora are available in multitag format.
+Content lines are tab separated; there is one column for the tokens plus one column per tagset.
 Sentence boundaries are marked by empty lines.
-The linguistic features include the lemmatized form of the word, its part-of-speech and chunk tags.
-Two named entity tags are included with each word: one is the most specific DBpedia category it belongs to, if any; the other is the ConLL named entity tag.
-While the named entity tags can be considered "silver", the linguistic features are provided on a "best-effort" basis.
-
-The corpus is available under the Creative Commons Attribution-Sharealike 3.0 Unported License (CC-BY-SA), the same license under which the text of Wikipedia is released.
-The data files can be freely downloaded from ...\footnote{To be filled in in the camera-ready version.}.
+The linguistic features include the lemmatized form of the word and its part-of-speech tag.
+Two NE tags are included with each word: one is the most specific DBpedia category it belongs to, if any; the other is the CoNLL NE tag.
+While the NE tags can be considered as a "silver standard", the linguistic features are provided on a "best-effort" basis.
 
@@
 We used the CoNLL standard method for evaluation.
-According to this, an automatic labeling is correct if it gives the same start and end position, and the same NE class, as the gold standard.
+According to this, an automatic labeling is correct if it gives the same start and end position, and the same NE class as the gold standard.
 Based on this, precision and recall can be calculated, and the F-measure, as usual, the harmonic mean of these two values.
@@
 The strong correlation between capitalization and named entities makes recognition of entities in text relatively simple: a word is part of an entity if it is capitalized.
+
+training corpora in absence of gold standard datasets
+
+training sets for domains differing from newswire
 

### 2012-03-25 paper/engNERwiki.tex
Filtering.

--- 2012-03-25 paper/engNERwiki.tex
@@
 
-The full mapping can be found in Appendix~\ref{sec:appa}.
-
+% The full mapping can be found in Appendix~\ref{sec:appa}.
+%
 Using the mapping described above, we compile an entity list that contains all entities in DBpedia tagged with the appropriate CoNLL named entity category.
@@
 
-\subsubsection{Filtering}
+\subsubsection{Sentence Filtering}
 
-There are many incomplete sentences in the Wikipedia text: image captions, enumerations items, contents of table cells, etc.
+Certain types of sentences were filtered from the corpus.
+As mentioned above, sentences with words tagged as \texttt{UNK} are discarded.
+Furthermore, there are many incomplete sentences in the Wikipedia text: image captions, enumerations items, contents of table cells, etc.
 On the one hand, these sentence fragments may prove to be of too low quality to be of any use in the traditional NER task.
-On the other hand, they could prove to be invaluable when training a NER tagger for User Generated Content, which is well known to be noisy and fragmented.
-We have added a post-filtering step to recognize and label these fragments, so that users of the corpus can decide what to do with them.
-XXX
-
-Manual inspection has found two conditions that can be used to detect fragments.
+On the other hand, they could prove to be invaluable when training a NER tagger for User Generated Content, which is known to be noisy and fragmented.
+As a compromise solution, we included these fragments in the corpus, but labelled them as "low quality", so that users of the corpus can decide whether they want to use them or not.
 A sentence is labelled as such if it either lacks a punctuation mark at the end, or it contains no finite verbs.
-Our final corpus contains XXXSTAT full sentences and STAT fragments.
-%Ezeket leméred, vagy mérjem én? Ha utóbbi, akkor melyikben kell mérni?
-% TODO: Last sentence to evaluation? But then this whole subsection is too short.
-
-In Wikipedia the nationalities (e.g.
-American and Australian) links to the corresponding places (America and Australia, respectively).
-However, in CoNLL scheme these types of entities take parts of MISC category.
-XXX Postprocessing: list look-up (nationalities, adjective forms of person who live in the corresponding country); if the name is POS-tagged as adjective, then it is not a NE
-
-%nem biztos, hogy ide kell
-Link boundaries need to be adjusted, e.g.~to remove excess punctuation.
 

### 2012-03-25 paper/engNERwiki.tex
Evaluation is kesz

--- 2012-03-25 paper/engNERwiki.tex
@@
 
-The full mapping can be found in Appendix~\ref{sec:appa}.
+%The full mapping can be found in Appendix~\ref{sec:appa}.
 
@@
 For the English cross-corpus evaluation the CoNLL-2003 dataset was chosen, mostly because of the similarity in tagset and format.
-It is well-studied (e.g.
-XXX) that cross-checking between different corpora does not lead to F-measure improvement.
+It is well-studied that cross-checking between different corpora leads to F-measure decrease.
 Stylistic and domain differences certainly affect NER performance, and the different annotation schemes pose several compatibility problems.
 Nothman et al.~\shortcite{Nothman:08} showed that each set of gold standard training data performs much better on corresponding test sets than on test sets from other sources.
-The situation here is similar: the NE-tagger trained on Wikipedia does not achieve as high performance tested against CoNLL test set as one trained on the own training set (see table \ref{enresults} and \ref{huresults} for results).
+The situation here is similar: the NE-tagger trained on Wikipedia does not achieve as high performance tested against CoNLL test set as one trained on the own training set (see table \ref{enresults} for results).
 
@@
 
-Results of experiments done on the Hungarian corpus look similar.
+Since in the subtask of NE labeling for Hungarian we followed the Szeged NER corpus annotation guidelines, it is obvious to perform the experiments on this dataset, as well.
+Hungarian results are quite similar to the English ones (see table \ref{huresults}), the only difference is that F-measures for Hungarian are significantly higher, than ones for English.
+It can be due to the fact that the MISC category for Hungarian contains less types of names, thus the inconsistency and diversity of this class is smaller (see section \ref{sec:hun}).
+In contrast to the CoNLL corpus, the Szeged NER corpus was highly accurately annotated with an inter-annotator agreement over 99\%.
 
-Although two human-annotated NER corpora exist for Hungarian, we chose the Szeged NER corpus \cite{}, which follows the CoNLL annotation scheme. It contains ca.
-
-The strong correlation between capitalization and named entities makes recognition of entities in text relatively simple: a word is part of an entity if it is capitalized.
-
-training corpora in absence of gold standard datasets
-
-training sets for domains differing from newswire
+Due to the quite good F-measure of training on the Hungarian Wikipedia train corpus and testing on the corresponding test set, our Hungarian corpus can serve as a training corpus to build NE-taggers for domains falling out of newswire.
 
@@
 
-\bibitem[\protect\citename{Balasuriya \bgroup et al.\egroup}2009]{Balasuriya:09}
-Dominic Balasuriya, Nicky Ringland, Joel Nothman, Tara Murphy, James R. Curran
-\newblock 2009.
-\newblock Named Entity Recognition in Wikipedia.
-\newblock In: {\em Proceedings of the 2009 Workshop on the People's Web Meets NLP, ACL-IJCNLP}, pages 10--18.
+%\bibitem[\protect\citename{Balasuriya \bgroup et al.\egroup}2009]{Balasuriya:09}
+%Dominic Balasuriya, Nicky Ringland, Joel Nothman, Tara Murphy, James R. Curran
+%\newblock 2009.
+%\newblock Named Entity Recognition in Wikipedia.
+%\newblock In: {\em Proceedings of the 2009 Workshop on the People's Web Meets NLP, ACL-IJCNLP}, pages 10--18.
 

### 2012-03-25 paper/engNERwiki.tex
Got rid of some XXXs.

--- 2012-03-25 paper/engNERwiki.tex
@@
 Of these, 3,903,467 are English and 212,120 are Hungarian.
-XXX %kell ez ide?
 
@@
 
-Wikipedia's knowledge may be incorporated as features in NER learning, e.g.~Kazama and Torisawa \shortcite{Kazama:07} showed that automatic extraction of category labels from Wikipedia improves the accuracy of a supervised NE-tagger.
+The knowledge in Wikipedia may be incorporated as features in NER learning, e.g.~Kazama and Torisawa \shortcite{Kazama:07} showed that automatic extraction of category labels from Wikipedia improves the accuracy of a supervised NE-tagger.
 
@@
 Since our method is mainly XXX language-independent, it can be easily applied for languages which are structurally differing from English.
-Hungarian is a highly agglutinative language, with free word order and other particularities XXX detailed later in section \ref{sec:hun}.
+Hungarian is a highly agglutinative language, with free word order and other particularities detailed later in section \ref{sec:hun}.
 There are human annotated CoNLL-style datasets, but the one presented here is the first automatically NE-annotated corpus for Hungarian.
@@
 
-Our goal is to create a large NE-annotated corpus, automatically generated from Wikipedia articles. We follow a similar path XXX to \cite{Nothman:08}, who broke up XXX the process into four steps:
+Our goal is to create a large NE-annotated corpus, automatically generated from Wikipedia articles. We follow a similar path to \cite{Nothman:08}, who broke down the process into four steps:
 
@@
 In this section, we describe how these steps were implemented.
-While the approach in essence is language-independent, the exact details XXX have to be accomodated to the target language.
+While the approach in essence is language-independent, the exact details have to be accomodated to the target language.
 This section explains the general idea and its execution for English; the next section describes how the idea is adapted to Hungarian.
@@
 From the available NER annotation standards we elected to use the CoNLL \cite{Tjong:03} NE types.
-It is not difficult to see the parallels XXX between the DBpedia sub-hierarchies \texttt{Person}, \texttt{Organisation} and \texttt{Place} and the CoNLL NE categories PERSON, ORGANIZATION and LOCATION.
+It is not difficult to see the parallels between the DBpedia sub-hierarchies \texttt{Person}, \texttt{Organisation} and \texttt{Place} and the CoNLL NE categories PERSON, ORGANIZATION and LOCATION.
 The fourth category, MISCELLANEOUS is more elusive; according to the ConLL NER annotation guide\footnote{http://www.cnts.ua.ac.be/conll2003/ner/annotation.txt}, the sub-hierarchies \texttt{Event} and \texttt{Work} belong to this category, as well as various other classes outside the main hierarchies.
@@
 Therefore, only the second, manual approach ensures correct tagging.
-XXX
 

### 2012-03-25 paper/engNERwiki.tex
Pfff.

@@
 corpora is given for less resourced languages, as well.
-Since our method is mainly XXX language-independent, it can be easily applied for languages which are structurally differing from English.
+Since our method is largely language-independent, it can be easily applied for languages which are structurally differing from English.
 Hungarian is a highly agglutinative language, with free word order and other particularities detailed later in section \ref{sec:hun}.
@@
 The ontology contains the 320 most frequent entity categories, arranged into a taxonomy under the base class \texttt{owl:Thing}.
-Most of the classes belong into the 5 largest sub-hierarchies (here named after their topmost classes XXX): \texttt{Agent} (which is further divided into \texttt{Person} and \texttt{Organisation}), \texttt{Event}, \texttt{Place}, \texttt{Species} and \texttt{Work}.
+Most of the classes belong into the 5 largest sub-hierarchies (here named after their topmost classes): \texttt{Agent} (which is further divided into \texttt{Person} and \texttt{Organisation}), \texttt{Event}, \texttt{Place}, \texttt{Species} and \texttt{Work}.
 The taxonomy is rather flat: the top level contains 44 classes and there are several nodes with a branching factor of 20.
@@
 
-While the correspondence described above holds for most classes in the XXX sub-hierarchies, there are some exceptions.
+While the correspondence described above holds for most classes in the sub-hierarchies, there are some exceptions.
 For instance, the class \texttt{SportsLeague} is part of the \texttt{Organisation} sub-hierarchy, but according to the CoNLL annotation scheme, they are tagged as MISC.
@@
 We created a file of DBpedia class--NE category mappings.
-Whenever an entity is evaluated, we look up its class, as well as the ancestors of its class, in the mapping, and assign the category of the class that matches it most closely to the entity.
+Whenever an entity is evaluated, we look up its class, as well as the ancestors of its class, in the mapping, and assign to it the category of the class that matches it most closely.
 If no match is found, the entity is tagged with \texttt{O}.
@@
 
-As of version 3.7, the DBpedia ontology allows multiple superclasses, making the hierarchy a directed acyclic graph (CITE).
+As of version 3.7, the DBpedia ontology allows multiple superclasses, making the hierarchy a directed acyclic graph\footnote{http://blog.dbpedia.org/2011/09/11/dbpedia-37-released-including-15-localized-editions}.
 This introduces the problem of selecting the right superclass, and hence, CoNLL mapping for classes with more than one parent.
-In version 3.7, the only such class is \textit{Library}, which can be traced back to both \textit{Place} and \textit{Organisation}.
-There are two ways to tackle this problem.
-The first is to follow the DBpedia guidelines and only retain the first parent.
-The second method is manually deciding the right NE mapping for the ambiguous XXX class and adding it to the mapping file.
+In version 3.7, the only such class is \texttt{Library}, which can be traced back to both \texttt{Place} and \texttt{Organisation}.
 
-The first approach, unfortunately, already fails to assign the correct ConLL NE tag to our only example.
+DBpedia guidelines suggest that only the first parent is retained.
+Unfortunately, this approach already fails to assign the correct ConLL NE tag to our only example.
 The reason behind this failure is that multiple parents in an ontology show an inherent ambiguity in the class in question: it is entirely natural to think of a library as a location as well as a public organization.
-DBpedia's guidelines prefers Organisation, while ConLL's one refers it as a Location.
-Therefore, only the second, manual approach ensures correct tagging.
+DBpedia's guidelines prefer \texttt{Organisation}; ConLL's prefer LOCATION.
+Hence, we picked the right NE mapping manually and added it to the mapping file.
 
@@
 This gives rise to two potential problems.
-One is that compared to English, the list is relatively shorter: the entity/page ratio is 12.12\%, as opposed to the 37.66\% of the English Wikipedia.
+One is that compared to English, the list is relatively short: the entity/page ratio is 12.12\%, as opposed to the 37.66\% of the English Wikipedia.
 The other, since mostly Hungarian people, places and organizations are missing, a NER tagger that takes the surface forms of words into account might be mislead as to the language model of entity names.
@@
 
-There is another special case in Hungarian: opposite to English, the number of compund words is quite large in Hungarian, and NEs can also be subjects of compounding.
+There is another special case in Hungarian: opposite to English, the number of compund words is quite large, and NEs can also be subjects of compounding.
 In this case the common noun following the NE is joint with a hyphen, so they constitute one token.
@@
 
-The corpora are available under the Creative Commons Attribution-Sharealike 3.0 Unported License (CC-BY-SA), the same license under which the text of Wikipedia is released.
+The corpora are released under the Creative Commons Attribution-Sharealike 3.0 Unported License (CC-BY-SA), the same license under which the text of Wikipedia is released.
 The data files can be freely downloaded from ...\footnote{To be filled in in the camera-ready version.}.

### 2012-03-25 paper/engNERwiki.tex
Conclusion

--- 2012-03-25 paper/engNERwiki.tex
@@
 Of these, 3,903,467 are English and 212,120 are Hungarian.
+XXX %kell ez ide?
 
@@
 
-The knowledge in Wikipedia may be incorporated as features in NER learning, e.g.~Kazama and Torisawa \shortcite{Kazama:07} showed that automatic extraction of category labels from Wikipedia improves the accuracy of a supervised NE-tagger.
+Wikipedia's knowledge may be incorporated as features in NER learning, e.g.~Kazama and Torisawa \shortcite{Kazama:07} showed that automatic extraction of category labels from Wikipedia improves the accuracy of a supervised NE-tagger.
 
@@
 Since our method is mainly XXX language-independent, it can be easily applied for languages which are structurally differing from English.
-Hungarian is a highly agglutinative language, with free word order and other particularities detailed later in section \ref{sec:hun}.
+Hungarian is a highly agglutinative language, with free word order and other particularities XXX detailed later in section \ref{sec:hun}.
 There are human annotated CoNLL-style datasets, but the one presented here is the first automatically NE-annotated corpus for Hungarian.
@@
 
-Our goal is to create a large NE-annotated corpus, automatically generated from Wikipedia articles. We follow a similar path to \cite{Nothman:08}, who broke down the process into four steps:
+Our goal is to create a large NE-annotated corpus, automatically generated from Wikipedia articles. We follow a similar path XXX to \cite{Nothman:08}, who broke up XXX the process into four steps:
 
@@
 In this section, we describe how these steps were implemented.
-While the approach in essence is language-independent, the exact details have to be accomodated to the target language.
+While the approach in essence is language-independent, the exact details XXX have to be accomodated to the target language.
 This section explains the general idea and its execution for English; the next section describes how the idea is adapted to Hungarian.
@@
 From the available NER annotation standards we elected to use the CoNLL \cite{Tjong:03} NE types.
-It is not difficult to see the parallels between the DBpedia sub-hierarchies \texttt{Person}, \texttt{Organisation} and \texttt{Place} and the CoNLL NE categories PERSON, ORGANIZATION and LOCATION.
+It is not difficult to see the parallels XXX between the DBpedia sub-hierarchies \texttt{Person}, \texttt{Organisation} and \texttt{Place} and the CoNLL NE categories PERSON, ORGANIZATION and LOCATION.
 The fourth category, MISCELLANEOUS is more elusive; according to the ConLL NER annotation guide\footnote{http://www.cnts.ua.ac.be/conll2003/ner/annotation.txt}, the sub-hierarchies \texttt{Event} and \texttt{Work} belong to this category, as well as various other classes outside the main hierarchies.
@@
 Therefore, only the second, manual approach ensures correct tagging.
+XXX
 
-% The full mapping can be found in Appendix~\ref{sec:appa}.
-%
+%The full mapping can be found in Appendix~\ref{sec:appa}.
+
 Using the mapping described above, we compile an entity list that contains all entities in DBpedia tagged with the appropriate CoNLL named entity category.
@@
 
-\subsubsection{Sentence Filtering}
+\subsubsection{Filtering}
 
-Certain types of sentences were filtered from the corpus.
-As mentioned above, sentences with words tagged as \texttt{UNK} are discarded.
-Furthermore, there are many incomplete sentences in the Wikipedia text: image captions, enumerations items, contents of table cells, etc.
+There are many incomplete sentences in the Wikipedia text: image captions, enumerations items, contents of table cells, etc.
 On the one hand, these sentence fragments may prove to be of too low quality to be of any use in the traditional NER task.
-On the other hand, they could prove to be invaluable when training a NER tagger for User Generated Content, which is known to be noisy and fragmented.
-As a compromise solution, we included these fragments in the corpus, but labelled them as "low quality", so that users of the corpus can decide whether they want to use them or not.
+On the other hand, they could prove to be invaluable when training a NER tagger for User Generated Content, which is well known to be noisy and fragmented.
+We have added a post-filtering step to recognize and label these fragments, so that users of the corpus can decide what to do with them.
+XXX
+
+Manual inspection has found two conditions that can be used to detect fragments.
 A sentence is labelled as such if it either lacks a punctuation mark at the end, or it contains no finite verbs.
+Our final corpus contains XXXSTAT full sentences and STAT fragments.
+%Ezeket leméred, vagy mérjem én? Ha utóbbi, akkor melyikben kell mérni?
+% TODO: Last sentence to evaluation? But then this whole subsection is too short.
+
+In Wikipedia the nationalities (e.g.
+American and Australian) links to the corresponding places (America and Australia, respectively).
+However, in CoNLL scheme these types of entities take parts of MISC category.
+XXX Postprocessing: list look-up (nationalities, adjective forms of person who live in the corresponding country); if the name is POS-tagged as adjective, then it is not a NE
+
+%nem biztos, hogy ide kell
+Link boundaries need to be adjusted, e.g.~to remove excess punctuation.
 
@@
 
-az eredeti DBpedia osztályokat is megtarthatnánk, hiszen nem mindig célravezető a CoNLL osztályok használata, mivel azok elég tág kategóriák, amikbe olyan dolgok is tartoznak, amik nagyon eltérőek is lehetnek
+We have presented freely available NE-tagged corpora for English and Hungarian, fully automatically generated from Wikipedia.
+In contrast to the methods used so far for automatic annotation of NEs in Wikipedia texts, we applied a new approach, namely mapping DBpedia ontology classes to standard CoNLL NE tags, and assigning them to Wikipedia entities.
+Following Nothman \shortcite{Nothman:08} the process can be divided into four main steps: (1) classifying all articles into entity classes; (2) splitting Wikipedia articles into sentences; (3) labeling NEs according to link targets; and (4) selecting sentences for inclusion in the corpus.
 
-később majd kirakjuk a scripteket is, és akár újra és újra frissíthető lesz
+Thanks for the huge amount of Wikipedia articles, the possibility of building large enough corpora is given for less resourced languages, like Hungarian, as well.
+Due to the particularities of the Hungarian language, some steps are slightly different, and special linguistics phenomena pose several problems related to the NER task to solve.
+
+Automatically generated, so called "silver standard corpora" can be very useful for improving NER in more ways.
+We showed that huge gazetteer lists extracted from our corpora, and training with extra features given by the model trained on our corpora, improves overall F-measure.
+Moreover, our Hungarian corpus can serve as a training corpus to build NE-taggers for more general domains, than the standard newswire.
+
+%az eredeti DBpedia osztályokat is megtarthatnánk, hiszen nem mindig célravezető a CoNLL osztályok használata, mivel azok elég tág kategóriák, amikbe olyan dolgok is tartoznak, amik nagyon eltérőek is lehetnek
+
+%később majd kirakjuk a scripteket is, és akár újra és újra frissíthető lesz
 

### 2012-03-25 paper/engNERwiki.tex
E.

--- 2012-03-25 paper/engNERwiki.tex
@@
 
-The paper is structured as follows. XXX In section~\ref{sec:hun}, we show how the algorithm was adapted to Hungarian.
+The paper is structured as follows.
+In section~\ref{sec:related} we overview related works.
+Section~\ret{sec:create} introduces our XXX algorithm; section~\ref{sec:hun}, shows how it was adapted to Hungarian.
+The corpus format is described in section~\ref{sec:data}.
+Section~\ref{sec:eval} contains the evaluation and section~\ref{sec:conclusion} our conclusions.
 
@@
 \section{Data Description}
+\label{sec:data}
 
@@
 \section{Evaluation}
+\label{sec:eval}
 
@@
 \section{Conclusions and Future Work}
+\label{sec:conclusion}
 

### 2012-03-25 paper/engNERwiki.tex
Paper structure.

--- 2012-03-25 paper/engNERwiki.tex
@@
 
-The paper is structured as follows. XXX In section~\ref{sec:hun}, we show how the algorithm was adapted to Hungarian.
+The paper is structured as follows.
+In Section~\ref{sec:related} we give an overview of related works.
+We introduce our algorithm in Section~\ref{sec:create}, and show how it was adapted to Hungarian in Section~\ref{sec:hun}.
+The corpus format is described in Section~\ref{sec:data}.
+Section~\ref{sec:eval} contains evaluations and Section~\ref{sec:conclusion} our conclusions.
 

### 2012-03-25 paper/engNERwiki.tex
abstract elkezdve

--- 2012-03-25 paper/engNERwiki.tex
@@
 
+Supervised Named Entity Recogizers require
+
 \end{abstract}
@@
 
-The paper is structured as follows.
-In section~\ref{sec:related} we overview related works.
-Section~\ret{sec:create} introduces our XXX algorithm; section~\ref{sec:hun}, shows how it was adapted to Hungarian.
-The corpus format is described in section~\ref{sec:data}.
-Section~\ref{sec:eval} contains the evaluation and section~\ref{sec:conclusion} our conclusions.
+The paper is structured as follows. XXX In section~\ref{sec:hun}, we show how the algorithm was adapted to Hungarian.
 
@@
 corpora is given for less resourced languages, as well.
-Since our method is largely language-independent, it can be easily applied for languages which are structurally differing from English.
-Hungarian is a highly agglutinative language, with free word order and other particularities detailed later in section \ref{sec:hun}.
+Since our method is mainly XXX language-independent, it can be easily applied for languages which are structurally differing from English.
+Hungarian is a highly agglutinative language, with free word order and other particularities XXX detailed later in section \ref{sec:hun}.
 There are human annotated CoNLL-style datasets, but the one presented here is the first automatically NE-annotated corpus for Hungarian.
@@
 The ontology contains the 320 most frequent entity categories, arranged into a taxonomy under the base class \texttt{owl:Thing}.
-Most of the classes belong into the 5 largest sub-hierarchies (here named after their topmost classes): \texttt{Agent} (which is further divided into \texttt{Person} and \texttt{Organisation}), \texttt{Event}, \texttt{Place}, \texttt{Species} and \texttt{Work}.
+Most of the classes belong into the 5 largest sub-hierarchies (here named after their topmost classes XXX): \texttt{Agent} (which is further divided into \texttt{Person} and \texttt{Organisation}), \texttt{Event}, \texttt{Place}, \texttt{Species} and \texttt{Work}.
 The taxonomy is rather flat: the top level contains 44 classes and there are several nodes with a branching factor of 20.
@@
 From the available NER annotation standards we elected to use the CoNLL \cite{Tjong:03} NE types.
-It is not difficult to see the parallels between the DBpedia sub-hierarchies \texttt{Person}, \texttt{Organisation} and \texttt{Place} and the CoNLL NE categories PERSON, ORGANIZATION and LOCATION.
+It is not difficult to see the parallels XXX between the DBpedia sub-hierarchies \texttt{Person}, \texttt{Organisation} and \texttt{Place} and the CoNLL NE categories PERSON, ORGANIZATION and LOCATION.
 The fourth category, MISCELLANEOUS is more elusive; according to the ConLL NER annotation guide\footnote{http://www.cnts.ua.ac.be/conll2003/ner/annotation.txt}, the sub-hierarchies \texttt{Event} and \texttt{Work} belong to this category, as well as various other classes outside the main hierarchies.
 
-While the correspondence described above holds for most classes in the sub-hierarchies, there are some exceptions.
+While the correspondence described above holds for most classes in the XXX sub-hierarchies, there are some exceptions.
 For instance, the class \texttt{SportsLeague} is part of the \texttt{Organisation} sub-hierarchy, but according to the CoNLL annotation scheme, they are tagged as MISC.
@@
 We created a file of DBpedia class--NE category mappings.
-Whenever an entity is evaluated, we look up its class, as well as the ancestors of its class, in the mapping, and assign to it the category of the class that matches it most closely.
+Whenever an entity is evaluated, we look up its class, as well as the ancestors of its class, in the mapping, and assign the category of the class that matches it most closely to the entity.
 If no match is found, the entity is tagged with \texttt{O}.
@@
 
-As of version 3.7, the DBpedia ontology allows multiple superclasses, making the hierarchy a directed acyclic graph\footnote{http://blog.dbpedia.org/2011/09/11/dbpedia-37-released-including-15-localized-editions}.
+As of version 3.7, the DBpedia ontology allows multiple superclasses, making the hierarchy a directed acyclic graph (CITE).
 This introduces the problem of selecting the right superclass, and hence, CoNLL mapping for classes with more than one parent.
-In version 3.7, the only such class is \texttt{Library}, which can be traced back to both \texttt{Place} and \texttt{Organisation}.
+In version 3.7, the only such class is \textit{Library}, which can be traced back to both \textit{Place} and \textit{Organisation}.
+There are two ways to tackle this problem.
+The first is to follow the DBpedia guidelines and only retain the first parent.
+The second method is manually deciding the right NE mapping for the ambiguous XXX class and adding it to the mapping file.
 
-DBpedia guidelines suggest that only the first parent is retained.
-Unfortunately, this approach already fails to assign the correct ConLL NE tag to our only example.
+The first approach, unfortunately, already fails to assign the correct ConLL NE tag to our only example.
 The reason behind this failure is that multiple parents in an ontology show an inherent ambiguity in the class in question: it is entirely natural to think of a library as a location as well as a public organization.
-DBpedia's guidelines prefer \texttt{Organisation}; ConLL's prefer LOCATION.
-Hence, we picked the right NE mapping manually and added it to the mapping file.
+DBpedia's guidelines prefers Organisation, while ConLL's one refers it as a Location.
+Therefore, only the second, manual approach ensures correct tagging.
+XXX
 
-% The full mapping can be found in Appendix~\ref{sec:appa}.
-%
+%The full mapping can be found in Appendix~\ref{sec:appa}.
+
 Using the mapping described above, we compile an entity list that contains all entities in DBpedia tagged with the appropriate CoNLL named entity category.
@@
 This gives rise to two potential problems.
-One is that compared to English, the list is relatively short: the entity/page ratio is 12.12\%, as opposed to the 37.66\% of the English Wikipedia.
+One is that compared to English, the list is relatively shorter: the entity/page ratio is 12.12\%, as opposed to the 37.66\% of the English Wikipedia.
 The other, since mostly Hungarian people, places and organizations are missing, a NER tagger that takes the surface forms of words into account might be mislead as to the language model of entity names.
@@
 
-There is another special case in Hungarian: opposite to English, the number of compund words is quite large, and NEs can also be subjects of compounding.
+There is another special case in Hungarian: opposite to English, the number of compund words is quite large in Hungarian, and NEs can also be subjects of compounding.
 In this case the common noun following the NE is joint with a hyphen, so they constitute one token.
@@
 \section{Data Description}
-\label{sec:data}
 
-The corpora are released under the Creative Commons Attribution-Sharealike 3.0 Unported License (CC-BY-SA), the same license under which the text of Wikipedia is released.
+The corpora are available under the Creative Commons Attribution-Sharealike 3.0 Unported License (CC-BY-SA), the same license under which the text of Wikipedia is released.
 The data files can be freely downloaded from ...\footnote{To be filled in in the camera-ready version.}.
@@
 \section{Evaluation}
-\label{sec:eval}
 
@@
 \section{Conclusions and Future Work}
-\label{sec:conclusion}
 

### 2012-03-25 paper/engNERwiki.tex
abstract

@@
 
-Supervised Named Entity Recogizers require
+Supervised Named Entity Recogizers require large amount of annotated texts.
+However, manual annotation is a highly costly procedure, thus reducing the annotation cost is essential.
+We present a fully automatic method to build NE-annotated corpora from Wikipedia.
+In contrast to recent work, we apply a new method, which maps the DBpedia classes into CoNLL NE types.
+Since our method is mainly language-independent, we used it for generating corpora for English and Hungarian, as well.
+The corpora are freely available.
 

### 2012-03-25 paper/engNERwiki.tex
Some fixes + deleted a paragraph.

--- 2012-03-25 paper/engNERwiki.tex
@@
 
-Supervised Named Entity Recogizers require
-
 \end{abstract}
@@
 One approach to reduce the manual intervention required in language resource production, and thus ultimately production costs is automatically generating resources.
-An other approach to achieve this aim is using collaborative annotation and/or collaboratively constructed resources, such as Wikipedia, Wiktionary or Linked Open Data.
+Another approach to achieve this aim is using collaborative annotation and/or collaboratively constructed resources, such as Wikipedia, Wiktionary or Linked Open Data.
 Researchers apply the Web-as-corpus approach, often dealing with such datasets as Google's Web 1T 5-gram collection or DBpedia.
@@
 
-Wikipedia is recognized as one of the largest available collection of entities, and also as an external knowledge to improve the accuracy of NER.
+Wikipedia is recognized as one of the largest available collection of entities, and also as a resource that can improve the accuracy of NER.
 The most obvious utilization of Wikipedia for NER is extracting gazetteers containing person names, locations or organizations (e.g.~Toral and Mu\~noz \shortcite{Toral:06}).
@@
 However, they followed a different way: the NEs were tagged by high quality open source softwares.
-Our corpora is available under the same license as Wikipedia's one (Creative Commons Attribution-Sharealike 3.0 Unported License).
+In contrast, our approach relies on the manual judgement of thousands of Wikipedia volunteers.
 
 Thanks for the huge amount of Wikipedia articles, the possibility of building large enough 
-%(XXX ez mit jelent?) elég nagy :)
 corpora is given for less resourced languages, as well.
-Since our method is mainly XXX language-independent, it can be easily applied for languages which are structurally differing from English.
-Hungarian is a highly agglutinative language, with free word order and other particularities XXX detailed later in section \ref{sec:hun}.
+Since our method is largely language-independent, it can be easily applied for languages which are structurally differing from English.
+Hungarian is a highly agglutinative language, with free word order and other particularities detailed later in section \ref{sec:hun}.
 There are human annotated CoNLL-style datasets, but the one presented here is the first automatically NE-annotated corpus for Hungarian.
@@
 In addition to in-article links, our method takes advantage of the redirect and interlanguage links, available as SQL dumps.
-
-Beside regular articles, Wikipedia also contains redirect and disambiguation pages, which require special treatment.
-Redirect pages are empty, and are safe to ignore altogether.
-Disambiguation pages contain text, usually in the form of an enumeration of page links with a short description.
-Full sentences are scarce, and the structure of these pages makes it hard to extract valid training sentences from them.
-Therefore, disambiguation pages are dropped from the corpus as well.
 
@@
 From the available NER annotation standards we elected to use the CoNLL \cite{Tjong:03} NE types.
-It is not difficult to see the parallels XXX between the DBpedia sub-hierarchies \texttt{Person}, \texttt{Organisation} and \texttt{Place} and the CoNLL NE categories PERSON, ORGANIZATION and LOCATION.
+It is not difficult to see the parallels between the DBpedia sub-hierarchies \texttt{Person}, \texttt{Organisation} and \texttt{Place} and the CoNLL NE categories PERSON, ORGANIZATION and LOCATION.
 The fourth category, MISCELLANEOUS is more elusive; according to the ConLL NER annotation guide\footnote{http://www.cnts.ua.ac.be/conll2003/ner/annotation.txt}, the sub-hierarchies \texttt{Event} and \texttt{Work} belong to this category, as well as various other classes outside the main hierarchies.
 
-While the correspondence described above holds for most classes in the XXX sub-hierarchies, there are some exceptions.
+While the correspondence described above holds for most classes in the sub-hierarchies, there are some exceptions.
 For instance, the class \texttt{SportsLeague} is part of the \texttt{Organisation} sub-hierarchy, but according to the CoNLL annotation scheme, they are tagged as MISC.
@@
 
-As of version 3.7, the DBpedia ontology allows multiple superclasses, making the hierarchy a directed acyclic graph (CITE).
+As of version 3.7, the DBpedia ontology allows multiple superclasses, making the hierarchy a directed acyclic graph\footnote{http://blog.dbpedia.org/2011/09/11/dbpedia-37-released-including-15-localized-editions}.
 This introduces the problem of selecting the right superclass, and hence, CoNLL mapping for classes with more than one parent.
-In version 3.7, the only such class is \textit{Library}, which can be traced back to both \textit{Place} and \textit{Organisation}.
-There are two ways to tackle this problem.
-The first is to follow the DBpedia guidelines and only retain the first parent.
-The second method is manually deciding the right NE mapping for the ambiguous XXX class and adding it to the mapping file.
+In version 3.7, the only such class is \texttt{Library}, which can be traced back to both \texttt{Place} and \texttt{Organisation}.
 
-The first approach, unfortunately, already fails to assign the correct ConLL NE tag to our only example.
+DBpedia guidelines suggest that only the first parent is retained.
+Unfortunately, this approach already fails to assign the correct ConLL NE tag to our only example.
 The reason behind this failure is that multiple parents in an ontology show an inherent ambiguity in the class in question: it is entirely natural to think of a library as a location as well as a public organization.
-DBpedia's guidelines prefers Organisation, while ConLL's one refers it as a Location.
-Therefore, only the second, manual approach ensures correct tagging.
-XXX
+DBpedia's guidelines prefer \texttt{Organisation}; ConLL's prefer LOCATION.
+Hence, we picked the right NE mapping manually and added it to the mapping file.
 
-%The full mapping can be found in Appendix~\ref{sec:appa}.
-
+% The full mapping can be found in Appendix~\ref{sec:appa}.
+%
 Using the mapping described above, we compile an entity list that contains all entities in DBpedia tagged with the appropriate CoNLL named entity category.

### 2012-03-25 paper/engNERwiki.tex
tablazatok felmozgatva
Re-modified the Filtering section.

@@
 
-There are many incomplete sentences in the Wikipedia text: image captions, enumerations items, contents of table cells, etc.
+Certain types of sentences were filtered from the corpus.
+As mentioned above, sentences with words tagged as \texttt{UNK} are discarded.
+Furthermore, there are many incomplete sentences in the Wikipedia text: image captions, enumerations items, contents of table cells, etc.
 On the one hand, these sentence fragments may prove to be of too low quality to be of any use in the traditional NER task.
-On the other hand, they could prove to be invaluable when training a NER tagger for User Generated Content, which is well known to be noisy and fragmented.
-We have added a post-filtering step to recognize and label these fragments, so that users of the corpus can decide what to do with them.
-XXX
-
-Manual inspection has found two conditions that can be used to detect fragments.
+On the other hand, they could prove to be invaluable when training a NER tagger for User Generated Content, which is known to be noisy and fragmented.
+As a compromise solution, we included these fragments in the corpus, but labelled them as "low quality", so that users of the corpus can decide whether they want to use them or not.
 A sentence is labelled as such if it either lacks a punctuation mark at the end, or it contains no finite verbs.
-Our final corpus contains XXXSTAT full sentences and STAT fragments.
-%Ezeket leméred, vagy mérjem én? Ha utóbbi, akkor melyikben kell mérni?
-% TODO: Last sentence to evaluation? But then this whole subsection is too short.
-
-In Wikipedia the nationalities (e.g.
-American and Australian) links to the corresponding places (America and Australia, respectively).
-However, in CoNLL scheme these types of entities take parts of MISC category.
-XXX Postprocessing: list look-up (nationalities, adjective forms of person who live in the corresponding country); if the name is POS-tagged as adjective, then it is not a NE
-
-%nem biztos, hogy ide kell
-Link boundaries need to be adjusted, e.g.~to remove excess punctuation.
 

### 2012-03-25 paper/engNERwiki.tex
Fixed the botched merging.

--- 2012-03-25 paper/engNERwiki.tex
@@
 \begin{abstract}
+
+Supervised Named Entity Recogizers require large amount of annotated texts.
+However, manual annotation is a highly costly procedure, thus reducing the annotation cost is essential.
+We present a fully automatic method to build NE-annotated corpora from Wikipedia.
+In contrast to recent work, we apply a new method, which maps the DBpedia classes into CoNLL NE types.
+Since our method is mainly language-independent, we used it for generating corpora for English and Hungarian, as well.
+The corpora are freely available.
 
@@
 This introduces the problem of selecting the right superclass, and hence, CoNLL mapping for classes with more than one parent.
-In version 3.7, the only such class is \texttt{Library}, which can be traced back to both \texttt{Place} and \texttt{Organisation}.
+In version 3.7, the only such class is \textit{Library}, which can be traced back to both \textit{Place} and \textit{Organisation}.
 
@@
 
-% The full mapping can be found in Appendix~\ref{sec:appa}.
-%
+%The full mapping can be found in Appendix~\ref{sec:appa}.
+
 Using the mapping described above, we compile an entity list that contains all entities in DBpedia tagged with the appropriate CoNLL named entity category.
@@
 
-There is another special case in Hungarian: opposite to English, the number of compund words is quite large, and NEs can also be subjects of compounding.
+There is another special case in Hungarian: opposite to English, the number of compund words is quite large and NEs can also be subjects of compounding.
 In this case the common noun following the NE is joint with a hyphen, so they constitute one token.
@@
 
-The corpora are released under the Creative Commons Attribution-Sharealike 3.0 Unported License (CC-BY-SA), the same license under which the text of Wikipedia is released.
+The corpora are available under the Creative Commons Attribution-Sharealike 3.0 Unported License (CC-BY-SA), the same license under which the text of Wikipedia is released.
 The data files can be freely downloaded from ...\footnote{To be filled in in the camera-ready version.}.
 
-The corpora are available in multitag format.
+The files are in multitag format.
 Content lines are tab separated; there is one column for the tokens plus one column per tagset.
@@
 
+\begin{table*}[ht]
+\begin{center}
+\begin{tabular}{lllllll}
+\hline  & \bf enwiki & \bf enwiki filtered & \bf CoNLL & \bf huwiki & \bf huwiki filtered  & \bf Szeged NER \\ \hline
+token & 60,520,819 & 21,718,854 & 302,811 &  19,108,027 & 3,512,249  & 225,963\\
+NE & 3,169,863 & 3,169,863 & 50,758 & 456,281 & 456,281  & 25,896\\
+NE density & 5.23\% & 14.59\% & 16.76\% & 2.38\% & 12.99\%  & 11.46\%\\
+\hline
+\end{tabular}
+\end{center}
+\caption{\label{size} Corpus size and NE-density.}
+\end{table*}
+
 \section{Evaluation}
 \label{sec:eval}
-
-% Corpus sizes, performance, tagging conventions
-% Hun -> not utf-8, loss of precision?
-%measuring Wikipedia's inconsistency using n-gram tag variation (\cite{Nothman:09})
-%ez jó mérőszám lehet a gold standard korpuszok összehasonlítására
-%használhatjuk fordítva is: ha egy adott nucleus ugyanolyan n-gram környezetben nagyon sokszor van X-ként taggelve, és kevésszer Y-ként, akkor az Y-okat is átcsinálhatnánk X-szé
 
@@
 However, after further investigation of the NE density in our Wikipedia corpora and gold standard corpora (see table \ref{size}), we decided not to include the sentences without NEs in evaluation datasets.
-
-\begin{table*}[ht]
-\begin{center}
-\begin{tabular}{lllllll}
-\hline  & \bf enwiki & \bf enwiki filtered & \bf CoNLL & \bf huwiki & \bf huwiki filtered  & \bf Szeged NER \\ \hline
-token & 60,520,819 & 21,718,854 & 302,811 &  19,108,027 & 3,512,249  & 225,963\\
-NE & 3,169,863 & 3,169,863 & 50,758 & 456,281 & 456,281  & 25,896\\
-NE density & 5.23\% & 14.59\% & 16.76\% & 2.38\% & 12.99\%  & 11.46\%\\
-\hline
-\end{tabular}
-\end{center}
-\caption{\label{size} Corpus size and NE-density.}
-\end{table*}
 

### 2012-03-25 paper/engNERwiki.tex
blabla

--- 2012-03-25 paper/engNERwiki.tex
@@
 
-\title{XXX}
+\title{Automatically generated NE tagged corpora for English and Hungarian}
 
@@
 However, manual annotation is a highly costly procedure, thus reducing the annotation cost is essential.
-We present a fully automatic method to build NE-annotated corpora from Wikipedia.
+We present a fully automatic method to build NE annotated corpora from Wikipedia.
 In contrast to recent work, we apply a new method, which maps the DBpedia classes into CoNLL NE types.
@@
 
-Named Entity Recognition (NER), the task of identifying Named Entities (NEs) in unstructured texts and classifying them into pre-selected classes, is one of the most important subtasks in many NLP tasks, such as Information Retrieval, Information Extraction or Machine Translation.
+Named Entity Recognition (NER), the task of identifying Named Entities (NEs) in unstructured texts and classifying them into pre-selected classes, is one of the most important subtasks in many NLP tasks, such as information retrieval, information extraction or machine translation.
 The NER task was introduced with the 6th Message Understanding Conference (MUC) in 1995 \cite{Grishman:96}.
-In MUC shared tasks the NER consists of three subtasks: entity names, temporal expressions and number expressions.
-Although there is a general agreement in the NER community about the inclusion of temporal expressions and some numerical expressions such as amounts of money and other types of units, the most studied types are names of persons, locations and organizations.
+In MUC shared tasks the NER consists of three subtasks: entity names, temporal and number expressions.
+Although there is a general agreement in the NER community about the inclusion of temporal expressions and some numerical expressions, the most studied types are names of persons, locations and organizations.
 The fourth type, called "miscellaneous'', was introduced in the CoNLL NER tasks in 2002 \cite{Tjong:02} and 2003 \cite{Tjong:03}, and includes proper names falling outside the three classic types.
@@
 
-The standard datasets are highly domain-specific (mostly newspaper articles about company activities and defense related activities), and are restricted in size, as well.
-Thus they are not large enough to train and test robust NE-taggers on them.
-Moreover, merging these datasets to get a bigger training corpus faces NLP researchers with the problem of combining different tagsets and annotation schemes.
+The standard datasets are highly domain-specific (mostly newswire), and are restricted in size, as well. 
+%Thus they are not large enough to train and test robust NE-taggers on them. Moreover, 
+Merging these datasets to get a bigger training corpus faces NLP researchers with the problem of combining different tagsets and annotation schemes.
 However, to build a robust supervised machine learning system for NER, large enough, accurately annotated corpora are essential.
@@
 
-%One of the main trends in NLP is the need for reducing the annotation cost.
-After examining the calls for papers of nowadays conferences and workshops there seems to be three major directions to achieve this aim.
-
 %The first approach is the application of unsupervised and semi-supervised learning techniques.
@@
 In this paper we present a kind of combination of the approaches mentioned above.
-We built freely available NE-tagged corpora for English and Hungarian -- fully automatically generated from Wikipedia.
+We built freely available NE tagged corpora for English and Hungarian -- fully automatically generated from Wikipedia.
 %Our goal was to create gold standard corpora which can serve as train, devel and test datasets for automatic NE-taggers.
@@
 The paper is structured as follows.
-In Section~\ref{sec:related} we give an overview of related works.
-We introduce our algorithm in Section~\ref{sec:create}, and show how it was adapted to Hungarian in Section~\ref{sec:hun}.
-The corpus format is described in Section~\ref{sec:data}.
-Section~\ref{sec:eval} contains evaluations and Section~\ref{sec:conclusion} our conclusions.
+In section~\ref{sec:related} we give an overview of related work.
+Our method is introduced in section~\ref{sec:create}, and showed how it is applied to Hungarian in section~\ref{sec:hun}.
+The corpus format is described in section~\ref{sec:data}.
+Section~\ref{sec:eval} contains evaluations and section~\ref{sec:conclusion} our conclusion.
 
@@
 It is a goldmine of information; at the time of writing, Wikipedia contains about 21 million interlinked articles.
-Of these, 3,903,467 are English and 212,120 are Hungarian.
+Of these, 3,903,467 are English, and 212,120 are Hungarian.
 XXX %kell ez ide?
 
-Wikipedia has been applied for several NLP tasks, e.g.~word sense disambiguation, ontology and thesaurus building, multilingual information retrieval, and question answering, as Medelyan et al.~\shortcite{Medelyan:09} point out in their fully detailed survey about mining meaning from Wikipedia.
+Wikipedia has been applied for several NLP tasks, e.g.~word sense disambiguation, ontology and thesaurus building, and question answering, as Medelyan et al.~\shortcite{Medelyan:09} point out in their fully detailed survey about mining meaning from Wikipedia.
 
@@
 They used the inherent category structure of Wikipedia to determine the NE-type of a proposed entity.
-Nothman et al.~\shortcite{Nothman:08} used a similar method to create a NE-annotated text in English.
+Nothman et al.~\shortcite{Nothman:08} used a similar method to create a NE annotated text in English.
 They transformed the Wikipedia links into NE-annotations by classifying the target articles into standard entity classes.
@@
 Hungarian is a highly agglutinative language, with free word order and other particularities detailed later in section \ref{sec:hun}.
-There are human annotated CoNLL-style datasets, but the one presented here is the first automatically NE-annotated corpus for Hungarian.
+There are human annotated CoNLL-style datasets, but the one presented here is the first automatically NE annotated corpus for Hungarian.
 
@@
 
-Our goal is to create a large NE-annotated corpus, automatically generated from Wikipedia articles. We follow a similar path XXX to \cite{Nothman:08}, who broke up XXX the process into four steps:
+Our goal is to create a large NE annotated corpus, automatically generated from Wikipedia articles. We follow a similar path XXX to \cite{Nothman:08}, who broke up XXX the process into four steps:
 
@@
 
-\section{Conclusions and Future Work}
+\section{Conclusion}
 \label{sec:conclusion}
 
-We have presented freely available NE-tagged corpora for English and Hungarian, fully automatically generated from Wikipedia.
+We have presented freely available NE tagged corpora for English and Hungarian, fully automatically generated from Wikipedia.
 In contrast to the methods used so far for automatic annotation of NEs in Wikipedia texts, we applied a new approach, namely mapping DBpedia ontology classes to standard CoNLL NE tags, and assigning them to Wikipedia entities.
@@
 We showed that huge gazetteer lists extracted from our corpora, and training with extra features given by the model trained on our corpora, improves overall F-measure.
-Moreover, our Hungarian corpus can serve as a training corpus to build NE-taggers for more general domains, than the standard newswire.
+Moreover, our Hungarian corpus can serve as a training corpus to build NE-taggers for more general domains, than the classic newswire.
 

### 2012-03-25 paper/engNERwiki.tex

@@
 
-Our goal is to create a large NE annotated corpus, automatically generated from Wikipedia articles. We follow a similar path XXX to \cite{Nothman:08}, who broke up XXX the process into four steps:
+Our goal is to create a large NE-annotated corpus, automatically generated from Wikipedia articles. We follow a similar path to \cite{Nothman:08}, who broke down the process into four steps:
 
@@
 
-As of version 3.7, the DBpedia ontology allows multiple superclasses, making the hierarchy a directed acyclic graph (CITE).
+As of version 3.7, the DBpedia ontology allows multiple superclasses, making the hierarchy a directed acyclic graph\footnote{http://blog.dbpedia.org/2011/09/11/dbpedia-37-released-including-15-localized-editions}.
 This introduces the problem of selecting the right superclass, and hence, CoNLL mapping for classes with more than one parent.
 In version 3.7, the only such class is \textit{Library}, which can be traced back to both \textit{Place} and \textit{Organisation}.
-There are two ways to tackle this problem.
-The first is to follow the DBpedia guidelines and only retain the first parent.
-The second method is manually deciding the right NE mapping for the ambiguous XXX class and adding it to the mapping file.
 
-The first approach, unfortunately, already fails to assign the correct ConLL NE tag to our only example.
+DBpedia guidelines suggest that only the first parent is retained.
+Unfortunately, this approach already fails to assign the correct ConLL NE tag to our only example.
 The reason behind this failure is that multiple parents in an ontology show an inherent ambiguity in the class in question: it is entirely natural to think of a library as a location as well as a public organization.
-DBpedia's guidelines prefers Organisation, while ConLL's one refers it as a Location.
-Therefore, only the second, manual approach ensures correct tagging.
-XXX
+DBpedia's guidelines prefer \texttt{Organisation}; ConLL's prefer LOCATION.
+Hence, we picked the right NE mapping manually and added it to the mapping file.
 
@@
 
-\subsubsection{Filtering}
+\subsubsection{Sentence Filtering}
 
@@
 
-There is another special case in Hungarian: opposite to English, the number of compund words is quite large in Hungarian, and NEs can also be subjects of compounding.
+There is another special case in Hungarian: opposite to English, the number of compund words is quite large and NEs can also be subjects of compounding.
 In this case the common noun following the NE is joint with a hyphen, so they constitute one token.
@@
 
-The corpora are available in multitag format.
+The files are in multitag format.
 Content lines are tab separated; there is one column for the tokens plus one column per tagset.

### 2012-03-25 paper/engNERwiki.tex
huzas

@@
 The paper is structured as follows.
-In section~\ref{sec:related} we give an overview of related work.
-Our method is introduced in section~\ref{sec:create}, and showed how it is applied to Hungarian in section~\ref{sec:hun}.
-The corpus format is described in section~\ref{sec:data}.
-Section~\ref{sec:eval} contains evaluations and section~\ref{sec:conclusion} our conclusion.
+In Section~\ref{sec:related} we give an overview of related work.
+Our method is introduced in Section~\ref{sec:create}, and showed how it is applied to Hungarian in Section~\ref{sec:hun}.
+The corpus format is described in Section~\ref{sec:data}.
+Section~\ref{sec:eval} contains evaluations and Section~\ref{sec:conclusion} our conclusion.
 
@@
 Of these, 3,903,467 are English, and 212,120 are Hungarian.
-XXX %kell ez ide?
 
@@
 The ontology contains the 320 most frequent entity categories, arranged into a taxonomy under the base class \texttt{owl:Thing}.
-Most of the classes belong into the 5 largest sub-hierarchies (here named after their topmost classes): \texttt{Agent} (which is further divided into \texttt{Person} and \texttt{Organisation}), \texttt{Event}, \texttt{Place}, \texttt{Species} and \texttt{Work}.
+Most of the classes belong into the 6 largest sub-hierarchies (here named after their topmost classes): \texttt{Person}, \texttt{Organisation}), \texttt{Event}, \texttt{Place}, \texttt{Species} and \texttt{Work}.
 The taxonomy is rather flat: the top level contains 44 classes and there are several nodes with a branching factor of 20.
@@
 It is not difficult to see the parallels between the DBpedia sub-hierarchies \texttt{Person}, \texttt{Organisation} and \texttt{Place} and the CoNLL NE categories PERSON, ORGANIZATION and LOCATION.
-The fourth category, MISCELLANEOUS is more elusive; according to the ConLL NER annotation guide\footnote{http://www.cnts.ua.ac.be/conll2003/ner/annotation.txt}, the sub-hierarchies \texttt{Event} and \texttt{Work} belong to this category, as well as various other classes outside the main hierarchies.
+The fourth category, MISCELLANEOUS is more elusive; according to the CoNLL NER annotation guide\footnote{http://www.cnts.ua.ac.be/conll2003/ner/annotation.txt}, the sub-hierarchies \texttt{Event} and \texttt{Work} belong to this category, as well as various other classes outside the main hierarchies.
 
@@
 For instance, the class \texttt{SportsLeague} is part of the \texttt{Organisation} sub-hierarchy, but according to the CoNLL annotation scheme, they are tagged as MISC.
-Similarly, the class \texttt{SportsTeamSeason} is under \texttt{Event}, by which account it would be classified as MISC; however, CoNLL does not recognize such entities.
 
@@
 In version 3.7, the only such class is \textit{Library}, which can be traced back to both \textit{Place} and \textit{Organisation}.
+Hence, we picked the right NE mapping manually, and added it to the mapping file.
 
-DBpedia guidelines suggest that only the first parent is retained.
-Unfortunately, this approach already fails to assign the correct ConLL NE tag to our only example.
+%DBpedia guidelines suggest that only the first parent is retained.
+Unfortunately, this approach already fails to assign the correct CoNLL NE tag to our only example.
 The reason behind this failure is that multiple parents in an ontology show an inherent ambiguity in the class in question: it is entirely natural to think of a library as a location as well as a public organization.
-DBpedia's guidelines prefer \texttt{Organisation}; ConLL's prefer LOCATION.
+DBpedia's guidelines prefer \texttt{Organisation}; CoNLL's prefer LOCATION.
 Hence, we picked the right NE mapping manually and added it to the mapping file.
@@
 The procedure described in the previous section was used to generate the Hungarian corpus, as well.
-However, linguistic diversity pose several problems even in the NER task.
-In this section we describe the differences between the two languages related to labeling NEs, and the changes they prompted in the algorithm.
+However, linguistic diversity pose several problems.
+In this section we describe the differences between the two languages related to labeling NEs, and the changes they prompted in the method.
 
@@
 For example: in the compound \textit{SAP-szoftvercég} ['SAP software company'] the common noun does not change the labeling, but in the case of the compound \textit{WorldCom-botrány} ['WorldCom scandal'] the NE tag changes from ORG to O.
+The solution of this problem is not obvious, and needs more investigation.
 
-Inflections of acronyms and foreign names ending with non-pronounced vowel have similar surface forms to the aforementioned compounds, e.g.~\textit{MTI-t} ['MTI'+ACC], \textit{Shakespeare-rel} ['with Shakespeare'].
+%Inflections of acronyms and foreign names ending with non-pronounced vowel have similar surface forms to the aforementioned compounds, e.g.~\textit{MTI-t} ['MTI'+ACC], \textit{Shakespeare-rel} ['with Shakespeare'].
 It is important to distinguish these types of hyphenated NEs, because inflections do not change NE labeling in contrast to some type of compounds.
 Due to the surface similarity of these phenomena the solution of this problem is not obvious, and needs more investigation.
-
-% Ezt hová?
-%most van: 102 millió szó a disambiguation pages-zel együtt
-%az angol: 1,7 milliárd
 
@@
 To evaluate our corpora we used a maximum entropy NE-tagger \cite{Varga:07}, which was originally developed for labeling NEs in Hungarian texts, but it can be tuned for different languages, as well.
-Corpus-specific features (e.g.~NP chunks, Wikipedia links) were removed to get better comparability, so the feature set was as follows:
+Corpus-specific features (e.g.~NP chunks, Wikipedia links) were removed to get better comparability, so the feature set consists of gazetteer features; sentence start and end position; Boolean-valued orthographic properties of the word form; string-valued surface properties of the word form; and morphological information.
 
-\begin{itemize}
-\item gazetteer features;
-\item sentence start and end position;
-\item Boolean-valued orthographic properties of the word form;
-\item string-valued surface properties of the word form;
-\item morphological information.
-\end{itemize}
+%\begin{itemize}
+%\item gazetteer features;
+%\item sentence start and end position;
+%\item Boolean-valued orthographic properties of the word form;
+%\item string-valued surface properties of the word form;
+%\item morphological information.
+%\end{itemize}
 
@@
 The rationale behind this is that we wanted to reserve the original distribution of names in Wikipedia as much as possible.
-However, after further investigation of the NE density in our Wikipedia corpora and gold standard corpora (see table \ref{size}), we decided not to include the sentences without NEs in evaluation datasets.
+However, after further investigation of the NE density in our Wikipedia corpora and gold standard corpora (see Table \ref{size}), we decided not to include the sentences without NEs in evaluation datasets.
 
@@
 Nothman et al.~\shortcite{Nothman:08} showed that each set of gold standard training data performs much better on corresponding test sets than on test sets from other sources.
-The situation here is similar: the NE-tagger trained on Wikipedia does not achieve as high performance tested against CoNLL test set as one trained on the own training set (see table \ref{enresults} for results).
+The situation here is similar: the NE-tagger trained on Wikipedia does not achieve as high performance tested against CoNLL test set as one trained on the own training set (see Table \ref{enresults} for results).
 
@@
 Since in the subtask of NE labeling for Hungarian we followed the Szeged NER corpus annotation guidelines, it is obvious to perform the experiments on this dataset, as well.
-Hungarian results are quite similar to the English ones (see table \ref{huresults}), the only difference is that F-measures for Hungarian are significantly higher, than ones for English.
+Hungarian results are quite similar to the English ones (see Table \ref{huresults}), the only difference is that F-measures for Hungarian are significantly higher, than ones for English.
 It can be due to the fact that the MISC category for Hungarian contains less types of names, thus the inconsistency and diversity of this class is smaller (see section \ref{sec:hun}).
@@
 %\appendix
-%\section{DBpedia to ConLL Mapping}
+%\section{DBpedia to CoNLL Mapping}
 %\label{sec:appa}

### 2012-03-25 paper/engNERwiki.tex
My nice paragraph, bruhuhu...

@@
 As of version 3.7, the DBpedia ontology allows multiple superclasses, making the hierarchy a directed acyclic graph\footnote{http://blog.dbpedia.org/2011/09/11/dbpedia-37-released-including-15-localized-editions}.
-This introduces the problem of selecting the right superclass, and hence, CoNLL mapping for classes with more than one parent.
+Since selecting the right superclass, and hence, CoNLL tag for classes with more than one parent cannot be reliably done automatically, the class-to-category mapping has to be determined manually.
 In version 3.7, the only such class is \textit{Library}, which can be traced back to both \textit{Place} and \textit{Organisation}.
-Hence, we picked the right NE mapping manually, and added it to the mapping file.
-
-%DBpedia guidelines suggest that only the first parent is retained.
-Unfortunately, this approach already fails to assign the correct CoNLL NE tag to our only example.
-The reason behind this failure is that multiple parents in an ontology show an inherent ambiguity in the class in question: it is entirely natural to think of a library as a location as well as a public organization.
-DBpedia's guidelines prefer \texttt{Organisation}; CoNLL's prefer LOCATION.
-Hence, we picked the right NE mapping manually and added it to the mapping file.
 

### 2012-03-25 paper/engNERwiki.tex
A little change.
ezmegaz

--- 2012-03-25 paper/engNERwiki.tex
@@
 The most obvious utilization of Wikipedia for NER is extracting gazetteers containing person names, locations or organizations (e.g.~Toral and Mu\~noz \shortcite{Toral:06}).
-Creating dictionaries of entities is also a common step of named entity disambiguation \cite{Bunescu:06,Cucerzan:07}.
+Creating dictionaries of entities is also a common step of NE disambiguation \cite{Bunescu:06,Cucerzan:07}.
 Not only supervised systems use lists generated from Wikipedia: Nadeau et al.~\shortcite{Nadeau:06} use similar lists in an unsupervised NER system.
@@
 
-Using the mapping described above, we compile an entity list that contains all entities in DBpedia tagged with the appropriate CoNLL named entity category.
+Using the mapping described above, we compile a NE list that contains all entities in DBpedia tagged with the appropriate CoNLL category.
 
@@
 
-Strictly speaking, our original assumption of equating Wikipedia articles with named entities is not valid: many pages describe common nouns (Book, Aircraft), calendar-related concepts (March 15, 2007), or other concepts that fall outside the scope of NER.
+Strictly speaking, our original assumption of equating Wikipedia articles with NEs is not valid: many pages describe common nouns (Book, Aircraft), calendar-related concepts (March 15, 2007), or other concepts that fall outside the scope of NER.
 To increase sentence coverage, we modified the algorithm to prevent it from misclassifying links to these pages as unknown entities and discarding the sentence.
@@
 \item[Lowercase links] Entities are sometimes referred to by common nouns, such as \textit{republic} to \textit{Roman Republic}, \texttt{LOC}, which results in a lowercase link.
-These references are not considered named entities and are ignored.
+These references are not considered NEs and are ignored.
 \end{description}
@@
 As mentioned earlier, in Hungarian, apart from the first word in the sentence, only proper nouns are capitalized.
-The following concepts are not proper nouns in Hungarian, and thus are not considered as NEs according to the Szeged NER guideline:
+The following concepts are not proper nouns in Hungarian, and thus are not considered as NEs according to the Szeged NER guideline: names of languages, nationalities, religions, political ideologies; adjectives derived from a NE; names of months, days, holidays; names of special events and wars.
 
-\begin{itemize}
-\item names of languages, nationalities, religions, political ideologies;
-\item adjectives derived from a NE;
-\item names of months, days, holidays;
-\item names of special events and wars.
-\end{itemize}
+%\begin{itemize}
+%\item names of languages, nationalities, religions, political ideologies;
+%\item adjectives derived from a NE;
+%\item names of months, days, holidays;
+%\item names of special events and wars.
+%\end{itemize}
 
@@
 In contrast to the methods used so far for automatic annotation of NEs in Wikipedia texts, we applied a new approach, namely mapping DBpedia ontology classes to standard CoNLL NE tags, and assigning them to Wikipedia entities.
-Following Nothman \shortcite{Nothman:08} the process can be divided into four main steps: (1) classifying all articles into entity classes; (2) splitting Wikipedia articles into sentences; (3) labeling NEs according to link targets; and (4) selecting sentences for inclusion in the corpus.
+Following Nothman \shortcite{Nothman:08}, the process can be divided into four main steps: (1) classifying all articles into entity classes; (2) splitting Wikipedia articles into sentences; (3) labeling NEs according to link targets; and (4) selecting sentences for inclusion in the corpus.
 
@@
 We showed that huge gazetteer lists extracted from our corpora, and training with extra features given by the model trained on our corpora, improves overall F-measure.
-Moreover, our Hungarian corpus can serve as a training corpus to build NE-taggers for more general domains, than the classic newswire.
+Moreover, our Hungarian corpus can serve as a training corpus to build NE taggers for more general domains, than the classic newswire.
 

### 2012-03-25 paper/engNERwiki.tex

@@
 
-As of version 3.7, the DBpedia ontology allows multiple superclasses, making the hierarchy a directed acyclic graph\footnote{http://blog.dbpedia.org/2011/09/11/dbpedia-37-released-including-15-localized-editions}.
-This introduces the problem of selecting the right superclass, and hence, CoNLL mapping for classes with more than one parent.
-In version 3.7, the only such class is \textit{Library}, which can be traced back to both \textit{Place} and \textit{Organisation}.
-Hence, we picked the right NE mapping manually, and added it to the mapping file.
-
 %DBpedia guidelines suggest that only the first parent is retained.
@@
 Hence, we picked the right NE mapping manually and added it to the mapping file.
+
+As of version 3.7, the DBpedia ontology allows multiple superclasses, making the hierarchy a directed acyclic graph\footnote{http://blog.dbpedia.org/2011/09/11/dbpedia-37-released-including-15-localized-editions}.
+Since selecting the right superclass, and hence, CoNLL tag for classes with more than one parent cannot be reliably done automatically, the class-to-category mapping has to be determined manually.
+The only such class in version 3.7, \texttt{Library}, can be traced back to both \texttt{Place} and \texttt{Organisation}; its CoNLL tag is \texttt{LOC}.
 

### 2012-03-25 paper/engNERwiki.tex
Cut a few words...

@@
 
-There are several ways to improve the method described above to increase precision, sentence coverage and to account for peculiarities in the CoNLL guidelines.
-These are detailed in the next sub-sections.
+The following sub-sections describe how the method explained above can be improved to increase precision, sentence coverage and to account for peculiarities in the CoNLL guidelines.
 

### 2012-03-25 paper/engNERwiki.tex
ezmegaz

@@
 The ontology contains the 320 most frequent entity categories, arranged into a taxonomy under the base class \texttt{owl:Thing}.
-Most of the classes belong into the 6 largest sub-hierarchies (here named after their topmost classes): \texttt{Person}, \texttt{Organisation}), \texttt{Event}, \texttt{Place}, \texttt{Species} and \texttt{Work}.
+Most of the classes belong into the 6 largest sub-hierarchies (here named after their topmost classes): \texttt{Person}, \texttt{Organisation}, \texttt{Event}, \texttt{Place}, \texttt{Species} and \texttt{Work}.
 The taxonomy is rather flat: the top level contains 44 classes and there are several nodes with a branching factor of 20.
@@
 From the available NER annotation standards we elected to use the CoNLL \cite{Tjong:03} NE types.
-It is not difficult to see the parallels between the DBpedia sub-hierarchies \texttt{Person}, \texttt{Organisation} and \texttt{Place} and the CoNLL NE categories PERSON, ORGANIZATION and LOCATION.
-The fourth category, MISCELLANEOUS is more elusive; according to the CoNLL NER annotation guide\footnote{http://www.cnts.ua.ac.be/conll2003/ner/annotation.txt}, the sub-hierarchies \texttt{Event} and \texttt{Work} belong to this category, as well as various other classes outside the main hierarchies.
+It is not difficult to see the parallels between the DBpedia sub-hierarchies \texttt{Person}, \texttt{Organisation} and \texttt{Place} and the CoNLL NE categories PER, ORG and LOC.
+The fourth category, MISC is more elusive; according to the CoNLL NER annotation guide\footnote{http://www.cnts.ua.ac.be/conll2003/ner/annotation.txt}, the sub-hierarchies \texttt{Event} and \texttt{Work} belong to this category, as well as various other classes outside the main hierarchies.
 
@@
 
-Using the mapping described above, we compile a NE list that contains all entities in DBpedia tagged with the appropriate CoNLL category.
+Using the mapping described above, we compile a list that contains all entities in DBpedia tagged with the appropriate CoNLL category.
 
@@
 A NER corpus consists of sentences where NEs are accurately tagged.
-In order to automatically prepare such sentences, two tasks need to be performed: identifying the entities in the sentence and tagging them with the correct tag.
+In order to automatically prepare such sentences, two tasks need to be performed: identifying entities in the sentence and tagging them with the correct tag.
 Sentences, for which accurate tagging could not be accomplished, must be removed from the corpus.
@@
 We assume that individual Wikipedia articles describe NEs.
-A link to an article can then be perceived as a mapping that identifies its anchor text with a particular entity.
+A link to an article can then be perceived as a mapping that identifies its anchor text with a particular NE.
 
-Discovered entities are tagged with the CoNLL category assigned to them in the entity list extracted from DBpedia.
-If the link target is not in the entity list, or the link points to a disambiguation page, we cannot determine the type of the entity and tag it as \texttt{UNK}.
+Discovered entities are tagged with the CoNLL label assigned to them in the entity list extracted from DBpedia.
+If the link target is not in the entity list, or the link points to a disambiguation page, we can not determine the type of the entity, and tag it as UNK.
 Links to redirect pages are resolved to point instead to the redirect target, after which they are handled as regular cross-references.
@@
 
-The following sub-sections describe how the method explained above can be improved to increase precision, sentence coverage and to account for peculiarities in the CoNLL guidelines.
+There are several ways to increase precision, sentence coverage and to account for peculiarities in the CoNLL guidelines. These are detailed in the next subsections.
 
@@
 We evaluated our English corpus against (a) itself (divided into train and test sets); and (b) a manually annotated English corpus.
-Since the filtered English Wikipedia corpus -- containing only the sentences with NEs -- is still very huge, and we are limited by time and memory, our experiments were performed with a portion of 3.5 million tokens (the size of our filtered Hungarian corpus).
+Since the filtered English Wikipedia corpus, containing only the sentences with NEs, is still very huge, and we are limited by time and memory, our experiments were performed with a portion of 3.5 million tokens (the size of our filtered Hungarian corpus).
 It was divided into train and test sets (90\%-10\%).
@@
 For the English cross-corpus evaluation the CoNLL-2003 dataset was chosen, mostly because of the similarity in tagset and format.
-It is well-studied that cross-checking between different corpora leads to F-measure decrease.
+It is well-studied that cross-checking between different corpora decrease F-measure.
 Stylistic and domain differences certainly affect NER performance, and the different annotation schemes pose several compatibility problems.

### 2012-03-25 paper/engNERwiki.tex

@@
 
-There are several ways to increase precision, sentence coverage and to account for peculiarities in the CoNLL guidelines. These are detailed in the next subsections.
+The following sub-sections describe how the method explained above can be improved to increase precision, sentence coverage and to account for peculiarities in the CoNLL guidelines.
 

### 2012-03-25 paper/engNERwiki.tex
Deleted a few words.
A.

@@
 
-%DBpedia guidelines suggest that only the first parent is retained.
-Unfortunately, this approach already fails to assign the correct CoNLL NE tag to our only example.
-The reason behind this failure is that multiple parents in an ontology show an inherent ambiguity in the class in question: it is entirely natural to think of a library as a location as well as a public organization.
-DBpedia's guidelines prefer \texttt{Organisation}; CoNLL's prefer LOCATION.
-Hence, we picked the right NE mapping manually and added it to the mapping file.
-
 As of version 3.7, the DBpedia ontology allows multiple superclasses, making the hierarchy a directed acyclic graph\footnote{http://blog.dbpedia.org/2011/09/11/dbpedia-37-released-including-15-localized-editions}.
@@
   \item Title of redirect pages that target the current page
-  \item Title of the current page in other languages, extracted from the interlanguage link file, as Wikipedia usually mentions the original name of foreign entities
+  \item Title of the current page in other languages, as Wikipedia usually mentions the original name of foreign entities
   \end{itemize}
@@
 \begin{description}
-\item[Derived words] According to the CoNLL guidelines, word derived from name entities must be tagged as \texttt{MISC}.
+\item[Derived words] According to the CoNLL guidelines, words derived from name entities must be tagged as \texttt{MISC}.
 We complied with this rule by tagging each entity whose head is not a noun, as well as when the link's anchor text is not contained in the entity's name, as \texttt{MISC}.
@@
 Our solution assigns the correct tag to these entities.
-\item[First word in a sentence] As first words are always capitalized, labelling them is difficult if they are unlinked and not contained in the entity alias set.
+\item[First word in a sentence] As first words are always capitalized, labeling them is difficult if they are unlinked and not contained in the entity alias set.
 We base the decision on the POS tag of the first word: if it is \texttt{NNP}, we tag it as \texttt{UNK}; otherwise, \texttt{O}.

### 2012-03-25 paper/engNERwiki.tex
jav

--- 2012-03-25 paper/engNERwiki.tex
@@
 \begin{description}
-\item[Derived words] According to the CoNLL guidelines, word derived from name entities must be tagged as \texttt{MISC}.
+\item[Derived words] According to the CoNLL guidelines, word derived from NEs are tagged as \texttt{MISC}.
 We complied with this rule by tagging each entity whose head is not a noun, as well as when the link's anchor text is not contained in the entity's name, as \texttt{MISC}.
-The most prominent example for such entities are ethnic groups, which can be linked to their home country, a \texttt{LOC}; e.g.
-\textit{Turkish} to \textit{Turkey}.
+The most prominent example for such entities are nationalities, which can be linked to their home country, a \texttt{LOC}; e.g.~\textit{Turkish} to \textit{Turkey}.
 Our solution assigns the correct tag to these entities.
-\item[First word in a sentence] As first words are always capitalized, labelling them is difficult if they are unlinked and not contained in the entity alias set.
-We base the decision on the POS tag of the first word: if it is \texttt{NNP}, we tag it as \texttt{UNK}; otherwise, \texttt{O}.
+\item[First word in a sentence] As first words are always capitalized, labeling them is difficult if they are unlinked and not contained in the entity alias set.
+We base the decision on the POS tag of the first word: if it is NNP, we tag it as UNK; otherwise, O.
 \item[Reference cleansing] Page titles and anchor texts may contain more than just the entity name.
@@
 
-Certain types of sentences were filtered from the corpus.
 As mentioned above, sentences with words tagged as \texttt{UNK} are discarded.
@@
 In this section we describe the differences between the two languages related to labeling NEs, and the changes they prompted in the method.
+
+\begin{table*}[ht]
+\begin{center}
+\begin{tabular}{lllll}
+\hline \bf train & \bf test & \bf precision & \bf recall & \bf F-measure \\ \hline
+Szeged NER & Szeged NER & 94.50 & 94.35 & 94.43 \\
+huwiki & huwiki & 90.64 & 88.91 &  89.76 \\
+huwiki & Szeged NER & 63.08 & 70.46 & 66.57 \\
+Szeged NER with wikilists & Szeged NER & 95.48 & 95.48 & 95.48 \\
+Szeged NER with wikitags & Szeged NER & 95.38 & 94.92 & 95.15 \\
+\hline
+\end{tabular}
+\end{center}
+\caption{\label{huresults} Hungarian results.}
+\end{table*}
 
@@
 
-To annotate our corpus with NE-tags, we chose to follow the annotation guidelines of the largest human-annotated NER corpus for Hungarian, Szeged NER corpus \cite{Szarvas:06}.
-It is quite similar to the CoNLL standards: it contains business newswire texts, it comprises ca.~200,000 tokens, and the text has been annotated with NE class labels in line with the CoNLL annotation scheme.
+To annotate our corpus with NE tags, we chose to follow the annotation guidelines of the largest human-annotated NER corpus for Hungarian, Szeged NER corpus \cite{Szarvas:06}.
+It is similar to CoNLL standards: contains newswire texts, comprises ca.~200,000 tokens, and is annotated with NE class labels in line with the CoNLL annotation scheme.
 However, the convention of what constitutes a NE is slightly different for Hungarian.
@@
 
+\begin{table*}[ht]
+\begin{center}
+\begin{tabular}{lllll}
+\hline \bf train & \bf test & \bf precision & \bf recall & \bf F-measure \\ \hline
+CoNLL & CoNLL & 85.13 & 85.13 & 85.13 \\
+enwiki & enwiki & 72.46 & 73.33 &  72.89 \\
+enwiki & CoNLL & 56.55 & 49.77 & 52.94 \\
+CoNLL with wikilists & CoNLL & 86.33 & 86.35 & 86.34 \\
+CoNLL with wikitags & CoNLL & 85.88 & 85.94 & 85.91 \\
+\hline
+\end{tabular}
+\end{center}
+\caption{\label{enresults} English results.}
+\end{table*}
+
 \section{Evaluation}
@@
 Having the obvious advantages, an automatically generated corpus can not serve as a gold standard dataset.
-Then what can we do with these so called "silver standard corpora" (SSC)?
+Then what can we do with silver standard corpora?
 They can be very useful for improving NER in more ways: (a) for less resourced languages, they can serve as training corpora in absence of gold standard datasets; (b) they can serve as (supplementary or independent) training sets for domains differing from newswire; (c) they can be sources of huge entity lists, and (d) feature extraction.
@@
 
-\begin{table*}[ht]
-\begin{center}
-\begin{tabular}{lllll}
-\hline \bf train & \bf test & \bf precision & \bf recall & \bf F-measure \\ \hline
-CoNLL & CoNLL & 85.13 & 85.13 & 85.13 \\
-enwiki & enwiki & 72.46 & 73.33 &  72.89 \\
-enwiki & CoNLL & 56.55 & 49.77 & 52.94 \\
-CoNLL with wikilists & CoNLL & 86.33 & 86.35 & 86.34 \\
-CoNLL with wikitags & CoNLL & 85.88 & 85.94 & 85.91 \\
-\hline
-\end{tabular}
-\end{center}
-\caption{\label{enresults} English results.}
-\end{table*}
-
-\begin{table*}[ht]
-\begin{center}
-\begin{tabular}{lllll}
-\hline \bf train & \bf test & \bf precision & \bf recall & \bf F-measure \\ \hline
-Szeged NER & Szeged NER & 94.50 & 94.35 & 94.43 \\
-huwiki & huwiki & 90.64 & 88.91 &  89.76 \\
-huwiki & Szeged NER & 63.08 & 70.46 & 66.57 \\
-Szeged NER with wikilists & Szeged NER & 95.48 & 95.48 & 95.48 \\
-Szeged NER with wikitags & Szeged NER & 95.38 & 94.92 & 95.15 \\
-\hline
-\end{tabular}
-\end{center}
-\caption{\label{huresults} Hungarian results.}
-\end{table*}
-
-We evaluated our English corpus against (a) itself (divided into train and test sets); and (b) a manually annotated English corpus.
+The English train corpus was evaluated against itself and a manually annotated English corpus.
 Since the filtered English Wikipedia corpus, containing only the sentences with NEs, is still very huge, and we are limited by time and memory, our experiments were performed with a portion of 3.5 million tokens (the size of our filtered Hungarian corpus).
@@
 
-For the English cross-corpus evaluation the CoNLL-2003 dataset was chosen, mostly because of the similarity in tagset and format.
+For English cross-corpus evaluation the CoNLL-2003 dataset was chosen, mostly because of the similarity in tagset and format.
 It is well-studied that cross-checking between different corpora decrease F-measure.
@@
 
-Since in the subtask of NE labeling for Hungarian we followed the Szeged NER corpus annotation guidelines, it is obvious to perform the experiments on this dataset, as well.
+Since in NE labeling for Hungarian we followed the Szeged NER corpus annotation guidelines, it is obvious to perform the experiments on this dataset, as well.
 Hungarian results are quite similar to the English ones (see Table \ref{huresults}), the only difference is that F-measures for Hungarian are significantly higher, than ones for English.
@@
 In contrast to the methods used so far for automatic annotation of NEs in Wikipedia texts, we applied a new approach, namely mapping DBpedia ontology classes to standard CoNLL NE tags, and assigning them to Wikipedia entities.
-Following Nothman \shortcite{Nothman:08}, the process can be divided into four main steps: (1) classifying all articles into entity classes; (2) splitting Wikipedia articles into sentences; (3) labeling NEs according to link targets; and (4) selecting sentences for inclusion in the corpus.
+Following Nothman \shortcite{Nothman:08}, the process can be divided into four main steps:  classifying all articles into entity classes; splitting Wikipedia articles into sentences; labeling NEs according to link targets; and selecting sentences for inclusion in the corpus.
 
@@
 
-Automatically generated, so called "silver standard corpora" can be very useful for improving NER in more ways.
+Automatically generated, silver standard corpora can be very useful for improving NER in more ways.
 We showed that huge gazetteer lists extracted from our corpora, and training with extra features given by the model trained on our corpora, improves overall F-measure.

### 2012-03-25 paper/engNERwiki.tex
E.

@@
 
-% Eszter: ehhez a felső listához hivatkozzam Nothmant? A fele tőle jön, viszont mi máshogy csináltuk kicsit.
 Our solution is similar to Nothman's \shortcite{Nothman:08}.
@@
 \begin{description}
-\item[Derived words] According to the CoNLL guidelines, word derived from NEs are tagged as \texttt{MISC}.
+\item[Derived words] According to the CoNLL guidelines, words derived from NEs are tagged as \texttt{MISC}.
 We complied with this rule by tagging each entity whose head is not a noun, as well as when the link's anchor text is not contained in the entity's name, as \texttt{MISC}.
@@
 \item[First word in a sentence] As first words are always capitalized, labeling them is difficult if they are unlinked and not contained in the entity alias set.
-We base the decision on the POS tag of the first word: if it is NNP, we tag it as UNK; otherwise, O.
+We base the decision on the POS tag of the first word: if it is \texttt{NNP}, we tag it as \texttt{UNK}; otherwise, \texttt{O}.
 \item[Reference cleansing] Page titles and anchor texts may contain more than just the entity name.

### 2012-03-25 paper/engNERwiki.tex
ez

--- 2012-03-25 paper/engNERwiki.tex
@@
 
-A NER corpus consists of sentences where NEs are accurately tagged.
-In order to automatically prepare such sentences, two tasks need to be performed: identifying entities in the sentence and tagging them with the correct tag.
+In order to automatically prepare sentences where NEs are accurately tagged, two tasks need to be performed: identifying entities in the sentence and tagging them with the correct tag.
 Sentences, for which accurate tagging could not be accomplished, must be removed from the corpus.
@@
 
-There are some special cases to the algorithm described thus far. These are detailed below.
+%There are some special cases to the method, which are detailed below.
 
@@
 
-For some of the well-resourced languages, DBpedia has internationalized "chapters" -- data dumps based on the Wikipedia of those particular languages.
+For well-resourced languages, DBpedia has internationalized chapters; data dumps based on Wikipedia.
 Hungarian does not have a separate chapter.
@@
 The linguistic features include the lemmatized form of the word and its part-of-speech tag.
-Two NE tags are included with each word: one is the most specific DBpedia category it belongs to, if any; the other is the CoNLL NE tag.
+Two NE tags are included with each word: the most specific DBpedia category it belongs to and the CoNLL NE tag.
 While the NE tags can be considered as a "silver standard", the linguistic features are provided on a "best-effort" basis.
@@
 The rationale behind this is that we wanted to reserve the original distribution of names in Wikipedia as much as possible.
-However, after further investigation of the NE density in our Wikipedia corpora and gold standard corpora (see Table \ref{size}), we decided not to include the sentences without NEs in evaluation datasets.
+However, after further investigation of the NE density in our corpora and gold standard corpora (see Table \ref{size}), we decided not to include the sentences without NEs in evaluation datasets.
 
@@
 
-For English cross-corpus evaluation the CoNLL-2003 dataset was chosen, mostly because of the similarity in tagset and format.
+For English cross-corpus evaluation the CoNLL-2003 corpus was chosen.
 It is well-studied that cross-checking between different corpora decrease F-measure.
-Stylistic and domain differences certainly affect NER performance, and the different annotation schemes pose several compatibility problems.
-Nothman et al.~\shortcite{Nothman:08} showed that each set of gold standard training data performs much better on corresponding test sets than on test sets from other sources.
-The situation here is similar: the NE-tagger trained on Wikipedia does not achieve as high performance tested against CoNLL test set as one trained on the own training set (see Table \ref{enresults} for results).
+Domain differences certainly affect NER performance, and the different annotation schemes pose several compatibility problems.
+Nothman et al.~\shortcite{Nothman:08} showed that each set of gold standard training data performs better on corresponding test sets than on test sets from other sources.
+The situation here is similar: the NE-tagger trained on Wikipedia does not achieve as high performance tested against CoNLL test set as one trained on the own train set (see Table \ref{enresults}).
 
-However, Wikipedia-derived corpora can also be used for improving NER performance in other ways.
-First, like e.g.~\cite{Toral:06} we collected huge gazetteer lists from the corpus for each NE category, which improved the overall F-measure given to the NE-tagger training and testing on CoNLL dataset.
+However, Wikipedia-derived corpora can also be used for improving NER accuracy in other ways.
+First, we collected gazetteer lists from the corpus for each NE category, which improved the overall F-measure given to the NE-tagger training and testing on CoNLL dataset.
 A second trial was labeling the CoNLL datasets by the model trained on Wikipedia corpus, and giving these labels as extra features to the next CoNLL train.
-Both methods result in improved overall F-measure on CoNLL test set.
+Both methods result in improved F-measure on CoNLL test set.
 
-Since in NE labeling for Hungarian we followed the Szeged NER corpus annotation guidelines, it is obvious to perform the experiments on this dataset, as well.
-Hungarian results are quite similar to the English ones (see Table \ref{huresults}), the only difference is that F-measures for Hungarian are significantly higher, than ones for English.
-It can be due to the fact that the MISC category for Hungarian contains less types of names, thus the inconsistency and diversity of this class is smaller (see section \ref{sec:hun}).
-In contrast to the CoNLL corpus, the Szeged NER corpus was highly accurately annotated with an inter-annotator agreement over 99\%.
+Since in Hungarian NE tagging we followed the Szeged NER corpus annotation guidelines, we performed the experiments on this dataset.
+Hungarian results are similar to the English ones (see Table \ref{huresults}), the only difference is that F-measures for Hungarian are significantly higher.
+It can be due to the fact that the MISC category for Hungarian contains less types of names, thus the inconsistency of this class is smaller (see section \ref{sec:hun}).
+In contrast to the CoNLL corpus, the Szeged NER corpus was accurately annotated with an inter-annotator agreement over 99\%.
 
-Due to the quite good F-measure of training on the Hungarian Wikipedia train corpus and testing on the corresponding test set, our Hungarian corpus can serve as a training corpus to build NE-taggers for domains falling out of newswire.
+Due to the quite good F-measure of training on our Hungarian train corpus and testing on the corresponding test set, our Hungarian corpus can serve as a training corpus to build NE-taggers for domains falling out of newswire.
 
@@
 Thanks for the huge amount of Wikipedia articles, the possibility of building large enough corpora is given for less resourced languages, like Hungarian, as well.
-Due to the particularities of the Hungarian language, some steps are slightly different, and special linguistics phenomena pose several problems related to the NER task to solve.
+Due to the particularities of Hungarian, some steps are slightly different, and special linguistics phenomena pose several problems related to the NER task to solve.
 
-Automatically generated, silver standard corpora can be very useful for improving NER in more ways.
-We showed that huge gazetteer lists extracted from our corpora, and training with extra features given by the model trained on our corpora, improves overall F-measure.
-Moreover, our Hungarian corpus can serve as a training corpus to build NE taggers for more general domains, than the classic newswire.
+Automatically generated corpora can be useful for improving NER in more ways.
+We showed that gazetteer lists extracted from our corpora, and training with extra features given by the model trained on our corpora, improves F-measure.
+Moreover, our Hungarian corpus can serve as a training corpus for more general domains.
+%than the classic newswire.
 

### 2012-04-25 paper/engNERwiki.tex

@@
 
-%\author{Dávid Márk Nemeskey \\
-%  Research Institute for Computer Science and Automation \\
-%  Hungarian Academy of Sciences \\
-%  H-1111 Lágymányosi út 11, Budapest \\
-%  {\tt nemeskey.david@sztaki.hu} \\\And
-%  Eszter Simon \\
-%  Research Institute for Linguistics \\
-%  Hungarian Academy of Sciences \\
-%  H-1068 Benczúr utca 33, Budapest \\
-%  {\tt eszter@nytud.hu} \\}
+\author{Dávid Márk Nemeskey \\
+  Research Institute for \\ Computer Science and Automation \\
+  Hungarian Academy of Sciences \\
+  H-1111 Lágymányosi út 11, Budapest \\
+  {\tt nemeskey.david@sztaki.hu} \\\And
+  Eszter Simon \\
+  Research Institute for Linguistics \\
+  Hungarian Academy of Sciences \\
+  H-1068 Benczúr utca 33, Budapest \\
+  {\tt simon.eszter@nytud.mta.hu} \\}
 
@@
 
-Supervised Named Entity Recogizers require large amount of annotated texts.
+Supervised Named Entity Recognizers require large amount of annotated texts.
 However, manual annotation is a highly costly procedure, thus reducing the annotation cost is essential.
@@
 The standard datasets are highly domain-specific (mostly newswire), and are restricted in size, as well. 
-%Thus they are not large enough to train and test robust NE-taggers on them. Moreover, 
+%Thus they are not large enough to train and test robust NE taggers on them. Moreover, 
 Merging these datasets to get a bigger training corpus faces NLP researchers with the problem of combining different tagsets and annotation schemes.
@@
 Another approach to achieve this aim is using collaborative annotation and/or collaboratively constructed resources, such as Wikipedia, Wiktionary or Linked Open Data.
-Researchers apply the Web-as-corpus approach, often dealing with such datasets as Google's Web 1T 5-gram collection or DBpedia.
+Researchers apply the Web-as-corpus approach, often dealing with such datasets as Google Web 1T 5-gram collection or DBpedia.
 
@@
 
-In this paper we present a kind of combination of the approaches mentioned above.
+In this paper we present a kind of combination of these approaches, namely automatically generating resources and using collaboratively constructed resources.
 We built freely available NE tagged corpora for English and Hungarian -- fully automatically generated from Wikipedia.
-%Our goal was to create gold standard corpora which can serve as train, devel and test datasets for automatic NE-taggers.
+%Our goal was to create gold standard corpora which can serve as train, devel and test datasets for automatic NE taggers.
 
@@
 In Section~\ref{sec:related} we give an overview of related work.
-Our method is introduced in Section~\ref{sec:create}, and showed how it is applied to Hungarian in Section~\ref{sec:hun}.
+Section \ref{sec:create} contains a description of our method, while Section \ref{sec:hun} shows how it is applied to Hungarian.
+The corpus format is described in Section~\ref{sec:data}.
+In Section \ref{sec:eval} we present experiments and results on the newly generated datasets.
+Section \ref{sec:conclusion} concludes the paper with a summary.
+
+%Our method is introduced in Section~\ref{sec:create}, and showed how it is applied to Hungarian in Section~\ref{sec:hun}.
 The corpus format is described in Section~\ref{sec:data}.
@@
 
-The knowledge in Wikipedia may be incorporated as features in NER learning, e.g.~Kazama and Torisawa \shortcite{Kazama:07} showed that automatic extraction of category labels from Wikipedia improves the accuracy of a supervised NE-tagger.
+The knowledge in Wikipedia may be incorporated as features in NER learning, e.g.~Kazama and Torisawa \shortcite{Kazama:07} showed that automatic extraction of category labels from Wikipedia improves the accuracy of a supervised NE tagger.
 
 A third approach to improve NER with Wikipedia is the automatic creation of training data.
-Richman and Schone \shortcite{Richman:08} built corpora for less commonly taught languages annotated with NE-tags.
-They used the inherent category structure of Wikipedia to determine the NE-type of a proposed entity.
+Richman and Schone \shortcite{Richman:08} built corpora for less commonly taught languages annotated with NE tags.
+They used the inherent category structure of Wikipedia to determine the NE type of a proposed entity.
 Nothman et al.~\shortcite{Nothman:08} used a similar method to create a NE annotated text in English.
-They transformed the Wikipedia links into NE-annotations by classifying the target articles into standard entity classes.
+They transformed the Wikipedia links into NE annotations by classifying the target articles into standard entity classes.
 Their approach to classification is based primarily on category head nouns and definitional opening sentences of articles.
@@
 In contrast to the aforementioned methods, we applied a new approach to recognize and classify NEs in corpora generated from Wikipedia.
-We mapped the DBpedia ontology classes to standard NE tags and assigned them to Wikipedia entities (see more details in section \ref{dbpedia}).
+We mapped the DBpedia ontology classes to standard NE tags and assigned them to Wikipedia entities (see more details in Section \ref{dbpedia}).
 
@@
 
-Thanks for the huge amount of Wikipedia articles, the possibility of building large enough 
+Thanks to the huge amount of Wikipedia articles, the possibility of building large enough 
 corpora is given for less resourced languages, as well.
 Since our method is largely language-independent, it can be easily applied for languages which are structurally differing from English.
-Hungarian is a highly agglutinative language, with free word order and other particularities detailed later in section \ref{sec:hun}.
+Hungarian is a highly agglutinative language, with free word order and other particularities detailed later in Section \ref{sec:hun}.
 There are human annotated CoNLL-style datasets, but the one presented here is the first automatically NE annotated corpus for Hungarian.
@@
 
-Our goal is to create a large NE-annotated corpus, automatically generated from Wikipedia articles. We follow a similar path to \cite{Nothman:08}, who broke down the process into four steps:
+Our goal is to create a large NE annotated corpus, automatically generated from Wikipedia articles.
+We follow a similar path to Nothman et al.~\shortcite{Nothman:08}, who broke down the process into four steps:
 
@@
 
-In this section, we describe how these steps were implemented.
-While the approach in essence is language-independent, the exact details XXX have to be accomodated to the target language.
-This section explains the general idea and its execution for English; the next section describes how the idea is adapted to Hungarian.
+%Subsectionöket valahogy ez alapján kéne.
+
+In this section, we describe how these steps were implemented. While the approach in essence is language-independent, the exact details XXX have to be accomodated to the target language. 
+%Ez utóbbi mondat mit jelent?
+
+%This section explains the general idea and its execution for English; the next section describes how the idea is adapted to Hungarian.
 
@@
 The English corpus is based on the Wikipedia snapshot as of January 15, 2011.
-The XML files were parsed with the \texttt{mwlib} parser\footnote{http://code.pediapress.com}.
-The raw text was tokenized with a modified version of the Punkt sentence and word tokenizers \cite{Kiss:06} in NLTK \cite{Loper:02}.
-The words were lemmatized with the Wordnet Lemmatizer, also in NLTK, and POS tagged with HunPos \cite{Halacsy:07}.
+The XML files were parsed by the mwlib parser\footnote{http://code.pediapress.com}.
+The raw text was tokenized by a modified version of the Punkt sentence and word tokenizers \cite{Kiss:06} in NLTK.
+%\cite{Loper:02}. 
+Lemmatization was performed using the Wordnet Lemmatizer, also in NLTK. Part-of-speech tagging was performed by HunPos \cite{Halacsy:07}.
+%The words were lemmatized with the Wordnet Lemmatizer, also in NLTK, and POS tagged with \texttt{HunPos} \cite{Halacsy:07}.
 
@@
 
-Many authors, such as Kazama and Torisawa \shortcite{Kazama:07} and Nothman et al.~\shortcite{Nothman:08} used semi-supervised methods based on Wikipedia categories and text to classify articles into entity categories.
+Many authors, such as Kazama and Torisawa \shortcite{Kazama:07} and Nothman et al.~\shortcite{Nothman:08} used semi-supervised methods based on Wikipedia categories and text to classify articles into NE types.
 To avoid the inevitable classification errors, we obtain entity type information from the DBpedia knowledge base \cite{Bizer:09}.
-This enables us to have high precision information about entity types at the expense of recall: out of the 3,903,467 English Wikipedia pages, 1,470,293 are covered by DBpedia.
+This enables us to have high precision information about entity types at the expense of recall: of the 3,903,467 English Wikipedia pages, 1,470,293 are covered by DBpedia.
 
@@
 The ontology contains the 320 most frequent entity categories, arranged into a taxonomy under the base class \texttt{owl:Thing}.
-Most of the classes belong into the 6 largest sub-hierarchies (here named after their topmost classes): \texttt{Person}, \texttt{Organisation}, \texttt{Event}, \texttt{Place}, \texttt{Species} and \texttt{Work}.
+Most of the classes belong to the 6 largest sub-hierarchies (here named after their topmost classes): \texttt{Person}, \texttt{Organisation}, \texttt{Event}, \texttt{Place}, \texttt{Species} and \texttt{Work}.
 The taxonomy is rather flat: the top level contains 44 classes and there are several nodes with a branching factor of 20.
@@
 From the available NER annotation standards we elected to use the CoNLL \cite{Tjong:03} NE types.
-It is not difficult to see the parallels between the DBpedia sub-hierarchies \texttt{Person}, \texttt{Organisation} and \texttt{Place} and the CoNLL NE categories PER, ORG and LOC.
-The fourth category, MISC is more elusive; according to the CoNLL NER annotation guide\footnote{http://www.cnts.ua.ac.be/conll2003/ner/annotation.txt}, the sub-hierarchies \texttt{Event} and \texttt{Work} belong to this category, as well as various other classes outside the main hierarchies.
+It is not difficult to see the parallels between the DBpedia sub-hierarchies \texttt{Person}, \texttt{Organisation} and \texttt{Place} and the CoNLL NE types \texttt{PER}, \texttt{ORG} and \texttt{LOC}.
+The fourth category, \texttt{MISC} is more elusive; according to the CoNLL NER annotation guide\footnote{http://www.cnts.ua.ac.be/conll2003/ner/annotation.txt}, the sub-hierarchies \texttt{Event} and \texttt{Work} belong to this category, as well as various other classes outside the main hierarchies.
 
 While the correspondence described above holds for most classes in the sub-hierarchies, there are some exceptions.
-For instance, the class \texttt{SportsLeague} is part of the \texttt{Organisation} sub-hierarchy, but according to the CoNLL annotation scheme, they are tagged as MISC.
+For instance, the class \texttt{SportsLeague} is part of the \texttt{Organisation} sub-hierarchy, but according to the CoNLL annotation scheme, they should be tagged as \texttt{MISC}.
 
-To account for these misclassification, as well as the miscellaneous classes outside the main hierarchies, we have introduced a mapping mechanism.
-We created a file of DBpedia class--NE category mappings.
-Whenever an entity is evaluated, we look up its class, as well as the ancestors of its class, in the mapping, and assign to it the category of the class that matches it most closely.
+To avoid such kind of misclassification, 
+%as well as the miscellaneous classes outside the main hierarchies, we have introduced a mapping mechanism. 
+we created a file of DBpedia class--NE category mappings.
+Whenever an entity is evaluated, we look up its class and the ancestors of its class, and assign to it the category of the class that matches it most closely.
 If no match is found, the entity is tagged with \texttt{O}.
-Since we take advantage of the inheritance hierarchy, the mapping list remains short: it contains only the root classes of the main hierarchies, exceptions like those mentioned above, and the various classes that belong to the MISC category according to the CoNLL annotation guideline.
+%Since we take advantage of the inheritance hierarchy, the mapping list remains short: it contains only the root classes of the main hierarchies, exceptions like those mentioned above, and the various classes that belong to the MISC category according to the CoNLL annotation guideline.
 
-As of version 3.7, the DBpedia ontology allows multiple superclasses, making the hierarchy a directed acyclic graph\footnote{http://blog.dbpedia.org/2011/09/11/dbpedia-37-released-including-15-localized-editions}.
+As of version 3.7, the DBpedia ontology allows multiple superclasses, making a directed acyclic graph\footnote{http://blog.dbpedia.org/2011/09/11/dbpedia-37-released-including-15-localized-editions}.
 Since selecting the right superclass, and hence, CoNLL tag for classes with more than one parent cannot be reliably done automatically, the class-to-category mapping has to be determined manually.
-The only such class in version 3.7, \texttt{Library}, can be traced back to both \texttt{Place} and \texttt{Organisation}; its CoNLL tag is \texttt{LOC}.
+The only such class in version 3.7 (\texttt{Library}) can be traced back to both \texttt{Place} and \texttt{Organisation}; its CoNLL tag is \texttt{LOC}.
 
@@
 We would like to note that even though we use CoNLL categories, our method can be modified to work with any tagset compatible with the DBpedia ontology.
-Indeed, the DBpedia classes define a named entity tagset themselves.
+Indeed, the DBpedia classes define a NE tagset themselves.
 We leave the exploration of these possibilities for future work.
@@
 \end{center}
-\caption{\label{size} Corpus size and NE-density.}
+\caption{\label{size} Corpus size and NE density.}
 \end{table*}
@@
 
-To evaluate our corpora we used a maximum entropy NE-tagger \cite{Varga:07}, which was originally developed for labeling NEs in Hungarian texts, but it can be tuned for different languages, as well.
+To evaluate our corpora we used a maximum entropy NE tagger \cite{Varga:07}, which was originally developed for labeling NEs in Hungarian texts, but it can be tuned for different languages, as well.
 Corpus-specific features (e.g.~NP chunks, Wikipedia links) were removed to get better comparability, so the feature set consists of gazetteer features; sentence start and end position; Boolean-valued orthographic properties of the word form; string-valued surface properties of the word form; and morphological information.
@@
 Nothman et al.~\shortcite{Nothman:08} showed that each set of gold standard training data performs better on corresponding test sets than on test sets from other sources.
-The situation here is similar: the NE-tagger trained on Wikipedia does not achieve as high performance tested against CoNLL test set as one trained on the own train set (see Table \ref{enresults}).
+The situation here is similar: the NE tagger trained on Wikipedia does not achieve as high performance tested against CoNLL test set as one trained on the own train set (see Table \ref{enresults}).
 
 However, Wikipedia-derived corpora can also be used for improving NER accuracy in other ways.
-First, we collected gazetteer lists from the corpus for each NE category, which improved the overall F-measure given to the NE-tagger training and testing on CoNLL dataset.
+First, we collected gazetteer lists from the corpus for each NE category, which improved the overall F-measure given to the NE tagger training and testing on CoNLL dataset.
 A second trial was labeling the CoNLL datasets by the model trained on Wikipedia corpus, and giving these labels as extra features to the next CoNLL train.
@@
 
-Due to the quite good F-measure of training on our Hungarian train corpus and testing on the corresponding test set, our Hungarian corpus can serve as a training corpus to build NE-taggers for domains falling out of newswire.
+Due to the quite good F-measure of training on our Hungarian train corpus and testing on the corresponding test set, our Hungarian corpus can serve as a training corpus to build NE taggers for domains falling out of newswire.
 

### 2012-04-26 paper/engNERwiki.tex
Acknowledgements.
NLTK citation modified.

@@
 Our goal is to create a large NE annotated corpus, automatically generated from Wikipedia articles.
-We follow a similar path to Nothman et al.~\shortcite{Nothman:08}, who broke down the process into four steps:
+We follow a similar path to Nothman et al.~\shortcite{Nothman:08}, and we broke down the process into four steps:
 
 \begin{enumerate}
-\item Classify all articles into entity classes.
-\item Split Wikipedia articles into sentences.
+\item Classify Wikipedia articles into entity classes.
+\item Parse Wikipedia and split articles into sentences.
 \item Label NEs according to link targets.
@@
 
-%Subsectionöket valahogy ez alapján kéne.
+In this section, we describe how these steps were implemented.
+This section explains the general approach and its execution for English; the next section describes how the idea is adapted to Hungarian.
 
-In this section, we describe how these steps were implemented. While the approach in essence is language-independent, the exact details XXX have to be accomodated to the target language. 
-%Ez utóbbi mondat mit jelent?
-
-%This section explains the general idea and its execution for English; the next section describes how the idea is adapted to Hungarian.
-
-\subsection{Parsing Wikipedia}
-
-Wikipedia is a very rich source of information; aside from the article text, a huge amount of data is encompassed in infoboxes, templates and the category structure.
-Our task requires the use of only the links between the articles and the article text.
-In addition to in-article links, our method takes advantage of the redirect and interlanguage links, available as SQL dumps.
-
-The English corpus is based on the Wikipedia snapshot as of January 15, 2011.
-The XML files were parsed by the mwlib parser\footnote{http://code.pediapress.com}.
-The raw text was tokenized by a modified version of the Punkt sentence and word tokenizers \cite{Kiss:06} in NLTK.
-%\cite{Loper:02}. 
-Lemmatization was performed using the Wordnet Lemmatizer, also in NLTK. Part-of-speech tagging was performed by HunPos \cite{Halacsy:07}.
-%The words were lemmatized with the Wordnet Lemmatizer, also in NLTK, and POS tagged with \texttt{HunPos} \cite{Halacsy:07}.
-
-%Chunking was performed with HunChunk (CITE). Experiments have shown that HunPos obtains STAT f-score on the Brown corpus; HunChunk performs at around STAT, CITE?.
+%While the approach in essence is language-independent, the exact details have to be accomodated to the target language.
 
@@
 We leave the exploration of these possibilities for future work.
+
+\subsection{Parsing Wikipedia}
+
+Wikipedia is a very rich source of information; aside from the article text, a huge amount of data is encompassed in infoboxes, templates and the category structure.
+Our task requires the use of only the links between the articles and the article text.
+In addition to in-article links, our method takes advantage of the redirect and interlanguage links, available as SQL dumps.
+
+The English corpus is based on the Wikipedia snapshot as of January 15, 2011.
+The XML files were parsed by the mwlib parser\footnote{http://code.pediapress.com}.
+The raw text was tokenized by a modified version of the Punkt sentence and word tokenizers \cite{Kiss:06}.
+Lemmatization was performed using the Wordnet Lemmatizer in NLTK \cite{Bird:09}.
+Part-of-speech tagging was performed by HunPos \cite{Halacsy:07}.
+
+%Chunking was performed by HunChunk (CITE). Experiments have shown that HunPos obtains STAT f-score on the Brown corpus; HunChunk performs at around STAT, CITE?.
 
@@
 
-\subsubsection{Sentence Filtering}
+\subsection{Sentence Filtering}
 
@@
 The solution of this problem is not obvious, and needs more investigation.
+
+% TODO: <NE>-<random kisbetűs szó> => nem NE (magyar). Lásd: "a tű nélküli mezoterápia alapja a <Nobel, PER>-díjas"
 
@@
 
-%\section*{Acknowledgments}
+\section*{Acknowledgments}
 
-%The authors are grateful to Attila Zséder and Gábor Recski for their respective work on Wikipedia parsing and {\tt hunner}.
-% + CE(A?)SAR
+% TODO: the corpora will be published in the META-SHARE repository.
+This research was supported by the CESAR project under the ICT Policy Support Programme (grant 271022). The authors are grateful to Attila Zséder for his work on Wikipedia parsing.
 
@@
 
-\bibitem[\protect\citename{Loper and Bird}2002]{Loper:02}
-Edward Loper, Steven Bird.
-\newblock 2002.
-\newblock NLTK: the Natural Language Toolkit. 
-\newblock In: {\em Proceedings of the ACL-02 Workshop on Effective tools and methodologies for teaching natural language processing and computational linguistics}, Volume 1, pages 63--70.
+\bibitem[\protect\citename{Bird et al.}2009]{Bird:09}
+Steven Bird, Ewan Klein, Edward Loper.
+\newblock 2009.
+\newblock {\em Natural Language Processing with Python}.
+\newblock O'Reilly Media Inc.
 

### 2012-04-26 paper/engNERwiki.tex

--- 2012-04-25 paper/engNERwiki.tex
@@
 It has been trained on the largest manually annotated Hungarian corpus \cite{Csendes:04}, and it handles the peculiarities of Hungarian ortography, such as the periods placed after numbers in date expressions.
-Lemmatization was performed by \texttt{hunmorph} \cite{Tron:05} and \texttt{hundisambig}, the latter being an in-house disambiguator to select the right analysis based on the word context.
+Lemmatization was performed by hunmorph \cite{Tron:05} and hundisambig, the latter being an in-house disambiguator to select the right analysis based on the word context.
 
@@
 
-For well-resourced languages, DBpedia has internationalized chapters; data dumps based on Wikipedia.
-Hungarian does not have a separate chapter.
-Instead, the Hungarian entity list comprises of the pages in the English list that have their equivalents in the Hungarian Wikipedia.
-Two consequences follow.
+For well-resourced languages, DBpedia has internationalized chapters, 
+%data dumps based on Wikipedia. 
+but not for Hungarian. Instead, the Hungarian entity list comprises of the pages in the English list that have their equivalents in the Hungarian Wikipedia. Two consequences follow.
 
@@
 As mentioned earlier, in Hungarian, apart from the first word in the sentence, only proper nouns are capitalized.
-The following concepts are not proper nouns in Hungarian, and thus are not considered as NEs according to the Szeged NER guideline: names of languages, nationalities, religions, political ideologies; adjectives derived from a NE; names of months, days, holidays; names of special events and wars.
+The following concepts are not proper nouns in Hungarian, and thus are not considered as NEs: names of languages, nationalities, religions, political ideologies; adjectives derived from a NE; names of months, days, holidays; names of special events and wars.
 

### 2012-04-26 paper/engNERwiki.tex
Email address changed to use the official .mta domain.

--- 2012-04-26 paper/engNERwiki.tex
@@
   H-1111 Lágymányosi út 11, Budapest \\
-  {\tt nemeskey.david@sztaki.hu} \\\And
+  {\tt nemeskey.david@sztaki.mta.hu} \\\And
   Eszter Simon \\
@@
 
-Wikipedia has been applied for several NLP tasks, e.g.~word sense disambiguation, ontology and thesaurus building, and question answering, as Medelyan et al.~\shortcite{Medelyan:09} point out in their fully detailed survey about mining meaning from Wikipedia.
+Wikipedia has been applied to several NLP tasks, e.g.~word sense disambiguation, ontology and thesaurus building, and question answering, as Medelyan et al.~\shortcite{Medelyan:09} point out in their fully detailed survey about mining meaning from Wikipedia.
 
@@
 Our goal is to create a large NE annotated corpus, automatically generated from Wikipedia articles.
-We follow a similar path to Nothman et al.~\shortcite{Nothman:08}, and we broke down the process into four steps:
+We follow a similar path to Nothman et al.~\shortcite{Nothman:08} and we broke down the process into four steps:
 
@@
 \item Parse Wikipedia and split articles into sentences.
-\item Label NEs according to link targets.
-\item Select sentences for inclusion in the corpus.
+\item Label named entities in the text.
+\item Select which sentences are included in the corpus.
 \end{enumerate}

### 2012-04-26 paper/engNERwiki.tex

--- 2012-04-26 paper/engNERwiki.tex
@@
 As mentioned earlier, in Hungarian, apart from the first word in the sentence, only proper nouns are capitalized.
-The following concepts are not proper nouns in Hungarian, and thus are not considered as NEs: names of languages, nationalities, religions, political ideologies; adjectives derived from a NE; names of months, days, holidays; names of special events and wars.
+The following concepts are not proper nouns in Hungarian, and thus are not considered as NEs: names of languages, nationalities, religions, political ideologies; adjectives derived from NEs; names of months, days, holidays; names of special events and wars.
 

### 2012-04-26 paper/engNERwiki.tex

@@
 As of version 3.7, the DBpedia ontology allows multiple superclasses, making a directed acyclic graph\footnote{http://blog.dbpedia.org/2011/09/11/dbpedia-37-released-including-15-localized-editions}.
-Since selecting the right superclass, and hence, CoNLL tag for classes with more than one parent cannot be reliably done automatically, the class-to-category mapping has to be determined manually.
+Since selecting the right superclass, and hence, CoNLL tag for classes with more than one parent can not be reliably done automatically, the class-to-category mapping has to be determined manually.
 The only such class in version 3.7 (\texttt{Library}) can be traced back to both \texttt{Place} and \texttt{Organisation}; its CoNLL tag is \texttt{LOC}.
@@
 Since in Hungarian only proper nouns are capitalized, we can usually decide whether a noun is proper or not based on the initial letter.
-However, this rule cannot be used if the noun is at the beginning of a sentence, as words at the beginning of sentences are always capitalized.
+However, this rule can not be used if the noun is at the beginning of a sentence, as words at the beginning of sentences are always capitalized.
 Hence, all sentences that begin with nouns have been removed from the corpus.
@@
 However, the joint common noun can modify the original sense of NE, depending on the semantics of the common noun.
-For example: in the compound \textit{SAP-szoftvercég} ['SAP software company'] the common noun does not change the labeling, but in the case of the compound \textit{WorldCom-botrány} ['WorldCom scandal'] the NE tag changes from ORG to O.
-The solution of this problem is not obvious, and needs more investigation.
+For example: in the compound \textit{Nobel-díj} ['Nobel Prize'] the common noun changes the labeling from PER to MISC, while
+%\textit{SAP-szoftvercég} ['SAP software company'] the common noun does not change the labeling, but 
+in the case of the compound \textit{WorldCom-botrány} ['WorldCom scandal'] the NE tag changes from ORG to O. The solution of this problem is not obvious, and needs more investigation.
 
@@
 The corpora are available under the Creative Commons Attribution-Sharealike 3.0 Unported License (CC-BY-SA), the same license under which the text of Wikipedia is released.
-The data files can be freely downloaded from ...\footnote{To be filled in in the camera-ready version.}.
+The data files can be freely downloaded from XXX.
+The corpora will also be distributed through META-SHARE network, which is an open, distributed facility for exchanging and sharing resources, and is one of the lines of action of META-NET, a Network of Excellence funded by the European Commission.
 
@@
 Sentence boundaries are marked by empty lines.
-The linguistic features include the lemmatized form of the word and its part-of-speech tag.
+The linguistic features include the lemmatized form of the word and its POS tag.
 Two NE tags are included with each word: the most specific DBpedia category it belongs to and the CoNLL NE tag.
@@
 The rationale behind this is that we wanted to reserve the original distribution of names in Wikipedia as much as possible.
-However, after further investigation of the NE density in our corpora and gold standard corpora (see Table \ref{size}), we decided not to include the sentences without NEs in evaluation datasets.
+However, after further investigation of the NE density in our corpora and gold standard corpora, we decided not to include the sentences without NEs in evaluation datasets.
+
+Table \ref{size} summarizes the data regarding corpus size and NE density.
+The English (enwiki) and the Hungarian Wikipedia (huwiki) corpora originally have the NE density of 5.23\% and 2.38\%, respectively.
+In comparison to the gold standard datasets (CoNLL, Szeged NER) these counts are quite low.
+It can be due to the difference between domains: newswire articles usually contain more NEs, typically ORG.
+The other reason might be that we discarded sentences containing unidentified NEs (cf.~Section \ref{sec:create}).
 

### 2012-04-27 paper/engNERwiki.tex
Changes.

@@
 Since selecting the right superclass, and hence, CoNLL tag for classes with more than one parent can not be reliably done automatically, the class-to-category mapping has to be determined manually.
-The only such class in version 3.7 (\texttt{Library}) can be traced back to both \texttt{Place} and \texttt{Organisation}; its CoNLL tag is \texttt{LOC}.
+The only such class in version 3.7, texttt{Library}, can be traced back to both \texttt{Place} and \texttt{Organisation}; its CoNLL tag is \texttt{LOC}.
 
@@
 
-Wikipedia is a very rich source of information; aside from the article text, a huge amount of data is encompassed in infoboxes, templates and the category structure.
+Wikipedia is a rich source of information; aside from the article text, a huge amount of data is encompassed in infoboxes, templates and the category structure.
 Our task requires the use of only the links between the articles and the article text.
@@
 
+\subsubsection{Nothman}
+
+First, many pages describe common nouns (Book, Aircraft), dates or other concepts that fall outside the scope of NER.
+Second, not all entities are marked with links: in a Wikipedia article, typically only the first occurrence of a particular entity is linked to the corresponding page.
+Subsequent mentions are unmarked and often incomplete -- e.g.~family names are used instead of full names.
+Finally,
+...
+Nothman et al.~\shortcite{Nothman:08}
+
 \subsubsection{Non-entity Links}
@@
 \end{description}
+
+\subsubsection{Unmarked Entities}
+
+In a Wikipedia article, typically only the first occurrence of a particular entity is linked to the corresponding page.
+Subsequent mentions are unmarked and often incomplete -- e.g.~family names are used instead of full names.
+To account for such mentions, we try to associate capitalized words in the article text with entities we have tagged thus far.
 

### 2012-04-27 paper/engNERwiki.tex
The section Unmarked Entities made shorter.

@@
 Subsequent mentions are unmarked and often incomplete -- e.g.~family names are used instead of full names.
-To account for such mentions, we try to associate capitalized words in the article text with entities we have tagged thus far.
 
-\subsubsection{Unmarked Entities}
+To account for such mentions, we apply Nothman's \shortcite{Nothman:08} solution.
+For each page, we maintain a list of entities discovered in the page so far and try to associate capitalized words in the article text with these entities.
+We augment the list with the aliases of every entity, such as titles of redirect pages that target it, the first and last names of a \texttt{PER} entity and any numbers in the name.
+If the current page is a named entity, the title and its aliases are added to the list as well; moreover, as Wikipedia usually includes the original name of foreign entities in the article text, localized versions of the title are also added to the list as aliases.
 
-Not all entities are marked with links.
-Authors may neglect to add references to their text.
-Furthermore, in a Wikipedia article, typically only the first occurrence of a particular entity is linked to the corresponding page.
-Subsequent mentions are unmarked and often incomplete -- e.g.~family names are used instead of full names.
-To account for such mentions, we try to associate capitalized words in the article text with entities we have tagged thus far.
-
-For each article, we build a set that contains all entities discovered so far, as well as their aliases:
-\begin{itemize}
-\item If the current page is a NE:
-  \begin{itemize}
-  \item The page title, with Wikipedia antics, such as a category name in parentheses removed
-  \item Title of redirect pages that target the current page
-  \item Title of the current page in other languages, as Wikipedia usually mentions the original name of foreign entities
-  \end{itemize}
-\item Entities marked with links in the text thus far
-\item Certain subsets of the above:
-  \begin{itemize}
-  \item First and last names of \texttt{PER} entities
-  \item Numbers, such as \textit{737} in \textit{Boeing 737}
-  \end{itemize}
-\end{itemize}
-
-When an unmarked entity is encountered, the set is consulted to find an item that contains it in its entirety; i.e.~all of its words, in the same order.
-If either exactly one such item is found, or the matching entities in the list all have the same NE tag, the unmarked entity is tagged accordingly; otherwise it is tagged as \texttt{UNK}.
-
-Our solution is similar to Nothman's \shortcite{Nothman:08}.
-Instead of a trie, however, we use a set, albeit with more alias types.
-We expect more precise tagging from our slightly more rigorous solution.
+Nothman's solution used a trie to store the entity list. Ours uses a set, albeit with more alias types than what he used. We expect more precise tagging from our slightly more rigorous solution.
 

### 2012-04-27 paper/engNERwiki.tex
A few words added and others removed.

@@
 Discovered entities are tagged with the CoNLL label assigned to them in the entity list extracted from DBpedia.
-If the link target is not in the entity list, or the link points to a disambiguation page, we can not determine the type of the entity, and tag it as UNK.
+If the link target is not in the entity list, or the link points to a disambiguation page, we cannot determine the type of the entity, and tag it as UNK.
 Links to redirect pages are resolved to point instead to the redirect target, after which they are handled as regular cross-references.
@@
 For each page, we maintain a list of entities discovered in the page so far and try to associate capitalized words in the article text with these entities.
-We augment the list with the aliases of every entity, such as titles of redirect pages that target it, the first and last names of a \texttt{PER} entity and any numbers in the name.
+We augment the list with the aliases of every entity, such as titles of redirect pages that target it, the first and last names in case of a \texttt{PER} entity and any numbers in the name.
 If the current page is a named entity, the title and its aliases are added to the list as well; moreover, as Wikipedia usually includes the original name of foreign entities in the article text, localized versions of the title are also added to the list as aliases.
@@
 Furthermore, there are many incomplete sentences in the Wikipedia text: image captions, enumerations items, contents of table cells, etc.
-On the one hand, these sentence fragments may prove to be of too low quality to be of any use in the traditional NER task.
+On the one hand, these sentence fragments may be of too low quality to be of any use in the traditional NER task.
 On the other hand, they could prove to be invaluable when training a NER tagger for User Generated Content, which is known to be noisy and fragmented.

### 2012-04-27 paper/engNERwiki.tex
Changes.

@@
 To avoid the inevitable classification errors, we obtain entity type information from the DBpedia knowledge base \cite{Bizer:09}.
-This enables us to have high precision information about entity types at the expense of recall: of the 3,903,467 English Wikipedia pages, 1,470,293 are covered by DBpedia.
+This enables us to have high precision information about entity types at the expense of recall: of the 3,903,467 English Wikipedia pages, 1,470,293 are covered by DBpedia\footnote{as of 18 March, 2012.}.
+% TODO as of?
 
@@
 
-Our approach to entity identification relies on Wikipedia cross-references found in the article text.
+Our approach is based on the work of Nothman et al.~\shortcite{Nothman:08}.
+The Wikipedia cross-references found in the article text are used to identification entities.
 We assume that individual Wikipedia articles describe NEs.
 A link to an article can then be perceived as a mapping that identifies its anchor text with a particular NE.
+% TODO identifies -> ? vagy with -> as
 
@@
 Finally, sentences with UNK links in them are removed from the corpus.
+%  TODO , and -> and?
 
 The following sub-sections describe how the method explained above can be improved to increase precision, sentence coverage and to account for peculiarities in the CoNLL guidelines.
-
-\subsubsection{Nothman}
-
-First, many pages describe common nouns (Book, Aircraft), dates or other concepts that fall outside the scope of NER.
-Second, not all entities are marked with links: in a Wikipedia article, typically only the first occurrence of a particular entity is linked to the corresponding page.
-Subsequent mentions are unmarked and often incomplete -- e.g.~family names are used instead of full names.
-Finally,
-...
-Nothman et al.~\shortcite{Nothman:08}
 
@@
 Nothman's solution used a trie to store the entity list. Ours uses a set, albeit with more alias types than what he used. We expect more precise tagging from our slightly more rigorous solution.
+% TODO: change at least use
 
@@
 
-To annotate our corpus with NE tags, we chose to follow the annotation guidelines of the largest human-annotated NER corpus for Hungarian, Szeged NER corpus \cite{Szarvas:06}.
+To annotate our corpus with NE tags, we chose to follow the annotation guidelines of the largest human-annotated NER corpus for Hungarian, the Szeged NER corpus \cite{Szarvas:06}.
 It is similar to CoNLL standards: contains newswire texts, comprises ca.~200,000 tokens, and is annotated with NE class labels in line with the CoNLL annotation scheme.
@@
 %\textit{SAP-szoftvercég} ['SAP software company'] the common noun does not change the labeling, but 
-in the case of the compound \textit{WorldCom-botrány} ['WorldCom scandal'] the NE tag changes from ORG to O. The solution of this problem is not obvious, and needs more investigation.
+in the case of the compound \textit{WorldCom-botrány} ['WorldCom scandal'] the NE tag changes from ORG to O. The solution to this problem is not obvious, and needs more investigation.
 

### 2012-04-27 paper/engNERwiki.tex

--- 2012-04-27 paper/engNERwiki.tex
@@
 Subsequent mentions are unmarked and often incomplete -- e.g.~family names are used instead of full names.
+To account for such mentions, we try to associate capitalized words in the article text with entities we have tagged thus far.
 
-To account for such mentions, we apply Nothman's \shortcite{Nothman:08} solution.
-For each page, we maintain a list of entities discovered in the page so far and try to associate capitalized words in the article text with these entities.
-We augment the list with the aliases of every entity, such as titles of redirect pages that target it, the first and last names of a \texttt{PER} entity and any numbers in the name.
-If the current page is a named entity, the title and its aliases are added to the list as well; moreover, as Wikipedia usually includes the original name of foreign entities in the article text, localized versions of the title are also added to the list as aliases.
+\subsubsection{Unmarked Entities}
 
-Nothman's solution used a trie to store the entity list. Ours uses a set, albeit with more alias types than what he used. We expect more precise tagging from our slightly more rigorous solution.
+Not all entities are marked with links.
+Authors may neglect to add references to their text.
+Furthermore, in a Wikipedia article, typically only the first occurrence of a particular entity is linked to the corresponding page.
+Subsequent mentions are unmarked and often incomplete -- e.g.~family names are used instead of full names.
+To account for such mentions, we try to associate capitalized words in the article text with entities we have tagged thus far.
+
+For each article, we build a set that contains all entities discovered so far, as well as their aliases:
+\begin{itemize}
+\item If the current page is a NE:
+  \begin{itemize}
+  \item The page title, with Wikipedia antics, such as a category name in parentheses removed
+  \item Title of redirect pages that target the current page
+  \item Title of the current page in other languages, as Wikipedia usually mentions the original name of foreign entities
+  \end{itemize}
+\item Entities marked with links in the text thus far
+\item Certain subsets of the above:
+  \begin{itemize}
+  \item First and last names of \texttt{PER} entities
+  \item Numbers, such as \textit{737} in \textit{Boeing 737}
+  \end{itemize}
+\end{itemize}
+
+When an unmarked entity is encountered, the set is consulted to find an item that contains it in its entirety; i.e.~all of its words, in the same order.
+If either exactly one such item is found, or the matching entities in the list all have the same NE tag, the unmarked entity is tagged accordingly; otherwise it is tagged as \texttt{UNK}.
+
+Our solution is similar to Nothman's \shortcite{Nothman:08}.
+Instead of a trie, however, we use a set, albeit with more alias types.
+We expect more precise tagging from our slightly more rigorous solution.
 
@@
 
-The English train corpus was evaluated against itself and a manually annotated English corpus.
+The English Wikipedia corpus was evaluated against itself and a manually annotated English corpus.
 Since the filtered English Wikipedia corpus, containing only the sentences with NEs, is still very huge, and we are limited by time and memory, our experiments were performed with a portion of 3.5 million tokens (the size of our filtered Hungarian corpus).
@@
 Nothman et al.~\shortcite{Nothman:08} showed that each set of gold standard training data performs better on corresponding test sets than on test sets from other sources.
-The situation here is similar: the NE tagger trained on Wikipedia does not achieve as high performance tested against CoNLL test set as one trained on the own train set (see Table \ref{enresults}).
+The situation here is similar (see Table \ref{enresults} for results): the NE tagger trained on Wikipedia does not achieve as high performance tested against CoNLL test set (enwiki-CoNLL) as one trained on the own train set (enwiki-enwiki).
 
 However, Wikipedia-derived corpora can also be used for improving NER accuracy in other ways.
-First, we collected gazetteer lists from the corpus for each NE category, which improved the overall F-measure given to the NE tagger training and testing on CoNLL dataset.
-A second trial was labeling the CoNLL datasets by the model trained on Wikipedia corpus, and giving these labels as extra features to the next CoNLL train.
+First, we collected gazetteer lists from the corpus for each NE category, which improved the overall F-measure given to the NE tagger training and testing on CoNLL dataset (CoNLL with wikilists).
+A second trial was labeling the CoNLL datasets by the model trained on Wikipedia corpus, and giving these labels as extra features to the next CoNLL train (CoNLL with wikitags).
 Both methods result in improved F-measure on CoNLL test set.
@@
 Hungarian results are similar to the English ones (see Table \ref{huresults}), the only difference is that F-measures for Hungarian are significantly higher.
-It can be due to the fact that the MISC category for Hungarian contains less types of names, thus the inconsistency of this class is smaller (see section \ref{sec:hun}).
+It can be due to the fact that the MISC category for Hungarian contains less types of names, thus the inconsistency of this class is smaller (cf.~Section \ref{sec:hun}).
 In contrast to the CoNLL corpus, the Szeged NER corpus was accurately annotated with an inter-annotator agreement over 99\%.
@@
 In contrast to the methods used so far for automatic annotation of NEs in Wikipedia texts, we applied a new approach, namely mapping DBpedia ontology classes to standard CoNLL NE tags, and assigning them to Wikipedia entities.
-Following Nothman \shortcite{Nothman:08}, the process can be divided into four main steps:  classifying all articles into entity classes; splitting Wikipedia articles into sentences; labeling NEs according to link targets; and selecting sentences for inclusion in the corpus.
+Following Nothman \shortcite{Nothman:08}, the process can be divided into four main steps:  classifying Wikipedia articles into entity classes; parsing Wikipedia and splitting articles into sentences; labeling NEs in the text; and selecting sentences for inclusion in the corpus.
 
-Thanks for the huge amount of Wikipedia articles, the possibility of building large enough corpora is given for less resourced languages, like Hungarian, as well.
+Thanks to the huge amount of Wikipedia articles, the possibility of building large enough corpora is given for less resourced languages, like Hungarian, as well.
 Due to the particularities of Hungarian, some steps are slightly different, and special linguistics phenomena pose several problems related to the NER task to solve.
@@
 We showed that gazetteer lists extracted from our corpora, and training with extra features given by the model trained on our corpora, improves F-measure.
-Moreover, our Hungarian corpus can serve as a training corpus for more general domains.
-%than the classic newswire.
+Moreover, our Hungarian corpus can serve as a training corpus for more general domains, than the classic newswire.
 
@@
 % TODO: the corpora will be published in the META-SHARE repository.
-This research was supported by the CESAR project under the ICT Policy Support Programme (grant 271022). The authors are grateful to Attila Zséder for his work on Wikipedia parsing.
+This research was supported by the CESAR project under the ICT Policy Support Programme (grant no. 271022). The authors are grateful to Attila Zséder for his work on Wikipedia parsing.
 
@@
 %\newblock In: {\em Proceedings of the 2009 Workshop on the People's Web Meets NLP, ACL-IJCNLP}, pages 10--18.
+
+\bibitem[\protect\citename{Bird et al.}2009]{Bird:09}
+Steven Bird, Ewan Klein, Edward Loper.
+\newblock 2009.
+\newblock {\em Natural Language Processing with Python}.
+\newblock O'Reilly Media Inc.
 
@@
 
-\bibitem[\protect\citename{Bird et al.}2009]{Bird:09}
-Steven Bird, Ewan Klein, Edward Loper.
-\newblock 2009.
-\newblock {\em Natural Language Processing with Python}.
-\newblock O'Reilly Media Inc.
-
 \bibitem[\protect\citename{Medelyan  \bgroup et al.\egroup}2009]{Medelyan:09}
@@
 \bibitem[\protect\citename{Nothman \bgroup et al.\egroup}2009]{Nothman:09}
-Joel Nothman and Tara Murphy and James R. Curran.
+Joel Nothman, Tara Murphy and James R. Curran.
 \newblock 2009.

### 2012-04-27 paper/engNERwiki.tex
Added the download URL; changed a few things here and there.

@@
 
-Our goal is to create a large NE annotated corpus, automatically generated from Wikipedia articles.
-We follow a similar path to Nothman et al.~\shortcite{Nothman:08} and we broke down the process into four steps:
+Our goal is to create a large NE annotated corpus, automatically generated from Wikipedia articles. Following Nothman et al.~\shortcite{Nothman:08}, we broke down the process into four steps:
 
 \begin{enumerate}
-\item Classify Wikipedia articles into entity classes.
-\item Parse Wikipedia and split articles into sentences.
-\item Label named entities in the text.
-\item Select which sentences are included in the corpus.
+\item Classify Wikipedia articles into entity classes
+%\item Parse Wikipedia and split articles into sentences
+\item Parse Wikipedia into sentences
+\item Label named entities in the text
+\item Select the sentences to include in the corpus
 \end{enumerate}
@@
 
-The following sub-sections describe how the method explained above can be improved to increase precision, sentence coverage and to account for peculiarities in the CoNLL guidelines.
+The following sub-sections describe how the method explained above can be improved to increase precision, sentence coverage and to account for peculiarities in English orthography and the CoNLL guidelines.
 
@@
 The corpora are available under the Creative Commons Attribution-Sharealike 3.0 Unported License (CC-BY-SA), the same license under which the text of Wikipedia is released.
-The data files can be freely downloaded from XXX.
+The data files can be freely downloaded from \texttt{http://hlt.sztaki.hu}.
 The corpora will also be distributed through META-SHARE network, which is an open, distributed facility for exchanging and sharing resources, and is one of the lines of action of META-NET, a Network of Excellence funded by the European Commission.

### 2012-04-27 paper/engNERwiki.tex
Kornai javításai

--- 2012-04-27 paper/engNERwiki.tex
@@
   Hungarian Academy of Sciences \\
-  H-1111 Lágymányosi út 11, Budapest \\
+  H-1111 Kende utca 13-17, Budapest \\
   {\tt nemeskey.david@sztaki.mta.hu} \\\And
@@
 
-Supervised Named Entity Recognizers require large amount of annotated texts.
-However, manual annotation is a highly costly procedure, thus reducing the annotation cost is essential.
-We present a fully automatic method to build NE annotated corpora from Wikipedia.
-In contrast to recent work, we apply a new method, which maps the DBpedia classes into CoNLL NE types.
-Since our method is mainly language-independent, we used it for generating corpora for English and Hungarian, as well.
-The corpora are freely available.
+Supervised Named Entity Recognizers require large amounts of annotated
+text. Since manual annotation is a highly costly procedure, reducing the
+annotation cost is essential. We present a fully automatic method to build NE
+annotated corpora from Wikipedia. In contrast to recent work, we apply a new
+method, which maps the DBpedia classes into CoNLL NE types. Since our method
+is mainly language-independent, we used it to generate corpora for English
+and Hungarian. The corpora are freely available.
 
@@
 Although there is a general agreement in the NER community about the inclusion of temporal expressions and some numerical expressions, the most studied types are names of persons, locations and organizations.
-The fourth type, called "miscellaneous'', was introduced in the CoNLL NER tasks in 2002 \cite{Tjong:02} and 2003 \cite{Tjong:03}, and includes proper names falling outside the three classic types.
+The fourth type, called ``miscellaneous'', was introduced in the CoNLL NER tasks in 2002 \cite{Tjong:02} and 2003 \cite{Tjong:03}, and includes proper names falling outside the three classic types.
 Since then, MUC and CoNLL datasets and annotation schemes have been the major standards applied in the field of NER.
 
-The standard datasets are highly domain-specific (mostly newswire), and are restricted in size, as well. 
-%Thus they are not large enough to train and test robust NE taggers on them. Moreover, 
-Merging these datasets to get a bigger training corpus faces NLP researchers with the problem of combining different tagsets and annotation schemes.
-However, to build a robust supervised machine learning system for NER, large enough, accurately annotated corpora are essential.
-Annotating large amount of texts with linguistic information is a time-consuming, highly skilled and delicate work.
-Mainly due to this fact, one of the main challenges in NLP is reducing the annotation cost.
+The standard datasets are highly domain-specific (mostly newswire) and are
+restricted in size. Researchers attempting to merge these datasets to get a bigger training corpus are faced with the
+problem of combining different tagsets and annotation schemes. Manually
+annotating large amounts of text with linguistic information is a
+time-consuming, highly skilled and delicate job, but large, accurately
+annotated corpora are essential for building robust supervised machine
+learning NER systems. Therefore, reducing the annotation cost is a key
+challenge.
 
-One approach to reduce the manual intervention required in language resource production, and thus ultimately production costs is automatically generating resources.
-Another approach to achieve this aim is using collaborative annotation and/or collaboratively constructed resources, such as Wikipedia, Wiktionary or Linked Open Data.
-Researchers apply the Web-as-corpus approach, often dealing with such datasets as Google Web 1T 5-gram collection or DBpedia.
+One approach is to generate the resources automatically, another one is to use
+collaborative annotation and/or collaboratively constructed resources, such as
+Wikipedia, Wiktionary, Linked Open Data, or DBpedia. In this paper we combine
+these approaches by automatically generating freely available NE tagged
+corpora from Wikipedia.
 
-%The first approach is the application of unsupervised and semi-supervised learning techniques.
-Researchers following this approach try to not only reduce, but avoid the annotation cost with minimizing the need for annotated data.
-These methods can be applied to any language or genre for which adequate raw text resources are available.
-
-In this paper we present a kind of combination of these approaches, namely automatically generating resources and using collaboratively constructed resources.
-We built freely available NE tagged corpora for English and Hungarian -- fully automatically generated from Wikipedia.
-%Our goal was to create gold standard corpora which can serve as train, devel and test datasets for automatic NE taggers.
-
-The paper is structured as follows.
-In Section~\ref{sec:related} we give an overview of related work.
-Section \ref{sec:create} contains a description of our method, while Section \ref{sec:hun} shows how it is applied to Hungarian.
-The corpus format is described in Section~\ref{sec:data}.
-In Section \ref{sec:eval} we present experiments and results on the newly generated datasets.
-Section \ref{sec:conclusion} concludes the paper with a summary.
-
-%Our method is introduced in Section~\ref{sec:create}, and showed how it is applied to Hungarian in Section~\ref{sec:hun}.
-The corpus format is described in Section~\ref{sec:data}.
-Section~\ref{sec:eval} contains evaluations and Section~\ref{sec:conclusion} our conclusion.
+The paper is structured as follows. In
+Section~\ref{sec:related} we give an overview of related work. Section
+\ref{sec:create} contains a description of our method, and Section
+\ref{sec:hun} shows how it is applied to Hungarian. The corpus format is
+described in Section~\ref{sec:data}. In Section \ref{sec:eval} we present
+experiments and results on the newly generated datasets. Section
+\ref{sec:conclusion} concludes the paper with a summary.
 
@@
 \label{sec:related}
-% Wikify!
 
-Wikipedia is a free, multilingual Internet encyclopedia, written collaboratively by volunteers\footnote{http://en.wikipedia.org/wiki/Wikipedia}.
-It is a goldmine of information; at the time of writing, Wikipedia contains about 21 million interlinked articles.
-Of these, 3,903,467 are English, and 212,120 are Hungarian.
+Wikipedia (WP, see {\tt http://wikipedia.org}), a free
+multilingual Internet encyclopedia, written collaboratively by volunteers, is
+a goldmine of information: at the time of writing, WP contains about 21
+million interlinked articles. Of these, 3,903,467 are English, and 212,120 are
+Hungarian. WP has been applied to several NLP tasks such as word sense
+disambiguation, ontology and thesaurus building, and question answering (see
+Medelyan et al.~\shortcite{Medelyan:09} for a survey). It is recognized as one
+of the largest available collections of entities, and also as a resource that
+can improve the accuracy of NER. The most obvious utilization of WP for NER is
+extracting gazetteers containing person names, locations or organizations
+(e.g.~Toral and Mu\~noz \shortcite{Toral:06}). Creating dictionaries of
+entities is also a common step of NE disambiguation
+\cite{Bunescu:06,Cucerzan:07}. Both supervised and unsupervised NER systems
+use such lists, see e.g.~Nadeau et al.~\shortcite{Nadeau:06} The knowledge
+embodied in WP may also be incorporated in NER learning as features,
+e.g.~Kazama and Torisawa \shortcite{Kazama:07} showed that automatic
+extraction of category labels from WP improves the accuracy of a supervised NE
+tagger.
 
-Wikipedia has been applied to several NLP tasks, e.g.~word sense disambiguation, ontology and thesaurus building, and question answering, as Medelyan et al.~\shortcite{Medelyan:09} point out in their fully detailed survey about mining meaning from Wikipedia.
+Another approach to improve NER with WP is the automatic creation of training
+data. Richman and Schone \shortcite{Richman:08} built corpora for less
+commonly taught languages annotated with NE tags. They used the inherent
+category structure of WP to determine the NE type of a proposed
+entity. Nothman et al.~\shortcite{Nothman:08} used a similar method to create
+a NE annotated text in English. They transformed the WP links into NE
+annotations by classifying the target articles into standard entity
+classes. Their approach to classification is based primarily on category head
+nouns and the opening sentences of articles where definitions are often given.
 
-Wikipedia is recognized as one of the largest available collection of entities, and also as a resource that can improve the accuracy of NER.
-The most obvious utilization of Wikipedia for NER is extracting gazetteers containing person names, locations or organizations (e.g.~Toral and Mu\~noz \shortcite{Toral:06}).
-Creating dictionaries of entities is also a common step of NE disambiguation \cite{Bunescu:06,Cucerzan:07}.
-Not only supervised systems use lists generated from Wikipedia: Nadeau et al.~\shortcite{Nadeau:06} use similar lists in an unsupervised NER system.
-
-The knowledge in Wikipedia may be incorporated as features in NER learning, e.g.~Kazama and Torisawa \shortcite{Kazama:07} showed that automatic extraction of category labels from Wikipedia improves the accuracy of a supervised NE tagger.
-
-A third approach to improve NER with Wikipedia is the automatic creation of training data.
-Richman and Schone \shortcite{Richman:08} built corpora for less commonly taught languages annotated with NE tags.
-They used the inherent category structure of Wikipedia to determine the NE type of a proposed entity.
-Nothman et al.~\shortcite{Nothman:08} used a similar method to create a NE annotated text in English.
-They transformed the Wikipedia links into NE annotations by classifying the target articles into standard entity classes.
-Their approach to classification is based primarily on category head nouns and definitional opening sentences of articles.
-
-In contrast to the aforementioned methods, we applied a new approach to recognize and classify NEs in corpora generated from Wikipedia.
-We mapped the DBpedia ontology classes to standard NE tags and assigned them to Wikipedia entities (see more details in Section \ref{dbpedia}).
-
-To our knowledge, such type of automatically built corpora are not freely available, with the one exception of the Semantically Annotated Snapshot of the English Wikipedia \cite{Zaragoza:07}.
-It provides a wide range of linguistic information: POS tags, dependency labels, WordNet super senses and NE annotation according to WSJ and CoNLL tagsets.
-However, they followed a different way: the NEs were tagged by high quality open source softwares.
-In contrast, our approach relies on the manual judgement of thousands of Wikipedia volunteers.
-
-Thanks to the huge amount of Wikipedia articles, the possibility of building large enough 
-corpora is given for less resourced languages, as well.
-Since our method is largely language-independent, it can be easily applied for languages which are structurally differing from English.
-Hungarian is a highly agglutinative language, with free word order and other particularities detailed later in Section \ref{sec:hun}.
-There are human annotated CoNLL-style datasets, but the one presented here is the first automatically NE annotated corpus for Hungarian.
+Our approach to recognize and classify NEs in corpora generated from WP was to
+map the DBpedia ontology classes to standard NE tags and assign these to WP
+entities (see more details in Section \ref{dbpedia}). Except for the
+Semantically Annotated Snapshot of the English WP (SASWP) \cite{Zaragoza:07},
+no such automatically built corpora are freely available. SASWP provides a
+wide range of linguistic information: POS tags, dependency labels, WordNet
+super senses and NE annotation according to WSJ and CoNLL tagsets. Even though
+the SASWP NEs were tagged by the best available open source taggers, the tags 
+provided here, being based on the manual judgement of thousands of WP
+volunteers, are more reliable. Given the huge number of WP articles we can
+build sufficiently large corpora for less resourced languages as well, as our
+method is largely language-independent. We demonstrate this on Hungarian, a
+highly agglutinative language, with free word order and other typological 
+characteristics detailed later in Section \ref{sec:hun}.
+There are smaller, manually annotated CoNLL-style datasets, but the one presented here is the first automatically NE annotated corpus for Hungarian.
 
@@
 
-Our goal is to create a large NE annotated corpus, automatically generated from Wikipedia articles.
-We follow a similar path to Nothman et al.~\shortcite{Nothman:08} and we broke down the process into four steps:
+Our goal is to create a large NE annotated corpus, automatically generated
+from WP articles. We followed a similar path to Nothman et
+al.~\shortcite{Nothman:08} and broke down the process into four steps:
 
 \begin{enumerate}
-\item Classify Wikipedia articles into entity classes.
-\item Parse Wikipedia and split articles into sentences.
+\item Classify WP articles into entity classes.
+\item Parse WP and split articles into sentences.
 \item Label named entities in the text.
-\item Select which sentences are included in the corpus.
+\item Select the sentences for inclusion in the corpus.
 \end{enumerate}
@@
 In this section, we describe how these steps were implemented.
-This section explains the general approach and its execution for English; the next section describes how the idea is adapted to Hungarian.
+This section explains the general approach and its execution for English; Section \ref{sec:hun} describes how the idea is adapted to Hungarian.
 
@@
 
-Many authors, such as Kazama and Torisawa \shortcite{Kazama:07} and Nothman et al.~\shortcite{Nothman:08} used semi-supervised methods based on Wikipedia categories and text to classify articles into NE types.
-To avoid the inevitable classification errors, we obtain entity type information from the DBpedia knowledge base \cite{Bizer:09}.
-This enables us to have high precision information about entity types at the expense of recall: of the 3,903,467 English Wikipedia pages, 1,470,293 are covered by DBpedia\footnote{as of 18 March, 2012.}.
-% TODO as of?
+Many authors, such as Kazama and Torisawa \shortcite{Kazama:07} and Nothman et
+al.~\shortcite{Nothman:08} used semi-supervised methods based on WP categories
+and text to classify articles into NE types. To avoid the inevitable
+classification errors, we obtain entity type information from the DBpedia
+knowledge base \cite{Bizer:09}, which presents type, properties, home pages,
+etc. information about pages in WP in structured form. With DBpedia we have
+high precision information about entity types at the expense of recall: of the
+3,903,467 English WP pages, 1,470,293 are covered by DBpedia (as of 18 March,
+2012).
 
-DBpedia is an open database that presents information about entities in Wikipedia in structured form.
-Each entity corresponds to a Wikipedia article.
-DBpedia provides various information about entities including type, properties, home pages, etc.
-Here we limit our interest to the type system.
+The types in DBpedia are organized into a class hierarchy, available as an OWL\footnote{http://www.w3.org/TR/owl-ref/}
+ontology containing 320 frequent
+entity categories, arranged into a taxonomy under the base class
+\texttt{owl:Thing}. Most of the classes belong to the 6 largest
+sub-hierarchies: \texttt{Person}, \texttt{Organisation}, \texttt{Event},
+\texttt{Place}, \texttt{Species} and \texttt{Work}. The taxonomy is rather
+flat: the top level contains 44 classes and there are several nodes with a
+branching factor of 20.
 
-The types in DBpedia are organized into a class hierarchy, available as an OWL\footnote{http://www.w3.org/TR/owl-ref/} ontology.
-The ontology contains the 320 most frequent entity categories, arranged into a taxonomy under the base class \texttt{owl:Thing}.
-Most of the classes belong to the 6 largest sub-hierarchies (here named after their topmost classes): \texttt{Person}, \texttt{Organisation}, \texttt{Event}, \texttt{Place}, \texttt{Species} and \texttt{Work}.
-The taxonomy is rather flat: the top level contains 44 classes and there are several nodes with a branching factor of 20.
-
-The type of entities is extracted automatically from Wikipedia categories.
-However, the mapping between Wikipedia categories and classes in the DBpedia ontology is manually defined.
-This, together with the fact that the existence of the reference ontology prevents the proliferation of categories observable in Wikipedia \cite{Bizer:09}, ensures that type information in DBpedia can be considered gold quality.
+The type of entities is extracted automatically from WP categories.
+However, the mapping between WP categories and classes in the DBpedia ontology is manually defined.
+This, together with the fact that the existence of the reference ontology prevents the proliferation of categories observable in WP \cite{Bizer:09}, ensures that type information in DBpedia can be considered gold quality.
 
@@
 
-While the correspondence described above holds for most classes in the sub-hierarchies, there are some exceptions.
-For instance, the class \texttt{SportsLeague} is part of the \texttt{Organisation} sub-hierarchy, but according to the CoNLL annotation scheme, they should be tagged as \texttt{MISC}.
-
-To avoid such kind of misclassification, 
+While the correspondence described above holds for most classes in the
+sub-hierarchies, there are some exceptions. For instance, the class
+\texttt{SportsLeague} is part of the \texttt{Organisation} sub-hierarchy, but
+according to the CoNLL annotation scheme, they should be tagged as
+\texttt{MISC}.  To avoid misclassification,
 %as well as the miscellaneous classes outside the main hierarchies, we have introduced a mapping mechanism. 
@@
 
-As of version 3.7, the DBpedia ontology allows multiple superclasses, making a directed acyclic graph\footnote{http://blog.dbpedia.org/2011/09/11/dbpedia-37-released-including-15-localized-editions}.
-Since selecting the right superclass, and hence, CoNLL tag for classes with more than one parent can not be reliably done automatically, the class-to-category mapping has to be determined manually.
-The only such class in version 3.7, texttt{Library}, can be traced back to both \texttt{Place} and \texttt{Organisation}; its CoNLL tag is \texttt{LOC}.
+As of version 3.7, the DBpedia ontology allows multiple superclasses, making a
+directed acyclic
+graph\footnote{http://blog.dbpedia.org/2011/09/11/dbpedia-37-released-including-15-localized-editions}. Since
+selecting the right superclass, and hence, CoNLL tag for classes with more
+than one parent, cannot be reliably done automatically, the class-to-category
+mapping had to be determined manually. The only such class in version 3.7,
+\texttt{Library}, can be traced back to both \texttt{Place} and
+\texttt{Organisation}; its CoNLL tag is \texttt{LOC}.  Using the mapping thus
+created, we compile a list that contains all entities in DBpedia tagged with
+the appropriate CoNLL category.
 
-%The full mapping can be found in Appendix~\ref{sec:appa}.
-
-Using the mapping described above, we compile a list that contains all entities in DBpedia tagged with the appropriate CoNLL category.
-
-We would like to note that even though we use CoNLL categories, our method can be modified to work with any tagset compatible with the DBpedia ontology.
-Indeed, the DBpedia classes define a NE tagset themselves.
-We leave the exploration of these possibilities for future work.
+We note here that our method can be trivially modified to work with any tagset
+compatible with the DBpedia ontology (indeed, the DBpedia classes define a NE
+tagset themselves), but we leave the exploration of these possibilities for
+future work.
 
@@
 
-Wikipedia is a rich source of information; aside from the article text, a huge amount of data is encompassed in infoboxes, templates and the category structure.
-Our task requires the use of only the links between the articles and the article text.
-In addition to in-article links, our method takes advantage of the redirect and interlanguage links, available as SQL dumps.
-
-The English corpus is based on the Wikipedia snapshot as of January 15, 2011.
-The XML files were parsed by the mwlib parser\footnote{http://code.pediapress.com}.
-The raw text was tokenized by a modified version of the Punkt sentence and word tokenizers \cite{Kiss:06}.
-Lemmatization was performed using the Wordnet Lemmatizer in NLTK \cite{Bird:09}.
-Part-of-speech tagging was performed by HunPos \cite{Halacsy:07}.
+WP is a rich source of information; in addition to the article text, a huge
+amount of data is embedded in infoboxes, templates, and the category
+structure. Our task requires only the links between the articles and the
+article text. In addition to in-article links, our method takes advantage of
+the redirect and interlanguage links, available as SQL dumps. The English
+corpus is based on the WP snapshot as of January 15, 2011. The XML files were
+parsed by the mwlib parser\footnote{http://code.pediapress.com}, the raw
+text was tokenized by a modified version of the Punkt sentence and word
+tokenizers \cite{Kiss:06}. For lemmatization we used the Wordnet Lemmatizer in
+NLTK \cite{Bird:09}, and for part-of-speech tagging the HunPOS tagger
+\cite{Halacsy:07}.
 
@@
 
-In order to automatically prepare sentences where NEs are accurately tagged, two tasks need to be performed: identifying entities in the sentence and tagging them with the correct tag.
-Sentences, for which accurate tagging could not be accomplished, must be removed from the corpus.
+In order to automatically prepare sentences where NEs are accurately tagged,
+two tasks need to be performed: identifying entities in the sentence and
+tagging them with the correct tag. Sentences for which accurate tagging could
+not be accomplished must be removed from the corpus. Our approach is based on
+the work of Nothman et al.~\shortcite{Nothman:08}. The WP cross-references
+found in the article text are used to identify entities. We assume that
+individual WP articles describe NEs. A link to an article can then be
+perceived as a mapping that identifies its anchor text with a particular NE.
 
-Our approach is based on the work of Nothman et al.~\shortcite{Nothman:08}.
-The Wikipedia cross-references found in the article text are used to identification entities.
-We assume that individual Wikipedia articles describe NEs.
-A link to an article can then be perceived as a mapping that identifies its anchor text with a particular NE.
-% TODO identifies -> ? vagy with -> as
+The discovered entities are tagged with the CoNLL label assigned to them in
+the entity list extracted from DBpedia. If the link target is not in the
+entity list, or the link points to a disambiguation page, we cannot determine
+the type of the entity, and tag it as UNK for subsequent removal from the
+corpus. Links to redirect pages are resolved to point instead to the redirect
+target, after which they are handled as regular cross-references. Finally, sentences with UNK links in them are removed from the corpus.
 
-Discovered entities are tagged with the CoNLL label assigned to them in the entity list extracted from DBpedia.
-If the link target is not in the entity list, or the link points to a disambiguation page, we cannot determine the type of the entity, and tag it as UNK.
-Links to redirect pages are resolved to point instead to the redirect target, after which they are handled as regular cross-references.
-Finally, sentences with UNK links in them are removed from the corpus.
-%  TODO , and -> and?
-
-The following sub-sections describe how the method explained above can be improved to increase precision, sentence coverage and to account for peculiarities in the CoNLL guidelines.
+The following sub-sections describe how the method explained above can be
+improved to increase precision, sentence coverage and to account for
+peculiarities in the CoNLL guidelines.
 
@@
 
-Strictly speaking, our original assumption of equating Wikipedia articles with NEs is not valid: many pages describe common nouns (Book, Aircraft), calendar-related concepts (March 15, 2007), or other concepts that fall outside the scope of NER.
+Strictly speaking, our original assumption of equating WP articles with NEs is not valid: many pages describe common nouns (Book, Aircraft), calendar-related concepts (March 15, 2007), or other concepts that fall outside the scope of NER.
 To increase sentence coverage, we modified the algorithm to prevent it from misclassifying links to these pages as unknown entities and discarding the sentence.
@@
 \begin{description}
-\item[Common nouns] Links to common nouns are filtered by POS tags; if a link contains no \texttt{NNP}s, it is ignored.
-\item[Dates] Time expressions require special attention, because dates and months are often linked to the respective Wikipedia pages.
-We circumvented this problem by compiling a list of calendar-related pages and adding them to the main entity list tagged with the CoNLL category \texttt{O}.
-\item[Lowercase links] Entities are sometimes referred to by common nouns, such as \textit{republic} to \textit{Roman Republic}, \texttt{LOC}, which results in a lowercase link.
-These references are not considered NEs and are ignored.
+\item[Common noun] links are filtered by POS tags; if a link contains no \texttt{NNP}s, it is ignored.
+\item[Time expression] links require special attention, because dates and
+  months are often linked to the respective WP pages. We circumvented this
+  problem by compiling a list of calendar-related pages and adding them to the
+  main entity list tagged with the CoNLL category \texttt{O}.
+\item[Lowercase] links for entities referred to by common nouns, such as \textit{republic} to \textit{Roman Republic} are not considered NEs and are ignored.
 \end{description}
@@
 
-In a Wikipedia article, typically only the first occurrence of a particular entity is linked to the corresponding page.
-Subsequent mentions are unmarked and often incomplete -- e.g.~family names are used instead of full names.
-
-To account for such mentions, we apply Nothman's \shortcite{Nothman:08} solution.
-For each page, we maintain a list of entities discovered in the page so far and try to associate capitalized words in the article text with these entities.
-We augment the list with the aliases of every entity, such as titles of redirect pages that target it, the first and last names in case of a \texttt{PER} entity and any numbers in the name.
-If the current page is a named entity, the title and its aliases are added to the list as well; moreover, as Wikipedia usually includes the original name of foreign entities in the article text, localized versions of the title are also added to the list as aliases.
-
-Nothman's solution used a trie to store the entity list. Ours uses a set, albeit with more alias types than what he used. We expect more precise tagging from our slightly more rigorous solution.
-% TODO: change at least use
+In a WP article, typically only the first occurrence of a particular entity is
+linked to the corresponding page. Subsequent mentions are unmarked and often
+incomplete -- e.g.~family names are used instead of full names. To account
+for such mentions, we apply Nothman's \shortcite{Nothman:08} solution. For
+each page, we maintain a list of entities discovered in the page so far and
+try to associate capitalized words in the article text with these entities. We
+augment the list with the aliases of every entity, such as titles of redirect
+pages that target it, the first and last names in case of a \texttt{PER}
+entity and any numbers in the name. If the current page is a NE, the
+title and its aliases are added to the list as well; moreover, as WP usually
+includes the original name of foreign entities in the article text, localized
+versions of the title are also added to the list as aliases. Nothman's
+solution used a trie to store the entity list, while we use a set, with more
+alias types than what he used. We expect more precise tagging from our
+slightly more rigorous solution.
 
@@
 We base the decision on the POS tag of the first word: if it is \texttt{NNP}, we tag it as \texttt{UNK}; otherwise, \texttt{O}.
-\item[Reference cleansing] Page titles and anchor texts may contain more than just the entity name.
-Personal titles are part of the entity name in Wikipedia, but not in CoNLL, and punctuation marks around the entity may become part of the link by mistake.
-We tag all punctuation marks after the entity name as \texttt{O}.
-
-To tackle personal titles, we extracted a list from the Wikipedia page \textit{List of titles}, which contains titles in many languages.
-We manually removed all titles that also function as given names, such as \textit{Regina}.
-If a link to a \texttt{PER} or \texttt{UNK} entity, or an unlinked entity starts with, or consists solely of a title in the list, we tag the words that make up the title as \texttt{O}.
-\item[Incidental capitalization] Various non-\texttt{NNP} words in English are capitalized: names of months, the \textit{I} pronoun and non-entity acronyms, such as \textit{RSVP}.
-While the latter two group is unlikely to appear in Wikipedia text, we assembled a list of these words and tag them as \texttt{O} unless they are part of the alias set.
+\item[Reference cleansing] Page titles and anchor texts may contain more than
+  just the entity name. Personal titles are part of the entity name in WP, but
+  not in CoNLL, and punctuation marks around the entity may become part of the
+  link by mistake. We tag all punctuation marks after the entity name as
+  \texttt{O}. To handle personal titles, we extracted a list from the WP page
+  \textit{List of titles}, which contains titles in many languages. We
+  manually removed all titles that also function as given names, such as
+  \textit{Regina}. If a link to a \texttt{PER} or \texttt{UNK} entity, or an
+  unlinked entity starts with, or consists solely of a title in the list, we
+  tag the words that make up the title as \texttt{O}.
+\item[Incidental capitalization] Various non-\texttt{NNP} words in English are
+  capitalized: names of months, the pronoun \textit{I}, and non-entity
+  acronyms such as \textit{RSVP}. While the latter two types are unlikely to
+  appear in WP text, we assembled a list of these words and tag them as
+  \texttt{O} unless they are part of the alias set.
 \end{description}
@@
 
-As mentioned above, sentences with words tagged as \texttt{UNK} are discarded.
-Furthermore, there are many incomplete sentences in the Wikipedia text: image captions, enumerations items, contents of table cells, etc.
-On the one hand, these sentence fragments may be of too low quality to be of any use in the traditional NER task.
-On the other hand, they could prove to be invaluable when training a NER tagger for User Generated Content, which is known to be noisy and fragmented.
-As a compromise solution, we included these fragments in the corpus, but labelled them as "low quality", so that users of the corpus can decide whether they want to use them or not.
-A sentence is labelled as such if it either lacks a punctuation mark at the end, or it contains no finite verbs.
+As mentioned above, sentences with words tagged as \texttt{UNK} are
+discarded. Furthermore, there are many incomplete sentences in the WP text:
+image captions, enumerations items, contents of table cells, etc. On the one
+hand, these sentence fragments may be of too low quality to be of any use in
+the traditional NER task. On the other hand, they could prove to be invaluable
+when training a NER tagger for User Generated Content, which is known to be
+noisy and fragmented. As a compromise we included these fragments in the
+corpus, but labelled them as ``low quality", so that users of the corpus can
+decide whether they want to use them or not. A sentence is labelled as such if
+it either lacks a punctuation mark at the end, or it contains no finite verb.
 
@@
 
-The procedure described in the previous section was used to generate the Hungarian corpus, as well.
-However, linguistic diversity pose several problems.
-In this section we describe the differences between the two languages related to labeling NEs, and the changes they prompted in the method.
+The procedure described in the previous section was used to generate the
+Hungarian corpus as well. However, typological differences posed several
+problems. In this section we describe the differences between the two
+languages related to labeling NEs, and the changes they prompted in the
+method.
 
@@
 
-Although Hungarian is reckoned to be a less resourced language, and it is not supperted in NLTK, several high quality language processing tools have been developed for Hungarian in recent years.
-For tokenization and sentence segmentation we used an in-house statistical tool tailored for Hungarian.
-It has been trained on the largest manually annotated Hungarian corpus \cite{Csendes:04}, and it handles the peculiarities of Hungarian ortography, such as the periods placed after numbers in date expressions.
-Lemmatization was performed by hunmorph \cite{Tron:05} and hundisambig, the latter being an in-house disambiguator to select the right analysis based on the word context.
+Although Hungarian is reckoned to be a less resourced language, and it is not
+supperted in NLTK, several high quality language processing tools have been
+developed for Hungarian in recent years. For tokenization and sentence
+segmentation we used an in-house statistical tool tailored for Hungarian. It
+has been trained on the largest manually annotated Hungarian corpus
+\cite{Csendes:04}, and it handles the peculiarities of Hungarian ortography,
+such as the periods placed after numbers in date expressions. Lemmatization
+was performed by HunMorph \cite{Tron:05} and HunDisambig, an in-house
+disambiguator to select the right analysis based on the word context.
 
@@
 
-Hungarian is called a synthetic language: for the most part, it expresses grammatical elements in a single word form using affixes, as opposed to isolating languages (as English), which tend to employ separate words, e.g.~prepositions, pronouns, auxiliaries, for expressing grammatical phenomena.
-The Hungarian morphological analyzer outputs KR-codes \cite{Kornai:04}, which, in addition to the POS category, also includes inflectional information, which makes it much better suited to agglutinative languages, than Penn Treebank POS tags.
-One shortcoming of the KR-code is that it does not differenciate between common and proper nouns.
-Since in Hungarian only proper nouns are capitalized, we can usually decide whether a noun is proper or not based on the initial letter.
-However, this rule can not be used if the noun is at the beginning of a sentence, as words at the beginning of sentences are always capitalized.
-Hence, all sentences that begin with nouns have been removed from the corpus.
+For the most part Hungarian expresses grammatical elements within a word form
+using affixes. HunMorph outputs KR-codes \cite{Kornai:04}, which, in addition
+to the POS category, also include inflectional information, making it much
+better suited to agglutinative languages than Penn Treebank POS tags. One
+shortcoming of the KR-code is that it does not differentiate between common
+and proper nouns. Since in Hungarian only proper nouns are capitalized, we can
+usually decide whether a noun is proper based on the initial letter. However,
+this rule can not be used if the noun is at the beginning of a sentence, so
+sentences that begin with nouns have been removed from the corpus.
 
@@
 For well-resourced languages, DBpedia has internationalized chapters, 
-%data dumps based on Wikipedia. 
-but not for Hungarian. Instead, the Hungarian entity list comprises of the pages in the English list that have their equivalents in the Hungarian Wikipedia. Two consequences follow.
+but not for Hungarian. Instead, the Hungarian entity list comprises of the
+pages in the English list that have their equivalents in the Hungarian WP. Two
+consequences follow.  First, in order to identify which pages denote entities
+in the Hungarian WP, an additional step is required, in which the Hungarian
+equivalents of the English pages are added to the entity list. The English
+titles are retained because (due to the medium size of the Hungarian WP)
+in-article links sometimes point to English articles.
 
-First, in order to identify which pages denote entities in the Hungarian Wikipedia, an additional step is required, in which the Hungarian equivalents of the English pages are added to the entity list.
-The English titles are retained, as due to the medium size of the Hungarian Wikipedia, in-article links sometimes point to English articles.
+Second, entities without a page in the English WP are absent from the entity
+list. This gives rise to two potential problems. One is that compared to
+English, the list is relatively shorter: the entity/page ratio is 12.12\%, as
+opposed to the 37.66\% of the English WP. The other, since mostly Hungarian
+people, places and organizations are missing, a NER tagger that takes the
+surface forms of words into account might be mislead as to the language model
+of entity names. To overcome these problems, the list has to be extended with
+Hungarian entity pages that do not have a corresponding English page. We leave
+this for future work.
 
-Second, entities without a page in the English Wikipedia are absent from the entity list.
-This gives rise to two potential problems.
-One is that compared to English, the list is relatively shorter: the entity/page ratio is 12.12\%, as opposed to the 37.66\% of the English Wikipedia.
-The other, since mostly Hungarian people, places and organizations are missing, a NER tagger that takes the surface forms of words into account might be mislead as to the language model of entity names.
-To overcome these problems, the list has to be extended with Hungarian entity pages that do not have a corresponding English page.
-We leave this for future work.
-
-To annotate our corpus with NE tags, we chose to follow the annotation guidelines of the largest human-annotated NER corpus for Hungarian, the Szeged NER corpus \cite{Szarvas:06}.
-It is similar to CoNLL standards: contains newswire texts, comprises ca.~200,000 tokens, and is annotated with NE class labels in line with the CoNLL annotation scheme.
-However, the convention of what constitutes a NE is slightly different for Hungarian.
+To annotate our corpus with NE tags, we chose to follow the annotation
+guidelines of the largest human-annotated NER corpus for Hungarian, the Szeged
+NER corpus \cite{Szarvas:06}. It is similar to CoNLL standards: contains
+newswire texts, comprises ca.~200,000 tokens, and is annotated with NE class
+labels in line with the CoNLL annotation scheme. However, the convention of
+what constitutes a NE is slightly different for Hungarian.
 
@@
 
-The Szeged NER guideline relies heavily on the rules of capitalization to decide which words should be marked as NEs.
-As mentioned earlier, in Hungarian, apart from the first word in the sentence, only proper nouns are capitalized.
-The following concepts are not proper nouns in Hungarian, and thus are not considered as NEs: names of languages, nationalities, religions, political ideologies; adjectives derived from NEs; names of months, days, holidays; names of special events and wars.
+The Szeged NER guideline relies heavily on the rules of capitalization to
+decide which words should be marked as NEs. The following concepts are not
+proper nouns in Hungarian, and thus are not considered as NEs: names of
+languages, nationalities, religions, political ideologies; adjectives derived
+from NEs; names of months, days, holidays; names of special events and wars.
 
@@
 
-There is another special case in Hungarian: opposite to English, the number of compund words is quite large and NEs can also be subjects of compounding.
-In this case the common noun following the NE is joint with a hyphen, so they constitute one token.
-However, the joint common noun can modify the original sense of NE, depending on the semantics of the common noun.
-For example: in the compound \textit{Nobel-díj} ['Nobel Prize'] the common noun changes the labeling from PER to MISC, while
-%\textit{SAP-szoftvercég} ['SAP software company'] the common noun does not change the labeling, but 
-in the case of the compound \textit{WorldCom-botrány} ['WorldCom scandal'] the NE tag changes from ORG to O. The solution to this problem is not obvious, and needs more investigation.
-
-% TODO: <NE>-<random kisbetűs szó> => nem NE (magyar). Lásd: "a tű nélküli mezoterápia alapja a <Nobel, PER>-díjas"
+There is another special case in Hungarian: unlike in English, the number of
+compund words is quite large, and NEs can also be subject to compounding. In
+this case the common noun following the NE is joined with a hyphen, so they
+constitute one token. However, the joint common noun can modify the original
+sense of NE, depending on the semantics of the common noun. For example in
+the compound \textit{Nobel-díj} [`Nobel Prize'] the common noun changes the
+labeling from \texttt{PER} to \texttt{MISC}, while
+%\textit{SAP-szoftvercég} ['SAP software company'] the common noun does not
+%change the labeling, but
+in the case of the compound \textit{WorldCom-botrány} [`WorldCom scandal'] the NE tag changes from \texttt{ORG} to \texttt{O}.
+The solution to this problem is not obvious, and needs more investigation.
 
@@
 
-The corpora are available under the Creative Commons Attribution-Sharealike 3.0 Unported License (CC-BY-SA), the same license under which the text of Wikipedia is released.
-The data files can be freely downloaded from XXX.
-The corpora will also be distributed through META-SHARE network, which is an open, distributed facility for exchanging and sharing resources, and is one of the lines of action of META-NET, a Network of Excellence funded by the European Commission.
+The corpora are available under the Creative Commons Attribution-Sharealike
+3.0 Unported License (CC-BY-SA), the same license under which the text of WP
+is released. The data files can be freely downloaded from \texttt{http://hlt.sztaki.hu}. The corpora
+will also be distributed through the META-SHARE network, which is an open,
+distributed facility for exchanging and sharing resources, and is one of the
+lines of action of META-NET, a Network of Excellence funded by the European
+Commission.
 
@@
 Two NE tags are included with each word: the most specific DBpedia category it belongs to and the CoNLL NE tag.
-While the NE tags can be considered as a "silver standard", the linguistic features are provided on a "best-effort" basis.
+While the NE tags can be considered as a ``silver standard", the linguistic features are provided on a ``best-effort" basis.
 
@@
 
-Having the obvious advantages, an automatically generated corpus can not serve as a gold standard dataset.
-Then what can we do with silver standard corpora?
-They can be very useful for improving NER in more ways: (a) for less resourced languages, they can serve as training corpora in absence of gold standard datasets; (b) they can serve as (supplementary or independent) training sets for domains differing from newswire; (c) they can be sources of huge entity lists, and (d) feature extraction.
+Having the obvious advantages, an automatically generated corpus can not serve as a gold standard dataset. Then what can we do with silver standard corpora? They can be very useful for
+improving NER in several ways: (a) for less resourced languages, they can
+serve as training corpora in lieu of gold standard datasets; (b) they can
+serve as supplementary or independent training sets for domains differing
+from newswire; (c) they can be sources of huge entity lists, and (d) feature
+extraction.
 
-To evaluate our corpora we used a maximum entropy NE tagger \cite{Varga:07}, which was originally developed for labeling NEs in Hungarian texts, but it can be tuned for different languages, as well.
-Corpus-specific features (e.g.~NP chunks, Wikipedia links) were removed to get better comparability, so the feature set consists of gazetteer features; sentence start and end position; Boolean-valued orthographic properties of the word form; string-valued surface properties of the word form; and morphological information.
-
-%\begin{itemize}
-%\item gazetteer features;
-%\item sentence start and end position;
-%\item Boolean-valued orthographic properties of the word form;
-%\item string-valued surface properties of the word form;
-%\item morphological information.
-%\end{itemize}
+To evaluate our corpora we used a maximum entropy NE tagger \cite{Varga:07},
+which was originally developed for labeling NEs in Hungarian texts, but can
+be tuned for different languages as well. Corpus-specific features (e.g.~NP
+chunks, WP links) were removed to get better comparability, so the feature set
+consists of gazetteer features; sentence start and end position;
+Boolean-valued orthographic properties of the word form; string-valued surface
+properties of the word form; and morphological information.
 
@@
 
-Our automatic annotation process retains all of the Wikipedia sentences, which remained after our two-fold filtering method.
-Thus sentences without NEs are also included in the corpus.
-The rationale behind this is that we wanted to reserve the original distribution of names in Wikipedia as much as possible.
-However, after further investigation of the NE density in our corpora and gold standard corpora, we decided not to include the sentences without NEs in evaluation datasets.
+Our automatic annotation process retains all of the WP sentences which
+remained after our two-step filtering method, so sentences without NEs are
+also included in the corpus. The rationale behind this is that we wanted to
+reserve the original distribution of names in WP as much as possible. However,
+after further investigation of the NE density in our corpora and gold standard
+corpora, we decided not to include the sentences without NEs in evaluation
+datasets.
 
-Table \ref{size} summarizes the data regarding corpus size and NE density.
-The English (enwiki) and the Hungarian Wikipedia (huwiki) corpora originally have the NE density of 5.23\% and 2.38\%, respectively.
-In comparison to the gold standard datasets (CoNLL, Szeged NER) these counts are quite low.
-It can be due to the difference between domains: newswire articles usually contain more NEs, typically ORG.
-The other reason might be that we discarded sentences containing unidentified NEs (cf.~Section \ref{sec:create}).
+Table \ref{size} summarizes the data regarding corpus size and NE density. The
+English (enwiki) and the Hungarian WP (huwiki) corpora originally have the NE
+density of 5.23\% and 2.38\%, respectively. In comparison to the gold standard
+datasets (CoNLL, Szeged NER) these counts are quite low. It can be due to the
+difference between domains: newswire articles usually contain more NEs,
+typically ORG. The other reason might be that we discarded sentences
+containing unidentified NEs (cf.~Section \ref{sec:create}).
 
@@
 
-The English Wikipedia corpus was evaluated against itself and a manually annotated English corpus.
-Since the filtered English Wikipedia corpus, containing only the sentences with NEs, is still very huge, and we are limited by time and memory, our experiments were performed with a portion of 3.5 million tokens (the size of our filtered Hungarian corpus).
-It was divided into train and test sets (90\%-10\%).
+The English WP corpus was evaluated against itself and a manually annotated
+English corpus. Since the filtered English WP corpus, containing only the
+sentences with NEs, is still very large, our experiments were performed with a
+sample of 3.5 million tokens, the size of our filtered Hungarian corpus,
+divided into train and test sets (90\%-10\%).
 
-For English cross-corpus evaluation the CoNLL-2003 corpus was chosen.
-It is well-studied that cross-checking between different corpora decrease F-measure.
-Domain differences certainly affect NER performance, and the different annotation schemes pose several compatibility problems.
-Nothman et al.~\shortcite{Nothman:08} showed that each set of gold standard training data performs better on corresponding test sets than on test sets from other sources.
-The situation here is similar (see Table \ref{enresults} for results): the NE tagger trained on Wikipedia does not achieve as high performance tested against CoNLL test set (enwiki-CoNLL) as one trained on the own train set (enwiki-enwiki).
+For English cross-corpus evaluation the CoNLL-2003 corpus was chosen. As is
+well known, training and testing across different corpora decreases
+F-measure. Domain differences certainly affect NER performance, and the
+different annotation schemes pose several compatibility problems. Nothman et
+al.~\shortcite{Nothman:08} showed that each set of gold standard training data
+performs better on corresponding test sets than on test sets from other
+sources. The situation here is similar (see Table \ref{enresults} for
+results): the NE tagger trained on WP does not achieve as high performance
+tested against CoNLL test set (enwiki-CoNLL) as one trained on its own train
+set (enwiki-enwiki).
 
-However, Wikipedia-derived corpora can also be used for improving NER accuracy in other ways.
-First, we collected gazetteer lists from the corpus for each NE category, which improved the overall F-measure given to the NE tagger training and testing on CoNLL dataset (CoNLL with wikilists).
-A second trial was labeling the CoNLL datasets by the model trained on Wikipedia corpus, and giving these labels as extra features to the next CoNLL train (CoNLL with wikitags).
-Both methods result in improved F-measure on CoNLL test set.
+WP-derived corpora can also be used for improving NER accuracy in other
+ways. First, we collected gazetteer lists from the corpus for each NE
+category, which improved the overall F-measure given to the NE tagger training
+and testing on CoNLL dataset (CoNLL with wikilists). A second trial was
+labeling the CoNLL datasets by the model trained on WP corpus, and giving
+these labels as extra features to the next CoNLL train (CoNLL with
+wikitags). Both methods result in improved F-measure on CoNLL test set.
 
-Since in Hungarian NE tagging we followed the Szeged NER corpus annotation guidelines, we performed the experiments on this dataset.
-Hungarian results are similar to the English ones (see Table \ref{huresults}), the only difference is that F-measures for Hungarian are significantly higher.
-It can be due to the fact that the MISC category for Hungarian contains less types of names, thus the inconsistency of this class is smaller (cf.~Section \ref{sec:hun}).
-In contrast to the CoNLL corpus, the Szeged NER corpus was accurately annotated with an inter-annotator agreement over 99\%.
+Since in Hungarian NE tagging we followed the Szeged NER corpus annotation
+guidelines, we performed the experiments on this dataset. Hungarian results
+are similar to the English ones (see Table \ref{huresults}), the only
+difference is that F-measures for Hungarian are significantly higher. This can
+be due to the fact that the MISC category for Hungarian contains less types of
+names, thus the inconsistency of this class is smaller (cf.~Section
+\ref{sec:hun}). In contrast to the CoNLL corpus, the Szeged NER corpus was
+accurately annotated with an inter-annotator agreement over 99\%.
 
-Due to the quite good F-measure of training on our Hungarian train corpus and testing on the corresponding test set, our Hungarian corpus can serve as a training corpus to build NE taggers for domains falling out of newswire.
+Due to the quite good F-measure of training on our Hungarian train corpus and
+testing on the corresponding test set, our Hungarian corpus can serve as a
+training corpus to build NE taggers for non-newswire domains.
 
@@
 
-We have presented freely available NE tagged corpora for English and Hungarian, fully automatically generated from Wikipedia.
-In contrast to the methods used so far for automatic annotation of NEs in Wikipedia texts, we applied a new approach, namely mapping DBpedia ontology classes to standard CoNLL NE tags, and assigning them to Wikipedia entities.
-Following Nothman \shortcite{Nothman:08}, the process can be divided into four main steps:  classifying Wikipedia articles into entity classes; parsing Wikipedia and splitting articles into sentences; labeling NEs in the text; and selecting sentences for inclusion in the corpus.
+We have presented freely available NE tagged corpora for English and
+Hungarian, fully automatically generated from WP. In contrast to the methods
+used so far for automatic annotation of NEs in WP texts, we applied a new
+approach, namely mapping DBpedia ontology classes to standard CoNLL NE tags,
+and assigning them to WP entities. 
+Following Nothman \shortcite{Nothman:08},
+the process can be divided into four main steps: classifying WP articles into
+entity classes; parsing WP and splitting articles into sentences; labeling NEs
+in the text; and selecting sentences for inclusion in the corpus.
 
-Thanks to the huge amount of Wikipedia articles, the possibility of building large enough corpora is given for less resourced languages, like Hungarian, as well.
-Due to the particularities of Hungarian, some steps are slightly different, and special linguistics phenomena pose several problems related to the NER task to solve.
+The huge amount of WP articles opens the possibility of building large enough
+corpora for otherwise less resourced languages such as Hungarian. Due to the
+particularities of Hungarian, some steps are slightly different, and special
+linguistic phenomena pose several problems related to the NER task to solve.
 
-Automatically generated corpora can be useful for improving NER in more ways.
-We showed that gazetteer lists extracted from our corpora, and training with extra features given by the model trained on our corpora, improves F-measure.
-Moreover, our Hungarian corpus can serve as a training corpus for more general domains, than the classic newswire.
+Automatically generated corpora can be useful for improving NER in more
+ways. We showed that gazetteer lists extracted from our corpora, and training
+with extra features given by the model trained on our corpora, improve
+F-measure. Moreover, our Hungarian corpus can serve as a training corpus for
+more general domains than the classic newswire.
 
@@
 
-\section*{Acknowledgments}
+\section*{Acknowledgements}
 
 % TODO: the corpora will be published in the META-SHARE repository.
-This research was supported by the CESAR project under the ICT Policy Support Programme (grant no. 271022). The authors are grateful to Attila Zséder for his work on Wikipedia parsing.
+This research was supported by the OTKA grant no.
+82333 and the CESAR project under the ICT Policy Support Programme (grant no.
+271022).
+The authors are grateful to Attila Zséder for his work on Wikipedia parsing and to András Kornai for his help.
 

### 2012-04-28 paper/engNERwiki.tex
A few modifications to the paper.
Final version.

@@
 \documentclass[11pt]{article}
+\usepackage[english]{babel}
 \usepackage{acl2012}
@@
 \setlength\titlebox{6.5cm}    % Expanding the titlebox
+\hyphenation{DBpedia}
 
@@
 
-%While the approach in essence is language-independent, the exact details have to be accomodated to the target language.
+%While the approach in essence is language-independent, the exact details have to be accommodated to the target language.
 
@@
 graph\footnote{http://blog.dbpedia.org/2011/09/11/dbpedia-37-released-including-15-localized-editions}. Since
-selecting the right superclass, and hence, CoNLL tag for classes with more
-than one parent, cannot be reliably done automatically, the class-to-category
+selecting the right superclass, and hence, CoNLL tag, for classes with more
+than one parent cannot be reliably done automatically, the class-to-category
 mapping had to be determined manually. The only such class in version 3.7,
@@
 improved to increase precision, sentence coverage and to account for
-peculiarities in the CoNLL guidelines.
+peculiarities in the English orthography and the CoNLL guidelines.
 
@@
   link by mistake. We tag all punctuation marks after the entity name as
-  \texttt{O}. To handle personal titles, we extracted a list from the WP page
+  \texttt{O}.
+
+  To handle personal titles, we extracted a list from the WP page
   \textit{List of titles}, which contains titles in many languages. We
@@
   tag the words that make up the title as \texttt{O}.
+  % TODO 2 paragraphs?
 \item[Incidental capitalization] Various non-\texttt{NNP} words in English are
@@
 Although Hungarian is reckoned to be a less resourced language, and it is not
-supperted in NLTK, several high quality language processing tools have been
+supported in NLTK, several high quality language processing tools have been
 developed for Hungarian in recent years. For tokenization and sentence
@@
 has been trained on the largest manually annotated Hungarian corpus
-\cite{Csendes:04}, and it handles the peculiarities of Hungarian ortography,
+\cite{Csendes:04}, and it handles the peculiarities of Hungarian orthography,
 such as the periods placed after numbers in date expressions. Lemmatization
@@
 There is another special case in Hungarian: unlike in English, the number of
-compund words is quite large, and NEs can also be subject to compounding. In
+compound words is quite large, and NEs can also be subject to compounding. In
 this case the common noun following the NE is joined with a hyphen, so they
@@
 
-%az eredeti DBpedia osztályokat is megtarthatnánk, hiszen nem mindig célravezető a CoNLL osztályok használata, mivel azok elég tág kategóriák, amikbe olyan dolgok is tartoznak, amik nagyon eltérőek is lehetnek
-
 %később majd kirakjuk a scripteket is, és akár újra és újra frissíthető lesz
@@
 
-% TODO: the corpora will be published in the META-SHARE repository.
-This research was supported by the OTKA grant no.
+This research was supported by OTKA grant no.
 82333 and the CESAR project under the ICT Policy Support Programme (grant no.
 271022).
-The authors are grateful to Attila Zséder for his work on Wikipedia parsing and to András Kornai for his help.
+The authors are grateful to Attila Zséder for his work on Wikipedia parsing and to András Kornai for his insightful comments.
 

### 2012-06-27 paper/engNERwiki.tex
Order of authors fixed.

@@
 
-\author{Dávid Márk Nemeskey \\
+\author{Eszter Simon \\
+  Research Institute for Linguistics \\
+  Hungarian Academy of Sciences \\
+  H-1068 Benczúr utca 33, Budapest \\
+  {\tt simon.eszter@nytud.mta.hu} \\\And
+  Dávid Márk Nemeskey \\
   Research Institute for \\ Computer Science and Automation \\
@@
   H-1111 Kende utca 13-17, Budapest \\
-  {\tt nemeskey.david@sztaki.mta.hu} \\\And
-  Eszter Simon \\
-  Research Institute for Linguistics \\
-  Hungarian Academy of Sciences \\
-  H-1068 Benczúr utca 33, Budapest \\
-  {\tt simon.eszter@nytud.mta.hu} \\}
+  {\tt nemeskey.david@sztaki.mta.hu} \\}
 

### 2012-06-28 paper/jeju_presentation.tex
Presentation template added.

% $Header: /home/vedranm/bitbucket/beamer/solutions/generic-talks/generic-ornate-15min-45min.en.tex,v 90e850259b8b 2007/01/28 20:48:30 tantau $

\documentclass[utf8x]{beamer}

\mode<presentation>
{
%  \usetheme[height=7mm]{Rochester}
  \usetheme{Rochester}
  \useinnertheme{rounded}
%  \useoutertheme{infolines}
  % or ...

  \setbeamercovered{transparent}
}
\usefonttheme[onlymath]{serif}

\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}

\usepackage{times}
\usepackage[T1]{fontenc}
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.

%% to draw trees without jpg figures
%\usepackage{synttree}
% For code blocks
\usepackage{algorithm}  % Conflict w/ synttree
\usepackage{algorithmic}
% for sane tabular handling
\usepackage{array}
% For strikethrough (normalem keeps it as emph)
\usepackage[normalem]{ulem}

\usepackage{mathrsfs}
\usepackage{bm}  % For bold math

\newcommand{\vitem}{\vfill \item}

\title % (optional, use only with long paper titles)
{Reinforcement Learning End of Term Presentation}

%\subtitle
%{Presentation Subtitle} % (optional)

\author % (optional, use only with lots of authors)
{Eszter ~Simon, David Márk ~Nemeskey}

% TODO: RILHAS
\institute[Computer and Automation Research Institue of the Hungarian Academy of Sciences] % (optional, but mostly needed)
{
  Data Mining and Search Research Group\\
  MTA SZTAKI
}
% - Use the \inst command only if there are several affiliations.
% - Keep it simple, no one is interested in your street address.

\date % (optional)
{2012.07.12.}

% \subject{Talks}
% This is only inserted into the PDF information catalog. Can be left
% out.

% If you have a file called "university-logo-filename.xxx", where xxx
% is a graphic format that can be processed by latex or pdflatex,
% resp., then you can add a logo as follows:

% \pgfdeclareimage[height=0.5cm]{university-logo}{university-logo-filename}
% \logo{\pgfuseimage{university-logo}}

\AtBeginSection[]
{
   \begin{frame}
       \frametitle{Outline}
       \tableofcontents[currentsection]
   \end{frame}
}

% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command: 
%\beamerdefaultoverlayspecification{<+->}

\begin{document}

% TODO: kell outline?
\begin{frame}{Outline}
  \titlepage
\end{frame}

\section{Creating the English Corpus}
\subsection*{Motivation}
\begin{frame}{Blabla}
\end{frame}
\subsection{Articles to Entities}
\begin{frame}{Problem Statement}
  In many real world applications the problem is to predict outputs with a nontrivial structure given some inputs.
  \vfill
  Example: \textit{\small It was love at first sight.}
  \vfill
\end{frame}

\section{Conclusions}
\begin{frame}{Results -- cont.}
  Single hold-out vs. cross-validation:
  \begin{itemize}
  \vitem Single hold-out is biased
  \vitem cross-validation gives more stable results
  \end{itemize}
  \vfill
\end{frame}

\end{document}

### 2012-06-29 paper/jeju_presentation.tex
Section about the Hungarian corpus added.
Draft of one slide added. Yay!
Draft of the Hungarian part is done.

@@
 
+\section{Creating the Hungarian Corpus}
+\begin{frame}{The Hungarian Corpus}
+  The algorithm is, in theory, language-agnostic. We tested this property by
+  applying it to a medium-sized language, Hungarian. \\
+  Difference from bigger, well-resourced languages:
+  \begin{itemize}
+  \vitem E % TODO: ide mit?
+  \end{itemize}
+\end{frame}
+%TODO introductory slide here
+\begin{frame}{Differences from English}
+  \begin{itemize}
+  \vitem Language processing tools
+    \begin{itemize}
+    \vitem Inflectional morphology: different set of tools were needed 
+    \vitem We developed a flexible framework that enables the plugging of any NLP tool
+    % TODO: what kind of linguistic tools are needed: table?
+    \end{itemize}
+  \vitem Wikipedia size and coverage
+    \begin{itemize}
+    \vitem Hungarian is medium-size language: ~200k articles vs. ~4M English pages
+    \vitem Cross-references may point to English pages
+    \vitem Templates are in Hungarian: per-language template configuration
+    \end{itemize}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Differences from English -- cont.}
+  \begin{itemize}
+  \vitem DBpedia coverage
+    \begin{itemize}
+    \vitem Internationalized chapters for big languages only
+    \vitem "Hungarian" DBpedia list contains English articles % Nem egyértelmű
+      \begin{itemize}
+      \vitem Hungarian equivalents must be added to the list % Added: lásd előző oldal 2/2
+      \vitem Hungarian entities without an English page are missing:
+             the NER language model might not be accurate
+      \vitem Future work: extend the list with the missing Hungarian entities
+      \end{itemize}
+    \end{itemize}
+  \vitem NE conventions
+    \begin{itemize}
+    \vitem Szeged NER guideline: similar to CoNLL
+    \vitem Common nouns in Hungarian: names of languages, religions, months, etc.
+    % TODO: compounding
+    \end{itemize}
+  \end{itemize}
+\end{frame}
+
 \section{Conclusions}

### 2012-06-29 paper/jeju_presentation.tex
Changes from Eszter merged.

@@
 \newcommand{\vitem}{\vfill \item}
+\newcommand{\nyil}{$\rightarrow$\ }
+
+\newtheorem{nix}{}[section]
 
 \title % (optional, use only with long paper titles)
-{Reinforcement Learning End of Term Presentation}
+{Automatically generated NE tagged corpora for English and Hungarian}
 
@@
 \author % (optional, use only with lots of authors)
-{Eszter ~Simon, David Márk ~Nemeskey}
+{Eszter Simon\inst{1}, Dávid Márk Nemeskey\inst{2}}
 
-% TODO: RILHAS
-\institute[Computer and Automation Research Institue of the Hungarian Academy of Sciences] % (optional, but mostly needed)
-{
-  Data Mining and Search Research Group\\
-  MTA SZTAKI
-}
+\institute{\inst{1} Research Institute for Linguistics, Hungarian Academy of Sciences \\
+\inst{2} Computer and Automation Research Institute, Hungarian Academy of Sciences} % (optional, but mostly needed)
 % - Use the \inst command only if there are several affiliations.
@@
 % TODO: kell outline?
-\begin{frame}{Outline}
+\begin{frame}{}
   \titlepage
@@
 
+\section{Introduction}
+
+\begin{frame}{Datasets}
+
+The standard datasets are
+
+\begin{itemize}
+\item highly domain-specific
+\item restricted in size
+\end{itemize}
+
+\bigskip
+
+To get a bigger training corpus:
+
+\begin{itemize}
+\item merging datasets \nyil problems of combining different tagsets
+\item manual annotation \nyil not cost-effective
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}{Reducing the annotation cost}
+
+Reducing the annotation cost:
+
+\begin{itemize}
+\item to generate the resource automatically 
+\item to use collaboratively constructed resources (e.g.~Wikipedia, DBpedia)
+\end{itemize}
+
+\bigskip
+
+\begin{nix}
+combining these approaches: \\ automatically generating NE tagged corpora from Wikipedia
+\end{nix}
+
+\end{frame}
+
+%\section{Wikipedia and NER}
+
+\begin{frame}{Wikipedia and NER}
+
+\begin{nix}
+Wikipedia is one of the largest available collections of entities
+\end{nix}
+
+\smallskip
+
+Improving the accuracy of NER with WP:
+
+\begin{itemize}
+\item extracting gazetteers containing person names, locations and organizations 
+\item extracting the knowledge embodied in WP and incorporating it in NER learning as features
+\item automatic creation of training data from WP
+\begin{itemize}
+\item by using the category structure of WP
+\item by using the category head nouns and the opening sentences of articles
+\end{itemize} 
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}{Our approach}
+
+\begin{nix}
+mapping the DBpedia ontology classes to standard NE tags, and assign them to WP entities
+\end{nix}
+
+\bigskip
+
+our method is largely language-independent \nyil sufficiently large corpora for less resourced languages can be built
+
+\bigskip
+
+\begin{nix}
+this is the first automatically annotated corpus for Hungarian 
+\end{nix}
+
+\end{frame}
+
 \section{Creating the English Corpus}
+
 \subsection*{Motivation}
@@
 \section{Creating the Hungarian Corpus}
+
 \begin{frame}{The Hungarian Corpus}
@@
 \end{frame}
-%TODO introductory slide here
+
 \begin{frame}{Differences from English}
@@
 
+\section{Data description}
+
+\begin{frame}{Data format}
+
+\begin{tabular}{llllll}
+\textbf{token} & \textbf{text/link} & \textbf{WP link} & \textbf{POS} & \textbf{lemma} & \textbf{NE} \\
+This & text & 0 &    DT &   This & O \\
+is &   text & 0 &    VBZ &  be &   O \\
+a &    text & 0 &    DT &   a  &    O\\
+list & text & 0 &    NN &   list  & O\\
+of &   text & 0 &    IN &   of &     O\\
+films  & B-link & films &  NNS &  film & O\\
+by &   text & 0 &    IN &   by &     O\\
+year & text & 0 &    NN &   year & O\\
+produced & text & 0 &    VBN &  produce & O\\
+in &   text & 0 &    IN &   in &     O\\
+the &  text & 0 &    DT &   the &   O\\
+country &text & 0 &    NN &   country &O\\
+of &   text & 0 &    IN &   of &     O\\
+South &  B-link & South Korea &  NNP &  South  &B-LOC\\
+Korea &  I-link & South Korea &  NNP &  Korea   & I-LOC\\
+\end{tabular}
+
+\end{frame}
+
+\begin{frame}{Data description}
+
+\begin{itemize}
+\item available under CC-BY-SA 3.0 Unported License
+\item freely downloadable from {\tt http://hlt.sztaki.hu/resources/} 
+\item also distributed through META-SHARE (from 1 August)
+\end{itemize}
+
+\end{frame}
+
+\section{Evaluation}
+
+\begin{frame}{Corpus size and NE density}
+
+English:
+
+\smallskip
+
+\begin{center}
+\begin{tabular}{llll}
+\hline  & \bf enwiki & \bf enwiki filtered & \bf CoNLL \\ \hline
+token & 60,520,819 & 21,718,854 & 302,811 \\
+NE & 3,169,863 & 3,169,863 & 50,758 \\
+NE density & 5.23\% & 14.59\% & 16.76\% \\ \hline
+\end{tabular}
+\end{center}
+
+\smallskip
+
+Hungarian:
+
+\smallskip
+
+\begin{center}
+\begin{tabular}{llll}
+\hline  & \bf huwiki & \bf huwiki filtered  & \bf Szeged NER \\ \hline
+token & 19,108,027 & 3,512,249  & 225,963 \\
+NE & 456,281 & 456,281  & 25,896 \\
+NE density & 2.38\% & 12.99\%  & 11.46\% \\ \hline 
+\end{tabular}
+\end{center}
+
+\end{frame}
+
+\begin{frame}{English results}
+
+\begin{center}
+\begin{tabular}{lllll}
+\hline \bf Train & \bf Test & \bf P (\%) & \bf R (\%) & \bf F (\%) \\ \hline
+CoNLL & CoNLL & 85.13 & 85.13 & 85.13 \\
+enwiki & enwiki & 72.46 & 73.33 &  72.89 \\
+enwiki & CoNLL & 56.55 & 49.77 & 52.94 \\
+CoNLL with wikilists & CoNLL & 86.33 & 86.35 & 86.34 \\
+CoNLL with wikitags & CoNLL & 85.88 & 85.94 & 85.91 \\
+\hline
+\end{tabular}
+\end{center}
+
+\end{frame}
+
+\begin{frame}{Hungarian results}
+
+\begin{center}
+\begin{tabular}{lllll}
+\hline \bf Train & \bf Test & \bf P (\%) & \bf R (\%) & \bf F (\%) \\ \hline
+Szeged  & Szeged  & 94.50 & 94.35 & 94.43 \\
+huwiki & huwiki & 90.64 & 88.91 &  89.76 \\
+huwiki & Szeged  & 63.08 & 70.46 & 66.57 \\
+Szeged  with wikilists & Szeged  & 95.48 & 95.48 & 95.48 \\
+Szeged  with wikitags & Szeged  & 95.38 & 94.92 & 95.15 \\
+\hline
+\end{tabular}
+\end{center}
+
+\end{frame}
+
 \section{Conclusions}

### 2012-06-29 paper/jeju_presentation.tex
The F-score in the Hungarian table has been changed to bold face.

@@
 Szeged  & Szeged  & 94.50 & 94.35 & 94.43 \\
-huwiki & huwiki & 90.64 & 88.91 &  89.76 \\
+huwiki & huwiki & 90.64 & 88.91 &  \textbf{89.76} \\
 huwiki & Szeged  & 63.08 & 70.46 & 66.57 \\

### 2012-06-29 paper/jeju_presentation.tex
Copied "Motivation" part from the paper.

@@
 \subsection*{Motivation}
-\begin{frame}{Blabla}
+\begin{frame}{Motivation}
+  \begin{itemize}
+  \vitem Our goal is to create a large NE annotated corpus, automatically
+        generated from WP articles.
+  \vitem Our method is similar to Nothman et al. (2008)
+  \vitem The process consists of four steps:
+    \begin{enumerate}
+    \vitem Classify WP articles into entity classes.
+    \vitem Parse WP and split articles into sentences.
+    \vitem Label named entities in the text.
+    \vitem Select the sentences for inclusion in the corpus.
+    \end{enumerate}
+  \end{itemize}
 \end{frame}
+
 \subsection{Articles to Entities}
@@
 \end{frame}
+
+\subsection{}
+\subsection{}
+\subsection{}
 

### 2012-06-30 paper/jeju_presentation.tex
"Creating the English Corpus" part done.

@@
 
-\documentclass[utf8x]{beamer}
+\documentclass[utf8x,t]{beamer}
 
@@
     \vitem Classify WP articles into entity classes.
-    \vitem Parse WP and split articles into sentences.
+    \vitem \textit{Parse WP and split articles into sentences.}
     \vitem Label named entities in the text.
@@
 \subsection{Articles to Entities}
-\begin{frame}{Problem Statement}
-  In many real world applications the problem is to predict outputs with a nontrivial structure given some inputs.
-  \vfill
-  Example: \textit{\small It was love at first sight.}
+\begin{frame}{Articles to Entities}
+  Our method is based on thinking of WP articles as NEs.
+  \begin{itemize}
+  \vitem Usual approach: semi-supervised learning (infoboxes, first sentence)
+  \vitem Our approach: create entity list from DBpedia
+    \begin{itemize}
+    \vitem OWL ontology with 320 classes
+    \vitem We take the most specific class
+    \vitem Manual mapping to CoNLL categories that exploits inheritence
+    \end{itemize}
+  \vitem Decreased coverage, increased precision
+  \end{itemize}
+\end{frame}
+
+\subsection{Named Entity Tagging}
+\begin{frame}{Named Entity Tagging}
+  The entity list from DBpedia is used for NE tagging.
+  \begin{itemize}
+  \vitem Entity identification:
+    \begin{itemize}
+    \vitem Look up linked pages in the entity list
+    \vitem References to the current page, if in the list
+    \end{itemize}
+  \vitem NE tagging: from the DBpedia - CoNLL category mapping
+  \vitem Exceptions:
+    \begin{itemize}
+    \vitem Not entities: common nouns, date expressions, etc.
+    \vitem Unmarked entities: link to already identified entities
+    \vitem Special cases: derived words, first word in a sentence, etc.
+    \end{itemize}
+  \end{itemize}
+\end{frame}
+
+\subsection{Sentence Filtering}
+\begin{frame}{Sentence Filtering}
+  We only include fully tagged sentences in the corpus.
+  \begin{itemize}
+  \vitem Sentences that contain unknown links / uppercase words
+  \vitem To filter incomplete sentences:
+    \begin{itemize}
+    \vitem Sentences without an EOS punctuation mark
+    \vitem Sentences without a predicate (verb)
+    \end{itemize}
+  \end{itemize}
   \vfill
 \end{frame}
-
-\subsection{}
-\subsection{}
-\subsection{}
 

### 2012-07-01 paper/jeju_presentation.tex
Incorporated Eszter's changes to the presentation.

@@
 \newcommand{\nyil}{$\rightarrow$\ }
+\newcommand{\nagytilde}{$\sim$}
 
@@
 \date % (optional)
-{2012.07.12.}
+{2012.07.12. \\ ACL NEWS2012 workshop}
 
@@
 \begin{itemize}
-\item highly domain-specific
-\item restricted in size
+\vitem highly domain-specific
+\vitem restricted in size
 \end{itemize}
+\vfill
 
@@
 \begin{itemize}
-\item merging datasets \nyil problems of combining different tagsets
-\item manual annotation \nyil not cost-effective
+\vitem merging datasets \nyil problems of combining different tagsets
+\vitem manual annotation \nyil not cost-effective
 \end{itemize}
+\vfill
 
@@
 \begin{frame}{Reducing the annotation cost}
+
+\bigskip
 
@@
     \end{itemize}
-  \vitem NE tagging: from the DBpedia - CoNLL category mapping
+  \vitem NE tagging: from the DBpedia--CoNLL category mapping
   \vitem Exceptions:
@@
 
-\begin{frame}{The Hungarian Corpus}
-  The algorithm is, in theory, language-agnostic. We tested this property by
-  applying it to a medium-sized language, Hungarian. \\
-  Difference from bigger, well-resourced languages:
-  \begin{itemize}
-  \vitem E % TODO: ide mit?
-  \end{itemize}
-\end{frame}
+%\begin{frame}{The Hungarian Corpus}
+%  The algorithm is, in theory, language-agnostic. We tested this property by
+%  applying it to a medium-sized language, Hungarian. \\
+%  Difference from bigger, well-resourced languages:
+%  \begin{itemize}
+%  \vitem E % TODO: ide mit?
+%  \end{itemize}
+%\end{frame}
 
+%TODO: the "Hungarian Corpus" phrase must appear in the frame title 
 \begin{frame}{Differences from English}
@@
     \begin{itemize}
-    \vitem Hungarian is medium-size language: ~200k articles vs. ~4M English pages
+    \vitem Hungarian is medium-size language: \nagytilde 200k articles vs.~\nagytilde 4M English pages
     \vitem Cross-references may point to English pages
@@
     \vitem Szeged NER guideline: similar to CoNLL
-    \vitem Common nouns in Hungarian: names of languages, religions, months, etc.
-    % TODO: compounding
+    \vitem Common nouns in Hungarian: names of languages, religions, months, etc. \nyil more consistent MISC category
+    \vitem NEs can be subject of compounding \nyil the joint common noun can modify the original sense of NE: \\ \textit{WorldCom} (\texttt{ORG}) \nyil \textit{WorldCom-botrány} ['WorldCom scandal'] (\texttt{O})
     \end{itemize}
@@
 \item freely downloadable from {\tt http://hlt.sztaki.hu/resources/} 
-\item also distributed through META-SHARE (from 1 August)
+\item also distributed through META-SHARE (from August)
 \end{itemize}
@@
 
+\begin{frame}{Silver standard corpora}
+
+Silver standard corpora can be useful for improving NER in several ways:
+
+\smallskip
+
+\begin{itemize}
+\item for less resourced languages they can serve as training corpora in lieu of gold standard datasets
+\item they can serve as training sets for domains differing from newswire
+\item they can be sources of huge entity lists
+\item they can be sources of feature extraction
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}{Evaluation}
+
+\smallskip
+
+A maxent NE tagger was used with the feature set consisting of
+
+\smallskip
+
+\begin{itemize}
+\item gazetteer features
+\item sentence start and end position
+\item Boolean-valued orthographic properties of the word form
+\item string-valued properties of the word form
+\item morphological information
+\end{itemize}
+
+\bigskip
+
+The CoNLL standard method was used for evaluation.
+
+\end{frame}
+
 \begin{frame}{Corpus size and NE density}
@@
 English:
-
-\smallskip
 
@@
 
-\smallskip
+%\smallskip
 
 Hungarian:
-
-\smallskip
 
@@
 \end{center}
+
+quite low NE density \nyil sentences without NEs were filtered out from the evaluation datasets
+
+\end{frame}
+
+\begin{frame}{Experiments}
+
+\begin{nix}
+our experiments were performed with a sample of 3.5 million tokens
+\end{nix}
+
+\bigskip
+
+\begin{itemize}
+\item \textit{CoNLL} train -- \textit{CoNLL} test
+\item \textit{enwiki} train -- \textit{enwiki} test
+\item \textit{enwiki} train -- \textit{CoNLL} test
+\item \textit{CoNLL} train with \textit{wikilists} (gazetteer lists extracted from WP) -- \textit{CoNLL} test
+\item \textit{CoNLL} train with \textit{wikitags} (labeling the CoNLL datasets by the model trained on WP, and giving these labels as extra features) -- \textit{CoNLL} test
+\end{itemize}
 
@@
 
+\begin{itemize}
+\item training and testing across different corpora decreases F-measure
+\item the NE tagger trained on WP does not achieve as high performance tested against CoNLL test set as one trained on its own train set
+\item gazetteer lists extracted from WP and training with extra features given by the model trained on WP improve F-measure
+\end{itemize}
+
 \end{frame}
@@
 
+\begin{itemize}
+\item F-measures for Hungarian are higher 
+\begin{itemize}
+\item the MISC category is more consistent
+\item the Szeged NER corpus is highly accurately tagged 
+\end{itemize}
+\item the Hungarian corpus can serve as a training corpus to build NE taggers for non-newswire domains
+\end{itemize}
+
 \end{frame}
 
+%\begin{frame}{Results -- analysis}
+
+%\begin{itemize}
+%\vitem training and testing across different corpora decreases F-measure
+%\vitem the NE tagger trained on WP does not achieve as high performance tested against CoNLL test set as one trained on its own train set
+%\vitem gazetteer lists extracted from WP and training with extra features given by the model trained on WP improve F-measure
+%\vitem F-measures for Hungarian are higher 
+%\begin{itemize}
+%\vitem the MISC category is more consistent
+%\vitem the Szeged NER corpus is highly accurately tagged 
+%\end{itemize}
+%\end{itemize}
+%  \vfill
+
+%\end{frame}
+
 \section{Conclusions}
-\begin{frame}{Results -- cont.}
-  Single hold-out vs. cross-validation:
+\begin{frame}{Conclusions}
   \begin{itemize}
-  \vitem Single hold-out is biased
-  \vitem cross-validation gives more stable results
+ \vitem we presented freely available NE tagged corpora for English and Hungarian, fully automatically generated from WP
+  \vitem we applied a new approach: mapping DBpedia ontology classes to standard CoNLL NE tags, and assigning them to WP entities
+  \vitem the huge amount of WP articles opens the possibility of building large enough corpora for less resourced languages such as Hungarian
+  \vitem silver standard corpora can improve NER accuracy in more ways
+  \vitem the Hungarian corpus can serve as a training corpus to build NE taggers for non-newswire domains
   \end{itemize}
@@
 
+\begin{frame}
+
+\bigskip
+
+\bigskip
+
+{\LARGE Thank you for your attention!}
+
+\bigskip
+\bigskip
+
+emails: {\tt simon.eszter@nytud.mta.hu \\ nemeskey.david@sztaki.mta.hu} \\
+
+\bigskip
+
+the corpora are freely downloadable from {\tt http://hlt.sztaki.hu/resources/}
+
+\end{frame}
+
 \end{document}

### 2012-07-02 paper/jeju_presentation.tex
Final version of the META-SHARE descriptor and penultimate version of the Jeju presentation.

@@
 
-\newcommand{\vitem}{\vfill \item}
+\newcommand{\vitem}{\item \vspace{4pt}}
 \newcommand{\nyil}{$\rightarrow$\ }
 \newcommand{\nagytilde}{$\sim$}
+\newtheorem{Examplee}{Example}[section]
 
@@
 
+\bigskip
+
 The standard datasets are
-
+\smallskip
 \begin{itemize}
-\vitem highly domain-specific
-\vitem restricted in size
+\item highly domain-specific
+\item restricted in size
 \end{itemize}
-\vfill
 
@@
 To get a bigger training corpus:
-
+\smallskip
 \begin{itemize}
-\vitem merging datasets \nyil problems of combining different tagsets
-\vitem manual annotation \nyil not cost-effective
+\item merging datasets \nyil problems of combining different tagsets
+\item manual annotation \nyil not cost-effective
 \end{itemize}
-\vfill
 
@@
 
-\begin{frame}{Reducing the annotation cost}
+\begin{frame}{Reducing the Annotation Cost}
 
@@
 Reducing the annotation cost:
-
+\smallskip
 \begin{itemize}
@@
 \begin{nix}
-combining these approaches: \\ automatically generating NE tagged corpora from Wikipedia
+Combining these approaches: \\ automatically generating NE tagged corpora from Wikipedia
 \end{nix}
@@
 Improving the accuracy of NER with WP:
-
+\smallskip
 \begin{itemize}
@@
 
-\begin{frame}{Our approach}
+\begin{frame}{Our Approach}
 
 \begin{nix}
-mapping the DBpedia ontology classes to standard NE tags, and assign them to WP entities
+Mapping the DBpedia ontology classes to standard NE tags, and assign them to WP entities
 \end{nix}
@@
 
-our method is largely language-independent \nyil sufficiently large corpora for less resourced languages can be built
+Our method is largely language-independent \nyil sufficiently large corpora for less resourced languages can be built
 
@@
 \begin{nix}
-this is the first automatically annotated corpus for Hungarian 
+This is the first automatically annotated corpus for Hungarian 
 \end{nix}
@@
 \begin{frame}{Motivation}
+  \bigskip
+
   \begin{itemize}
-  \vitem Our goal is to create a large NE annotated corpus, automatically
+  \item Our goal is to create a large NE annotated corpus, automatically
         generated from WP articles.
+  \smallskip
   \vitem Our method is similar to Nothman et al. (2008)
@@
 \begin{frame}{Articles to Entities}
+  \bigskip
   Our method is based on thinking of WP articles as NEs.
+  \smallskip
   \begin{itemize}
@@
     \end{itemize}
-  \vitem Decreased coverage, increased precision
+  \item Decreased coverage, increased precision
   \end{itemize}
@@
 \begin{frame}{Named Entity Tagging}
+  \bigskip
   The entity list from DBpedia is used for NE tagging.
+  \smallskip
   \begin{itemize}
@@
 \begin{frame}{Sentence Filtering}
+  \bigskip
   We only include fully tagged sentences in the corpus.
+  \smallskip
   \begin{itemize}
@@
   \end{itemize}
-  \vfill
 \end{frame}
@@
 \begin{frame}{Differences from English}
+  \bigskip
   \begin{itemize}
@@
 \begin{frame}{Differences from English -- cont.}
+  \bigskip
   \begin{itemize}
   \vitem DBpedia coverage
+    \smallskip
     \begin{itemize}
@@
     \end{itemize}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Differences from English -- cont.}
+  \bigskip
+  \begin{itemize}
   \vitem NE conventions
+    \smallskip
     \begin{itemize}
@@
     \vitem Common nouns in Hungarian: names of languages, religions, months, etc. \nyil more consistent MISC category
-    \vitem NEs can be subject of compounding \nyil the joint common noun can modify the original sense of NE: \\ \textit{WorldCom} (\texttt{ORG}) \nyil \textit{WorldCom-botrány} ['WorldCom scandal'] (\texttt{O})
+    \vitem NEs can be subject of compounding \nyil the joint common noun can modify the original sense and category of the NE:
     \end{itemize}
   \end{itemize}
+  \bigskip
+  \begin{Examplee}
+  \small{WorldCom {\normalfont(\texttt{ORG})} \nyil WorldCom-botrány ['WorldCom scandal'] {\normalfont(\texttt{O})}}
+  \end{Examplee}
 \end{frame}
 
-\section{Data description}
+\section{Data Description}
 
-\begin{frame}{Data format}
+\begin{frame}{Data Format}
 
@@
 
-\begin{frame}{Data description}
-
+\begin{frame}{Data Availability}
+\bigskip
+The corpora are
 \begin{itemize}
-\item available under CC-BY-SA 3.0 Unported License
-\item freely downloadable from {\tt http://hlt.sztaki.hu/resources/} 
-\item also distributed through META-SHARE (from August)
+\vitem available under CC-BY-SA 3.0 Unported License
+\vitem freely downloadable from \url{http://hlt.sztaki.hu/resources/} 
+\vitem also distributed through META-SHARE (from August)
+  \begin{itemize}
+  \vitem an open, distributed facility for exchanging and sharing resources
+  \vitem one of the lines of action of META-NET, a Network of Excellence funded
+        by the European Commission
+  \vitem \url{http://www.meta-net.eu/meta-share}
+  \end{itemize}
 \end{itemize}
@@
 
-\begin{frame}{Silver standard corpora}
+\begin{frame}{Silver Standard Corpora}
+
+\bigskip
 
@@
 \begin{itemize}
-\item for less resourced languages they can serve as training corpora in lieu of gold standard datasets
-\item they can serve as training sets for domains differing from newswire
-\item they can be sources of huge entity lists
-\item they can be sources of feature extraction
+\vitem for less resourced languages they can serve as training corpora in lieu of gold standard datasets
+\vitem they can serve as training sets for domains differing from newswire
+\vitem they can be sources of huge entity lists
+\vitem they can be sources of feature extraction
 \end{itemize}
@@
 
-\smallskip
+\bigskip
 
@@
 \begin{itemize}
-\item gazetteer features
-\item sentence start and end position
-\item Boolean-valued orthographic properties of the word form
-\item string-valued properties of the word form
-\item morphological information
+\vitem gazetteer features
+\vitem sentence start and end position
+\vitem Boolean-valued orthographic properties of the word form
+\vitem string-valued properties of the word form
+\vitem morphological information
 \end{itemize}
@@
 
-\begin{frame}{Corpus size and NE density}
+\begin{frame}{Corpus Size and NE Density}
 
@@
 
-quite low NE density \nyil sentences without NEs were filtered out from the evaluation datasets
+Quite low NE density \nyil sentences without NEs were filtered out from the evaluation datasets
 
@@
 \begin{nix}
-our experiments were performed with a sample of 3.5 million tokens
+Our experiments were performed with a sample of 3.5 million tokens
 \end{nix}
@@
 
-\begin{frame}{English results}
+\begin{frame}{English Results}
 
@@
 
-\begin{frame}{Hungarian results}
+\begin{frame}{Hungarian Results}
 
@@
 \begin{itemize}
-\item F-measures for Hungarian are higher 
+\vitem F-measures for Hungarian are higher 
 \begin{itemize}
-\item the MISC category is more consistent
-\item the Szeged NER corpus is highly accurately tagged 
+\vitem the MISC category is more consistent
+\vitem the Szeged NER corpus is highly accurately tagged 
 \end{itemize}
-\item the Hungarian corpus can serve as a training corpus to build NE taggers for non-newswire domains
+\vitem the Hungarian corpus can serve as a training corpus to build NE taggers for non-newswire domains
 \end{itemize}
@@
   \begin{itemize}
- \vitem we presented freely available NE tagged corpora for English and Hungarian, fully automatically generated from WP
-  \vitem we applied a new approach: mapping DBpedia ontology classes to standard CoNLL NE tags, and assigning them to WP entities
-  \vitem the huge amount of WP articles opens the possibility of building large enough corpora for less resourced languages such as Hungarian
-  \vitem silver standard corpora can improve NER accuracy in more ways
-  \vitem the Hungarian corpus can serve as a training corpus to build NE taggers for non-newswire domains
+  \vitem We presented freely available NE tagged corpora for English and Hungarian, fully automatically generated from WP
+  \vitem We applied a new approach: mapping DBpedia ontology classes to standard CoNLL NE tags, and assigning them to WP entities
+  \vitem The huge amount of WP articles opens the possibility of building large enough corpora for less resourced languages such as Hungarian
+  \vitem Silver standard corpora can improve NER accuracy in more ways
+  \vitem The Hungarian corpus can serve as a training corpus to build NE taggers for non-newswire domains
   \end{itemize}
@@
 
-{\LARGE Thank you for your attention!}
+{\huge Thank you for your attention!}
 
@@
 \bigskip
+\bigskip
 
-emails: {\tt simon.eszter@nytud.mta.hu \\ nemeskey.david@sztaki.mta.hu} \\
+emails: \\ {\tt simon.eszter@nytud.mta.hu \\ nemeskey.david@sztaki.mta.hu} \\
 
+\bigskip
 \bigskip

### 2012-07-03 paper/engNERwiki.tex
The examples for Common Noun pages in chapter 3 have been italicized.

@@
 
-Strictly speaking, our original assumption of equating WP articles with NEs is not valid: many pages describe common nouns (Book, Aircraft), calendar-related concepts (March 15, 2007), or other concepts that fall outside the scope of NER.
+Strictly speaking, our original assumption of equating WP articles with NEs is not valid: many pages describe common nouns (\textit{Book}, \textit{Aircraft}), calendar-related concepts (\textit{March 15}, \textit{2007}), or other concepts that fall outside the scope of NER.
 To increase sentence coverage, we modified the algorithm to prevent it from misclassifying links to these pages as unknown entities and discarding the sentence.

### 2012-10-01 paper/MSZNY2013/mszny2013.tex
Organized the papers into folders.

\documentclass{llncs}

\usepackage[a4paper,includeheadfoot,top=1.65in,bottom=1.65in,left=1.73in,hcentering]{geometry}
\usepackage[pdftex]{graphicx}

% Ezek egyikÈnek bekapcsol·s·val a magyar Èkezetes karakterek
% kˆzvetlen¸l is felhaszn·lhatÛk:
%\usepackage[latin2]{inputenc} 
\usepackage[utf8]{inputenc}

\usepackage[hungarian]{babel}
\selectlanguage{hungarian}

\begin{document}

\pagestyle{myheadings}
\def\leftmark{{\rm IV. Magyar Sz\'am\'\i t\'og\'epes Nyelv\'eszeti Konferencia}}
\def\rightmark{{\rm Szeged, d·tum}}

\setcounter{page}{3}

\title{\ \break Programcsomag inform\'aci\'okinyer\'esi\break
kutat\'asok t\'amogat\'as\'ara}
\author{Alexin Zolt\'an\inst{1}, Gyim\'othy Tibor\inst{1}}
\institute{Szegedi Tudom\'anyegyetem, TTK, Informatikai Tansz\'ekcsoport,\break
Szeged \'Arp\'ad t\'er 2., e-mail:\{alexin,gyimothy\}@inf.u-szeged.hu}

\maketitle

\begin{abstract}
A publik\'aci\'oban bemutat\'asra ker\"ul egy inform\'aci\'okinyer\'esi
kutat\'asokat t\'amogat\'o programcsomag, amelynek moduljai a
nyers sz\"oveg beolvas\'as\'at\'ol kezdve a v\'egeredm\'eny
webes megjelen\'\i t\'es\'eig minden sz\"uk- s\'eges funkci\'ot
megval\'os\'\i tanak.
\footnote{A szerz\H{o}k k\"osz\"onet\"uket fejezik ki az Oktat\'asi
Miniszt\'eriumnak, amely az NKFP 2/017/2001 projekt keret\'eben az itt
ismertet\'esre ker\"ul\H{o} kutat\'ast t\'amogatta.}
\\[2mm]
{\bf Kulcsszavak:} inform\'aci\'okinyer\'es, term\'eszetesnyelv-feldolgoz\'as,
felsz\'\i ni szintaktikai elemz\'es
\end{abstract}

%
\section{Bevezet\'es}
%
Az inform\'aci\'okinyer\'es (IE, Information Extraction) technol\'ogi\'aj\'anak
kutat\'asa dinamikusan fejl\H{o}d\H{o} ter\"ulet a term\'eszetesnyelv-feldolgoz\'asban.
Az Interneten megjelen\H{o} hatalmas inform\'aci\'ot\"omeg g\'epi feldolgoz\'asa \'es a
k\'\i v\'ant inform\'aci\'o t\"om\"or form\'aban t\"ort\'en\H{o} \"osszegy\H{u}jt\'ese
napi sz\"uks\'eglet, amelyre a gazdas\'ag, a tudom\'any, a politika, de ak\'ar a
h\'\i rszerz\'es ter\"ulet\'en is van ig\'eny. M\'\i g az inform\'aci\'o visszakeres\'es
(IR, Information Retrieval), amely a webes keres\H{o} prog- ramok jellemz\H{o}
tev\'ekenys\'ege, arra ir\'anyul, hogy a felhaszn\'al\'o ig\'enyeinek megfelel\H{o}
dokumentumokat v\'altozatlan form\'aban bocs\'assa rendelkez\'esre, addig
az inform\'aci\'okinyer\'es c\'elja a megtal\'alt dokumentumokban a l\'enyeges
inform\'aci\'o megjel\"ol\'ese, majd \"osszegy\H{u}jt\'ese. A sz\'am\'\i t\'og\'eppel
t\'amogatott sz\"ovegt\"om\"or\'\i t\'es, kivonatol\'as \'es az inform\'aci\'okinyer\'es
szoros kapcsolatban \'all egym\'assal.

\section{Matematikai k\'epletek \'es formul\'ak}

\begin{theorem}
Tegy\"uk fel, hogy $H$ = $C^{2}$ \'es
$\left(a_{\infty},b_{\infty}\right)$-subquadratkus a v\'egtelenben. Legyen
$\xi_{1},\allowbreak\dots,\allowbreak\xi_{N}$  egyens\'ulyi pontok,
azaz a $H' (\xi ) = 0$ megold\'asai.
Jel\"olje $\omega_{k}$ a $H'' \left(\xi_{k}\right)$ legkisebb saj\'at\'ert\'ek\'et,
\'es legyen:
\begin{equation}
  \omega : = {\rm Min\,} \left\{\omega_{1},\dots,\omega_{k}\right\}\ .
\end{equation}
Ha:
\begin{equation}
  \frac{T}{2\pi} b_{\infty} <
  - E \left[- \frac{T}{2\pi}a_{\infty}\right] <
  \frac{T}{2\pi}\omega
  \label{eq:three}
\end{equation}
akkor $\psi$ minimaliz\'al\'asa egy nem-konstant $T$-periodikus megold\'ast eredm\'enyez
$\overline{x}$.
\end{theorem}
%

Eml\'ekeztet\"unk r\'a m\'egegyszer, hogy $\alpha \in \bbbr$ eset\'en az $E [\alpha ]$ 
eg\'eszr\'esz\'en az $a\in \bbbz$ \'ertj\"uk, ahol $a< \alpha \le a+1$. P\'eld\'aul,
ha $a_{\infty} = 0$, akkor a t\'etel azt \'all\'itja sz\'amunkra, hogy $\overline{x}$
l\'etezik \'es nem konstans, felt\'eve, hogy teljes\"ul:

\begin{equation}
  \frac{T}{2\pi} b_{\infty} < 1 < \frac{T}{2\pi}
\end{equation}
vagy
\begin{equation}
  T\in \left(\frac{2\pi}{\omega},\frac{2\pi}{b_{\infty}}\right)\ .
  \label{eq:four}
\end{equation}

\subsection{Az inform\'aci\'okinyer\'eshez kapcsol\'od\'o modulok fejleszt\'es\'et 
t\'amogat\'o adatb\'azis: a Szeged Korpusz 2.0}

Az Oktat\'asi Miniszt\'erum \'altal t\'amogatott IKTA 27/2000 projekt
keret\'eben k\'esz\"ult el a Szeged Korpusz 1.0-s v\'altozata\cite{alexin:eacl2003},
amely egy sz\'ofajilag elemzett,
majd k\'ezzel egy\'ertem\H{u}s\'\i tett adatb\'azis volt. Ezt az inform\'aci\'okinyer\'esi
kutat\'asok t\'amogat\'as\'ara az MTA Nyelvtudom\'anyi Int\'ezettel \'es a
MorphoLogic Kft.-vel k\"oz\"os konzorcium jelent\H{o}sen tov\'abbfejlesztette.
A Szeged Korpusz \'ujabb v\'alto- zata
\footnote{SZTE, Informatikai Tansz\'ekcsoport, Nyelvtechnol\'ogiai
Csoport: \url{http://www.inf.u-szeged.hu/hlt}}
hat k\"ul\"onb\"oz\H{o} t\'emak\"orben
gy\H{u}jt\"ott, \"osszesen 1,2 milli\'o sz\'ot tartalmaz\'o, sz\'am\'\i t\'og\'eppel
feldolgozhat\'o sz\"oveg. Ennek az \'allom\'anynak egy mintegy 200 ezer szavas
r\'esz\'et k\'epezi a bevezet\H{o}ben m\'ar eml\'\i tett 6453 MTI r\"ovidh\'\i rt
tartalmaz\'o anyag.

\begin{figure}[ht]
\begin{verbatim}
<xml>
  <sentence id="1.1">
    <word>A<mscat>NE</mscat></word>
    <word>kutya<mscat>FN</mscat></word>
    <word>ugat<mscat>IGE</mscat></word>
    <punctuation>.</punctuation>
  </sentence>
</xml>
\end{verbatim}	
	\caption{Egy XML \'allom\'any r\'eszlete}
	\label{fig:xmlsample}
\end{figure}

\subsection{A beolvasott sz\"oveg szegment\'al\'as\'at v\'egz\H{o} modul}

A beolvasott sz\"oveg XML adatb\'aziss\'a alak\'\i t\'asa \'es az alapvet\H{o}
metainform\'aci\'ok (fejezet-, bekezd\'es-, mondat-, sz\'ostrukt\'ura)
meghat\'aroz\'asa a feldolgoz\'as els\H{o} l\'epcs\H{o}je\cite{bibok:mszny2003}
\cite{mihaczi:mszny2003}. A term\'eszetes nyelvi sz\"ovegekben sz\'amos
k\"ul\"onb\"oz\H{o} fajta sz\'o jelleg\H{u}, de a sz\'ot\'arakban nem szerepl\H{o}
lexikai elem tal\'alhat\'o (sz\'am, d\'atum, g\'epkocsirendsz\'am, e-mail c\'\i m, stb.),
amelyek felismer\'es\'ere valamint a mondat- hat\'arok meg\'allap\'\i t\'as\'ara
egy form\'alis-nyelvi eszk\"oz\"oket alkalmaz\'o modul k\'esz\"ult. A modul a GNU Flex
\footnote{A GNU Flex honlapja: \url{http://www.gnu.org/software/flex}}
regul\'aris automatagener\'ator eszk\"ozt haszn\'alja. Ebben regul\'aris
kifejez\'esek \'\i rj\'ak le az \'un. tokeneket (\ref{fig:regkif}. \'abra).
A {\it flex} program a regul\'aris kifejez\'esekb\H{o}l C programot k\'esz\'\i t,
amely a szegment\'al\'o modul magj\'at alkotja. A mondatokra \'es a szavakra bont\'as
hat\'asfoka igen j\'o, a hib\'asan felismert tokenek ar\'anya nem t\"obb, mint 0,5\%.

\begin{figure}[ht]
\raggedright
\begin{tt}
%/* sz\'okoz\"okkel tagolt sz\'amok, pl. 12 000 */\\
%NUMSPACE [0-9]\{1,3\}((\{SPACE\}|"\&thinsp;")[0-9]\{3\})+\\
%\ \\
/* ponttal tagolt sz\'amok, pl. 12.000 */\\
NUMDOT [0-9]\{1,3\}("."[0-9]\{3\})+\\
NUMDIGIT ([0-9]+",")?[0-9]+
\end{tt}	
\vspace{-2mm}
	\caption{Regul\'aris kifejez\'esek a Flex defin\'\i ci\'os f\'ajlj\'aban}
	\label{fig:regkif}
\end{figure}
\vspace{-3mm}

\section{A felhaszn\'al\'oi fel\"ulet \'es a webes megjelen\'\i t\H{o} modul}

A programcsomagban az utols\'o modul a felhaszn\'al\'oi fel\"ulet, amely
egy HTML nyelv\H{u} weblapot k\'esz\'\i t. Ez egyr\'eszt tartalmazza
a beolvasott nyers sz\"oveget, m\'asr\'eszt a programcsomag \'altal hozz\'aadott
metainform\'aci\'okat. Ez ut\'obbiakat grafikus eszk\"oz\"okkel jelen\'\i ti meg.
A modatokban azonos\'\i  tott szerepl\H{o}k k\"ul\"onb\"oz\H{o} sz\'\i nekkel,
a szerepek nevei a weblapon lebeg\H{o} \"uzenetablakokban ({\em tooltip}ekben) jelennek
meg. A mondatokra illesztett szemantikus keret \"osszes szerepl\H{o}j\'enek
sz\'ama \'es az azonos\'\i tott szerepl\H{o}k sz\'ama a mondatok ut\'an
tal\'alhat\'o. A \ref{fig:weboutput}. \'abr\'an a webes megjelen\'\i t\H{o} modul
egy k\'eperny\H{o}je l\'athat\'o.

\begin{figure}
	\begin{center}
		\includegraphics[width=\linewidth]{weboutput.png}
	\end{center}
  \vspace{-3mm}
	\caption{A webes megjelen\'\i t\H{o} modul egy k\'eperny\H{o}je}
  \vspace{-4mm}
	\label{fig:weboutput}
\end{figure}

\begin{table}
\caption{Ezt a p\'eld\'at a {\it The \TeX{}book,} 246. oldal\'ar\'ol vett\"uk}
\vspace{2pt}
\begin{tabular}{r@{\quad}rl}
\hline
\multicolumn{1}{l}{\rule{0pt}{12pt}
  \'Ev}&\multicolumn{2}{l}{A vil\'ag n\'epess\'ege}\\[2pt]
\hline\rule{0pt}{12pt}
8000 B.C.  &     5,000,000& \\
  50 A.D.  &   200,000,000& \\
1650 A.D.  &   500,000,000& \\
1945 A.D.  & 2,300,000,000& \\
1980 A.D.  & 4,400,000,000& \\[2pt]
\hline
\end{tabular}
\end{table}

Egy alternat\'\i v megjelen\'\i t\H{o} modul az eredm\'enyeket nem
weblapon, hanem Excel-ablakban jelen\'\i ti meg. Az azonos esem\'enyeket
egy munkalapon, a mondatokat egy-egy sorban, az azonos szerepl\H{o}ket pedig
azonos oszlopokban jelen\'\i ti meg. Ez a t\'abl\'azat tov\'abbi
feldolgoz\'asok kiindul\'opontja lehet.

\section{Eredm\'enyek}

A cikkben bemutatott programcsomag tesztel\'es\'ere a kutat\'ok egy keretrend- szert
(benchmark) k\'esz\'\i tettek. Ez k\'ezzel el\H{o}re annot\'alt, a rendszer
sz\'am\'ara ismeretlen mondatokat tartalmaz k\'et el\H{o}re kiv\'alasztott
t\'emak\"orben: a tulajdonos-v\'alt\'as \'es az \'uj telephely
nyit\'asa t\'emak\"or\'eben.
\footnote{A benchmarkban 176 r\"ovidh\'\i r, illetve 285 mondat szerepel.}
Erre a k\'et t\'emak\"orre megfelel\H{o} sz\'am\'u
szemantikuskeret-defin\'\i ci\'o \'es r\"ovidh\'\i r \'allt rendelkez\'esre.
A tesztmondatok sz\"oveges alakj\'ara lefuttatt\'ak a programcsomag egyes
komponenseit, majd \"osszehasonl\'\i tott\'ak a k\'ezzel k\'esz\'\i tett
\'es a g\'ep \'altal el\H{o}\'all\'\i tott k\'et \'allom\'any metainform\'aci\'oit.
Tekintve, hogy a programcsomag t\"obb elemb\H{o}l \'all, az egyes modulok hib\'aja
kumul\'al\'odik a v\'egeredm\'enyben. Az eredm\'enyek megb\'\i zhat\'obb
\'ert\'ekel\'ese \'erdek\'eben arra is van lehet\H{o}s\'eg, hogy az egyes
modulok \'altal adott r\'eszeredm\'enyeket k\"ul\"on \'ert\'ekelj\'ek, \'es
\"osszehasonl\'\i ts\'ak az etalonf\'ajllal.

\section{K\"osz\"onetnyilv\'an\'\i t\'as}
A szerz\H{o}k ez\'uton fejezik ki k\"osz\"onet\"uket az OM NKFP 2/017/2001
projektbeli konzorciumi partnereiknek, az MTA Nyelvtudom\'anyi Int\'ezet
Korpusznyelv\'eszeti Oszt\'aly\'anak \'es a MorphoLogic Kft.-nek, akikkel
a tudom\'anyos \'es szakmai kap- csolatokon t\'ul szoros, szem\'elyes
kapcsolatot alak\'\i tottak ki.
%
% ---- Bibliography ----
%
\begin{thebibliography}{15}
%
\bibitem{alexin:eacl2003}
Alexin Z., Csirik, J., Gyim\'othy, T., Bibok K., Hatvani, Cs.,
Pr\'osz\'eky, G., Tihanyi, L.:
Manually Annotated Hungarian Corpus.
in Proc. of the Research Note Sessions of the 10th Conference
of the European Chapter of the Association for Computational
Linguistics EACL'03, Budapest, Hungary, 53--56 (2003).
%
\bibitem{bibok:mszny2003}
Bibok K.:
A sz\'or\'ol \'es a sz\'ofajokr\'ol (a sz\'am\'\i t\'og\'epes nyelvfeldolgoz\'as kap- cs\'an),
Magyar Sz\'am\'\i t\'og\'epes Nyelv\'eszeti Konferencia (MSZNY 2003),
Szeged, Magyarorsz\'ag, 31--36, (2003).
%
\bibitem{farkas:mszny2004}
Farkas R., Konczer K., Szarvas Gy.:
Szemantikus keretilleszt\'es \'es az IE rendszer automatikus ki\'ert\'ekel\'se
Magyar Sz\'am\'\i t\'og\'epes Nyelv\'eszeti Konferencia (MSZNY 2004),
bek\"uldve, Szeged, Magyarorsz\'ag, (2004).
%
\bibitem{hocza:acta2004}
H\'ocza, A.:
Noun Phrase Recognition with Tree Patterns
elfogadva az Acta Cybernetica c. lapban t\"ort\'en\H{o} megjelen\'esre (2004).
%
\bibitem{mihaczi:mszny2003}
Mih\'aczi Andr\'as, N\'emeth L\'aszl\'o, R\'acz Mikl\'os:
Magyar sz\"ovegek tem\'eszetes nyelvi el\H{o}feldolgoz\'asa
Magyar Sz\'am\'\i t\'og\'epes Nyelv\'eszeti Konferencia (MSZNY 2003),
Szeged, Magyarorsz\'ag, 38--43, (2003).
%
\bibitem{proszeky:mszny2003}
Pr\'osz\'eky G.:
Automatikus inform\'aci\'oszerz\'es gazdas\'agi r\"ovidh\'\i rekb\H{o}l.
Magyar Sz\'am\'\i t\'og\'epes Nyelv\'eszeti Konferencia (MSZNY 2003),
Szeged, Magyarorsz\'ag, 161--166, (2003).
%
\bibitem{proszeky:neumann2003}
Pr\'osz\'eky G.:
Automatikus inform\'aci\'oszerz\'es gazdas\'agi-politikai r\"ovidh\'\i rekb\H{o}l.
VIII. Orsz\'agos (Centen\'ariumi) Neumann Kongresszus kiadv\'anya,
Budapest, Magyarorsz\'ag, 359--367, (2003).
%
\end{thebibliography}
\end{document}

### 2012-10-01 paper/MSZNY2013/abstract.tex
mszny absztrakt kezdes

--- 2012-10-01 paper/MSZNY2013/mszny2013.tex
@@
 
-% Ezek egyikÈnek bekapcsol·s·val a magyar Èkezetes karakterek
-% kˆzvetlen¸l is felhaszn·lhatÛk:
-%\usepackage[latin2]{inputenc} 
 \usepackage[utf8]{inputenc}
@@
 \pagestyle{myheadings}
-\def\leftmark{{\rm IV. Magyar Sz\'am\'\i t\'og\'epes Nyelv\'eszeti Konferencia}}
-\def\rightmark{{\rm Szeged, d·tum}}
+\def\leftmark{{\rm IX. Magyar Sz\'am\'\i t\'og\'epes Nyelv\'eszeti Konferencia}}
+\def\rightmark{{\rm Szeged, 2013. január 7-8.}}
 
@@
 
-\title{\ \break Programcsomag inform\'aci\'okinyer\'esi\break
-kutat\'asok t\'amogat\'as\'ara}
-\author{Alexin Zolt\'an\inst{1}, Gyim\'othy Tibor\inst{1}}
-\institute{Szegedi Tudom\'anyegyetem, TTK, Informatikai Tansz\'ekcsoport,\break
-Szeged \'Arp\'ad t\'er 2., e-mail:\{alexin,gyimothy\}@inf.u-szeged.hu}
+\title{\ \break Automatikusan tulajdonnév-annotált korpuszok előállítása a Wikipédiából}
+\author{Nemeskey Dávid Márk\inst{1}, Simon Eszter\inst{2}}
+\institute{
+\inst{}%
+MTA SZTAKI \break
+1111 Budapest, Lágymányosi utca 11., e-mail:nemeskey@sztaki.hu \break
+\and
+\inst{}%
+MTA Nyelvtudományi Intézet \break
+1068 Budapest, Benczúr u. 33., e-mail: simon.eszter@nytud.mta.hu}
 
@@
 
-\begin{abstract}
-A publik\'aci\'oban bemutat\'asra ker\"ul egy inform\'aci\'okinyer\'esi
-kutat\'asokat t\'amogat\'o programcsomag, amelynek moduljai a
-nyers sz\"oveg beolvas\'as\'at\'ol kezdve a v\'egeredm\'eny
-webes megjelen\'\i t\'es\'eig minden sz\"uk- s\'eges funkci\'ot
-megval\'os\'\i tanak.
-\footnote{A szerz\H{o}k k\"osz\"onet\"uket fejezik ki az Oktat\'asi
-Miniszt\'eriumnak, amely az NKFP 2/017/2001 projekt keret\'eben az itt
-ismertet\'esre ker\"ul\H{o} kutat\'ast t\'amogatta.}
-\\[2mm]
-{\bf Kulcsszavak:} inform\'aci\'okinyer\'es, term\'eszetesnyelv-feldolgoz\'as,
-felsz\'\i ni szintaktikai elemz\'es
-\end{abstract}
+\section{Introduction}
 
-%
-\section{Bevezet\'es}
-%
-Az inform\'aci\'okinyer\'es (IE, Information Extraction) technol\'ogi\'aj\'anak
-kutat\'asa dinamikusan fejl\H{o}d\H{o} ter\"ulet a term\'eszetesnyelv-feldolgoz\'asban.
-Az Interneten megjelen\H{o} hatalmas inform\'aci\'ot\"omeg g\'epi feldolgoz\'asa \'es a
-k\'\i v\'ant inform\'aci\'o t\"om\"or form\'aban t\"ort\'en\H{o} \"osszegy\H{u}jt\'ese
-napi sz\"uks\'eglet, amelyre a gazdas\'ag, a tudom\'any, a politika, de ak\'ar a
-h\'\i rszerz\'es ter\"ulet\'en is van ig\'eny. M\'\i g az inform\'aci\'o visszakeres\'es
-(IR, Information Retrieval), amely a webes keres\H{o} prog- ramok jellemz\H{o}
-tev\'ekenys\'ege, arra ir\'anyul, hogy a felhaszn\'al\'o ig\'enyeinek megfelel\H{o}
-dokumentumokat v\'altozatlan form\'aban bocs\'assa rendelkez\'esre, addig
-az inform\'aci\'okinyer\'es c\'elja a megtal\'alt dokumentumokban a l\'enyeges
-inform\'aci\'o megjel\"ol\'ese, majd \"osszegy\H{u}jt\'ese. A sz\'am\'\i t\'og\'eppel
-t\'amogatott sz\"ovegt\"om\"or\'\i t\'es, kivonatol\'as \'es az inform\'aci\'okinyer\'es
-szoros kapcsolatban \'all egym\'assal.
+\section{Wikipedia and NER}
 
-\section{Matematikai k\'epletek \'es formul\'ak}
+\section{Creating the English Corpus}
 
-\begin{theorem}
-Tegy\"uk fel, hogy $H$ = $C^{2}$ \'es
-$\left(a_{\infty},b_{\infty}\right)$-subquadratkus a v\'egtelenben. Legyen
-$\xi_{1},\allowbreak\dots,\allowbreak\xi_{N}$  egyens\'ulyi pontok,
-azaz a $H' (\xi ) = 0$ megold\'asai.
-Jel\"olje $\omega_{k}$ a $H'' \left(\xi_{k}\right)$ legkisebb saj\'at\'ert\'ek\'et,
-\'es legyen:
-\begin{equation}
-  \omega : = {\rm Min\,} \left\{\omega_{1},\dots,\omega_{k}\right\}\ .
-\end{equation}
-Ha:
-\begin{equation}
-  \frac{T}{2\pi} b_{\infty} <
-  - E \left[- \frac{T}{2\pi}a_{\infty}\right] <
-  \frac{T}{2\pi}\omega
-  \label{eq:three}
-\end{equation}
-akkor $\psi$ minimaliz\'al\'asa egy nem-konstant $T$-periodikus megold\'ast eredm\'enyez
-$\overline{x}$.
-\end{theorem}
-%
+\section{Creating the Hungarian Corpus}
 
-Eml\'ekeztet\"unk r\'a m\'egegyszer, hogy $\alpha \in \bbbr$ eset\'en az $E [\alpha ]$ 
-eg\'eszr\'esz\'en az $a\in \bbbz$ \'ertj\"uk, ahol $a< \alpha \le a+1$. P\'eld\'aul,
-ha $a_{\infty} = 0$, akkor a t\'etel azt \'all\'itja sz\'amunkra, hogy $\overline{x}$
-l\'etezik \'es nem konstans, felt\'eve, hogy teljes\"ul:
+\section{Data description}
 
-\begin{equation}
-  \frac{T}{2\pi} b_{\infty} < 1 < \frac{T}{2\pi}
-\end{equation}
-vagy
-\begin{equation}
-  T\in \left(\frac{2\pi}{\omega},\frac{2\pi}{b_{\infty}}\right)\ .
-  \label{eq:four}
-\end{equation}
+\section{Evaluation}
 
-\subsection{Az inform\'aci\'okinyer\'eshez kapcsol\'od\'o modulok fejleszt\'es\'et 
-t\'amogat\'o adatb\'azis: a Szeged Korpusz 2.0}
+\section{Conclusion}
 
-Az Oktat\'asi Miniszt\'erum \'altal t\'amogatott IKTA 27/2000 projekt
-keret\'eben k\'esz\"ult el a Szeged Korpusz 1.0-s v\'altozata\cite{alexin:eacl2003},
-amely egy sz\'ofajilag elemzett,
-majd k\'ezzel egy\'ertem\H{u}s\'\i tett adatb\'azis volt. Ezt az inform\'aci\'okinyer\'esi
-kutat\'asok t\'amogat\'as\'ara az MTA Nyelvtudom\'anyi Int\'ezettel \'es a
-MorphoLogic Kft.-vel k\"oz\"os konzorcium jelent\H{o}sen tov\'abbfejlesztette.
-A Szeged Korpusz \'ujabb v\'alto- zata
-\footnote{SZTE, Informatikai Tansz\'ekcsoport, Nyelvtechnol\'ogiai
-Csoport: \url{http://www.inf.u-szeged.hu/hlt}}
-hat k\"ul\"onb\"oz\H{o} t\'emak\"orben
-gy\H{u}jt\"ott, \"osszesen 1,2 milli\'o sz\'ot tartalmaz\'o, sz\'am\'\i t\'og\'eppel
-feldolgozhat\'o sz\"oveg. Ennek az \'allom\'anynak egy mintegy 200 ezer szavas
-r\'esz\'et k\'epezi a bevezet\H{o}ben m\'ar eml\'\i tett 6453 MTI r\"ovidh\'\i rt
-tartalmaz\'o anyag.
-
-\begin{figure}[ht]
-\begin{verbatim}
-<xml>
-  <sentence id="1.1">
-    <word>A<mscat>NE</mscat></word>
-    <word>kutya<mscat>FN</mscat></word>
-    <word>ugat<mscat>IGE</mscat></word>
-    <punctuation>.</punctuation>
-  </sentence>
-</xml>
-\end{verbatim}	
-	\caption{Egy XML \'allom\'any r\'eszlete}
-	\label{fig:xmlsample}
-\end{figure}
-
-\subsection{A beolvasott sz\"oveg szegment\'al\'as\'at v\'egz\H{o} modul}
-
-A beolvasott sz\"oveg XML adatb\'aziss\'a alak\'\i t\'asa \'es az alapvet\H{o}
-metainform\'aci\'ok (fejezet-, bekezd\'es-, mondat-, sz\'ostrukt\'ura)
-meghat\'aroz\'asa a feldolgoz\'as els\H{o} l\'epcs\H{o}je\cite{bibok:mszny2003}
-\cite{mihaczi:mszny2003}. A term\'eszetes nyelvi sz\"ovegekben sz\'amos
-k\"ul\"onb\"oz\H{o} fajta sz\'o jelleg\H{u}, de a sz\'ot\'arakban nem szerepl\H{o}
-lexikai elem tal\'alhat\'o (sz\'am, d\'atum, g\'epkocsirendsz\'am, e-mail c\'\i m, stb.),
-amelyek felismer\'es\'ere valamint a mondat- hat\'arok meg\'allap\'\i t\'as\'ara
-egy form\'alis-nyelvi eszk\"oz\"oket alkalmaz\'o modul k\'esz\"ult. A modul a GNU Flex
-\footnote{A GNU Flex honlapja: \url{http://www.gnu.org/software/flex}}
-regul\'aris automatagener\'ator eszk\"ozt haszn\'alja. Ebben regul\'aris
-kifejez\'esek \'\i rj\'ak le az \'un. tokeneket (\ref{fig:regkif}. \'abra).
-A {\it flex} program a regul\'aris kifejez\'esekb\H{o}l C programot k\'esz\'\i t,
-amely a szegment\'al\'o modul magj\'at alkotja. A mondatokra \'es a szavakra bont\'as
-hat\'asfoka igen j\'o, a hib\'asan felismert tokenek ar\'anya nem t\"obb, mint 0,5\%.
-
-\begin{figure}[ht]
-\raggedright
-\begin{tt}
-%/* sz\'okoz\"okkel tagolt sz\'amok, pl. 12 000 */\\
-%NUMSPACE [0-9]\{1,3\}((\{SPACE\}|"\&thinsp;")[0-9]\{3\})+\\
-%\ \\
-/* ponttal tagolt sz\'amok, pl. 12.000 */\\
-NUMDOT [0-9]\{1,3\}("."[0-9]\{3\})+\\
-NUMDIGIT ([0-9]+",")?[0-9]+
-\end{tt}	
-\vspace{-2mm}
-	\caption{Regul\'aris kifejez\'esek a Flex defin\'\i ci\'os f\'ajlj\'aban}
-	\label{fig:regkif}
-\end{figure}
-\vspace{-3mm}
-
-\section{A felhaszn\'al\'oi fel\"ulet \'es a webes megjelen\'\i t\H{o} modul}
-
-A programcsomagban az utols\'o modul a felhaszn\'al\'oi fel\"ulet, amely
-egy HTML nyelv\H{u} weblapot k\'esz\'\i t. Ez egyr\'eszt tartalmazza
-a beolvasott nyers sz\"oveget, m\'asr\'eszt a programcsomag \'altal hozz\'aadott
-metainform\'aci\'okat. Ez ut\'obbiakat grafikus eszk\"oz\"okkel jelen\'\i ti meg.
-A modatokban azonos\'\i  tott szerepl\H{o}k k\"ul\"onb\"oz\H{o} sz\'\i nekkel,
-a szerepek nevei a weblapon lebeg\H{o} \"uzenetablakokban ({\em tooltip}ekben) jelennek
-meg. A mondatokra illesztett szemantikus keret \"osszes szerepl\H{o}j\'enek
-sz\'ama \'es az azonos\'\i tott szerepl\H{o}k sz\'ama a mondatok ut\'an
-tal\'alhat\'o. A \ref{fig:weboutput}. \'abr\'an a webes megjelen\'\i t\H{o} modul
-egy k\'eperny\H{o}je l\'athat\'o.
-
-\begin{figure}
-	\begin{center}
-		\includegraphics[width=\linewidth]{weboutput.png}
-	\end{center}
-  \vspace{-3mm}
-	\caption{A webes megjelen\'\i t\H{o} modul egy k\'eperny\H{o}je}
-  \vspace{-4mm}
-	\label{fig:weboutput}
-\end{figure}
-
-\begin{table}
-\caption{Ezt a p\'eld\'at a {\it The \TeX{}book,} 246. oldal\'ar\'ol vett\"uk}
-\vspace{2pt}
-\begin{tabular}{r@{\quad}rl}
-\hline
-\multicolumn{1}{l}{\rule{0pt}{12pt}
-  \'Ev}&\multicolumn{2}{l}{A vil\'ag n\'epess\'ege}\\[2pt]
-\hline\rule{0pt}{12pt}
-8000 B.C.  &     5,000,000& \\
-  50 A.D.  &   200,000,000& \\
-1650 A.D.  &   500,000,000& \\
-1945 A.D.  & 2,300,000,000& \\
-1980 A.D.  & 4,400,000,000& \\[2pt]
-\hline
-\end{tabular}
-\end{table}
-
-Egy alternat\'\i v megjelen\'\i t\H{o} modul az eredm\'enyeket nem
-weblapon, hanem Excel-ablakban jelen\'\i ti meg. Az azonos esem\'enyeket
-egy munkalapon, a mondatokat egy-egy sorban, az azonos szerepl\H{o}ket pedig
-azonos oszlopokban jelen\'\i ti meg. Ez a t\'abl\'azat tov\'abbi
-feldolgoz\'asok kiindul\'opontja lehet.
-
-\section{Eredm\'enyek}
-
-A cikkben bemutatott programcsomag tesztel\'es\'ere a kutat\'ok egy keretrend- szert
-(benchmark) k\'esz\'\i tettek. Ez k\'ezzel el\H{o}re annot\'alt, a rendszer
-sz\'am\'ara ismeretlen mondatokat tartalmaz k\'et el\H{o}re kiv\'alasztott
-t\'emak\"orben: a tulajdonos-v\'alt\'as \'es az \'uj telephely
-nyit\'asa t\'emak\"or\'eben.
-\footnote{A benchmarkban 176 r\"ovidh\'\i r, illetve 285 mondat szerepel.}
-Erre a k\'et t\'emak\"orre megfelel\H{o} sz\'am\'u
-szemantikuskeret-defin\'\i ci\'o \'es r\"ovidh\'\i r \'allt rendelkez\'esre.
-A tesztmondatok sz\"oveges alakj\'ara lefuttatt\'ak a programcsomag egyes
-komponenseit, majd \"osszehasonl\'\i tott\'ak a k\'ezzel k\'esz\'\i tett
-\'es a g\'ep \'altal el\H{o}\'all\'\i tott k\'et \'allom\'any metainform\'aci\'oit.
-Tekintve, hogy a programcsomag t\"obb elemb\H{o}l \'all, az egyes modulok hib\'aja
-kumul\'al\'odik a v\'egeredm\'enyben. Az eredm\'enyek megb\'\i zhat\'obb
-\'ert\'ekel\'ese \'erdek\'eben arra is van lehet\H{o}s\'eg, hogy az egyes
-modulok \'altal adott r\'eszeredm\'enyeket k\"ul\"on \'ert\'ekelj\'ek, \'es
-\"osszehasonl\'\i ts\'ak az etalonf\'ajllal.
-
-\section{K\"osz\"onetnyilv\'an\'\i t\'as}
-A szerz\H{o}k ez\'uton fejezik ki k\"osz\"onet\"uket az OM NKFP 2/017/2001
-projektbeli konzorciumi partnereiknek, az MTA Nyelvtudom\'anyi Int\'ezet
-Korpusznyelv\'eszeti Oszt\'aly\'anak \'es a MorphoLogic Kft.-nek, akikkel
-a tudom\'anyos \'es szakmai kap- csolatokon t\'ul szoros, szem\'elyes
-kapcsolatot alak\'\i tottak ki.
 %

### 2012-10-01 paper/MSZNY2013/abstract.tex
Added bibliography to the article.

@@
 
-\section{Creating the English Corpus}
-
-\section{Creating the Hungarian Corpus}
+\section{Creating the Corpus}
+\subsection{Creating the English Corpus}
 
@@
 
+\cite{Nothman:08} \cite{Szarvas:06} \cite{Medelyan:09}
+
 %
@@
 %
-\begin{thebibliography}{15}
-%
-\bibitem{alexin:eacl2003}
-Alexin Z., Csirik, J., Gyim\'othy, T., Bibok K., Hatvani, Cs.,
-Pr\'osz\'eky, G., Tihanyi, L.:
-Manually Annotated Hungarian Corpus.
-in Proc. of the Research Note Sessions of the 10th Conference
-of the European Chapter of the Association for Computational
-Linguistics EACL'03, Budapest, Hungary, 53--56 (2003).
-%
-\bibitem{bibok:mszny2003}
-Bibok K.:
-A sz\'or\'ol \'es a sz\'ofajokr\'ol (a sz\'am\'\i t\'og\'epes nyelvfeldolgoz\'as kap- cs\'an),
-Magyar Sz\'am\'\i t\'og\'epes Nyelv\'eszeti Konferencia (MSZNY 2003),
-Szeged, Magyarorsz\'ag, 31--36, (2003).
-%
-\bibitem{farkas:mszny2004}
-Farkas R., Konczer K., Szarvas Gy.:
-Szemantikus keretilleszt\'es \'es az IE rendszer automatikus ki\'ert\'ekel\'se
-Magyar Sz\'am\'\i t\'og\'epes Nyelv\'eszeti Konferencia (MSZNY 2004),
-bek\"uldve, Szeged, Magyarorsz\'ag, (2004).
-%
-\bibitem{hocza:acta2004}
-H\'ocza, A.:
-Noun Phrase Recognition with Tree Patterns
-elfogadva az Acta Cybernetica c. lapban t\"ort\'en\H{o} megjelen\'esre (2004).
-%
-\bibitem{mihaczi:mszny2003}
-Mih\'aczi Andr\'as, N\'emeth L\'aszl\'o, R\'acz Mikl\'os:
-Magyar sz\"ovegek tem\'eszetes nyelvi el\H{o}feldolgoz\'asa
-Magyar Sz\'am\'\i t\'og\'epes Nyelv\'eszeti Konferencia (MSZNY 2003),
-Szeged, Magyarorsz\'ag, 38--43, (2003).
-%
-\bibitem{proszeky:mszny2003}
-Pr\'osz\'eky G.:
-Automatikus inform\'aci\'oszerz\'es gazdas\'agi r\"ovidh\'\i rekb\H{o}l.
-Magyar Sz\'am\'\i t\'og\'epes Nyelv\'eszeti Konferencia (MSZNY 2003),
-Szeged, Magyarorsz\'ag, 161--166, (2003).
-%
-\bibitem{proszeky:neumann2003}
-Pr\'osz\'eky G.:
-Automatikus inform\'aci\'oszerz\'es gazdas\'agi-politikai r\"ovidh\'\i rekb\H{o}l.
-VIII. Orsz\'agos (Centen\'ariumi) Neumann Kongresszus kiadv\'anya,
-Budapest, Magyarorsz\'ag, 359--367, (2003).
-%
-\end{thebibliography}
+\bibliographystyle{huplain}
+\bibliography{abstract}
+
 \end{document}

### 2012-10-01 paper/MSZNY2013/abstract.tex
intro

--- 2012-10-01 paper/MSZNY2013/abstract.tex
@@
 MTA SZTAKI \break
-1111 Budapest, Lágymányosi utca 11., e-mail:nemeskey@sztaki.hu \break
+1111 Budapest, Lágymányosi utca 11., e-mail:nemeskey@sztaki.mta.hu \break
 \and
@@
 
+Az automatikus tulajdonnév-felismerés (Named Entity Recognition, NER) a természetes nyelv feldolgozását célzó alkalmazások közül az egyik legnépszerűbb, mivel hatékonyan automatizálható, és eredménye hasznos bemenete különböző magasabb szintű információkinyerő és -feldolgozó rendszereknek.
+A feladat során strukturálatlan szövegben kell azonosítani és az előre definiált osztályok valamelyikébe besorolni a neveket.
+Az egyik leggyakrabban alkalmazott nemzetközi sztenderd a CoNLL-2003 verseny kiírását követi, amely szerint a négy névkategória a következő: személy-, intézmény-, földrajzi név és az ezek egyikébe sem tartozó egyéb kategória.
+
+%Named Entity Recognition (NER), the task of identifying Named Entities (NEs) in unstructured texts and classifying them into pre-selected classes, is one of the most important subtasks in many NLP tasks, such as information retrieval, information extraction or machine translation.
+MUC and CoNLL datasets and annotation schemes have been the major standards applied in the field of NER.
+
+A gold standard adathalmazok
+
+The standard datasets are highly domain-specific (mostly newswire) and are
+restricted in size. Researchers attempting to merge these datasets to get a bigger training corpus are faced with the
+problem of combining different tagsets and annotation schemes. Manually
+annotating large amounts of text with linguistic information is a
+time-consuming, highly skilled and delicate job, but large, accurately
+annotated corpora are essential for building robust supervised machine
+learning NER systems. Therefore, reducing the annotation cost is a key
+challenge.
+
+One approach is to generate the resources automatically, another one is to use
+collaborative annotation and/or collaboratively constructed resources, such as
+Wikipedia, Wiktionary, Linked Open Data, or DBpedia. In this paper we combine
+these approaches by automatically generating freely available NE tagged
+corpora from Wikipedia.
+
 \section{Wikipedia and NER}
+
+Wikipedia (WP, see {\tt http://wikipedia.org}), a free
+multilingual Internet encyclopedia, written collaboratively by volunteers, is
+a goldmine of information.
+WP has been applied to several NLP tasks such as word sense
+disambiguation, ontology and thesaurus building, and question answering (see
+Medelyan et al.~\cite{Medelyan:09} for a survey). It is recognized as one
+of the largest available collections of entities, and also as a resource that
+can improve the accuracy of NER. The most obvious utilization of WP for NER is
+extracting gazetteers containing person names, locations or organizations
+(e.g.~Toral and Mu\~noz \cite{Toral:06}). Creating dictionaries of
+entities is also a common step of NE disambiguation
+\cite{Bunescu:06,Cucerzan:07}. Both supervised and unsupervised NER systems
+use such lists, see e.g.~Nadeau et al.~\cite{Nadeau:06} The knowledge
+embodied in WP may also be incorporated in NER learning as features,
+e.g.~Kazama and Torisawa \cite{Kazama:07} showed that automatic
+extraction of category labels from WP improves the accuracy of a supervised NE
+tagger.
+Another approach to improve NER with WP is the automatic creation of training
+data. Nothman et al.~\cite{Nothman:08} used a similar method to create
+a NE annotated text in English. They transformed the WP links into NE
+annotations by classifying the target articles into standard entity
+classes. Their approach to classification is based primarily on category head
+nouns and the opening sentences of articles where definitions are often given.
+
+Our approach to recognize and classify NEs in corpora generated from WP was to
+map the DBpedia ontology classes to standard NE tags and assign these to WP
+entities. the one presented here is the first automatically NE annotated corpus for Hungarian.
 

### 2012-10-01 paper/MSZNY2013/abstract.tex
Data description added.

@@
 
+A korpusz a Wikipediához hasonlóan a Creative Commons Attribution-Sharealike
+3.0 Unported (CC-BY-SA) licensz alatt érhető el. Szabadon letölthető a
+\texttt{http://hlt.sztaki.hu} oldalról és a META-SHARE tárhelyről. Utóbbi egy nyílt,
+elosztott hálózat, amelyet a nyelvi erőforrások megosztására hoztak létre a META-NET-en,
+egy, az Európai Bizottság által támogatott Kiválósági Hálózaton belül.
+
+A korpusz vertikális formátumú (\texttt{.tsv}) szövegfájlok alkotják.
+Egy sor külön mezőkben tartalmazza a tokent, annak lemmatizált formájat és szófaját, illetve a NER címkéjét.
+Az utóbbi három mező közül csak a NER címkék ``ezüst sztenderd" minőségűek --- a többi nyelvi információt statisztikai eszközökkel generáltuk.
+
 \section{Evaluation}

### 2012-10-01 paper/MSZNY2013/abstract.tex
wiki es ner

--- 2012-10-01 paper/MSZNY2013/abstract.tex
@@
 
-A gold standard adathalmazok
+Az automatikus tulajdonnév-felismerés (NER) során alkalmazott gold standard korpuszok általában műfajspecifikusak (főleg rövidhírek), és korlátozott méretűek.
+Egy kellően robusztus alkalmazás építéséhez viszont nagy méretű és változatos nyelvi anyagot tartalmazó adathalmazra van szükség.
+A tanítókorpuszok előállítása hagyományosan kézi annotálás útján történik, ami meglehetősen sok időt, szakértelmet és pénzt kíván, vagyis kardinális kérdés az erőforrások minimalizálása.
+Nagyobb tanítóhalmazhoz juthatunk, ha már meglevő korpuszokat gyúrunk össze; ebben az esetben viszont azzal kell szembesülnünk, hogy ezek más és más annotációs útmutató alapján készültek, továbbá nem azonos címkekészletet használnak.
 
-The standard datasets are highly domain-specific (mostly newswire) and are
-restricted in size. Researchers attempting to merge these datasets to get a bigger training corpus are faced with the
-problem of combining different tagsets and annotation schemes. Manually
-annotating large amounts of text with linguistic information is a
-time-consuming, highly skilled and delicate job, but large, accurately
-annotated corpora are essential for building robust supervised machine
-learning NER systems. Therefore, reducing the annotation cost is a key
-challenge.
+Ezen problémák kiküszöbölésére egy lehetséges módszer, ha automatikusan állítunk elő  
+adathalmazokat, vagy pedig olyan közösségi tartalmakat használunk fel, mint pl.
+a Wikipedia, a Wiktionary vagy a DBPedia.
+Cikkünkben ezek kombinációját mutatjuk be: a Wikipediából előállított, automatikusan tulajdonnév-annotált korpuszt.
 
-One approach is to generate the resources automatically, another one is to use
-collaborative annotation and/or collaboratively constructed resources, such as
-Wikipedia, Wiktionary, Linked Open Data, or DBpedia. In this paper we combine
-these approaches by automatically generating freely available NE tagged
-corpora from Wikipedia.
+%The standard datasets are highly domain-specific (mostly newswire) and are
+%restricted in size. Researchers attempting to merge these datasets to get a bigger training corpus are faced with the
+%problem of combining different tagsets and annotation schemes. Manually
+%annotating large amounts of text with linguistic information is a
+%time-consuming, highly skilled and delicate job, but large, accurately
+%annotated corpora are essential for building robust supervised machine
+%learning NER systems. Therefore, reducing the annotation cost is a key
+%challenge.
+
+%One approach is to generate the resources automatically, another one is to use
+%collaborative annotation and/or collaboratively constructed resources, such as
+%Wikipedia, Wiktionary, Linked Open Data, or DBpedia. In this paper we combine
+%these approaches by automatically generating freely available NE tagged
+%corpora from Wikipedia.
 

### 2012-10-01 paper/MSZNY2013/abstract.tex
Deleted a sentence.

@@
 3.0 Unported (CC-BY-SA) licensz alatt érhető el. Szabadon letölthető a
-\texttt{http://hlt.sztaki.hu} oldalról és a META-SHARE tárhelyről. Utóbbi egy nyílt,
-elosztott hálózat, amelyet a nyelvi erőforrások megosztására hoztak létre a META-NET-en,
-egy, az Európai Bizottság által támogatott Kiválósági Hálózaton belül.
+\footnote{\texttt{http://hlt.sztaki.hu}} oldalról és a META-SHARE tárhelyről
+\footnote{\texttt{http://nlp.ipipan.waw.pl/metashare}}.
 
-A korpusz vertikális formátumú (\texttt{.tsv}) szövegfájlok alkotják.
-Egy sor külön mezőkben tartalmazza a tokent, annak lemmatizált formájat és szófaját, illetve a NER címkéjét.
-Az utóbbi három mező közül csak a NER címkék ``ezüst sztenderd" minőségűek --- a többi nyelvi információt statisztikai eszközökkel generáltuk.
+A korpusz vertikális formátumú (\texttt{.tsv}) szövegfájlok alkotják. Egy sor külön
+mezőkben tartalmazza a tokent, annak lemmatizált formájat és szófaját, illetve a NER
+címkéjét. Az utóbbi három mező közül csak a NER címkék ``ezüst sztenderd" minőségűek
+--- a többi nyelvi információt statisztikai eszközökkel generáltuk.
 

### 2012-10-01 paper/MSZNY2013/abstract.tex
wiki es ner

--- 2012-10-01 paper/MSZNY2013/abstract.tex
@@
 
-\section{Introduction}
+%\section{Introduction}
 
-Az automatikus tulajdonnév-felismerés (Named Entity Recognition, NER) a természetes nyelv feldolgozását célzó alkalmazások közül az egyik legnépszerűbb, mivel hatékonyan automatizálható, és eredménye hasznos bemenete különböző magasabb szintű információkinyerő és -feldolgozó rendszereknek.
+%Az automatikus tulajdonnév-felismerés (Named Entity Recognition, NER) a természetes nyelv feldolgozását célzó alkalmazások közül az egyik legnépszerűbb, mivel hatékonyan automatizálható, és eredménye hasznos bemenete különböző magasabb szintű információkinyerő és -feldolgozó rendszereknek.
 A feladat során strukturálatlan szövegben kell azonosítani és az előre definiált osztályok valamelyikébe besorolni a neveket.
@@
 
-\section{Wikipedia and NER}
+%\section{Wikipedia and NER}
 
-Wikipedia (WP, see {\tt http://wikipedia.org}), a free
-multilingual Internet encyclopedia, written collaboratively by volunteers, is
-a goldmine of information.
-WP has been applied to several NLP tasks such as word sense
-disambiguation, ontology and thesaurus building, and question answering (see
-Medelyan et al.~\cite{Medelyan:09} for a survey). It is recognized as one
-of the largest available collections of entities, and also as a resource that
-can improve the accuracy of NER. The most obvious utilization of WP for NER is
-extracting gazetteers containing person names, locations or organizations
-(e.g.~Toral and Mu\~noz \cite{Toral:06}). Creating dictionaries of
-entities is also a common step of NE disambiguation
-\cite{Bunescu:06,Cucerzan:07}. Both supervised and unsupervised NER systems
-use such lists, see e.g.~Nadeau et al.~\cite{Nadeau:06} The knowledge
-embodied in WP may also be incorporated in NER learning as features,
-e.g.~Kazama and Torisawa \cite{Kazama:07} showed that automatic
-extraction of category labels from WP improves the accuracy of a supervised NE
-tagger.
-Another approach to improve NER with WP is the automatic creation of training
-data. Nothman et al.~\cite{Nothman:08} used a similar method to create
-a NE annotated text in English. They transformed the WP links into NE
-annotations by classifying the target articles into standard entity
-classes. Their approach to classification is based primarily on category head
-nouns and the opening sentences of articles where definitions are often given.
+A Wikipediát, mely a legismertebb, önkéntesek által szerkesztett nyílt internetes enciklopédia, számos NLP feladathoz sikeresen alkalmazták már.
+Mivel ez az egyik legnagyobb elérhető forrása az entitásoknak, az automatikus tulajdonnév-felismerés teljesítményének növelésére több módon is kiválóan használható.
+Nyilvánvaló felhasználási módja különböző névlisták kinyerése (pl.
+\cite{Toral_2006.a}); de a Wikipediában található tudás a tanuló algoritmus számára használható jegyek formájában is beépíthető felügyelt tanulási rendszerekbe (pl.
+\cite{KaTo07}).
 
-Our approach to recognize and classify NEs in corpora generated from WP was to
-map the DBpedia ontology classes to standard NE tags and assign these to WP
-entities. the one presented here is the first automatically NE annotated corpus for Hungarian.
+Számos példát találunk arra is a nemzetközi irodalomban, amikor a Wikipediából építenek korpuszt, amelyben az entitásokra vonatkozó információt valamilyen formában megőrzik és feldolgozzák.
+Ezt a megközelítést követtük mi is, de az eddig alkalmazott módszerek helyett, melyek elsősorban a cikkek definíciójából nyerték ki az entitás kategóriájára vonatkozó információt, mi a DBPedia ontológiai osztályait képeztük le a sztenderd CoNLL névosztályokra, majd kötöttük őket Wikipédia-entitásokhoz.
+Az általunk létrehozott feldolgozólánc elemei nagyrészt nyelvfüggetlenek, vagyis módszerünkkel létrehozhatók tanítókorpuszok kevesebb erőforrással rendelkező nyelvekre is.
+Cikkünkben a feldolgozás folyamatát és eredményét: egy magyar és egy angol nyelvű automatikusan tulajdonnév-annotált korpuszt mutatunk be.
+
+%Wikipedia (WP, see {\tt http://wikipedia.org}), a free
+%multilingual Internet encyclopedia, written collaboratively by volunteers, is
+%a goldmine of information.
+%WP has been applied to several NLP tasks such as word sense
+%disambiguation, ontology and thesaurus building, and question answering (see
+%Medelyan et al.~\cite{Medelyan:09} for a survey). It is recognized as one
+%of the largest available collections of entities, and also as a resource that
+%can improve the accuracy of NER. The most obvious utilization of WP for NER is
+%extracting gazetteers containing person names, locations or organizations
+%(e.g.~Toral and Mu\~noz \cite{Toral:06}). Creating dictionaries of
+%entities is also a common step of NE disambiguation
+%\cite{Bunescu:06,Cucerzan:07}. Both supervised and unsupervised NER systems
+%use such lists, see e.g.~Nadeau et al.~\cite{Nadeau:06} The knowledge
+%embodied in WP may also be incorporated in NER learning as features,
+%e.g.~Kazama and Torisawa \cite{Kazama:07} showed that automatic
+%extraction of category labels from WP improves the accuracy of a supervised NE
+%tagger.
+
+%Another approach to improve NER with WP is the automatic creation of training
+%data. Nothman et al.~\cite{Nothman:08} used a similar method to create
+%a NE annotated text in English. They transformed the WP links into NE
+%annotations by classifying the target articles into standard entity
+%classes. Their approach to classification is based primarily on category head
+%nouns and the opening sentences of articles where definitions are often given.
+
+%Our approach to recognize and classify NEs in corpora generated from WP was to
+%map the DBpedia ontology classes to standard NE tags and assign these to WP
+%entities. the one presented here is the first automatically NE annotated corpus for Hungarian.
 

### 2012-10-01 paper/MSZNY2013/abstract.tex
Stuff added.

--- 2012-10-01 paper/MSZNY2013/abstract.tex
@@
 
+A korpuszt, hasonlóan Nothman et al.~\shortcite{Nothman:08} munkájához, a következő
+lépésekkel állítottuk elő:
+
+\begin{enumerate}
+\item a WP cikkeket entitás osztályokba soroltuk
+\item a cikkeket mondatokra bontottuk
+\item megjelöltük a névelemeket a szövegben
+\item kiszűrtük a rossz minőségű mondatokat
+\end{enumerate}
+
+Bár Nothman az első feladatot gépi tanulással oldotta meg, mi a nagyobb
+pontosság érdekében a DBpedia~\cite{Bizer:09} tudásbázis alapján osztályoztuk az
+entitásokat. A névelemeket WP kereszthivatkozások segítségével azonosítottuk. Minden
+olyan mondatot, ahol ismeretlen névelem fordult elő, eldobtunk.
+
+A fenti lépések mindegyike automatikus; emberi beavatkozást egyedül a DBpedia típusok
+névelemcímkékre történő leképezése jelent.
+
 \section{Data description}

### 2012-10-01 paper/MSZNY2013/abstract.tex
Sth. 	modified:   abstract.bib 	modified:   abstract.tex
Deleted an unneeded paragraph. 	modified:   abstract.tex

@@
 
-A korpuszt, hasonlóan Nothman et al.~\shortcite{Nothman:08} munkájához, a következő
+A korpuszt, hasonlóan Nothman et al.~\cite{Nothman:08} munkájához, a következő
 lépésekkel állítottuk elő:
@@
 
-A korpusz vertikális formátumú (\texttt{.tsv}) szövegfájlok alkotják. Egy sor külön
-mezőkben tartalmazza a tokent, annak lemmatizált formájat és szófaját, illetve a NER
-címkéjét. Az utóbbi három mező közül csak a NER címkék ``ezüst sztenderd" minőségűek
---- a többi nyelvi információt statisztikai eszközökkel generáltuk.
-
 \section{Evaluation}

### 2012-10-01 paper/MSZNY2013/abstract.tex
English + Hungarian. 	modified:   abstract.tex
evaluation

@@
 
+Az automatikus annotálás nyilvánvaló előnyei mellett meg kell vizsgálnunk azt is, hogy az eredményül kapott  ún.
+,,silver standard'' korpuszok hogyan hasznosíthatók a NER teljesítményének növelésére.
+A már említett lehetőségek (névlisták kinyerése, jegyként való beépítés) mellett ezek a korpuszok önmagukban vagy más adathalmazokkal együttesen tanítókorpuszként szolgálhatnak a kevesebb erőforrással rendelkező nyelvekre, illetve a rövidhírektől eltérő műfajokra is.
+A magyar és az angol korpuszon végzett kísérletek egyaránt azt igazolják, hogy a Wikipédiából automatikusan előállított adathalmazok javítják a NER teljesítményét.
+Ugyanakkor számos problémával szembesültünk, melyekre a jövőben megoldást kell találnunk.
+
 \section{Conclusion}

### 2012-10-01 paper/MSZNY2013/abstract.tex
sectionok kikommentelve
Done. 	modified:   abstract.tex

@@
 
-Az automatikus tulajdonnév-felismerés (NER) során alkalmazott gold standard korpuszok általában műfajspecifikusak (főleg rövidhírek), és korlátozott méretűek.
-Egy kellően robusztus alkalmazás építéséhez viszont nagy méretű és változatos nyelvi anyagot tartalmazó adathalmazra van szükség.
-A tanítókorpuszok előállítása hagyományosan kézi annotálás útján történik, ami meglehetősen sok időt, szakértelmet és pénzt kíván, vagyis kardinális kérdés az erőforrások minimalizálása.
-Nagyobb tanítóhalmazhoz juthatunk, ha már meglevő korpuszokat gyúrunk össze; ebben az esetben viszont azzal kell szembesülnünk, hogy ezek más és más annotációs útmutató alapján készültek, továbbá nem azonos címkekészletet használnak.
+Az automatikus tulajdonnév-felismerés (Named Entity Recognition, NER) során alkalmazott
+gold standard korpuszok általában műfajspecifikusak (főleg rövidhírek), és korlátozott
+méretűek. Egy kellően robusztus alkalmazás építéséhez viszont nagy méretű és változatos
+nyelvi anyagot tartalmazó adathalmazra van szükség. A tanítókorpuszok előállítása
+hagyományosan kézi annotálás útján tör\-té\-nik, ami meglehetősen sok időt, szakértelmet és
+pénzt kíván, vagyis kardinális kérdés az erőforrások minimalizálása. Nagyobb
+tanítóhalmazhoz juthatunk, ha már meglevő korpuszokat egyesítünk; ebben az esetben
+viszont azzal kell szembesülnünk, hogy ezek különböző annotációs sémát követnek.
 
 Ezen problémák kiküszöbölésére egy lehetséges módszer, ha automatikusan állítunk elő  
-adathalmazokat, vagy pedig olyan közösségi tartalmakat használunk fel, mint pl.
-a Wikipedia, a Wiktionary vagy a DBPedia.
-Cikkünkben ezek kombinációját mutatjuk be: a Wikipediából előállított, automatikusan tulajdonnév-annotált korpuszt.
+adathalmazokat, vagy pedig olyan közösségi tartalmakat használunk fel, mint pl. a
+Wikipédia, a Wiktionary vagy a DBpedia. Cikkünkben olyan módszert mutatunk be, mely ezek
+kombinációja, és amely a Wikipédiából automatikusan állít elő tulajdonnév-annotált
+korpuszokat.
 
@@
 %collaborative annotation and/or collaboratively constructed resources, such as
-%Wikipedia, Wiktionary, Linked Open Data, or DBpedia. In this paper we combine
+%Wikipédia, Wiktionary, Linked Open Data, or DBpedia. In this paper we combine
 %these approaches by automatically generating freely available NE tagged
-%corpora from Wikipedia.
+%corpora from Wikipédia.
 
-%\section{Wikipedia and NER}
+%\section{Wikipédia and NER}
 
-A Wikipediát, mely a legismertebb, önkéntesek által szerkesztett nyílt internetes enciklopédia, számos NLP feladathoz sikeresen alkalmazták már.
-Mivel ez az egyik legnagyobb elérhető forrása az entitásoknak, az automatikus tulajdonnév-felismerés teljesítményének növelésére több módon is kiválóan használható.
-Nyilvánvaló felhasználási módja különböző névlisták kinyerése (pl.
-\cite{Toral_2006.a}); de a Wikipediában található tudás a tanuló algoritmus számára használható jegyek formájában is beépíthető felügyelt tanulási rendszerekbe (pl.
-\cite{KaTo07}).
+A Wikipédia a legismertebb, önkéntesek által szerkesztett nyílt internetes
+enciklopédia. Mivel ez az egyik
+legnagyobb elérhető forrása az entitásoknak, a NER teljesítményének növelésére több
+módon is kiválóan használható. Nyilvánvaló felhasználási
+módja különböző névlisták kinyerése (pl. \cite{Toral_2006.a}); de a Wikipédiában található
+tudás a tanuló algoritmus számára használható jegyek formájában is beépíthető felügyelt
+tanulási rendszerekbe (pl. \cite{KaTo07}).
 
-Számos példát találunk arra is a nemzetközi irodalomban, amikor a Wikipediából építenek korpuszt, amelyben az entitásokra vonatkozó információt valamilyen formában megőrzik és feldolgozzák.
-Ezt a megközelítést követtük mi is, de az eddig alkalmazott módszerek helyett, melyek elsősorban a cikkek definíciójából nyerték ki az entitás kategóriájára vonatkozó információt, mi a DBPedia ontológiai osztályait képeztük le a sztenderd CoNLL névosztályokra, majd kötöttük őket Wikipédia-entitásokhoz.
-Az általunk létrehozott feldolgozólánc elemei nagyrészt nyelvfüggetlenek, vagyis módszerünkkel létrehozhatók tanítókorpuszok kevesebb erőforrással rendelkező nyelvekre is.
-Cikkünkben a feldolgozás folyamatát és eredményét: egy magyar és egy angol nyelvű automatikusan tulajdonnév-annotált korpuszt mutatunk be.
+Számos példát találunk arra is a nemzetközi irodalomban, amikor a Wikipédiából építenek
+korpuszt, amelyben az entitásokra vonatkozó információt valamilyen formában megőrzik és
+feldolgozzák. Ezt az irányt követtük mi is, de az eddig alkalmazott módszerek
+helyett új megközelítést alkalmaztunk. Cikkünkben bemutatjuk a feldolgozás
+folyamatát és eredményét: egy magyar és egy angol nyelvű automatikusan
+tulajdonnév-annotált korpuszt.
 
-%Wikipedia (WP, see {\tt http://wikipedia.org}), a free
+%Wikipédia (WP, see {\tt http://wikipedia.org}), a free
 %multilingual Internet encyclopedia, written collaboratively by volunteers, is
@@
 
-\section{Creating the Corpus}
-\subsection{Creating the English Corpus}
+%\section{Creating the Corpus}
+%\subsection{Creating the English Corpus}
 
-A korpuszt, hasonlóan Nothman et al.~\cite{Nothman:08} munkájához, a következő
+A korpuszokat, hasonlóan Nothman et al.~\cite{Nothman:08} munkájához, a következő
 lépésekkel állítottuk elő:
@@
 \begin{enumerate}
-\item a WP cikkeket entitás osztályokba soroltuk
-\item a cikkeket mondatokra bontottuk
-\item megjelöltük a névelemeket a szövegben
-\item kiszűrtük a rossz minőségű mondatokat
+\item a Wikipédia-cikkeket entitásosztályokba soroltuk;
+\item a cikkeket mondatokra bontottuk;
+\item felcímkéztük a tulajdonneveket a szövegben;
+\item kiszűrtük a rossz minőségű mondatokat.
 \end{enumerate}
@@
 pontosság érdekében a DBpedia~\cite{Bizer:09} tudásbázis alapján osztályoztuk az
-entitásokat. A névelemeket WP kereszthivatkozások segítségével azonosítottuk. Minden
-olyan mondatot, ahol ismeretlen névelem fordult elő, eldobtunk.
+entitásokat. A tulajdonneveket Wikipédia kereszthivatkozások segítségével azonosítottuk.
+Minden olyan mondatot, ahol ismeretlen név fordult elő, eldobtunk.
 
-A fenti lépések mindegyike automatikus; emberi beavatkozást egyedül a DBpedia típusok
-névelemcímkékre történő leképezése jelent.
+A fenti lépések mindegyike automatikus; emberi beavatkozást egyedül a DBpedia-típusok
+sztenderd CoNLL névkategóriákra történő leképezése jelent. A módszer jórészt
+nyelvfüggetlen; ennek alátámasztására egy magyar és egy angol nyelvű korpuszt is
+előállítottunk. A cikkben részletesen is kitérünk a két Wikipédia sajátosságaira, a
+nyelvi és NER konvenciók különbségéből fakadó eltérésekre.
 
-\section{Data description}
+%\section{Evaluation}
 
-A korpusz a Wikipediához hasonlóan a Creative Commons Attribution-Sharealike
-3.0 Unported (CC-BY-SA) licensz alatt érhető el. Szabadon letölthető a
-\footnote{\texttt{http://hlt.sztaki.hu}} oldalról és a META-SHARE tárhelyről
-\footnote{\texttt{http://nlp.ipipan.waw.pl/metashare}}.
+Az automatikus annotálás nyilvánvaló előnyei mellett meg kell vizsgálnunk azt is, hogy az
+eredményül kapott  ún. ,,silver standard'' korpuszok hogyan hasznosíthatók a NER
+teljesítményének növelésére. A már említett lehetőségek (névlisták kinyerése, jegyként
+való beépítés) mellett ezek a korpuszok önmagukban vagy más adathalmazokkal együttesen
+tanítókorpuszként szolgálhatnak a kevesebb erőforrással rendelkező nyelvekre, illetve a
+rövidhírektől eltérő műfajokra is. A magyar és az angol korpuszon végzett kísérletek
+egyaránt azt igazolják, hogy a Wikipédiából automatikusan előállított adathalmazok
+javítják a NER teljesítményét. Cikkünkben részletes hibaelemzést adunk, és megvizsgáljuk
+a magyar Wikipédia és a Szeged NE korpusz\cite{Szarvas:06} annotációsémája közötti
+eltéréseket.
 
-\section{Evaluation}
+%\section{Data description}
 
-Az automatikus annotálás nyilvánvaló előnyei mellett meg kell vizsgálnunk azt is, hogy az eredményül kapott  ún.
-,,silver standard'' korpuszok hogyan hasznosíthatók a NER teljesítményének növelésére.
-A már említett lehetőségek (névlisták kinyerése, jegyként való beépítés) mellett ezek a korpuszok önmagukban vagy más adathalmazokkal együttesen tanítókorpuszként szolgálhatnak a kevesebb erőforrással rendelkező nyelvekre, illetve a rövidhírektől eltérő műfajokra is.
-A magyar és az angol korpuszon végzett kísérletek egyaránt azt igazolják, hogy a Wikipédiából automatikusan előállított adathalmazok javítják a NER teljesítményét.
-Ugyanakkor számos problémával szembesültünk, melyekre a jövőben megoldást kell találnunk.
+A korpusz a Wikipédiához hasonlóan a Creative Commons Attribution-Sharealike
+3.0 Unported (CC-BY-SA) licensz alatt érhető el. Szabadon letölthető a SZTAKI HLT
+csoportjának\footnote{\texttt{http://hlt.sztaki.hu}} oldaláról és a META-SHARE
+tárhelyről\footnote{\texttt{http://nlp.ipipan.waw.pl/metashare}}.
 
-\section{Conclusion}
-
-\cite{Nothman:08} \cite{Szarvas:06} \cite{Medelyan:09}
+%\section{Conclusion}
 

### 2012-10-08 paper/MSZNY2013/abstract.tex
Hungarian hyphenation now works. 	modified:   abstract.tex 	new file:   huhyphn.tex 	new file:   magyar.ldf
Some corrections in the text.

@@
 
-\usepackage[hungarian]{babel}
-\selectlanguage{hungarian}
+\usepackage[magyar]{babel}
+\selectlanguage{magyar} 
+\usepackage{t1enc}
+%\input huhyphn.tex
+\hyphenation{DBpedia Sharealike}
 
@@
 MUC and CoNLL datasets and annotation schemes have been the major standards applied in the field of NER.
+
+% So that DBpedia-típus and Sharealike do not generate overfull hbox.
+\pretolerance=1000
 
@@
 
-Számos példát találunk arra is a nemzetközi irodalomban, amikor a Wikipédiából építenek
+Találunk arra is példát a nemzetközi irodalomban, amikor a Wikipédiából építenek
 korpuszt, amelyben az entitásokra vonatkozó információt valamilyen formában megőrzik és
@@
 helyett új megközelítést alkalmaztunk. Cikkünkben bemutatjuk a feldolgozás
-folyamatát és eredményét: egy magyar és egy angol nyelvű automatikusan
+folyamatát és eredményét: egy magyar és egy angol nyelvű, automatikusan
 tulajdonnév-annotált korpuszt.

### 2012-10-08 paper/MSZNY2013/abstract.tex
Added the word "ontology" to the paper. :)

@@
 
-A fenti lépések mindegyike automatikus; emberi beavatkozást egyedül a DBpedia-típusok
+E lépések mindegyike automatikus; emberi beavatkozást egyedül a DBpedia ontológiájának
+fogalmi osztályait
 sztenderd CoNLL névkategóriákra történő leképezése jelent. A módszer jórészt

### 2012-10-08 paper/MSZNY2013/abstract.tex
The last modification to the abstract.

@@
 javítják a NER teljesítményét. Cikkünkben részletes hibaelemzést adunk, és megvizsgáljuk
-a magyar Wikipédia és a Szeged NE korpusz\cite{Szarvas:06} annotációsémája közötti
+a magyar Wikipédia és a Szeged NE korpusz\cite{Szarvas:06} annotációs konvenciója közötti
 eltéréseket.

### 2012-11-13 paper/MSZNY2013/cikk.tex
MSZNY cikk elkezdve

\documentclass{llncs}

\usepackage[a4paper,includeheadfoot,top=1.65in,bottom=1.65in,left=1.73in,hcentering]{geometry}
\usepackage[pdftex]{graphicx}

% Ezek egyikÈnek bekapcsol·s·val a magyar Èkezetes karakterek
% kˆzvetlen¸l is felhaszn·lhatÛk:
%\usepackage[latin2]{inputenc} 
\usepackage[utf8]{inputenc}

\usepackage[hungarian]{babel}
\selectlanguage{hungarian}

\begin{document}

\pagestyle{myheadings}
\def\leftmark{{\rm IX. Magyar Sz\'am\'\i t\'og\'epes Nyelv\'eszeti Konferencia}}
\def\rightmark{{\rm Szeged, 2013. január 7-8.}}

\setcounter{page}{3}

\title{\ \break Automatikusan tulajdonnév-annotált korpuszok előállítása a Wikipédiából}
%XXX a címet majd meg kéne változtatni, tényleg elég bénán hangzik magyarul
%Automatikus korpuszépítés tulajdonnév-felismerés céljára
\author{Nemeskey Dávid Márk\inst{1}, Simon Eszter\inst{2}}
\institute{
\inst{}%
MTA SZTAKI \break
1111 Budapest, Lágymányosi utca 11., e-mail:nemeskey@sztaki.mta.hu \break
\and
\inst{}%
MTA Nyelvtudományi Intézet \break
1068 Budapest, Benczúr u. 33., e-mail: simon.eszter@nytud.mta.hu}

\maketitle

\begin{abstract}
A felügyelt gépi tanulási módszerek alkalmazásához nagy méretű annotált korpuszokra van szükség, amelyek előállítása rendkívül emberierőforrás-igényes. Több lehetőség van az annotációs költségek csökkentésére, ezek közül az egyik az automatikus annotálás. Cikkünkben egy nyelvfüggetlen módszert és azzal előállított magyar és angol nyelvű korpuszokat mutatunk be. A korpuszok szövege a Wikipédiából származik, és tulajdonnévi annotálást is tartalmaz. Az automatikus annotáláshoz egy új módszert alkalmaztunk: a DBpedia ontológiai kategóriáit képeztük le a CoNLL névosztályokra. A korpuszok szabadon felhasználhatóak további fejlesztésekhez.
\\[2mm]
{\bf Kulcsszavak:} tulajdonnév-felismerés, korpuszépítés,
automatikus annotáció, Wikipédia
\end{abstract}

\section{Bevezet\'es}

Az automatikus tulajdonnév-felismerés (Named Entity Recognition, NER) a természetes nyelv feldolgozását célzó alkalmazások közül az egyik legnépszerűbb, mivel hatékonyan automatizálható, és eredménye hasznos bemenete különböző magasabb szintű információkinyerő és -feldolgozó rendszereknek. A feladat során strukturálatlan szövegben kell azonosítani és az előre definiált osztályok valamelyikébe besorolni a neveket. A tulajdonnév-felismerés feladata a 6. Message Understanding Conference (MUC) egyik versenykiírásában jelent meg először 1995-ben \cite{}. Itt három alfeladatot különítettek el: tulajdonneveket (ENAMEX), temporális (TIMEX) és különböző numerikus (NUMEX) kifejezéseket kellett felismerni. A NER-közösségen belül a temporális és a numerikus kifejezések annotálása is elfogadott, de a leginkább vizsgált típusok a személy-, földrajzi és intézménynevek. Ezek mellé vezettek be a CoNLL versenyeken \cite{} egy negyedik típust, amely az előző háromba nem tartozó egyéb tulajdonneveket foglalja magában. Az azóta eltelt időben ezek az annotációs sémák váltak nemzetközileg elfogadottá.

A versenyekre épített és aztán közzétett tulajdonnév-annotált korpuszok képezik azokat a sztenderdeket, amelyek összemérhetővé teszik az egyes tulajdonnév-felismerő rendszereket. Ezek a korpuszok viszont meglehetősen korlátozott méretűek és témaspecifikusak. Kellően robusztus tulajdonnév-felismerő rendszerek építéséhez viszont nagyméretű, a téma tekintetében heterogén korpuszokra van szükség. A kézi annotálás rendkívül idő-, erőforrás- és szakértelemigényes feladat, ezért az elmúlt időkben különösen nagy hangsúly került az annotált erőforrások automatikus előállítására. Ennek egy módja, ha már rendelkezésre álló korpuszokat dolgozunk össze; ekkor a különböző annotációs sémák és címkekészletek összeillesztése állít elénk sokszor megoldhatatlan problémákat. Egy másik lehetőség az olyan, önkéntesek által írt, szerkesztett közösségi tartalmak felhasználása, mint például a Wikipédia, a Wiktionary vagy a DBpedia. Megint másik megközelítés az annotáció automatizálása.

Cikkünkben egy olyan megközelítést mutatunk be, mely ezen lehetőségeket kombinálja: teljesen automatikus eszközökkel olyan korpuszokat építettünk, melyek Wikipédia-szócikkekből állnak, és tulajdonnévi annotációt tartalmaznak. Munkánk során új módszert alkalmaztunk: a DBpedia ontológiai kategóriáit képeztük le a CoNLL-névosztályokra. A létrehozott magyar és angol nyelvű korpuszok szabadon felhasználhatóak.

A cikk a következőképpen épül fel. %XXX

\section{Wikipédia és tulajdonnév-felismerés}

A Wikipédia egy többnyelvű, nyílt tartalmú, a nyílt közösség által fejlesztett webes világenciklopédia \footnote{http://wikipedia.org}. Jelenlegi XXX számú szócikkével aranybánya a különböző természetesnyelv-feldolgozó fejlesztések számára; használták már többek között jelentésegyértelműsítére, ontológia- és tezauruszépítésre, valamint kérdésmegválaszoló rendszerekhez (további alkalmazási lehetőségekhez lásd \cite{Medelyan}). Mivel a Wikipédia-címszavak jelentős része tulajdonnév, adja magát a lehetőség, hogy az automatikus tulajdonnév-felismeréshez is felhasználjuk.

A Wikipédia legkézenfekvőbb felhasználási módja a nagyméretű névlisták előállítása az általánosan alkalmazott névosztályokra, melyek javítják az általános célú névfelismerők hatékonyságát, felügyelt és felügyelet nélküli módszerek esetében is (pl. \cite{Toral:Munoz, Nadeau}), továbbá a névegyértelműsítésben is fontos szerepet játszanak (pl. \cite{Bunescu:Pasca}). A Wikipédiában található tudás jegyek formájában is beépíthető tulajdonnév-felismerő rendszerekbe: például Kazama és Torisawa \cite{KaTo} kísérlete azt bizonyítja, hogy a Wikipédia kategóriacímkéinek automatikus kinyerése növeli egy felügyelt névfelismerő rendszer pontosságát.

A Wikipédia alkalmazására a tulajdonnév-felismerés területén egy másik lehetőség magának a szövegbázisnak a felhasználása. Richman és Schone \cite{Richman} kevés erőforrással rendelkező nyelvek Wikipédia-szócikkeiből építettek korpuszokat, amelyekben a Wikipédia inherens kategóriastruktúráját használták fel a tulajdonnevek annotálásához. Nothman et al. \cite{Nothman:2008} a szócikkek definíciójának első mondatából kiindulva címkézte fel a szövegben belinkelt neveket, így építve automatikusan tulajdonnév-annotált korpuszt.

Az általunk alkalmazott módszer az említettekétől annyiban tér el, hogy mi a DBpedia ontológiai osztályait képeztük le a sztenderd CoNLL névosztályokra, majd ezeket Wikipédia-entitásokhoz kötöttük. Az így létrehozott korpuszokat szabadon elérhetővé tettük. Tudomásunk szerint jelenleg csak egy automatikusan tulajdonnév-annotált korpusz létezik, amely szabadon felhasználható, a Semantically Annotated Snapshot of the English (SASWP) \cite{Zaragoza}. Bár a névfelismerést a jelenlegi legjobb szabadon hozzáférhető rendszerekkel végezték, azt gondoljuk, hogy a Wikipédia többezer szerkesztőjének döntésein alapuló kategorizálás megbízhatóbb eredményt ad. A Wikipédiából, tekintve a szócikkek nagy számát, kevés erőforrással rendelkező nyelvekre is tudunk kellően nagy méretű korpuszokat építeni, amelyek bemenetül szolgálhatnak névfelismerő rendszerek tanításához és teszteléséhez. Az általunk létrehozott korpusz az első magyar nyelvű automatikusan tulajdonnév-annotált korpusz. 
%XXX ezt jobban magyarra kell szabni, az SASWP-t ki is lehet dobni, mert a magyar szempontjából nem releváns

\section{A korpuszépítés}

\section{Problémás esetek, hibaelemzés}

\section{A korpusz leírása}

\section{Kiértékelés}

\subsection{Az adatok}

\subsection{Kísérletek és eredmények}

\section{Konklúzió, összegzés}

\section*{Köszönetnyilvánítás}
A fejlesztés az OTKA 82333 számú projektjén belül valósult meg. A fejlesztést támogatta továbbá a CESAR projekt (No. 271022). A szerzők ezúton fejezik ki köszönetüket Zséder Attilának a Wikipédia-szövegek feldolgozásában végzett munkájáért, és Kornai Andrásnak támogatásáért. %vagy mi

%
% ---- Bibliography ----
%
\begin{thebibliography}{15}
%
\bibitem{alexin:eacl2003}
Alexin Z., Csirik, J., Gyim\'othy, T., Bibok K., Hatvani, Cs.,
Pr\'osz\'eky, G., Tihanyi, L.:
Manually Annotated Hungarian Corpus.
in Proc. of the Research Note Sessions of the 10th Conference
of the European Chapter of the Association for Computational
Linguistics EACL'03, Budapest, Hungary, 53--56 (2003).
%
\bibitem{bibok:mszny2003}
Bibok K.:
A sz\'or\'ol \'es a sz\'ofajokr\'ol (a sz\'am\'\i t\'og\'epes nyelvfeldolgoz\'as kap- cs\'an),
Magyar Sz\'am\'\i t\'og\'epes Nyelv\'eszeti Konferencia (MSZNY 2003),
Szeged, Magyarorsz\'ag, 31--36, (2003).
%
\bibitem{farkas:mszny2004}
Farkas R., Konczer K., Szarvas Gy.:
Szemantikus keretilleszt\'es \'es az IE rendszer automatikus ki\'ert\'ekel\'se
Magyar Sz\'am\'\i t\'og\'epes Nyelv\'eszeti Konferencia (MSZNY 2004),
bek\"uldve, Szeged, Magyarorsz\'ag, (2004).
%
\bibitem{hocza:acta2004}
H\'ocza, A.:
Noun Phrase Recognition with Tree Patterns
elfogadva az Acta Cybernetica c. lapban t\"ort\'en\H{o} megjelen\'esre (2004).
%
\bibitem{mihaczi:mszny2003}
Mih\'aczi Andr\'as, N\'emeth L\'aszl\'o, R\'acz Mikl\'os:
Magyar sz\"ovegek tem\'eszetes nyelvi el\H{o}feldolgoz\'asa
Magyar Sz\'am\'\i t\'og\'epes Nyelv\'eszeti Konferencia (MSZNY 2003),
Szeged, Magyarorsz\'ag, 38--43, (2003).
%
\bibitem{proszeky:mszny2003}
Pr\'osz\'eky G.:
Automatikus inform\'aci\'oszerz\'es gazdas\'agi r\"ovidh\'\i rekb\H{o}l.
Magyar Sz\'am\'\i t\'og\'epes Nyelv\'eszeti Konferencia (MSZNY 2003),
Szeged, Magyarorsz\'ag, 161--166, (2003).
%
\bibitem{proszeky:neumann2003}
Pr\'osz\'eky G.:
Automatikus inform\'aci\'oszerz\'es gazdas\'agi-politikai r\"ovidh\'\i rekb\H{o}l.
VIII. Orsz\'agos (Centen\'ariumi) Neumann Kongresszus kiadv\'anya,
Budapest, Magyarorsz\'ag, 359--367, (2003).
%
\end{thebibliography}
\end{document}

### 2012-11-26 paper/MSZNY2013/cikk.tex
I have begun writing chapter 3. 	modified:   cikk.tex

@@
 
-\section{A korpuszépítés}
+\section{Korpuszépítés}
+
+% TODO: lame wording, angol korpusz?
+A korpuszépítő algoritmus, hasonlóan Nothman et al.~\cite{Nothman:08} által
+leírthoz, a következő lépésekből áll:
+
+\begin{enumerate}
+\item a Wikipédia-cikkeket entitásosztályokba soroltuk;
+\item a cikkeket mondatokra bontottuk;
+\item felcímkéztük a tulajdonneveket a szövegben;
+\item kiszűrtük a rossz minőségű mondatokat.
+\end{enumerate}
+
+Az algoritmus alapjaiban nyelvfüggetlen: bármelyik nyelvre alkalmazható, amely
+rendelkezik megfelelő méretű Wikipédiával. Egyedül a harmadik lépés az, ahol
+figyelembe kell venni a nyelv, illetve a használt NER konvenció sajátosságait.
+Ennek oka, hogy az egyes nyelvek, illetve NER konvenciók eltérnek abban, hogy
+mit tekintenek névelemnek: pl. a \textit{római} szó a Szeged korpuszban nem
+számít annak, míg angol megfelelője, a \textit{Roman} a CoNLL annotációs sémában
+\textit{MISC} címkét kapna. A továbbiakban röviden ismertetjük a fenti lépéseket,
+a magyar nyelvű korpuszra koncentrálva. Részletes leírás, illetve az angol nyelvű
+korpuszban felmerülő problémák kifejtése \cite{simon-nemeskey:2012:NEWS2012}-ben
+található.
+
+% Argh, de nehéz magyarul írni
+
+\subsection{Wikipédia cikkek, mint entitások}
+
+A DBpedia nyelve
+
+\subsection{A cikkek feldolgozása}
+\subsection{Tulajdonnevek címkézése}
+
+\subsection{Szűrés}
+
+Bár Nothman az első feladatot gépi tanulással oldotta meg, mi a nagyobb
+pontosság érdekében a DBpedia~\cite{Bizer:09} tudásbázis alapján osztályoztuk az
+entitásokat. A tulajdonneveket Wikipédia kereszthivatkozások segítségével azonosítottuk.
+Minden olyan mondatot, ahol ismeretlen név fordult elő, eldobtunk.
 

### 2012-11-26 paper/MSZNY2013/cikk.tex
a reszem nagyjabol kesz

@@
 \selectlanguage{hungarian}
+
+\usepackage{booktabs}
 
@@
 A tulajdonnév-felismerés feladata a 6.
-Message Understanding Conference (MUC) egyik versenykiírásában jelent meg először 1995-ben \cite{}.
+Message Understanding Conference (MUC) egyik versenykiírásában jelent meg először 1995-ben \cite{MUC6}.
 Itt három alfeladatot különítettek el: tulajdonneveket (ENAMEX), temporális (TIMEX) és különböző numerikus (NUMEX) kifejezéseket kellett felismerni.
 A NER-közösségen belül a temporális és a numerikus kifejezések annotálása is elfogadott, de a leginkább vizsgált típusok a személy-, földrajzi és intézménynevek.
-Ezek mellé vezettek be a CoNLL versenyeken \cite{} egy negyedik típust, amely az előző háromba nem tartozó egyéb tulajdonneveket foglalja magában.
+Ezek mellé vezettek be a CoNLL versenyeken \cite{tksintro2002conll} \cite{conll2003intro} egy negyedik típust, amely az előző háromba nem tartozó egyéb tulajdonneveket foglalja magában.
 Az azóta eltelt időben ezek az annotációs sémák váltak nemzetközileg elfogadottá.
@@
 A versenyekre épített és aztán közzétett tulajdonnév-annotált korpuszok képezik azokat a sztenderdeket, amelyek összemérhetővé teszik az egyes tulajdonnév-felismerő rendszereket.
-Ezek a korpuszok viszont meglehetősen korlátozott méretűek és témaspecifikusak.
+Ezek a korpuszok meglehetősen korlátozott méretűek és témaspecifikusak.
 Kellően robusztus tulajdonnév-felismerő rendszerek építéséhez viszont nagyméretű, a téma tekintetében heterogén korpuszokra van szükség.
 A kézi annotálás rendkívül idő-, erőforrás- és szakértelemigényes feladat, ezért az elmúlt időkben különösen nagy hangsúly került az annotált erőforrások automatikus előállítására.
-Ennek egy módja, ha már rendelkezésre álló korpuszokat dolgozunk össze; ekkor a különböző annotációs sémák és címkekészletek összeillesztése állít elénk sokszor megoldhatatlan problémákat.
-Egy másik lehetőség az olyan, önkéntesek által írt, szerkesztett közösségi tartalmak felhasználása, mint például a Wikipédia, a Wiktionary vagy a DBpedia.
+Ennek egy módja, ha már rendelkezésre álló korpuszokat dolgozunk össze; ekkor a különböző annotációs sémák és címkekészletek összeillesztése állít elénk problémákat.
+Egy másik lehetőség az olyan webes közösségi tartalmak felhasználása, mint például a Wikipédia, a Wiktionary vagy a DBpedia.
 Megint másik megközelítés az annotáció automatizálása.
 
-Cikkünkben egy olyan megközelítést mutatunk be, mely ezen lehetőségeket kombinálja: teljesen automatikus eszközökkel olyan korpuszokat építettünk, melyek Wikipédia-szócikkekből állnak, és tulajdonnévi annotációt tartalmaznak.
+Cikkünkben egy olyan megközelítést mutatunk be, mely ezen lehetőségeket kombinálja: automatikus eszközökkel tulajdonnév-annotált korpuszokat építettünk Wikipédia szócikkekből.
 Munkánk során új módszert alkalmaztunk: a DBpedia ontológiai kategóriáit képeztük le a CoNLL-névosztályokra.
@@
 
-A cikk a következőképpen épül fel. %XXX
+A cikk a következőképpen épül fel.
+A \ref{wiki:ner}.~fejezetben bemutatjuk a Wikipédia eddigi felhasználási módjait a tulajdonnév-felismerés területén.
+A \ref{corpusbuilding}.~fejezetben leírjuk a korpuszépítési módszert, elsősorban a magyar nyelvű adatokra koncentrálva.
+Az alkalmazott módszer részletes hibaelemzését a \ref{probl}., a korpuszok leírását a \ref{desc}., míg a kiértékelést és az eredményeket a \ref{results}.~fejezet adja.
+Cikkünket az elért eredmények rövid összefoglalása zárja (\ref{conc}.~fejezet).
 
 \section{Wikipédia és tulajdonnév-felismerés}
+\label{wiki:ner}
 
-A Wikipédia egy többnyelvű, nyílt tartalmú, a nyílt közösség által fejlesztett webes világenciklopédia \footnote{http://wikipedia.org}.
-Jelenlegi XXX számú szócikkével aranybánya a különböző természetesnyelv-feldolgozó fejlesztések számára; használták már többek között jelentésegyértelműsítére, ontológia- és tezauruszépítésre, valamint kérdésmegválaszoló rendszerekhez (további alkalmazási lehetőségekhez lásd \cite{Medelyan}).
-Mivel a Wikipédia-címszavak jelentős része tulajdonnév, adja magát a lehetőség, hogy az automatikus tulajdonnév-felismeréshez is felhasználjuk.
+A Wikipédia egy többnyelvű, nyílt tartalmú, az internetes közösség által fejlesztett webes világenciklopédia \footnote{http://wikipedia.org}.
+Jelenlegi XXX számú szócikkével aranybánya a különböző természetesnyelv-feldolgozó fejlesztések számára; használták már többek között jelentésegyértelműsítére, ontológia- és tezauruszépítésre, valamint kérdésmegválaszoló rendszerekhez (további alkalmazási lehetőségekhez lásd \cite{Medelyan:09}).
+Mivel a Wikipédia címszavak jelentős része tulajdonnév, adja magát a lehetőség, hogy az automatikus tulajdonnév-felismeréshez is felhasználjuk.
 
-A Wikipédia legkézenfekvőbb felhasználási módja a nagyméretű névlisták előállítása az általánosan alkalmazott névosztályokra, melyek javítják az általános célú névfelismerők hatékonyságát, felügyelt és felügyelet nélküli módszerek esetében is (pl.
-\cite{Toral:Munoz, Nadeau}), továbbá a névegyértelműsítésben is fontos szerepet játszanak (pl.
-\cite{Bunescu:Pasca}).
-A Wikipédiában található tudás jegyek formájában is beépíthető tulajdonnév-felismerő rendszerekbe: például Kazama és Torisawa \cite{KaTo} kísérlete azt bizonyítja, hogy a Wikipédia kategóriacímkéinek automatikus kinyerése növeli egy felügyelt névfelismerő rendszer pontosságát.
+A Wikipédia legkézenfekvőbb alkalmazási módja nagyméretű névlisták előállítása, melyek javítják az általános célú névfelismerők hatékonyságát, felügyelt és felügyelet nélküli módszerek esetében is (pl.~\cite{Toral:06} és \cite{Nadeau:06}), továbbá a névegyértelműsítésben is fontos szerepet játszanak (pl.~\cite{Bunescu:06}).
+A Wikipédiában található tudás jegyek formájában is beépíthető tulajdonnév-felismerő rendszerekbe: például Kazama és Torisawa \cite{KaTo07} kísérlete azt bizonyítja, hogy a Wikipédia kategóriacímkéinek automatikus kinyerése növeli egy felügyelt névfelismerő rendszer pontosságát.
 
 A Wikipédia alkalmazására a tulajdonnév-felismerés területén egy másik lehetőség magának a szövegbázisnak a felhasználása.
-Richman és Schone \cite{Richman} kevés erőforrással rendelkező nyelvek Wikipédia-szócikkeiből építettek korpuszokat, amelyekben a Wikipédia inherens kategóriastruktúráját használták fel a tulajdonnevek annotálásához.
+Richman és Schone \cite{Richman:08} kevés erőforrással rendelkező nyelvek Wikipédia szócikkeiből építettek korpuszokat, amelyekben a Wikipédia inherens kategóriastruktúráját használták fel a tulajdonnevek annotálásához.
 Nothman et al.
-\cite{Nothman:2008} a szócikkek definíciójának első mondatából kiindulva címkézte fel a szövegben belinkelt neveket, így építve automatikusan tulajdonnév-annotált korpuszt.
+\cite{Nothman:08} a szócikkek definíciójának első mondatából kiindulva címkézte fel a szövegben belinkelt neveket, így építve automatikusan tulajdonnév-annotált korpuszt.
 
-Az általunk alkalmazott módszer az említettekétől annyiban tér el, hogy mi a DBpedia ontológiai osztályait képeztük le a sztenderd CoNLL névosztályokra, majd ezeket Wikipédia-entitásokhoz kötöttük.
+Az általunk alkalmazott módszer az említettekétől annyiban tér el, hogy mi a DBpedia ontológiai osztályait képeztük le a sztenderd CoNLL névosztályokra, majd ezeket Wikipédia entitásokhoz kötöttük.
 Az így létrehozott korpuszokat szabadon elérhetővé tettük.
-Tudomásunk szerint jelenleg csak egy automatikusan tulajdonnév-annotált korpusz létezik, amely szabadon felhasználható, a Semantically Annotated Snapshot of the English (SASWP) \cite{Zaragoza}.
+Tudomásunk szerint jelenleg csak egy angol nyelvű automatikusan tulajdonnév-annotált korpusz létezik, amely szabadon felhasználható, a Semantically Annotated Snapshot of the English Wikipedia \cite{Atserias:08}.
 Bár a névfelismerést a jelenlegi legjobb szabadon hozzáférhető rendszerekkel végezték, azt gondoljuk, hogy a Wikipédia többezer szerkesztőjének döntésein alapuló kategorizálás megbízhatóbb eredményt ad.
@@
 Az általunk létrehozott korpusz az első magyar nyelvű automatikusan tulajdonnév-annotált korpusz.
-%XXX ezt jobban magyarra kell szabni, az SASWP-t ki is lehet dobni, mert a magyar szempontjából nem releváns
+%XXX ezt kábé mind leírtam a conclusionben is, szóval innen akár ki is lehet venni
 
 \section{Korpuszépítés}
+\label{corpusbuilding}
 
@@
 \begin{enumerate}
-\item a Wikipédia-cikkeket entitásosztályokba soroltuk;
+\item a Wikipédia cikkeket entitásosztályokba soroltuk;
 \item a cikkeket mondatokra bontottuk;
@@
 Ennek oka, hogy az egyes nyelvek, illetve NER konvenciók eltérnek abban, hogy
-mit tekintenek névelemnek: pl. a \textit{római} szó a Szeged korpuszban nem
+mit tekintenek névelemnek: pl.~a \textit{római} szó a Szeged korpuszban nem
 számít annak, míg angol megfelelője, a \textit{Roman} a CoNLL annotációs sémában
-\textit{MISC} címkét kapna. A továbbiakban röviden ismertetjük a fenti lépéseket,
+\texttt{Misc} címkét kapna. A továbbiakban röviden ismertetjük a fenti lépéseket,
 a magyar nyelvű korpuszra koncentrálva. Részletes leírás, illetve az angol nyelvű
@@
 
-% Argh, de nehéz magyarul írni
+% Argh, de nehéz magyarul írni :)
 
-\subsection{Wikipédia cikkek, mint entitások}
+\subsection{Wikipédia cikkek mint entitások}
 
@@
 \section{Problémás esetek, hibaelemzés}
+\label{probl}
 
-\section{A korpusz leírása}
+A magyar Wikipédia korpusz gépi annotálását kézzel ellenőriztük a korpusz egy kis részén.
+A 19 milliós teljes korpuszt vettük alapul, amelyből már ki vannak szűrve a rossz minőségű mondatok, vagyis azok, amelyek nem nagybetűvel kezdődnek, és nincs írásjel a végükön.
+Ebből véletlenszerű mondatválogatással csináltunk egy kisebb korpuszt, amely kb.~19000 tokent tartalmaz.
+Ezt kézzel felannotáltuk, és összehasonlítottuk a kézi és a gépi annotálás eredményét.
+A következő tipikus hibák rajzolódtak ki.
+
+Bizonyos esetekben a név környezetében szereplő értelmező szerepű nyelvi egység is a névvel együtt annotálva lett.
+Ennek az az oka, hogy maga a Wikipédia címszó is tartalmazza ezt a szót, mely leszűkíti a név referenciáját, pl.~\textit{ókori Róma, magyar Wikipédia}.
+Emberekre utaló linkek esetében ezek nagyrészt rangjelzők, pl.~\textit{Szent István király, Anna brit királynő}.
+%Ez azért marad benne, mert a magyar List of titles stopword listánk nincs?
+
+Egy másik tipikus hibaforrás az automatikus tokenizálás és mondatra bontás.
+Sok esetben a rövidítést tartalmazó név (pl.~\textit{Warner Bros.}) utolsó eleme, vagyis a pont mondatvégi írásjelként értelmeződik, így nem annotálódik a névvel együtt.
+Máskor a mondatrabontás során a szövegközbeni link szétszakad, így a maradék elem nem kapja meg a megfelelő címkét.
+Ezenkívül rengeteg töredékes mondat maradt a szövegben, amelyeket egy erős szűrővel utólag is ki lehet szűrni, pl.~ha minden olyan mondatszerű egységet kidobunk, amelyben nincs ragozott ige.
+Ezzel persze azok is kiesnek, amelyek szabályos jelen idejű, kijelentő módú, létigét (nem) tartalmazó mondatok.
+
+Tipikus jelenség a magyarban, amikor egy név összetétel eleme lesz, pl.~\textit{Bizánc-ellenes}.
+Ilyen esetekben a köznév az összetétel alaptagja, vagyis az határozza meg a referenciát.
+A referenciaváltozást természetesen a címkézés változásának kell követnie, ami viszont nem, vagy nehezen kezelhető automatikusan, mivel nagyjából bármilyen köznév kapcsolódhat névhez.
+A helyzetet bonyolítja az is, hogy a mozaikszavakhoz, rövidítésekhez, valamint nem ejtett magánhangzóra végződő idegen nevekhez is kötőjellel kapcsoljuk a toldalékokat, amelynek a felszíni szerkezete nagyon hasonlít az összetételekéhez.
+Ennek a problémának a megoldása még további vizsgálatokat követel.
+
+A felsorolt problémás esetek ellenére összességében azt mondhatjuk, hogy a magyar Wikipédia korpusz annotálási pontossága közelíti a gold standard korpuszokét. %na ide jöhetnek majd pontos adatok
+
+\section{A korpuszok leírása}
+\label{desc}
+
+A korpuszokat Creative Commons Attribution-Sharealike 3.0 licensz alatt publikáltuk, vagyis ugyanolyan feltételekkel adjuk tovább, ahogy a Wikipédiából letöltöttük.
+Szabadon elérhetőek a \texttt{http://hlt.sztaki.hu} oldalon keresztül, valamint a META-SHARE tárhelyről (http://www.meta-net.eu/).
+(A META-SHARE célja egy nyílt rendszer létrehozása, amely lehetővé teszi a nyelvi erőforrások megosztását; létrehozója a META-NET, az Európai Bizottság által alapított kiválósági hálózat.)
+
+A fájlok ún.~\emph{multitag} formátumban vannak, amelyben a tartalmas sorok tabulátorral vannak elválasztva.
+Az első oszlop tartalmazza magukat a szövegszavakat, az egyes oszlopokban pedig a különböző szintű annotációk találhatók.
+A mondathatárokat üres sorok jelölik.
+A névcímkéken kívül minden token mellett szerepel a töve és a hozzá tartozó teljes morfológiai elemzése KR-kódokkal.
+Két további oszlopban közöljük, hogy sima szöveg vagy Wikipédia link-e az adott token, és ha utóbbi, akkor melyik szócikkre utal.
 
 \section{Kiértékelés}
+\label{results}
+
+Az automatikusan generált korpuszok nyilvánvaló előnyei mellett azonban megvan az a hátrányuk, hogy nem használhatók gold standard adatként.
+Kiértékelésünkben megmutatjuk, hogy a létrehozott silver standard adathalmazok viszont kiválóan használhatók a tulajdonnév-felismerés teljesítményének növelésére több módon is.
+
+A kiértékeléshez Varga és Simon \cite{Varga} tulajdonnév-felismerő rendszerét használtuk.
+A csak az egyes korpuszokra jellemző jegyeket (pl.~főnévi csoportok jelölése, Wikipédia linkek) kidobtuk, hogy növeljük a korpuszok összehasonlíthatóságát.
+Így a következő jegykészlettel dolgoztunk: mondatkezdő és -vég pozíciók, szóalakon alapuló jegyek, morfológiai információ és listajegyek.
+
+A kiértékeléshez a sztenderd CoNLL módszert alkalmaztuk, vagyis az annotációt csak akkor vettük helyesnek, ha a kezdő- és végpozíció is stimmelt, és a rendszer által kibocsátott címke megegyezett a gold standard címkével.
+Ezen alapulva a szokásos pontosságot, fedést és F-mértéket számoltuk.
 
@@
 
+A korpuszok a fent leírt szűrő eljárások után maradt mondatokat tartalmazzák, így azokat is, amelyekben nincs egy név sem.
+Ezeket azért tartottuk meg, hogy amennyire lehetséges, megőrizzük a nevek eredeti, Wikipédia-beli eloszlását.
+Viszont amikor megvizsgáltuk az egyes korpuszok telítettségét a nevek szempontjából, arra jutottunk, hogy a gold standard adathalmazokkal való összevetésükkor inkább sűrítjük a szövegeket, vagyis kivesszük azokat a mondatokat, amelyekben nincs név.
+A \ref{tab:size_en}.~és \ref{tab:size_hu}.~táblázatok mutatják az egyes korpuszokra vonatkozó számszerű adatokat, melyekből jól látható, hogy a Wikipédiából generált korpuszok telítettsége meglehetősen alacsony, a nevet nem tartalmazó mondatok kiszűrése után viszont hasonlóak lettek a számok.
+Mivel a Wikipédia az entitások kincsesbányájaként van számon tartva, ezek az adatok kissé meglepőek lehetnek.
+A szövegeknek ez a hígsága valószínűleg annak köszönhető, hogy a módszerünk szigorú, vagyis inkább minden olyan mondatot eltávolítottunk, amelyben nem lehetett beazonosítani a nevet, minthogy rosszul annotált nevek maradjanak benne.
+
+\begin{table}[ht]
+\begin{center}
+\begin{tabular}{lccc}
+\toprule  
+& \textbf{enwiki} & \textbf{enwiki filter} & \textbf{CoNLL} \\ 
+\midrule
+token & 60,520,819 & 21,718,854 & 301,418 \\
+NE & 3,169,863 & 3,169,863 & 35,089 \\
+telítettség (\%) & 5.23 & 14.59 & 11.64 \\
+\hline
+\end{tabular}
+\end{center}
+\caption{Az angol korpuszok mérete és telítettsége.}
+\label{tab:size_en}
+\end{table}
+
+\begin{table}[ht]
+\begin{center}
+\begin{tabular}{lccc}
+\toprule
+ & \textbf{huwiki} & \textbf{huwiki filter} & \textbf{Szeged NER} \\
+\midrule
+token & 19,108,027 & 3,512,249  & 225,963 \\
+NE & 456,281 & 456,281  & 25,896 \\
+telítettség (\%) & 2.38 & 12.99 & 11.46 \\
+\bottomrule
+\end{tabular}
+\end{center}
+\caption{A magyar korpuszok mérete és telítettsége.}
+\label{tab:size_hu}
+\end{table}
+
 \subsection{Kísérletek és eredmények}
 
-\section{Konklúzió, összegzés}
+A magyar és az angol korpuszok is hasonlóan lettek kiértékelve: először saját magukon, aztán egy választott gold standard adathalmazon.
+Mivel az angol korpusz még a nevet nem tartalmazó mondatok eltávolítása után is nagyon nagy, kísérleteinket egy kb.~3,5 millió tokenes részén végeztük (a szűrt magyar korpusz is ekkora), 90-10\%-os arányban vágva tanító és kiértékelő korpuszra.
+
+Az angol korpusz kiértékeléséhez a 2003-as CoNLL korpuszt, a magyarhoz pedig a Szeged NER korpuszt választottuk.
+Többek által (pl.
+\cite{Nothman:08} és \cite{ciaramita2005}) bizonyított tény, hogy a korpuszok közötti kiértékelés sokkal rosszabb eredményt ad, mint a saját kiértékelő halmazon való mérés.
+Különböző típusú szövegek esetén a különbség 20-30\% is lehet.
+A helyzet a mi esetünkben is nagyon hasonló (lásd a \ref{tab:enresults}.~és \ref{tab:huresults}.~táblázatokat az eredményekért): a Wikipédián tanított rendszerek teljesítménye közel sem olyan jó a gold standard korpuszok kiértékelő halmazán mérve, mint a saját kiértékelő halmazukon.
+
+\begin{table}[ht]
+\begin{center}
+\begin{tabular}{lcccc}
+\toprule 
+\textbf{tanítás} & \textbf{teszt} & \textbf{pontosság} & \textbf{fedés} & \textbf{F-mérték} \\ 
+\midrule
+CoNLL & CoNLL & 85.13 & 85.13 & 85.13 \\
+enwiki & enwiki & 72.46 & 73.33 &  72.89 \\
+enwiki & CoNLL & 56.55 & 49.77 & 52.94 \\
+CoNLL\_wikilisták & CoNLL & 86.33 & 86.35 & 86.34 \\
+CoNLL\_wikitag & CoNLL & 85.88 & 85.94 & 85.91 \\
+\bottomrule
+\end{tabular}
+\end{center}
+\caption{Eredmények az angol Wikipédia korpuszon.}
+\label{tab:enresults}
+\end{table}
+
+\begin{table}[ht]
+\begin{center}
+\begin{tabular}{lcccc}
+\toprule 
+\bf tanítás & \bf teszt & \bf pontosság & \bf fedés & \bf F-mérték \\ 
+\midrule
+Szeged & Szeged & 94.50 & 94.35 & 94.43 \\
+huwiki & huwiki & 90.64 & 88.91 &  89.76 \\
+huwiki & Szeged & 63.08 & 70.46 & 66.57 \\
+Szeged\_wikilisták & Szeged & 95.48 & 95.48 & 95.48 \\
+Szeged\_wikitag & Szeged & 95.38 & 94.92 & 95.15 \\
+\bottomrule
+\end{tabular}
+\end{center}
+\caption{Eredmények a magyar Wikipédia korpuszon.}
+\label{tab:huresults}
+\end{table}
+
+A Wikipédia korpuszokat más módon is használhatjuk a tulajdonnév-felismerés teljesítményének növelése érdekében.
+Egy kézenfekvő megoldás nagyméretű névlisták kinyerése a Wikipédiából, és azok hozzáadása gazetteer listaként a tanításhoz.
+Ez a módszer mindkét nyelv esetében növelte az F-mértéket.
+Egy másik kísérletünkben a rendszert a Wikipédia korpuszon tanítottuk, majd az általa kibocsátott címkéket jegyként hozzáadtuk a gold standard korpuszon való tanításhoz és teszteléshez.
+Ezzel a módszerrel is sikerült javítani a rendszer teljesítményét.
+
+Ha összehasonlítjuk a magyar és az angol eredményeket, azt látjuk, hogy a magyar korpuszokon szignifikánsan jobb eredményeket értünk el.
+Ennek több oka is lehet.
+Az egyik, hogy a magyar nyelv esetében a \texttt{Misc} kategóriába sokkal kevesebb dolog tartozik, vagyis kevésbé heterogén, így kevesebb hibát lehet elkövetni az automatikus annotálás során.
+És ahogy a magyar korpuszon végzett kézi hibaelemzés eredménye is mutatja, tényleg sikerült elkerülni az esetek nagy részében a félrecímkézéseket.
+A másik ok az lehet, hogy a magyar esetében használt silver és gold standard adathalmazok egyaránt sokkal tisztábbak, a NE-annotálás pontosabb, mint az angol erőforrások esetében.
+Mivel az automatikusan épített magyar korpuszunk közelíti a gold standard minőséget, mondhatjuk, hogy olyan erőforrást sikerült létrehoznunk, amely használható tulajdonnév-felismerő rendszerek tanító és kiértékelő korpuszaként, akár önállóan, akár kiegészítve a gold standard korpuszokat.
+
+\section{Összegzés}
+\label{conc}
+
+Cikkünkben egy olyan új módszert mutattunk be, amellyel magyar és angol nyelvű, automatikusan tulajdonnév-annotált korpuszokat hoztunk létre a Wikipédiából.
+Az eddig alkalmazottakkal ellentétben a mi metódusunk egy leképezést valósít meg a DBpedia ontológiai osztályairól a hagyományos címkekészletekre.
+Az így generált címkéket aztán a rendszer hozzárendeli a Wikipédiában szereplő entitásokhoz.
+
+Módszerünk nyilvánvaló előnyei, hogy nagyban csökkenti az annotálás költségeit, valamint hogy sokkal nagyobb adathalmazokat állíthatunk elő általa, mint kézi annotációval.
+Egy másik előnye, hogy bármely Wikipédiával rendelkező nyelvre alkalmazható, így kevés erőforrással rendelkező nyelvekre is előállíthatunk silver standard korpuszokat.
+A létrehozott korpuszok a továbbiakban számos módon alkalmazhatók a tulajdonnév-felismerő rendszerek hatékonyságának növelésére.
+Amennyiben kellően tiszta a korpusz, vagy az adott nyelvre nem létezik gold standard tisztaságú adathalmaz, felügyelt gépi tanulási rendszerekhez használható tanításhoz és kiértékeléshez.
+Továbbá erőforrásokkal bővebben ellátott nyelvek esetében is hasznosítható a klasszikus sajtó stílustól eltérő szövegek tulajdonnév-annotálásához.
+
+További, újdonságnak számító eredményünk, hogy az általunk előállított korpuszok szabadon elérhetőek és felhasználhatóak.
+Tudomásunk szerint ez az első magyar nyelvű automatikusan előállított tulajdonnév-annotált korpusz.
+Az angol erőforrások tekintetében is hasonló a helyzet: jelenleg a Semantically Annotated Snapshot of English Wikipedia mellett az itt publikált korpusz az egyetlen szabadon felhasználható tulajdonnév-annotált korpusz.
+
+Jelen cikkünkben a DBpedia ontológiai kategóriáit a sztenderd tulajdonnév-címkékre képeztük le, de a módszerben benne rejlik a lehetőség finomabbra hangolt tulajdonnév-hierarchiák támogatására is.
+Az internetes közösség által létrehozott tartalmak, mint a Wikipédia és a DBpedia, folyamatosan növekszenek, ezáltal egyre több információ felhasználását teszik lehetővé.
+A módszer frissítésével egyre nagyobb és finomabban annotált korpuszokat tudunk létrehozni a jövőben.
 
@@
 271022).
-A szerzők ezúton fejezik ki köszönetüket Zséder Attilának a Wikipédia-szövegek feldolgozásában végzett munkájáért, és Kornai Andrásnak támogatásáért.
+A szerzők ezúton fejezik ki köszönetüket Zséder Attilának a Wikipédia szövegek feldolgozásában végzett munkájáért, és Kornai Andrásnak támogatásáért.
 %vagy mi
 
-%
-% ---- Bibliography ----
-%
-\begin{thebibliography}{15}
-%
-\bibitem{alexin:eacl2003}
-Alexin Z., Csirik, J., Gyim\'othy, T., Bibok K., Hatvani, Cs.,
-Pr\'osz\'eky, G., Tihanyi, L.:
-Manually Annotated Hungarian Corpus.
-in Proc. of the Research Note Sessions of the 10th Conference
-of the European Chapter of the Association for Computational
-Linguistics EACL'03, Budapest, Hungary, 53--56 (2003).
-%
-\bibitem{bibok:mszny2003}
-Bibok K.:
-A sz\'or\'ol \'es a sz\'ofajokr\'ol (a sz\'am\'\i t\'og\'epes nyelvfeldolgoz\'as kap- cs\'an),
-Magyar Sz\'am\'\i t\'og\'epes Nyelv\'eszeti Konferencia (MSZNY 2003),
-Szeged, Magyarorsz\'ag, 31--36, (2003).
-%
-\bibitem{farkas:mszny2004}
-Farkas R., Konczer K., Szarvas Gy.:
-Szemantikus keretilleszt\'es \'es az IE rendszer automatikus ki\'ert\'ekel\'se
-Magyar Sz\'am\'\i t\'og\'epes Nyelv\'eszeti Konferencia (MSZNY 2004),
-bek\"uldve, Szeged, Magyarorsz\'ag, (2004).
-%
-\bibitem{hocza:acta2004}
-H\'ocza, A.:
-Noun Phrase Recognition with Tree Patterns
-elfogadva az Acta Cybernetica c. lapban t\"ort\'en\H{o} megjelen\'esre (2004).
-%
-\bibitem{mihaczi:mszny2003}
-Mih\'aczi Andr\'as, N\'emeth L\'aszl\'o, R\'acz Mikl\'os:
-Magyar sz\"ovegek tem\'eszetes nyelvi el\H{o}feldolgoz\'asa
-Magyar Sz\'am\'\i t\'og\'epes Nyelv\'eszeti Konferencia (MSZNY 2003),
-Szeged, Magyarorsz\'ag, 38--43, (2003).
-%
-\bibitem{proszeky:mszny2003}
-Pr\'osz\'eky G.:
-Automatikus inform\'aci\'oszerz\'es gazdas\'agi r\"ovidh\'\i rekb\H{o}l.
-Magyar Sz\'am\'\i t\'og\'epes Nyelv\'eszeti Konferencia (MSZNY 2003),
-Szeged, Magyarorsz\'ag, 161--166, (2003).
-%
-\bibitem{proszeky:neumann2003}
-Pr\'osz\'eky G.:
-Automatikus inform\'aci\'oszerz\'es gazdas\'agi-politikai r\"ovidh\'\i rekb\H{o}l.
-VIII. Orsz\'agos (Centen\'ariumi) Neumann Kongresszus kiadv\'anya,
-Budapest, Magyarorsz\'ag, 359--367, (2003).
-%
-\end{thebibliography}
+\bibliographystyle{plain}
+\bibliography{cikk}
+
 \end{document}

### 2012-11-27 paper/MSZNY2013/cikk.tex
Modified a few lines. modified:   cikk.tex

@@
 mit tekintenek névelemnek: pl.~a \textit{római} szó a Szeged korpuszban nem
-számít annak, míg angol megfelelője, a \textit{Roman} a CoNLL annotációs sémában
-\texttt{Misc} címkét kapna. A továbbiakban röviden ismertetjük a fenti lépéseket,
-a magyar nyelvű korpuszra koncentrálva. Részletes leírás, illetve az angol nyelvű
-korpuszban felmerülő problémák kifejtése \cite{simon-nemeskey:2012:NEWS2012}-ben
-található.
+számít annak, míg angol megfelelője, a \textit{Roman} a CoNLL annotációs séma
+szerint \texttt{Misc} címkét kapna.
+
+A továbbiakban röviden ismertetjük a fenti lépéseket, a magyar nyelvű korpuszra
+koncentrálva. Részletes leírás, illetve az angol nyelvű korpuszban felmerülő
+problémák kifejtése \cite{simon-nemeskey:2012:NEWS2012}-ben található.
 
@@
 
+% TODO: első részt máshova
+A korpuszon a minőség javítása érdekében tartalmi és formai szűrést is
+végeztünk. Értelemszerűen eldobtunk minden mondatot, ami azonosítatlan típusú
+tulajdonnevet tartalmazott. Kiszűrtük továbbá a rossz minőségű, töredék
+mondatokat, vagyis azokat, amelyek nem nagybetűvel kezdődnek, és nincs írásjel
+a végükön.
+
 Bár Nothman az első feladatot gépi tanulással oldotta meg, mi a nagyobb
@@
 A magyar Wikipédia korpusz gépi annotálását kézzel ellenőriztük a korpusz egy kis részén.
-A 19 milliós teljes korpuszt vettük alapul, amelyből már ki vannak szűrve a rossz minőségű mondatok, vagyis azok, amelyek nem nagybetűvel kezdődnek, és nincs írásjel a végükön.
+A 19 milliós teljes korpuszt vettük alapul.
 Ebből véletlenszerű mondatválogatással csináltunk egy kisebb korpuszt, amely kb.~19000 tokent tartalmaz.

### 2012-11-27 paper/MSZNY2013/cikk.tex
All that remains on my side is 3/3. Also, file set to use magyarlatex. 	modified:   cikk.tex

@@
 
-\usepackage[hungarian]{babel}
-\selectlanguage{hungarian}
+\usepackage{t1enc}
+\usepackage[magyar]{babel}
+\selectlanguage{magyar}
 
@@
 % TODO: lame wording, angol korpusz?
-A korpuszépítő algoritmus, hasonlóan Nothman et al.~\cite{Nothman:08} által
-leírthoz, a következő lépésekből áll:
+A korpuszépítő algoritmus a Wikipédia két feltételezett tulajdonságán alapszik:
+
+\begin{enumerate}
+\item a szócikkek többsége nevesített entitásokról szól;
+\item a folyó szövegben előforduló tulajdonneveket a kereszthivatkozások
+      azonosítják.
+\end{enumerate}
+
+Algoritmusunk e két feltételezés következményeit használja ki. Hasonlóan Nothman
+et al.~\cite{Nothman:08} által leírthoz, a következő lépésekből áll:
 
@@
 
-% Argh, de nehéz magyarul írni :)
-
 \subsection{Wikipédia cikkek mint entitások}
 
-A DBpedia nyelve
+Ahhoz, hogy a folyó szövegben lévő kereszthivatkozásokat felhasználhassuk a
+NER céljaira, a hivatkozott Wikipédia cikkeket tulajdonnév-osztályokba kell
+sorolnunk. A Wikipédia saját kategória-rendszere e célra a kategóriák nagy
+száma, illetve a rendszerezettség teljes hiánya miatt nem alkalmas. Egyes
+szerzők, mint Kazama és Torisawa \cite{KaTo07}, vagy Nothman at
+al.~\cite{Nothman:08} felügyelt tanulással oldották meg ezt a feladatot. Mi
+azonban a klasszifikációs hibák elkerülése miatt a DBpedia~\cite{Bizer:09}
+típushierarchiájának használata mellett döntöttünk.
+
+A DBpediában\footnote{A korpusz elkészítéséhez az akkor aktuális 3.7-es verziót
+használtuk.} a típusokat egy OWL\footnote{http://www.w3.org/TR/owl-ref/}
+ontológiába rendezték. A tudásbázis a Wikipédia entitások egy részhalmazát
+tartalmazza, és mindegyik entitáshoz hozzárendeli -- többet között -- azon
+ontológia-osztályokat, amikbe az tartozik. Mivel a DBpedia alapvetően az angol
+nyelvre koncentrál, a magyar entitáslista olyan angol oldalak adatait
+tartalmazza, melyeknek létezik megfelelője a magyar Wikipédiában. Az első
+lépésben ezért ennek a 58 337 oldalnak a magyar megfelelőit határoztuk meg.
+
+A kategorizálás az ontológia-osztályok alapján történt. Minden osztályt
+leképeztünk a neki megfelelő NER címkére (pl. \texttt{Organisation}:
+\texttt{Org}, de \texttt{Organisation/SportsLeague}: \texttt{Misc}).  %TODO: ez magyarul is?
+Ezután a listán szereplő minden entitáshoz hozzárendeltük az őt tartalmazó
+legszűkebb osztály címkéjét, vagy \texttt{O}-t, ha az osztály nem NE a Szeged
+NER sémában. Így egy 46 461 elemű, felcímkézett tulajdonnév-listát kaptunk,
+melyet gold standardként használtunk a továbbiakban.
 
 \subsection{A cikkek feldolgozása}
+
+A korpusz építéséhez a magyar Wikipédia 2012 március 9-ei állapotát vettük
+alapul. Az XML adatból az mwlib\footnote{http://code.pediapress.com}
+könyvtár segítségével nyertük ki a nyers szöveget. Ezt egy in-house statisztikai %TODO: in-house?, idézet (Csendes)
+eszközzel tokenizáltuk; a szavak lemmáit és szófajait a
+HunMorphot~\cite{Tron:05} és az erre épülő egyértelműsítőt, a HunDisambigot
+használtuk. A szófajokat KR-kódban~\cite{Kornai:04} ábrázoljuk.
+
 \subsection{Tulajdonnevek címkézése}
+
+% TODO
 
@@
 a végükön.
-
-Bár Nothman az első feladatot gépi tanulással oldotta meg, mi a nagyobb
-pontosság érdekében a DBpedia~\cite{Bizer:09} tudásbázis alapján osztályoztuk az
-entitásokat. A tulajdonneveket Wikipédia kereszthivatkozások segítségével azonosítottuk.
-Minden olyan mondatot, ahol ismeretlen név fordult elő, eldobtunk.
 

### 2012-11-27 paper/MSZNY2013/cikk.tex
Almost done, almost... 	modified:   cikk.tex

@@
 
+%TODO: itt kicsit kifejteni
 Algoritmusunk e két feltételezés következményeit használja ki. Hasonlóan Nothman
@@
 \subsection{Wikipédia cikkek mint entitások}
+\label{corpusbuilding:entities}
 
@@
 
-% TODO
+A címkézés két feladatot foglal magába: egyrészt azonosítani kell az
+entitásokat a folyó szövegben, másrészt besorolni őket a megfelelő NE
+kategóriába. Módszerünk Nothman et al.~\cite{Nothman:08} munkáján alapszik:
+a szövegben előforduló kereszthivatkozásokat tekintjük entitásoknak. Minden
+entitásnál megnézzük, hogy az oldal, amire mutat, szerepel-e a
+\ref{corpusbuilding:entities} fejezetben előállított listában. Ha igen,
+felcímkézzük a megfelelő címkével; ha nem, típusa \texttt{Unk} lesz. Végül
+minden mondatot, amiben \texttt{Unk} entitások szerepelnek, eldobunk.
+
+% TODO argh
+A fenti tiszta, még nyelvfüggetlen algoritmus többhelyütt is finomításra szorul.
+
+A Wikipédia nem minden szócikke szól entitásokról: köznevek, dátumok, stb. is
+alkalmasint saját oldalt kapnak. Annak érdekében, hogy az algoritmus ne
+kezelje ezeket \texttt{Unk} típusú entitásként, majd dobja el a mondatokat,
+amikben szerepelnek, szigorítottuk az entitásfelismerés szabályait: csak olyan
+kereszthivatkozást tekintünk entitásnak, melynek szavai nagybetűvel kezdődnek.
+Itt és később is kihasználtuk a magyar nyelv azon tulajdonságát, hogy a
+tulajdonneveket, és csak azokat kezdjük nagybetűvel.
+
+% TODO alias?!
+Az előző probléma inverze is gyakori: nem minden tulajdonnéven találunk ugyanis
+kereszthivatkozást. Ennek oka kettős: ha egy Wikipédia szócikkben adott entitás
+többször is előfordul, csak az első alkalommal linkelik a saját oldalához. De
+előfordulhat az is, hogy az entitás nem rendelkezik saját szócikkel. Az ilyen
+esetek kezelésére minden szócikknél fenntartunk egy listát, ahol az abban
+felismert entitásokat, illetve aliasaikat gyűjtjük, mint pl. a rájuk mutató
+átirányító és egyértelműsítő lapok címei. Ha ezután egy olyan mondattal
+találkozunk, amiben nagybetűs szócsoport szerepel, ellenőrizzük, hogy a csoport
+egyezik-e a listában szereplő egyik entitás nevével. Amennyiben igen,
+felcímkézzük; ha nem, a nagybetűs szavakat ismeretlen entitásnak tekintjük.
+
+Szigorúan véve, a fenti módszer minden mondatot eldobna, hiszen a mondat első
+szavát mindig nagybetűvel írjuk. Így annyit módosítottunk rajta, hogy ha a
+mondatkezdő szó egyedül alkot csoportot, csak akkor tekintjük potenciális
+entitásnak, ha szófaja főnév.
+
+% TODO: 3-ba
+Mivel célunk egy silver standard korpusz előállítása volt, ha választanunk
+kellett egy-egy lépésnél a pontosság (precision) vagy a teljesség (recall)
+között, mindig az utóbbi mellett döntöttünk. Ezt tükrözik a fenti, alkalmasint
+kicsit szigorú szabályok, illetve a DBpedia választása is. A Wikipédia mérete
+lehetővé teszi, hogy szigorú szűrések mellett is korábban nem látott méretű
+korpuszt állítsunk elő.
 
@@
 
-% TODO: első részt máshova
 A korpuszon a minőség javítása érdekében tartalmi és formai szűrést is

### 2012-11-28 paper/MSZNY2013/cikk.tex
Finished section 3. All that's left is to clear the TODOs. 	modified:   cikk.tex
jav

--- 2012-11-27 paper/MSZNY2013/cikk.tex
@@
 
-\usepackage[hungarian]{babel}
-\selectlanguage{hungarian}
+\usepackage{t1enc}
+\usepackage[magyar]{babel}
+\selectlanguage{magyar}
 
 \usepackage{booktabs}
+\usepackage{hyperref}
 
@@
 \begin{abstract}
-A felügyelt gépi tanulási módszerek alkalmazásához nagy méretű annotált korpuszokra van szükség, amelyek előállítása rendkívül emberierőforrás-igényes.
+A felügyelt gépi tanulási módszerek alkalmazásához nagyméretű annotált korpuszokra van szükség, amelyek előállítása rendkívül emberierőforrás-igényes.
 Több lehetőség van az annotációs költségek csökkentésére, ezek közül az egyik az automatikus annotálás.
@@
 A korpuszok szövege a Wikipédiából származik, és tulajdonnévi annotálást is tartalmaz.
-Az automatikus annotáláshoz egy új módszert alkalmaztunk: a DBpedia ontológiai kategóriáit képeztük le a CoNLL névosztályokra.
+Az automatikus annotáláshoz egy új módszert alkalmaztunk: a DBpedia ontológiai kategóriáit képeztük le CoNLL névosztályokra.
 A korpuszok szabadon felhasználhatóak további fejlesztésekhez.
@@
 Cikkünkben egy olyan megközelítést mutatunk be, mely ezen lehetőségeket kombinálja: automatikus eszközökkel tulajdonnév-annotált korpuszokat építettünk Wikipédia szócikkekből.
-Munkánk során új módszert alkalmaztunk: a DBpedia ontológiai kategóriáit képeztük le a CoNLL-névosztályokra.
+Munkánk során új módszert alkalmaztunk: a DBpedia ontológiai kategóriáit képeztük le CoNLL-névosztályokra.
 A létrehozott magyar és angol nyelvű korpuszok szabadon felhasználhatóak.
@@
 A Wikipédia egy többnyelvű, nyílt tartalmú, az internetes közösség által fejlesztett webes világenciklopédia \footnote{http://wikipedia.org}.
-Jelenlegi XXX számú szócikkével aranybánya a különböző természetesnyelv-feldolgozó fejlesztések számára; használták már többek között jelentésegyértelműsítére, ontológia- és tezauruszépítésre, valamint kérdésmegválaszoló rendszerekhez (további alkalmazási lehetőségekhez lásd \cite{Medelyan:09}).
-Mivel a Wikipédia címszavak jelentős része tulajdonnév, adja magát a lehetőség, hogy az automatikus tulajdonnév-felismeréshez is felhasználjuk.
+Több mint 22 millió szócikkével aranybánya a különböző természetesnyelv-feldolgozó fejlesztések számára; használták már többek között jelentésegyértelműsítére, ontológia- és tezauruszépítésre, valamint kérdésmegválaszoló rendszerekhez (további alkalmazási lehetőségekhez lásd \cite{Medelyan:09}).
+Mivel a Wikipédia címszavak jelentős része tulajdonnév, adja magát a lehetőség, hogy a tulajdonnév-felismeréshez is használjuk.
 
@@
 
-Az általunk alkalmazott módszer az említettekétől annyiban tér el, hogy mi a DBpedia ontológiai osztályait képeztük le a sztenderd CoNLL névosztályokra, majd ezeket Wikipédia entitásokhoz kötöttük.
+Az általunk alkalmazott módszer az említettektől annyiban tér el, hogy mi a DBpedia ontológiai osztályait képeztük le a sztenderd CoNLL névosztályokra, majd ezeket Wikipédia entitásokhoz kötöttük.
 Az így létrehozott korpuszokat szabadon elérhetővé tettük.
-Tudomásunk szerint jelenleg csak egy angol nyelvű automatikusan tulajdonnév-annotált korpusz létezik, amely szabadon felhasználható, a Semantically Annotated Snapshot of the English Wikipedia \cite{Atserias:08}.
+%Tudomásunk szerint jelenleg csak egy angol nyelvű automatikusan tulajdonnév-annotált korpusz létezik, amely szabadon felhasználható, a Semantically Annotated Snapshot of the English Wikipedia \cite{Atserias:08}.
 Bár a névfelismerést a jelenlegi legjobb szabadon hozzáférhető rendszerekkel végezték, azt gondoljuk, hogy a Wikipédia többezer szerkesztőjének döntésein alapuló kategorizálás megbízhatóbb eredményt ad.
@@
 Az általunk létrehozott korpusz az első magyar nyelvű automatikusan tulajdonnév-annotált korpusz.
-%XXX ezt kábé mind leírtam a conclusionben is, szóval innen akár ki is lehet venni
 
@@
 % TODO: lame wording, angol korpusz?
-A korpuszépítő algoritmus, hasonlóan Nothman et al.~\cite{Nothman:08} által
-leírthoz, a következő lépésekből áll:
+%\begin{enumerate}
+%\item a szócikkek többsége nevesített entitásokról szól;
+%\item a folyó szövegben előforduló entitásokat a kereszthivatkozások
+%      azonosítják.
+%\end{enumerate}
+
+%TODO: itt kicsit kifejteni
+
+A korpuszépítő algoritmus a Wikipédia két feltételezett tulajdonságán alapszik, miszerint a szócikkek többsége megnevezett entitásokról szól, valamint a folyó szövegben előforduló entitásokat a kereszthivatkozások azonosítják.
+Algoritmusunk e két feltételezés következményeit használja ki.
+Hasonlóan a Nothman et al.~\cite{Nothman:08} által leírtakhoz, a következő lépésekből áll:
 
 \begin{enumerate}
-\item a Wikipédia cikkeket entitásosztályokba soroltuk;
-\item a cikkeket mondatokra bontottuk;
-\item felcímkéztük a tulajdonneveket a szövegben;
-\item kiszűrtük a rossz minőségű mondatokat.
+\item a Wikipédia cikkeket entitásosztályokba soroljuk;
+\item a cikkeket mondatokra bontjuk;
+\item felcímkézzük a tulajdonneveket a szövegben;
+\item kiszűrjük a rossz minőségű mondatokat.
 \end{enumerate}
@@
 rendelkezik megfelelő méretű Wikipédiával. Egyedül a harmadik lépés az, ahol
-figyelembe kell venni a nyelv, illetve a használt NER konvenció sajátosságait.
-Ennek oka, hogy az egyes nyelvek, illetve NER konvenciók eltérnek abban, hogy
-mit tekintenek névelemnek: pl.~a \textit{római} szó a Szeged korpuszban nem
-számít annak, míg angol megfelelője, a \textit{Roman} a CoNLL annotációs séma
+figyelembe kell venni a nyelv, illetve a használt annotációs séma sajátosságait.
+Ennek oka, hogy az egyes nyelvek, illetve sémák eltérnek abban, hogy
+mit tekintenek annotálandó elemnek: pl.~a \textit{római} szó a magyarban nem
+számít névnek, míg angol megfelelője, a \textit{Roman} a CoNLL annotációs séma
 szerint \texttt{Misc} címkét kapna.
 
-A továbbiakban röviden ismertetjük a fenti lépéseket, a magyar nyelvű korpuszra
+Mivel célunk egy silver standard korpusz előállítása volt, ha választanunk
+kellett egy-egy lépésnél a pontosság (precision) vagy a teljesség (recall)
+között, mindig az előbbi mellett döntöttünk. A Wikipédia mérete lehetővé teszi,
+hogy szigorú szűrések mellett is korábban nem látott méretű korpuszt állítsunk elő.
+
+A továbbiakban röviden ismertetjük a fenti lépéseket, kizárólag a magyar nyelvű korpuszra
 koncentrálva. Részletes leírás, illetve az angol nyelvű korpuszban felmerülő
@@
 
-% Argh, de nehéz magyarul írni :)
+\subsection{Wikipédia cikkek mint entitások}
+\label{corpusbuilding:entities}
 
-\subsection{Wikipédia cikkek mint entitások}
+Ahhoz, hogy a folyó szövegben lévő kereszthivatkozásokat felhasználhassuk a
+tulajdonnevek azonosítására, a hivatkozott Wikipédia cikkeket névosztályokba kell
+sorolnunk. A Wikipédia saját kategóriarendszere a kategóriák nagy
+száma, illetve a rendszerezettség teljes hiánya miatt nem alkalmas erre a célra. Egyes
+szerzők, mint Kazama és Torisawa \cite{KaTo07} vagy Nothman et
+al.~\cite{Nothman:08} felügyelt tanulással oldották meg ezt a feladatot. Mi
+azonban a klasszifikációs hibák elkerülése céljából a DBpedia~\cite{Bizer:09}
+típushierarchiájának felhasználása mellett döntöttünk.
 
-A DBpedia nyelve
+A DBpediában\footnote{A korpusz elkészítéséhez az akkor aktuális 3.7-es verziót
+használtuk.} a típusok egy OWL\footnote{http://www.w3.org/TR/owl-ref/}
+ontológia részei. A tudásbázis a Wikipédia entitások egy részhalmazát
+tartalmazza, és mindegyik entitáshoz hozzárendeli -- többet között -- azon
+ontológiaosztályokat, amelyekbe az tartozik. Mivel a DBpediának nincs magyar változata, 
+a magyar entitáslista csak olyan angol oldalak adatait
+tartalmazza, melyeknek létezik megfelelője a magyar Wikipédiában. 
+Ezáltal a feldolgozás köréből kiesnek kifejezetten magyarspecifikus oldalak, de 
+így is 58.337 oldal anyagát dolgoztuk fel.
+
+A magyar nevek kategorizálásához a Szeged NER \cite{Szarvas:06} korpuszban alkalmazott névosztályokat vettük alapul, ezekre képeztük le a DBpedia egyes ontológiaosztályait. 
+%leképeztünk a neki megfelelő NER címkére (pl. \texttt{Organisation}:
+%\texttt{Org}, de \texttt{Organisation/SportsLeague}: \texttt{Misc}).  %TODO: ez magyarul is?
+Ezután minden entitáshoz hozzárendeltük az őt tartalmazó
+legszűkebb osztály címkéjét, vagy \texttt{O}-t, ha az osztály nem minősül annotálandónak a Szeged
+NER korpusz sémája szerint, így egy 46.461 elemű, felcímkézett tulajdonnév-listát kaptunk.
 
 \subsection{A cikkek feldolgozása}
+
+A korpusz építéséhez a magyar Wikipédia 2012. március 9-i állapotát vettük
+alapul. Az XML fájlokból az mwlib\footnote{http://code.pediapress.com}
+könyvtár segítségével nyertük ki a nyers szöveget. 
+A tokenizáláshoz egy házon belül előállított statisztikai eszközt használtunk, amelyet a Szeged korpuszon \cite{csendes} tanítottunk. 
+A lemmatizálást és a morfológiai elemzést a
+HunMorph~\cite{Tron:05} és az erre épülő egyértelműsítő, a HunDisambig
+segítségével végeztük. %A szófajokat KR-kódban~\cite{Kornai:04} ábrázoljuk.
+
 \subsection{Tulajdonnevek címkézése}
+
+A címkézés két feladatot foglal magába: egyrészt azonosítani kell az
+entitásokat a folyó szövegben, másrészt besorolni őket a megfelelő
+névosztályba. 
+A szövegben előforduló kereszthivatkozásokat potenciális neveknek tekintjük, és minden
+linknél megnézzük, hogy az oldal, amire mutat, szerepel-e a
+korábban előállított listában (lásd a \ref{corpusbuilding:entities}.~fejezetben). Ha igen,
+felcímkézzük a megfelelő címkével; ha nem, típusa \texttt{Unk} lesz. Végül
+minden mondatot, amelyben \texttt{Unk}-ként címkézett elem szerepel, eldobunk.
+
+% TODO argh
+Kiinduló alapfeltételezésünk, miszerint az egyes Wikipédia cikkek megnevezett entitásokról szólnak, és a szövegben előforduló entitásokat kereszthivatkozások azonosítják, nem minden esetben állja meg a helyét, ezért az algoritmusunk
+többhelyütt is finomításra szorul.
+
+Először, a Wikipédia nem minden szócikke szól megnevezett entitásokról: köznevek, dátumok és egyéb, nem tulajdonnévi elemek is
+alkalmasint saját oldalt kapnak. Annak érdekében, hogy az algoritmus ne
+kezelje ezeket \texttt{Unk} típusú entitásként, majd dobja el a mondatokat,
+amikben szerepelnek, szigorítottuk az entitásfelismerés szabályait: csak olyan
+kereszthivatkozást tekintünk potenciális névre való utalásnak, melynek szavai nagybetűvel kezdődnek.
+Itt és később is kihasználtuk a magyar nyelv azon tulajdonságát, hogy a
+tulajdonneveket, és csak azokat kezdjük nagybetűvel. 
+Kivételt képeznek ez alól természetesen a mondatkezdő pozícióban szereplő szavak.
+Szigorúan véve, ez a módszer minden mondatot eldobna, így annyit módosítottunk rajta, hogy a mondatkezdő szót csak akkor tekintjük potenciális annotálandó elemnek, ha a szófaja főnév.
+(A magyarban a morfológiai címkékre csak korlátozottan támaszkodhattunk ebben a feladatban, hiszen a KR-kódolás nem különbözteti meg a tulajdonneveket és a közneveket.)
+
+Másodszor, nem minden tulajdonnéven találunk
+kereszthivatkozást. Ennek oka kettős: ha egy Wikipédia szócikkben adott entitásra
+többször is névvel utalnak, csak az első alkalommal linkelik a saját oldalához. De
+előfordulhat az is, hogy az entitás nem rendelkezik saját szócikkel. Az ilyen
+esetek kezelésére minden szócikknél fenntartunk egy listát, ahol az abban
+felismert entitásokat, illetve egyéb nagybetűs említéseit, mint például a rájuk mutató
+átirányító és egyértelműsítő lapok címeit gyűjtjük. Ha ezután egy olyan mondattal
+találkozunk, amelyben nagybetűs szócsoport szerepel, ellenőrizzük, hogy a csoport
+egyezik-e a listában szereplő egyik entitás nevével. Amennyiben igen,
+felcímkézzük; ha nem, a nagybetűs szavakat ismeretlen entitásnak tekintjük.
 
@@
 
-% TODO: első részt máshova
-A korpuszon a minőség javítása érdekében tartalmi és formai szűrést is
-végeztünk. Értelemszerűen eldobtunk minden mondatot, ami azonosítatlan típusú
-tulajdonnevet tartalmazott. Kiszűrtük továbbá a rossz minőségű, töredék
-mondatokat, vagyis azokat, amelyek nem nagybetűvel kezdődnek, és nincs írásjel
-a végükön.
+Ahogy már említettük, az azonosítatlan entitást tartalmazó mondatokat kidobtuk a korpuszból, hogy növeljük az annotálás pontosságát.
 
-Bár Nothman az első feladatot gépi tanulással oldotta meg, mi a nagyobb
-pontosság érdekében a DBpedia~\cite{Bizer:09} tudásbázis alapján osztályoztuk az
-entitásokat. A tulajdonneveket Wikipédia kereszthivatkozások segítségével azonosítottuk.
-Minden olyan mondatot, ahol ismeretlen név fordult elő, eldobtunk.
+A hagyományos névfelismerő rendszerek tiszta, viszonylag pontosan annotált korpuszokból tanulják ki a megfelelő paramétereiket, és a tesztelésükhöz is hasonló adathalmazra van szükség.
+Ezért a rossz minőségű, töredékes mondatokat, amelyek nem nagybetűvel kezdődnek és nincs mondatzáró írásjel a végükön, szintén kiszűrtük.
+(Az így maradt 19 milliós adathalmazzal dolgoztunk tovább; a következőkben erre referálunk teljes korpuszként.) De a minőség javítása érdekében további szűrő lépéseket is lehet tenni, így például eldobhatjuk az olyan mondatokat is, amelyek nem tartalmaznak ragozott igét.
+Ezzel ugyan azok is kiesnek, amelyek szabályos jelen idejű, kijelentő módú, létigét (nem) tartalmazó mondatok, de az automatikus tokenizálás és mondatra bontás hibáiból származó töredékes mondatokat eltávolíthatjuk.
+
+Ha viszont közösségi tartalmak (User Generated Content) feldolgozására kívánja valaki használni a korpuszokat, amelyek köztudottan sokkal zajosabbak, mint a kézzel annotált adatok, és sok töredékes mondatot tartalmaznak, hasznos lehet minél kevesebb szűrő használata.
+Ezért a rossznak minősített mondatokat nem dobtuk el végleg, hanem a korpusz mellett ezt az adathalmazt is elérhetővé tettük.
 
@@
 A magyar Wikipédia korpusz gépi annotálását kézzel ellenőriztük a korpusz egy kis részén.
-A 19 milliós teljes korpuszt vettük alapul.
-Ebből véletlenszerű mondatválogatással csináltunk egy kisebb korpuszt, amely kb.~19000 tokent tartalmaz.
+A 19 milliós teljes korpuszt vettük alapul, amelyből véletlenszerű mondatválogatással csináltunk egy kb.~19000 tokent tartalmazó kisebb korpuszt.
 Ezt kézzel felannotáltuk, és összehasonlítottuk a kézi és a gépi annotálás eredményét.
@@
 Emberekre utaló linkek esetében ezek nagyrészt rangjelzők, pl.~\textit{Szent István király, Anna brit királynő}.
-%Ez azért marad benne, mert a magyar List of titles stopword listánk nincs?
+Ez utóbbiakat egy rangjelző szavakat tartalmazó tiltólistával a későbbiekben ki lehet szűrni.
 
@@
 Máskor a mondatrabontás során a szövegközbeni link szétszakad, így a maradék elem nem kapja meg a megfelelő címkét.
-Ezenkívül rengeteg töredékes mondat maradt a szövegben, amelyeket egy erős szűrővel utólag is ki lehet szűrni, pl.~ha minden olyan mondatszerű egységet kidobunk, amelyben nincs ragozott ige.
-Ezzel persze azok is kiesnek, amelyek szabályos jelen idejű, kijelentő módú, létigét (nem) tartalmazó mondatok.
+Erre esetleg megoldást jelenthet a tokenizáló teljesítményének javítása, vagy másik tokenizáló alkalmazása.
 
@@
 Az automatikusan generált korpuszok nyilvánvaló előnyei mellett azonban megvan az a hátrányuk, hogy nem használhatók gold standard adatként.
-Kiértékelésünkben megmutatjuk, hogy a létrehozott silver standard adathalmazok viszont kiválóan használhatók a tulajdonnév-felismerés teljesítményének növelésére több módon is.
+Kiértékelésünkben megmutatjuk, hogy a létrehozott silver standard adathalmaz viszont kiválóan használható a tulajdonnév-felismerés teljesítményének növelésére több módon is.
 
-A kiértékeléshez Varga és Simon \cite{Varga} tulajdonnév-felismerő rendszerét használtuk.
+A kiértékeléshez a Hunner \cite{Varga} tulajdonnév-felismerő rendszert használtuk.
 A csak az egyes korpuszokra jellemző jegyeket (pl.~főnévi csoportok jelölése, Wikipédia linkek) kidobtuk, hogy növeljük a korpuszok összehasonlíthatóságát.
@@
 
-A korpuszok a fent leírt szűrő eljárások után maradt mondatokat tartalmazzák, így azokat is, amelyekben nincs egy név sem.
+A korpusz a fent leírt szűrő eljárások után maradt mondatokat tartalmazza, így azokat is, amelyekben nincs egy név sem.
 Ezeket azért tartottuk meg, hogy amennyire lehetséges, megőrizzük a nevek eredeti, Wikipédia-beli eloszlását.
-Viszont amikor megvizsgáltuk az egyes korpuszok telítettségét a nevek szempontjából, arra jutottunk, hogy a gold standard adathalmazokkal való összevetésükkor inkább sűrítjük a szövegeket, vagyis kivesszük azokat a mondatokat, amelyekben nincs név.
-A \ref{tab:size_en}.~és \ref{tab:size_hu}.~táblázatok mutatják az egyes korpuszokra vonatkozó számszerű adatokat, melyekből jól látható, hogy a Wikipédiából generált korpuszok telítettsége meglehetősen alacsony, a nevet nem tartalmazó mondatok kiszűrése után viszont hasonlóak lettek a számok.
-Mivel a Wikipédia az entitások kincsesbányájaként van számon tartva, ezek az adatok kissé meglepőek lehetnek.
-A szövegeknek ez a hígsága valószínűleg annak köszönhető, hogy a módszerünk szigorú, vagyis inkább minden olyan mondatot eltávolítottunk, amelyben nem lehetett beazonosítani a nevet, minthogy rosszul annotált nevek maradjanak benne.
+Viszont amikor megvizsgáltuk az egyes korpuszok telítettségét a nevek szempontjából, arra jutottunk, hogy a gold standard adathalmazzal való összevetéskor inkább sűrítjük a szöveget, vagyis kivesszük azokat a mondatokat, amelyekben nincs név.
+A \ref{tab:size_hu}.~táblázat mutatja a magyar korpuszokra vonatkozó számszerű adatokat, melyekből jól látható, hogy a Wikipédiából generált korpusz telítettsége meglehetősen alacsony, a nevet nem tartalmazó mondatok kiszűrése után viszont hasonlóak lettek a számok.
+Mivel a Wikipédia az entitások kincsesbányájaként van számon tartva, ez az adat kissé meglepő lehet.
+A szövegnek ez a hígsága valószínűleg annak köszönhető, hogy a módszerünk szigorú, vagyis inkább minden olyan mondatot eltávolítottunk, amelyben nem lehetett beazonosítani a nevet, minthogy rosszul annotált nevek maradjanak benne.
 
-\begin{table}[ht]
-\begin{center}
-\begin{tabular}{lccc}
-\toprule  
-& \textbf{enwiki} & \textbf{enwiki filter} & \textbf{CoNLL} \\ 
-\midrule
-token & 60,520,819 & 21,718,854 & 301,418 \\
-NE & 3,169,863 & 3,169,863 & 35,089 \\
-telítettség (\%) & 5.23 & 14.59 & 11.64 \\
-\hline
-\end{tabular}
-\end{center}
-\caption{Az angol korpuszok mérete és telítettsége.}
-\label{tab:size_en}
-\end{table}
+%\begin{table}[ht]
+%\begin{center}
+%\begin{tabular}{lccc}
+%\toprule  
+%& \textbf{enwiki} & \textbf{enwiki filter} & \textbf{CoNLL} \\ 
+%\midrule
+%token & 60,520,819 & 21,718,854 & 301,418 \\
+%NE & 3,169,863 & 3,169,863 & 35,089 \\
+%telítettség (\%) & 5.23 & 14.59 & 11.64 \\
+%\hline
+%\end{tabular}
+%\end{center}
+%\caption{Az angol korpuszok mérete és telítettsége.}
+%\label{tab:size_en}
+%\end{table}
 
@@
 \midrule
-token & 19,108,027 & 3,512,249  & 225,963 \\
-NE & 456,281 & 456,281  & 25,896 \\
-telítettség (\%) & 2.38 & 12.99 & 11.46 \\
+token & 19.108.027 & 3.512.249  & 225.963 \\
+NE & 456.281 & 456.281  & 25.896 \\
+telítettség (\%) & 2,38 & 12,99 & 11,46 \\
 \bottomrule
@@
 
-A magyar és az angol korpuszok is hasonlóan lettek kiértékelve: először saját magukon, aztán egy választott gold standard adathalmazon.
-Mivel az angol korpusz még a nevet nem tartalmazó mondatok eltávolítása után is nagyon nagy, kísérleteinket egy kb.~3,5 millió tokenes részén végeztük (a szűrt magyar korpusz is ekkora), 90-10\%-os arányban vágva tanító és kiértékelő korpuszra.
+Jelen cikkben csak a magyar korpuszon elért eredményeket közöljük, az angolra vonatkozó részletes adatokért lásd korábbi cikkünket \cite{simon-nemeskey:2012:NEWS2012}. 
+A korpusz kétféleképpen lett kiértékelve: először saját magán, aztán egy választott gold standard adathalmazon.
+A nevet nem tartalmazó mondatok kiszűrése után maradt 3,5 millió tokenes korpuszt 90-10\%-os arányban tanító és kiértékelő halmazra osztottuk.
+%Mivel az angol korpusz még a nevet nem tartalmazó mondatok eltávolítása után is nagyon nagy, kísérleteinket egy kb.~3,5 millió tokenes részén végeztük (a szűrt magyar korpusz is ekkora), 90-10\%-os arányban vágva tanító és kiértékelő korpuszra.
 
-Az angol korpusz kiértékeléséhez a 2003-as CoNLL korpuszt, a magyarhoz pedig a Szeged NER korpuszt választottuk.
+Mivel a névkategóriák leképezésénél a Szeged NER korpusz címkekészletét használtuk, ezért adta magát, hogy a magyar korpusz kiértékeléséhez is ugyanezt alkalmazzuk.
 Többek által (pl.
@@
 Különböző típusú szövegek esetén a különbség 20-30\% is lehet.
-A helyzet a mi esetünkben is nagyon hasonló (lásd a \ref{tab:enresults}.~és \ref{tab:huresults}.~táblázatokat az eredményekért): a Wikipédián tanított rendszerek teljesítménye közel sem olyan jó a gold standard korpuszok kiértékelő halmazán mérve, mint a saját kiértékelő halmazukon.
+A helyzet a mi esetünkben is nagyon hasonló (lásd a \ref{tab:huresults}.~táblázatot az eredményekért): a Wikipédián tanított rendszer teljesítménye közel sem olyan jó a gold standard korpusz kiértékelő halmazán mérve, mint a saját kiértékelő halmazán.
 
-\begin{table}[ht]
-\begin{center}
-\begin{tabular}{lcccc}
-\toprule 
-\textbf{tanítás} & \textbf{teszt} & \textbf{pontosság} & \textbf{fedés} & \textbf{F-mérték} \\ 
-\midrule
-CoNLL & CoNLL & 85.13 & 85.13 & 85.13 \\
-enwiki & enwiki & 72.46 & 73.33 &  72.89 \\
-enwiki & CoNLL & 56.55 & 49.77 & 52.94 \\
-CoNLL\_wikilisták & CoNLL & 86.33 & 86.35 & 86.34 \\
-CoNLL\_wikitag & CoNLL & 85.88 & 85.94 & 85.91 \\
-\bottomrule
-\end{tabular}
-\end{center}
-\caption{Eredmények az angol Wikipédia korpuszon.}
-\label{tab:enresults}
-\end{table}
+%\begin{table}[ht]
+%\begin{center}
+%\begin{tabular}{lcccc}
+%\toprule 
+%\textbf{tanítás} & \textbf{teszt} & \textbf{pontosság} & \textbf{fedés} & \textbf{F-mérték} \\ 
+%\midrule
+%CoNLL & CoNLL & 85.13 & 85.13 & 85.13 \\
+%enwiki & enwiki & 72.46 & 73.33 &  72.89 \\
+%enwiki & CoNLL & 56.55 & 49.77 & 52.94 \\
+%CoNLL\_wikilisták & CoNLL & 86.33 & 86.35 & 86.34 \\
+%CoNLL\_wikitag & CoNLL & 85.88 & 85.94 & 85.91 \\
+%\bottomrule
+%\end{tabular}
+%\end{center}
+%\caption{Eredmények az angol Wikipédia korpuszon.}
+%\label{tab:enresults}
+%\end{table}
 
@@
 \midrule
-Szeged & Szeged & 94.50 & 94.35 & 94.43 \\
-huwiki & huwiki & 90.64 & 88.91 &  89.76 \\
-huwiki & Szeged & 63.08 & 70.46 & 66.57 \\
-Szeged\_wikilisták & Szeged & 95.48 & 95.48 & 95.48 \\
-Szeged\_wikitag & Szeged & 95.38 & 94.92 & 95.15 \\
+Szeged & Szeged & 94,50 & 94,35 & 94,43 \\
+huwiki & huwiki & 90,64 & 88,91 &  89,76  \\
+huwiki & Szeged & 63,08 & 70,46 & 66,57  \\
+Szeged\_wikilisták & Szeged & 95,48 & 95,48 & 95,48  \\
+Szeged\_wikitag & Szeged & 95,38 & 94,92 & 95,15 \\
 \bottomrule
@@
 
-A Wikipédia korpuszokat más módon is használhatjuk a tulajdonnév-felismerés teljesítményének növelése érdekében.
+A Wikipédia korpuszt további módokon is használhatjuk a tulajdonnév-felismerés teljesítményének növelése érdekében.
 Egy kézenfekvő megoldás nagyméretű névlisták kinyerése a Wikipédiából, és azok hozzáadása gazetteer listaként a tanításhoz.
-Ez a módszer mindkét nyelv esetében növelte az F-mértéket.
+Ez a módszer több mint 1\%-kal növelte az F-mértéket.
 Egy másik kísérletünkben a rendszert a Wikipédia korpuszon tanítottuk, majd az általa kibocsátott címkéket jegyként hozzáadtuk a gold standard korpuszon való tanításhoz és teszteléshez.
@@
 
-Ha összehasonlítjuk a magyar és az angol eredményeket, azt látjuk, hogy a magyar korpuszokon szignifikánsan jobb eredményeket értünk el.
-Ennek több oka is lehet.
+%na ezt még majd át kell írni a végleges eredmények ismeretében. A Misc-es fejtegetést lehet, hogy inkább a hibaelemzés részbe kéne átrakni. 
+Ha összehasonlítjuk a magyar és az angol eredményeket, azt látjuk, hogy a magyar korpuszokon szignifikánsan jobb eredményeket értünk el. Ennek több oka is lehet. 
 Az egyik, hogy a magyar nyelv esetében a \texttt{Misc} kategóriába sokkal kevesebb dolog tartozik, vagyis kevésbé heterogén, így kevesebb hibát lehet elkövetni az automatikus annotálás során.

### 2012-11-29 paper/MSZNY2013/cikk.tex
Tables added. 	modified:   cikk.tex
Added tables and wrote a few lines. 	modified:   cikk.tex

@@
 HunMorph~\cite{Tron:05} és az erre épülő egyértelműsítő, a HunDisambig
-segítségével végeztük. %A szófajokat KR-kódban~\cite{Kornai:04} ábrázoljuk.
+segítségével végeztük. %A szófajokat KR-kódban~\cite{Kornai:04} ábrázoljuk. % TODO ezzel mi a baj?
 
@@
 felismert entitásokat, illetve egyéb nagybetűs említéseit, mint például a rájuk mutató
-átirányító és egyértelműsítő lapok címeit gyűjtjük. Ha ezután egy olyan mondattal
+átirányító és egyértelműsítő lapok címeit gyűjtjük. Ha ezután egy olyan mondattal  % TODO: + idegen nyelvi megfelelők!
 találkozunk, amelyben nagybetűs szócsoport szerepel, ellenőrizzük, hogy a csoport
@@
 
-A magyar Wikipédia korpusz gépi annotálását kézzel ellenőriztük a korpusz egy kis részén.
-A 19 milliós teljes korpuszt vettük alapul, amelyből véletlenszerű mondatválogatással csináltunk egy kb.~19000 tokent tartalmazó kisebb korpuszt.
-Ezt kézzel felannotáltuk, és összehasonlítottuk a kézi és a gépi annotálás eredményét.
+A magyar Wikipédia korpusz gépi annotálását kézzel ellenőriztük a korpusz egy kis
+részén. A 19 milliós teljes korpuszt vettük alapul, amelyből véletlenszerű
+mondatválogatással csináltunk egy kb.~19.000 tokent tartalmazó kisebb korpuszt.
+Ezt kézzel felannotáltuk, és összehasonlítottuk a kézi és a gépi annotálás
+eredményét. Az eredményeket a \ref{tab:f_score} táblázat mutatja. Az összesített
+F-mérték 92.94 százaléknak adódott, ami % TODO ami mi?
+
 A következő tipikus hibák rajzolódtak ki.
 
-Bizonyos esetekben a név környezetében szereplő értelmező szerepű nyelvi egység is a névvel együtt annotálva lett.
-Ennek az az oka, hogy maga a Wikipédia címszó is tartalmazza ezt a szót, mely leszűkíti a név referenciáját, pl.~\textit{ókori Róma, magyar Wikipédia}.
-Emberekre utaló linkek esetében ezek nagyrészt rangjelzők, pl.~\textit{Szent István király, Anna brit királynő}.
-Ez utóbbiakat egy rangjelző szavakat tartalmazó tiltólistával a későbbiekben ki lehet szűrni.
+Típushiba kevés, ennél gyakoribb az álnegatív/álpozitív, még gyakoribb pedig a határok nem pontos felismerése. %TODO: részletesen, referálva a táblázatokra.
 
-Egy másik tipikus hibaforrás az automatikus tokenizálás és mondatra bontás.
-Sok esetben a rövidítést tartalmazó név (pl.~\textit{Warner Bros.}) utolsó eleme, vagyis a pont mondatvégi írásjelként értelmeződik, így nem annotálódik a névvel együtt.
-Máskor a mondatrabontás során a szövegközbeni link szétszakad, így a maradék elem nem kapja meg a megfelelő címkét.
-Erre esetleg megoldást jelenthet a tokenizáló teljesítményének javítása, vagy másik tokenizáló alkalmazása.
+Bizonyos esetekben a név környezetében lévő értelmező szerepű nyelvi egység is a
+névvel együtt annotálva lett. Ennek az az oka, hogy maga a Wikipédia címszó is
+tartalmazza ezt a szót, mely leszűkíti a név referenciáját,
+pl.~\textit{ókori Róma}, \textit{magyar Wikipédia}. Emberekre utaló linkek esetében
+ezek nagyrészt rangjelzők, pl.~\textit{Szent István király}, \textit{Anna brit királynő}.
+Ez utóbbiakat egy rangjelző szavakat tartalmazó tiltólistával a későbbiekben ki
+lehet szűrni.
 
-Tipikus jelenség a magyarban, amikor egy név összetétel eleme lesz, pl.~\textit{Bizánc-ellenes}.
-Ilyen esetekben a köznév az összetétel alaptagja, vagyis az határozza meg a referenciát.
-A referenciaváltozást természetesen a címkézés változásának kell követnie, ami viszont nem, vagy nehezen kezelhető automatikusan, mivel nagyjából bármilyen köznév kapcsolódhat névhez.
-A helyzetet bonyolítja az is, hogy a mozaikszavakhoz, rövidítésekhez, valamint nem ejtett magánhangzóra végződő idegen nevekhez is kötőjellel kapcsoljuk a toldalékokat, amelynek a felszíni szerkezete nagyon hasonlít az összetételekéhez.
-Ennek a problémának a megoldása még további vizsgálatokat követel.
+A hibák egy részét az alkalmazott nyelvfeldolgozó eszközök tévesztései okozzák.
+Az automatikus tokenizálás és mondatra bontás hibás működésére példa, amikor a
+rövidítést tartalmazó név (pl.~\textit{Warner Bros.}) utolsó eleme, vagyis a pont
+mondatvégi írásjelként értelmeződik, így nem annotálódik a névvel együtt. Máskor
+a mondatrabontás során a szövegközbeni link szétszakad, így a maradék elem nem
+kapja meg a megfelelő címkét. Mivel pedig a mondat első szavát csak akkor tekintjük
+potenciális entitásnak, ha főnév, a szófajmeghatározás hibája folytán előfordul,
+hogy átsiklunk egy mondatkezdő néven. Pl. a \textit{Hél visszaengedte volna.}
+mondatban a \textit{Hél} szót igeként ismerte fel a HunDisambig, így nem
+tekintettük tulajdonnév-jelöltnek. E problémákra esetleg megoldást jelenthet az
+alkalmazott eszközök teljesítményének javítása, vagy más eszközök alkalmazása.
 
-A felsorolt problémás esetek ellenére összességében azt mondhatjuk, hogy a magyar Wikipédia korpusz annotálási pontossága közelíti a gold standard korpuszokét. %na ide jöhetnek majd pontos adatok
+Tipikus jelenség a magyarban, amikor egy név összetétel eleme lesz,
+pl.~\textit{Bizánc-ellenes}. Ilyen esetekben a köznév az összetétel alaptagja,
+vagyis az határozza meg a referenciát. A referenciaváltozást természetesen a
+címkézés változásának kell követnie, ami viszont nem, vagy nehezen kezelhető
+automatikusan, mivel nagyjából bármilyen köznév kapcsolódhat névhez. A helyzetet
+bonyolítja az is, hogy a mozaikszavakhoz, rövidítésekhez, valamint nem ejtett
+magánhangzóra végződő idegen nevekhez is kötőjellel kapcsoljuk a toldalékokat,
+amelynek a felszíni szerkezete nagyon hasonlít az összetételekéhez. Ennek a
+problémának a megoldása még további vizsgálatokat követel.
+
+A felsorolt problémás esetek ellenére összességében azt mondhatjuk, hogy a magyar
+Wikipédia korpusz annotálási pontossága közelíti a gold standard korpuszokét. %na ide jöhetnek majd pontos adatok
+
+\begin{table}[ht]
+\begin{center}
+\begin{tabular}{l@{\hspace{1em}}c@{\hspace{2em}}c@{\hspace{2em}}c@{\hspace{2em}}c}
+\toprule
+& Precizitás(\%) &  Felidézés(\%) & F$_{\beta=1}$(\%) & Entitások száma \\
+\midrule
+LOC     &   98.72 &  95.65 &  97.16 & 161 \\
+MISC    &   95.24 &  76.92 &  85.11 & 26 \\
+ORG     &   89.66 &  89.66 &  89.66 & 29 \\
+PER     &   88.30 &  89.25 &  88.77 & 93 \\
+\midrule
+Összesítve &   94.33\% &  91.59\% &  92.94 & 309 \\
+\bottomrule
+\end{tabular}
+\end{center}
+\caption{Az inter-annotátor egyezés értékelése.}  % TODO: magyarul
+\label{tab:f_score}
+\end{table}
+
+\begin{table}[ht]
+\begin{center}
+	\begin{tabular}{l@{\hspace{0.5em}}|@{\hspace{0.5em}}cccc}
+\toprule
+Auto$\downarrow$ / Gold$\rightarrow$ & PER & ORG & LOC & MISC \\
+\midrule
+PER & 83 & 1 & & 2 \\
+ORG &  & 26 & 1 & 1 \\
+LOC &  & 1 & 154 & \\
+MISC &  &  & 1 & 20 \\
+\bottomrule
+\end{tabular}
+\end{center}
+\caption{A manuálisan annotált minta igazságmátrixa.}
+\label{tab:conf_matrix}
+\end{table}
+
+\begin{table}[ht]
+\begin{center}
+\begin{tabular}{l@{\hspace{0.5em}}cccc}
+\toprule
+	& PER & ORG & LOC & MISC \\
+\midrule
+Álpozitív& 83 & 1 & & 2 \\
+Álnegatív &  & 26 & 1 & 1 \\
+Részlegesen felismert nevek & & & & \\
+\bottomrule
+\end{tabular}
+\end{center}
+\caption{A hibásan tulajdonnévként felismert szavak (álpozitív), illetve a
+         fel nem ismert tulajdonnevek száma}  % TODO: these should be the row headers in the table
+\label{tab:fp_fn}
+\end{table}
 

### 2012-11-30 paper/MSZNY2013/cikk.tex
Tables filled with the correct data. 	modified:   cikk.tex

@@
 Ezután minden entitáshoz hozzárendeltük az őt tartalmazó
-legszűkebb osztály címkéjét, vagy \texttt{O}-t, ha az osztály nem minősül annotálandónak a Szeged
-NER korpusz sémája szerint, így egy 46.461 elemű, felcímkézett tulajdonnév-listát kaptunk.
+legszűkebb osztály címkéjét, vagy \texttt{O}-t, ha az osztály nem minősül
+annotálandónak a Szeged NER korpusz sémája szerint. Így egy 46.461 elemű,
+felcímkézett, angol nyelvű tulajdonnév-listát kaptunk. Utolsó lépésként
+meghatározzuk ezen oldalak magyar és egyéb nyelvű megfelelőit, és ezeket
+is felvesszük a listára. Az idegen nyelvű oldalak meghagyásának oka, hogy
+a magyar Wikipédia alkalmasint ezekre is hivatkozik.
 
@@
 felismert entitásokat, illetve egyéb nagybetűs említéseit, mint például a rájuk mutató
-átirányító és egyértelműsítő lapok címeit gyűjtjük. Ha ezután egy olyan mondattal  % TODO: + idegen nyelvi megfelelők!
+átirányító és egyértelműsítő lapok címeit gyűjtjük. Ha ezután egy olyan mondattal
 találkozunk, amelyben nagybetűs szócsoport szerepel, ellenőrizzük, hogy a csoport
@@
 
+\begin{table}[ht]
+\begin{center}
+\begin{tabular}{l@{\hspace{1em}}c@{\hspace{2em}}c@{\hspace{2em}}c@{\hspace{2em}}c}
+\toprule
+& Precizitás(\%) &  Felidézés(\%) & F$_{\beta=1}$(\%) & Entitások száma \\
+\midrule
+LOC     &   98.72 &  95.65 &  97.16 & 161 \\
+MISC    &   95.24 &  76.92 &  85.11 & 26 \\
+ORG     &   89.66 &  89.66 &  89.66 & 29 \\
+PER     &   88.30 &  89.25 &  88.77 & 93 \\
+\midrule
+Összesítve &   94.33\% &  91.59\% &  92.94 & 309 \\
+\bottomrule
+\end{tabular}
+\end{center}
+\caption{Az inter-annotátor egyezés értékelése.}  % TODO: magyarul annotátorok közötti, vagy van inter-?
+\label{tab:f_score}
+\end{table}
+
 A következő tipikus hibák rajzolódtak ki.
 
-Típushiba kevés, ennél gyakoribb az álnegatív/álpozitív, még gyakoribb pedig a határok nem pontos felismerése. %TODO: részletesen, referálva a táblázatokra.
+Típustévesztés kevés, ennél gyakoribb az álnegatív/álpozitív, illetve a határok nem pontos felismerése. %TODO: részletesen, referálva a táblázatokra.
 
+A típustévesztés alapvetően két okra vezethetők vissza. Az első esetben a DBpediában
+lévő típusinformáció helytelen. Pl.~ az \textit{MTA} a Szeged NER séma szerint
+\texttt{Org}, viszont a DBpediabeli osztálya, \texttt{WorldHeritageSite}, az
+épületre utal. Szintén ide tartozó probléma, amikor egy többértelmű névnek
+(pl.~\textit{South Park}, ami utalhat mind a rajzfilmsorozatra, mind az abban % TODO: ezek metonimok?
+szereplő városra) csak az egyik jelentése szerepel a DBpediában, így használattól
+függetlenül mindig ugyanúgy címkézzük.
+
+A típustévesztések másik részét a Wikipédiában lévő rossz kereszthivatkozások okozzák.
+Erre példa a \textit{Micimackó (regényszereplő)} című oldal, ahol szó esik a
+\textit{Walt Disney Co.} cégről. Azonban a kereszthivatkozás, amiből a típust
+megállapítjuk, csak a név első két szavát foglalja magába, és a cég alapítójának
+cikkére mutat\footnote{2012. november 30-ai állapot szerint.}, ezért ezt és a
+\textit{Disney} minden további említését tévesen \texttt{Per}-nek jelöltük. % TODO még Ikarus?
+
+A részlegesen felismert nevek (\ref{tab:fp_fn} táblázat) 
 Bizonyos esetekben a név környezetében lévő értelmező szerepű nyelvi egység is a
@@
 tekintettük tulajdonnév-jelöltnek. E problémákra esetleg megoldást jelenthet az
-alkalmazott eszközök teljesítményének javítása, vagy más eszközök alkalmazása.
+alkalmazott eszközök teljesítményének javítása, vagy más eszközök használata.
 
@@
 Wikipédia korpusz annotálási pontossága közelíti a gold standard korpuszokét. %na ide jöhetnek majd pontos adatok
-
-\begin{table}[ht]
-\begin{center}
-\begin{tabular}{l@{\hspace{1em}}c@{\hspace{2em}}c@{\hspace{2em}}c@{\hspace{2em}}c}
-\toprule
-& Precizitás(\%) &  Felidézés(\%) & F$_{\beta=1}$(\%) & Entitások száma \\
-\midrule
-LOC     &   98.72 &  95.65 &  97.16 & 161 \\
-MISC    &   95.24 &  76.92 &  85.11 & 26 \\
-ORG     &   89.66 &  89.66 &  89.66 & 29 \\
-PER     &   88.30 &  89.25 &  88.77 & 93 \\
-\midrule
-Összesítve &   94.33\% &  91.59\% &  92.94 & 309 \\
-\bottomrule
-\end{tabular}
-\end{center}
-\caption{Az inter-annotátor egyezés értékelése.}  % TODO: magyarul
-\label{tab:f_score}
-\end{table}
 
@@
 \midrule
-Álpozitív& 83 & 1 & & 2 \\
-Álnegatív &  & 26 & 1 & 1 \\
-Részlegesen felismert nevek & & & & \\
+Álpozitív & 1 &  0  &  1  &  0 \\
+Álnegatív & 3 &  0  &  5  &  4 \\
+Részlegesen felismert nevek & 7 & 1 & 0 & 0 \\
 \bottomrule

### 2012-11-30 paper/MSZNY2013/cikk.tex
Error analysis part done. 	modified:   cikk.tex

@@
 
-%TODO: itt kicsit kifejteni
-
 A korpuszépítő algoritmus a Wikipédia két feltételezett tulajdonságán alapszik, miszerint a szócikkek többsége megnevezett entitásokról szól, valamint a folyó szövegben előforduló entitásokat a kereszthivatkozások azonosítják.
@@
 
-% TODO argh
-Kiinduló alapfeltételezésünk, miszerint az egyes Wikipédia cikkek megnevezett entitásokról szólnak, és a szövegben előforduló entitásokat kereszthivatkozások azonosítják, nem minden esetben állja meg a helyét, ezért az algoritmusunk
+Kiinduló alapfeltételezésünk, miszerint az egyes Wikipédia cikkek megnevezett
+entitásokról szólnak, és a szövegben előforduló entitásokat kereszthivatkozások
+azonosítják, nem minden esetben állja meg a helyét. Algoritmusunk ezért
 többhelyütt is finomításra szorul.
@@
 
-A következő tipikus hibák rajzolódtak ki.
+A négy kategória igazságmátrixát a \ref{tab:conf_matrix}., a fel nem ismert
+tulajdonnevek, illetve hibásan tulajdonnévnek címkézett szavak számát pedig
+a \ref{tab:fp_fn}. táblázat mutatja. Az igazságtáblából látható, hogy a
+típustévesztés aránya elhanyagolható; ha az annotátorok közötti egyezést
+mérő Cohen kappát kiszámítjuk, 0.967 adódik ki, mely bőven a 0.8-as, tökéletes annotációs határ feletti. % TODO, ez kell? + hülye szósorrend
+Jóval gyakoribbak az entitásfelismerés első- és másodfajú hibái, illetve az
+entitáshatárok pontatlan felismerése. Az elemzés során a következő tipikus
+hibák rajzolódtak ki.
 
-Típustévesztés kevés, ennél gyakoribb az álnegatív/álpozitív, illetve a határok nem pontos felismerése. %TODO: részletesen, referálva a táblázatokra.
+\begin{table}[ht]
+\begin{center}
+	\begin{tabular}{l@{\hspace{0.5em}}|@{\hspace{0.5em}}cccc}
+\toprule
+Auto$\downarrow$ / Gold$\rightarrow$ & PER & ORG & LOC & MISC \\
+\midrule
+PER & 83 & 1 & & 2 \\
+ORG &  & 26 & 1 & 1 \\
+LOC &  & 1 & 154 & \\
+MISC &  &  & 1 & 20 \\
+\bottomrule
+\end{tabular}
+\end{center}
+\caption{A manuálisan annotált minta igazságmátrixa.}
+\label{tab:conf_matrix}
+\end{table}
 
@@
 
-A részlegesen felismert nevek (\ref{tab:fp_fn} táblázat) 
-Bizonyos esetekben a név környezetében lévő értelmező szerepű nyelvi egység is a
+\begin{table}[ht]
+\begin{center}
+\begin{tabular}{l@{\hspace{0.5em}}cccc}
+\toprule
+	& PER & ORG & LOC & MISC \\
+\midrule
+Hibásan tulajdonnévként felismert szavak (álpozitív) & 1 &  0  &  1  &  0 \\
+Fel nem ismert tulajdonnevek (álnegatív) & 3 &  0  &  5  &  4 \\
+Részlegesen felismert nevek & 7 & 1 & 0 & 0 \\
+\bottomrule
+\end{tabular}
+\end{center}
+\caption{Az entitásfelismerés első- és másodfajú hibái}
+\label{tab:fp_fn}
+\end{table}
+
+Az entitáshatárok pontatlan felismerésének (\ref{tab:fp_fn}. táblázat, 3. sor) leggyakoribb
+esete az, amikor a név környezetében lévő értelmező szerepű nyelvi egység is a
 névvel együtt annotálva lett. Ennek az az oka, hogy maga a Wikipédia címszó is
-tartalmazza ezt a szót, mely leszűkíti a név referenciáját,
-pl.~\textit{ókori Róma}, \textit{magyar Wikipédia}. Emberekre utaló linkek esetében
-ezek nagyrészt rangjelzők, pl.~\textit{Szent István király}, \textit{Anna brit királynő}.
-Ez utóbbiakat egy rangjelző szavakat tartalmazó tiltólistával a későbbiekben ki
+tartalmazza ezt a szót, mely leszűkíti a név referenciáját. Emberekre utaló linkek esetében
+ezek nagyrészt rangok, pl.~\textit{Szent István király}, \textit{I. Benedek pápa}.
+Ezeket egy rangjelző szavakat tartalmazó tiltólistával a későbbiekben ki
 lehet szűrni.
 
-A hibák egy részét az alkalmazott nyelvfeldolgozó eszközök tévesztései okozzák.
+A Wikipédia címekben szereplő értelmező szavak alkalmanként a teljes entitás
+felismerését is megakadályozzák. Ez akkor fordul elő, ha a hivatkozott oldal
+teljes címe nem tulajdonnév, viszont tartalmaz egy általunk ismert tulajdonnevet:
+pl.~\textit{ókori Róma}, \textit{magyar Wikipédia}. Az ilyen esetek helyes
+kezeléséhez módosítanunk kell az algoritmust, hogy a hivatkozásokat ne csak  % TODO kell? fogjuk?
+mint bonthatatlan entitásokat, hanem mint közönséges szöveget is kezelje.
+
+A hibák egy további részét az alkalmazott nyelvfeldolgozó eszközök tévesztései okozzák.
 Az automatikus tokenizálás és mondatra bontás hibás működésére példa, amikor a
@@
 A felsorolt problémás esetek ellenére összességében azt mondhatjuk, hogy a magyar
-Wikipédia korpusz annotálási pontossága közelíti a gold standard korpuszokét. %na ide jöhetnek majd pontos adatok
-
-\begin{table}[ht]
-\begin{center}
-	\begin{tabular}{l@{\hspace{0.5em}}|@{\hspace{0.5em}}cccc}
-\toprule
-Auto$\downarrow$ / Gold$\rightarrow$ & PER & ORG & LOC & MISC \\
-\midrule
-PER & 83 & 1 & & 2 \\
-ORG &  & 26 & 1 & 1 \\
-LOC &  & 1 & 154 & \\
-MISC &  &  & 1 & 20 \\
-\bottomrule
-\end{tabular}
-\end{center}
-\caption{A manuálisan annotált minta igazságmátrixa.}
-\label{tab:conf_matrix}
-\end{table}
-
-\begin{table}[ht]
-\begin{center}
-\begin{tabular}{l@{\hspace{0.5em}}cccc}
-\toprule
-	& PER & ORG & LOC & MISC \\
-\midrule
-Álpozitív & 1 &  0  &  1  &  0 \\
-Álnegatív & 3 &  0  &  5  &  4 \\
-Részlegesen felismert nevek & 7 & 1 & 0 & 0 \\
-\bottomrule
-\end{tabular}
-\end{center}
-\caption{A hibásan tulajdonnévként felismert szavak (álpozitív), illetve a
-         fel nem ismert tulajdonnevek száma}  % TODO: these should be the row headers in the table
-\label{tab:fp_fn}
-\end{table}
+Wikipédia korpusz annotálási pontossága közelíti a gold standard korpuszokét.
 

### 2012-11-30 paper/MSZNY2013/cikk.tex
Description of two addition error types added. Now it's REALLY done. 	modified:   cikk.tex

@@
 
+További javulást érhetünk el, ha algoritmusunkat felkészítjük a Wikipédia
+listaoldalainak kezelésére. A fel nem ismert MISC típusú tulajdonnevek mindegyike
+írók műveit felsoroló oldalakon fordul elő. Ezen műveknek nincs saját szócikkük,
+ezért címkézésük nehéz; a jövőben az ilyen oldalakat kihagyjuk a korpuszból. % TODO: így ok?
+
 A hibák egy további részét az alkalmazott nyelvfeldolgozó eszközök tévesztései okozzák.
@@
 problémának a megoldása még további vizsgálatokat követel.
+
+A fenti, elvi hibák mellett felfedeztünk implementációs hibákat is, amelyek
+a \ref{tab:fp_fn}. táblázat kb. 5-6 eleméért felelősek. Ezeket a korpusz  % TODO na ez de hülyén hangzik
+második verziójára javítani fogjuk.
 

### 2012-11-30 paper/MSZNY2013/cikk.tex
reszemrol kesz

--- 2012-11-30 paper/MSZNY2013/cikk.tex
@@
 
-% Ezek egyikÈnek bekapcsol·s·val a magyar Èkezetes karakterek
-% kˆzvetlen¸l is felhaszn·lhatÛk:
 %\usepackage[latin2]{inputenc} 
@@
 
-\title{\ \break Automatikusan tulajdonnév-annotált korpuszok előállítása a Wikipédiából}
+\title{\ \break Automatikus korpuszépítés tulajdonnév-felismerés céljára}
 %XXX a címet majd meg kéne változtatni, tényleg elég bénán hangzik magyarul
-%Automatikus korpuszépítés tulajdonnév-felismerés céljára
+%Automatikusan tulajdonnév-annotált korpuszok előállítása a Wikipédiából
 \author{Nemeskey Dávid Márk\inst{1}, Simon Eszter\inst{2}}
@@
 Több lehetőség van az annotációs költségek csökkentésére, ezek közül az egyik az automatikus annotálás.
-Cikkünkben egy nyelvfüggetlen módszert és azzal előállított magyar és angol nyelvű korpuszokat mutatunk be.
-A korpuszok szövege a Wikipédiából származik, és tulajdonnévi annotálást is tartalmaz.
-Az automatikus annotáláshoz egy új módszert alkalmaztunk: a DBpedia ontológiai kategóriáit képeztük le CoNLL névosztályokra.
-A korpuszok szabadon felhasználhatóak további fejlesztésekhez.
+Cikkünkben egy nyelvfüggetlen módszert mutatunk be, mellyel bármely Wikipédiával rendelkező nyelvre előállítható tulajdonnévi címkéket tartalmazó korpusz.
+Az automatikus annotálás során a DBpedia ontológiai kategóriáit képeztük le CoNLL-névosztályokra.
+Cikkünkben a magyar korpusz részletes hibaelemzését és kiértékelését adjuk.
 \\[2mm]
@@
 A feladat során strukturálatlan szövegben kell azonosítani és az előre definiált osztályok valamelyikébe besorolni a neveket.
-A tulajdonnév-felismerés feladata a 6.
-Message Understanding Conference (MUC) egyik versenykiírásában jelent meg először 1995-ben \cite{MUC6}.
-Itt három alfeladatot különítettek el: tulajdonneveket (ENAMEX), temporális (TIMEX) és különböző numerikus (NUMEX) kifejezéseket kellett felismerni.
+A tulajdonnév-felismerés feladata a 6.~Message Understanding Conference (MUC) egyik versenykiírásában jelent meg először 1995-ben \cite{MUC6}.
+Itt három alfeladatot különítettek el: tulajdonneveket, temporális és különböző numerikus kifejezéseket kellett felismerni.
 A NER-közösségen belül a temporális és a numerikus kifejezések annotálása is elfogadott, de a leginkább vizsgált típusok a személy-, földrajzi és intézménynevek.
-Ezek mellé vezettek be a CoNLL versenyeken \cite{tksintro2002conll} \cite{conll2003intro} egy negyedik típust, amely az előző háromba nem tartozó egyéb tulajdonneveket foglalja magában.
+Ezek mellé vezettek be a CoNLL-versenyeken \cite{tksintro2002conll} \cite{conll2003intro} egy negyedik típust, amely az előző háromba nem tartozó egyéb tulajdonneveket foglalja magában.
 Az azóta eltelt időben ezek az annotációs sémák váltak nemzetközileg elfogadottá.
 
-A versenyekre épített és aztán közzétett tulajdonnév-annotált korpuszok képezik azokat a sztenderdeket, amelyek összemérhetővé teszik az egyes tulajdonnév-felismerő rendszereket.
+A versenyekre épített és aztán közzétett tulajdonnév-annotált korpuszok képezik azokat a sztenderdeket, amelyek összemérhetővé teszik az egyes névfelismerő rendszereket.
 Ezek a korpuszok meglehetősen korlátozott méretűek és témaspecifikusak.
@@
 Ennek egy módja, ha már rendelkezésre álló korpuszokat dolgozunk össze; ekkor a különböző annotációs sémák és címkekészletek összeillesztése állít elénk problémákat.
-Egy másik lehetőség az olyan webes közösségi tartalmak felhasználása, mint például a Wikipédia, a Wiktionary vagy a DBpedia.
-Megint másik megközelítés az annotáció automatizálása.
+Egy másik lehetőség az olyan webes közösségi tartalmak felhasználása korpuszépítéshez, mint például a Wikipédia, a Wiktionary vagy a DBpedia.
+Megint másik megközelítés az annotáció automatizálása, ami az esetek nagy részében egy már rendelkezésre álló adathalmazon tanított rendszer új szövegen való futtatását jelenti.
 
@@
 Munkánk során új módszert alkalmaztunk: a DBpedia ontológiai kategóriáit képeztük le CoNLL-névosztályokra.
-A létrehozott magyar és angol nyelvű korpuszok szabadon felhasználhatóak.
+A módszert egyelőre a magyar és az angol Wikipédiára alkalmaztuk.
 
@@
 A Wikipédia egy többnyelvű, nyílt tartalmú, az internetes közösség által fejlesztett webes világenciklopédia \footnote{http://wikipedia.org}.
-Több mint 22 millió szócikkével aranybánya a különböző természetesnyelv-feldolgozó fejlesztések számára; használták már többek között jelentésegyértelműsítére, ontológia- és tezauruszépítésre, valamint kérdésmegválaszoló rendszerekhez (további alkalmazási lehetőségekhez lásd \cite{Medelyan:09}).
-Mivel a Wikipédia címszavak jelentős része tulajdonnév, adja magát a lehetőség, hogy a tulajdonnév-felismeréshez is használjuk.
+Több mint 22 millió szócikkével kincsesbánya a különböző természetesnyelv-feldolgozó fejlesztések számára; használták már többek között jelentésegyértelműsítére, ontológia- és tezauruszépítésre, valamint kérdésmegválaszoló rendszerekhez (további alkalmazási lehetőségekért lásd \cite{Medelyan:09}).
+Mivel a Wikipédia címszavainak jelentős része tulajdonnév, adja magát a lehetőség, hogy a tulajdonnév-felismeréshez is használjuk.
 
@@
 
-A Wikipédia alkalmazására a tulajdonnév-felismerés területén egy másik lehetőség magának a szövegbázisnak a felhasználása.
-Richman és Schone \cite{Richman:08} kevés erőforrással rendelkező nyelvek Wikipédia szócikkeiből építettek korpuszokat, amelyekben a Wikipédia inherens kategóriastruktúráját használták fel a tulajdonnevek annotálásához.
-Nothman et al.
-\cite{Nothman:08} a szócikkek definíciójának első mondatából kiindulva címkézte fel a szövegben belinkelt neveket, így építve automatikusan tulajdonnév-annotált korpuszt.
+A Wikipédia alkalmazására a névfelismerés területén egy másik lehetőség magának a szövegbázisnak a felhasználása.
+Richman és Schone \cite{Richman:08} kevés erőforrással rendelkező nyelvek Wikipédia-cikkeiből épített korpuszokat, amelyekben a Wikipédia inherens kategóriastruktúráját használták fel a tulajdonnevek annotálásához.
+Nothman et al.~\cite{Nothman:08} a szócikkek első mondatából kiindulva címkézte fel a szövegben belinkelt neveket, így építve automatikusan tulajdonnév-annotált korpuszt.
 
-Az általunk alkalmazott módszer az említettektől annyiban tér el, hogy mi a DBpedia ontológiai osztályait képeztük le a sztenderd CoNLL névosztályokra, majd ezeket Wikipédia entitásokhoz kötöttük.
-Az így létrehozott korpuszokat szabadon elérhetővé tettük.
+Az általunk alkalmazott módszer az említettektől annyiban tér el, hogy mi a DBpedia ontológiai osztályait képeztük le a sztenderd CoNLL-névosztályokra, majd ezeket Wikipédia-entitásokhoz kötöttük. 
 %Tudomásunk szerint jelenleg csak egy angol nyelvű automatikusan tulajdonnév-annotált korpusz létezik, amely szabadon felhasználható, a Semantically Annotated Snapshot of the English Wikipedia \cite{Atserias:08}.
@@
 A Wikipédiából, tekintve a szócikkek nagy számát, kevés erőforrással rendelkező nyelvekre is tudunk kellően nagy méretű korpuszokat építeni, amelyek bemenetül szolgálhatnak névfelismerő rendszerek tanításához és teszteléséhez.
-Az általunk létrehozott korpusz az első magyar nyelvű automatikusan tulajdonnév-annotált korpusz.
+Tudomásunk szerint az általunk létrehozott korpusz az első magyar nyelvű automatikusan tulajdonnév-annotált korpusz, amely szabadon elérhető és felhasználható további kutatásokhoz.
 
@@
 
-%TODO: itt kicsit kifejteni
-
 A korpuszépítő algoritmus a Wikipédia két feltételezett tulajdonságán alapszik, miszerint a szócikkek többsége megnevezett entitásokról szól, valamint a folyó szövegben előforduló entitásokat a kereszthivatkozások azonosítják.
 Algoritmusunk e két feltételezés következményeit használja ki.
-Hasonlóan a Nothman et al.~\cite{Nothman:08} által leírtakhoz, a következő lépésekből áll:
+Hasonlóan a Nothman et al.~\cite{Nothman:08} által leírtakhoz, a korpuszépítés a következő lépésekből áll:
 
 \begin{enumerate}
-\item a Wikipédia cikkeket entitásosztályokba soroljuk;
+\item a Wikipédia-cikkeket entitásosztályokba soroljuk;
 \item a cikkeket mondatokra bontjuk;
@@
 
-Az algoritmus alapjaiban nyelvfüggetlen: bármelyik nyelvre alkalmazható, amely
+Az algoritmus alapjaiban nyelvfüggetlen: bármely nyelvre alkalmazható, amely
 rendelkezik megfelelő méretű Wikipédiával. Egyedül a harmadik lépés az, ahol
@@
 mit tekintenek annotálandó elemnek: pl.~a \textit{római} szó a magyarban nem
-számít névnek, míg angol megfelelője, a \textit{Roman} a CoNLL annotációs séma
+számít névnek, míg angol megfelelője, a \textit{Roman} a CoNLL-séma
 szerint \texttt{Misc} címkét kapna.
 
-Mivel célunk egy silver standard korpusz előállítása volt, ha választanunk
+Mivel célunk egy minél tisztább, vagyis a gold standard színvonalat közelítő korpusz előállítása volt, ha választanunk
 kellett egy-egy lépésnél a pontosság (precision) vagy a teljesség (recall)
@@
 
-\subsection{Wikipédia cikkek mint entitások}
+\subsection{Wikipédia-cikkek mint entitások}
 \label{corpusbuilding:entities}
@@
 Ahhoz, hogy a folyó szövegben lévő kereszthivatkozásokat felhasználhassuk a
-tulajdonnevek azonosítására, a hivatkozott Wikipédia cikkeket névosztályokba kell
+tulajdonnevek azonosítására, a hivatkozott Wikipédia-cikkeket névosztályokba kell
 sorolnunk. A Wikipédia saját kategóriarendszere a kategóriák nagy
@@
 használtuk.} a típusok egy OWL\footnote{http://www.w3.org/TR/owl-ref/}
-ontológia részei. A tudásbázis a Wikipédia entitások egy részhalmazát
+ontológia részei. A tudásbázis a Wiki\-pé\-dia-entitások egy részhalmazát
 tartalmazza, és mindegyik entitáshoz hozzárendeli -- többet között -- azon
@@
 is felvesszük a listára. Az idegen nyelvű oldalak meghagyásának oka, hogy
-a magyar Wikipédia alkalmasint ezekre is hivatkozik.
+a magyar Wikipédia időnként ezekre is hivatkozik.
 
@@
 A korpusz építéséhez a magyar Wikipédia 2012. március 9-i állapotát vettük
-alapul. Az XML fájlokból az mwlib\footnote{http://code.pediapress.com}
+alapul. Az XML-fájlokból az mwlib\footnote{http://code.pediapress.com}
 könyvtár segítségével nyertük ki a nyers szöveget. 
@@
 
-% TODO argh
-Kiinduló alapfeltételezésünk, miszerint az egyes Wikipédia cikkek megnevezett entitásokról szólnak, és a szövegben előforduló entitásokat kereszthivatkozások azonosítják, nem minden esetben állja meg a helyét, ezért az algoritmusunk
-többhelyütt is finomításra szorul.
+Kiinduló alapfeltételezésünk, miszerint az egyes Wikipédia-cikkek megnevezett
+entitásokról szólnak, és a szövegben előforduló entitásokat kereszthivatkozások
+azonosítják, nem minden esetben állja meg a helyét. Algoritmusunk ezért
+több helyen finomításra szorul.
 
 Először, a Wikipédia nem minden szócikke szól megnevezett entitásokról: köznevek, dátumok és egyéb, nem tulajdonnévi elemek is
-alkalmasint saját oldalt kapnak. Annak érdekében, hogy az algoritmus ne
+saját oldalt kaptak. Annak érdekében, hogy az algoritmus ne
 kezelje ezeket \texttt{Unk} típusú entitásként, majd dobja el a mondatokat,
@@
 Másodszor, nem minden tulajdonnéven találunk
-kereszthivatkozást. Ennek oka kettős: ha egy Wikipédia szócikkben adott entitásra
+kereszthivatkozást. Ennek oka kettős: ha egy Wikipédia-cikkben adott entitásra
 többször is névvel utalnak, csak az első alkalommal linkelik a saját oldalához. De
@@
 
-A hagyományos névfelismerő rendszerek tiszta, viszonylag pontosan annotált korpuszokból tanulják ki a megfelelő paramétereiket, és a tesztelésükhöz is hasonló adathalmazra van szükség.
+A hagyományos névfelismerő rendszerek tiszta, viszonylag pontosan annotált korpuszokból tanulják meg a megfelelő paramétereiket, és a tesztelésükhöz is hasonló adathalmazra van szükség.
 Ezért a rossz minőségű, töredékes mondatokat, amelyek nem nagybetűvel kezdődnek és nincs mondatzáró írásjel a végükön, szintén kiszűrtük.
 (Az így maradt 19 milliós adathalmazzal dolgoztunk tovább; a következőkben erre referálunk teljes korpuszként.) De a minőség javítása érdekében további szűrő lépéseket is lehet tenni, így például eldobhatjuk az olyan mondatokat is, amelyek nem tartalmaznak ragozott igét.
-Ezzel ugyan azok is kiesnek, amelyek szabályos jelen idejű, kijelentő módú, létigét (nem) tartalmazó mondatok, de az automatikus tokenizálás és mondatra bontás hibáiból származó töredékes mondatokat eltávolíthatjuk.
+Ezzel ugyan azok is kiesnek, amelyek szabályos jelen idejű, kijelentő módú, létigét (nem) tartalmazó mondatok, de az automatikus tokenizálás és mondatra bontás hibáiból származó töredékes mondatokat eltávolíthatjuk, és így az ebből fakadó annotációs hibákat is kiszűrhetjük (lásd a \ref{probl}.~fejezetet).
 
 Ha viszont közösségi tartalmak (User Generated Content) feldolgozására kívánja valaki használni a korpuszokat, amelyek köztudottan sokkal zajosabbak, mint a kézzel annotált adatok, és sok töredékes mondatot tartalmaznak, hasznos lehet minél kevesebb szűrő használata.
-Ezért a rossznak minősített mondatokat nem dobtuk el végleg, hanem a korpusz mellett ezt az adathalmazt is elérhetővé tettük.
+Ezért a rossznak minősített mondatokat nem dobtuk el végleg, hanem ezt az adathalmazt is elérhetővé tettük.
 
@@
 részén. A 19 milliós teljes korpuszt vettük alapul, amelyből véletlenszerű
-mondatválogatással csináltunk egy kb.~19.000 tokent tartalmazó kisebb korpuszt.
+mondatválogatással csináltunk egy 18.830 tokent tartalmazó mintakorpuszt.
 Ezt kézzel felannotáltuk, és összehasonlítottuk a kézi és a gépi annotálás
-eredményét. Az eredményeket a \ref{tab:f_score} táblázat mutatja. Az összesített
-F-mérték 92.94 százaléknak adódott, ami % TODO ami mi?
+eredményét, amely az \ref{tab:f_score}.~táblázatban látható. Ha a gépi módszert egy annotátornak tekintjük, akkor az F-mérték az annotátorok közötti egyetértést mutatja.
 
@@
 \toprule
-& Precizitás(\%) &  Felidézés(\%) & F$_{\beta=1}$(\%) & Entitások száma \\
+& Pontosság(\%) &  Fedés(\%) & F$_{\beta=1}$(\%) & Entitások száma \\
 \midrule
@@
 \end{center}
-\caption{Az inter-annotátor egyezés értékelése.}  % TODO: magyarul annotátorok közötti, vagy van inter-?
+\caption{A gépi és a kézi annotálás közötti egyetértést mutató eredmények a mintakorpuszon.}  
 \label{tab:f_score}
@@
 
-A következő tipikus hibák rajzolódtak ki.
-
-Típustévesztés kevés, ennél gyakoribb az álnegatív/álpozitív, illetve a határok nem pontos felismerése. %TODO: részletesen, referálva a táblázatokra.
-
-A típustévesztés alapvetően két okra vezethetők vissza. Az első esetben a DBpediában
-lévő típusinformáció helytelen. Pl.~ az \textit{MTA} a Szeged NER séma szerint
-\texttt{Org}, viszont a DBpediabeli osztálya, \texttt{WorldHeritageSite}, az
-épületre utal. Szintén ide tartozó probléma, amikor egy többértelmű névnek
-(pl.~\textit{South Park}, ami utalhat mind a rajzfilmsorozatra, mind az abban % TODO: ezek metonimok?
-szereplő városra) csak az egyik jelentése szerepel a DBpediában, így használattól
-függetlenül mindig ugyanúgy címkézzük.
-
-A típustévesztések másik részét a Wikipédiában lévő rossz kereszthivatkozások okozzák.
-Erre példa a \textit{Micimackó (regényszereplő)} című oldal, ahol szó esik a
-\textit{Walt Disney Co.} cégről. Azonban a kereszthivatkozás, amiből a típust
-megállapítjuk, csak a név első két szavát foglalja magába, és a cég alapítójának
-cikkére mutat\footnote{2012. november 30-ai állapot szerint.}, ezért ezt és a
-\textit{Disney} minden további említését tévesen \texttt{Per}-nek jelöltük. % TODO még Ikarus?
-
-A részlegesen felismert nevek (\ref{tab:fp_fn} táblázat) 
-Bizonyos esetekben a név környezetében lévő értelmező szerepű nyelvi egység is a
-névvel együtt annotálva lett. Ennek az az oka, hogy maga a Wikipédia címszó is
-tartalmazza ezt a szót, mely leszűkíti a név referenciáját,
-pl.~\textit{ókori Róma}, \textit{magyar Wikipédia}. Emberekre utaló linkek esetében
-ezek nagyrészt rangjelzők, pl.~\textit{Szent István király}, \textit{Anna brit királynő}.
-Ez utóbbiakat egy rangjelző szavakat tartalmazó tiltólistával a későbbiekben ki
-lehet szűrni.
-
-A hibák egy részét az alkalmazott nyelvfeldolgozó eszközök tévesztései okozzák.
-Az automatikus tokenizálás és mondatra bontás hibás működésére példa, amikor a
-rövidítést tartalmazó név (pl.~\textit{Warner Bros.}) utolsó eleme, vagyis a pont
-mondatvégi írásjelként értelmeződik, így nem annotálódik a névvel együtt. Máskor
-a mondatrabontás során a szövegközbeni link szétszakad, így a maradék elem nem
-kapja meg a megfelelő címkét. Mivel pedig a mondat első szavát csak akkor tekintjük
-potenciális entitásnak, ha főnév, a szófajmeghatározás hibája folytán előfordul,
-hogy átsiklunk egy mondatkezdő néven. Pl. a \textit{Hél visszaengedte volna.}
-mondatban a \textit{Hél} szót igeként ismerte fel a HunDisambig, így nem
-tekintettük tulajdonnév-jelöltnek. E problémákra esetleg megoldást jelenthet az
-alkalmazott eszközök teljesítményének javítása, vagy más eszközök használata.
-
-Tipikus jelenség a magyarban, amikor egy név összetétel eleme lesz,
-pl.~\textit{Bizánc-ellenes}. Ilyen esetekben a köznév az összetétel alaptagja,
-vagyis az határozza meg a referenciát. A referenciaváltozást természetesen a
-címkézés változásának kell követnie, ami viszont nem, vagy nehezen kezelhető
-automatikusan, mivel nagyjából bármilyen köznév kapcsolódhat névhez. A helyzetet
-bonyolítja az is, hogy a mozaikszavakhoz, rövidítésekhez, valamint nem ejtett
-magánhangzóra végződő idegen nevekhez is kötőjellel kapcsoljuk a toldalékokat,
-amelynek a felszíni szerkezete nagyon hasonlít az összetételekéhez. Ennek a
-problémának a megoldása még további vizsgálatokat követel.
-
-A felsorolt problémás esetek ellenére összességében azt mondhatjuk, hogy a magyar
-Wikipédia korpusz annotálási pontossága közelíti a gold standard korpuszokét. %na ide jöhetnek majd pontos adatok
+A négy kategória igazságmátrixa (\ref{tab:conf_matrix}.~táblázat) jól mutatja, hogy a
+típustévesztés aránya elhanyagolható. Ezekből az értékekből kiszámítottuk az annotátorok közötti egyetértést
+mérő Cohen kappát is. A 0,967-es összesített érték Landis és Koch \cite{Landis} skáláján elhelyezve arról tanúskodik, hogy a korpusz annotációja megfelel a gold standard színvonalnak.
 
@@
 \end{center}
-\caption{A manuálisan annotált minta igazságmátrixa.}
+\caption{A manuálisan annotált mintakorpusz igazságmátrixa.}
 \label{tab:conf_matrix}
 \end{table}
+
+A típustévesztés alapvetően két okra vezethető vissza. Az első esetben a DBpediában
+lévő típusinformáció helytelen, például az \textit{MTA} DBpedia-beli osztálya \texttt{WorldHeritageSite}, ami miatt \texttt{Loc} címkét kap \texttt{Org} helyett. 
+Hasonló eset, amikor egy több referenciával rendelkező névnek csak egyik referense szerepel a DBpediában, így használattól függetlenül mindig ugyanazt a címkét kapja.
+
+A típustévesztések másik részét a Wikipédia rossz kereszthivatkozásai okozzák.
+Például az egyik cikk szerkesztője a \textit{Walt Disney Co.} cégnévnek csak egy részét linkelte be, mégpedig a személyről szóló oldalra.
+Ezért ebben a cikkben ennek a cégnévnek a különböző változatai (\textit{Disney}, \textit{Walt Disney} stb.) is mind személynévként lettek jelölve.
+%Erre példa a \textit{Walt Disney Co.} cég, amelynek csak a személyre vonatkozó része 
+%\textit{Micimackó (regényszereplő)} című oldal, ahol szó esik a
+%\textit{Walt Disney Co.} cégről. Azonban a kereszthivatkozás, amiből a típust
+%megállapítjuk, csak a név első két szavát foglalja magába, és a cég alapítójának
+%cikkére mutat\footnote{2012. november 30-ai állapot szerint.}, ezért ezt és a
+%\textit{Disney} minden további említését tévesen \texttt{Per}-nek jelöltük. % TODO még Ikarus?
 
@@
 \midrule
-Álpozitív & 1 &  0  &  1  &  0 \\
-Álnegatív & 3 &  0  &  5  &  4 \\
+Hibásan névként felismert szavak (álpozitív) & 1 &  0  &  1  &  0 \\
+Fel nem ismert nevek (álnegatív) & 3 &  0  &  5  &  4 \\
 Részlegesen felismert nevek & 7 & 1 & 0 & 0 \\
@@
 \end{center}
-\caption{A hibásan tulajdonnévként felismert szavak (álpozitív), illetve a
-         fel nem ismert tulajdonnevek száma}  % TODO: these should be the row headers in the table
+\caption{A névfelismerés további hibái.}
 \label{tab:fp_fn}
 \end{table}
+
+Jóval gyakoribbak a névfelismerés további hibái, vagyis bizonyos szavak hibásan névként való azonosítása és egyes nevek felismerésének elmulasztása, 
+illetve az
+entitáshatárok pontatlan felismerése. Az egyes névtípusokra lebontott hibák számát a \ref{tab:fp_fn}.~táblázat mutatja.
+
+Az entitáshatárok pontatlan felismerésének oka leggyakrabban
+az, hogy a név környezetében lévő értelmező szerepű nyelvi egység is a Wikipédia-címszó része. Emberekre utaló linkek esetében
+ezek a szavak nagyrészt rangjelölők, pl.~\textit{Szent István király}, \textit{I. Benedek pápa}.
+Ezeket egy tematikus tiltólistával a későbbiekben ki
+lehet szűrni.
+
+A Wikipédia-címszavakban szereplő értelmező szavak alkalmanként a teljes entitás
+felismerését is megakadályozzák. Ez akkor fordul elő, ha a hivatkozott oldal
+teljes címe nem tulajdonnév, viszont tartalmaz egy általunk ismert tulajdonnevet:
+pl.~\textit{ókori Róma}, \textit{magyar Wikipédia}. 
+%Az ilyen esetek helyes
+%kezeléséhez módosítanunk kell az algoritmust, hogy a hivatkozásokat ne csak  % TODO kell? fogjuk? nem
+%mint bonthatatlan entitásokat, hanem mint közönséges szöveget is kezelje.
+
+A fel nem ismert \texttt{Misc} típusú tulajdonnevek mindegyike
+írók műveit felsoroló oldalakon fordul elő. Ezen műveknek nincs saját szócikkük,
+ezért címkézésük nehéz.
+Mivel egy műcímben bármilyen nyelvi elem előfordulhat, az általunk alkalmazott szűrők együttese sem képes kiszűrni ezeket.
+Megoldást jelenthetne az, ha ezeket a kizárólag műcímeket felsoroló oldalakat külön kezelnénk, és egy komplex rendszert építenénk a feldolgozásukra.
+Mivel ez egy külön nyelvfeldolgozási feladat, jelen fejlesztésen belül nem vállalkozunk rá.
+%Ezért a jövőben az ilyen oldalakat kihagyjuk a korpuszból.
+% TODO: így ok?
+
+A hibák egy további részét az alkalmazott nyelvfeldolgozó eszközök tévesztései okozzák.
+Az automatikus tokenizálás és mondatra bontás hibás működésére példa, amikor a
+rövidítést tartalmazó név (pl.~\textit{Warner Bros.}) utolsó eleme, vagyis a pont
+mondatvégi írásjelként értelmeződik, így nem annotálódik a névvel együtt. Máskor
+a mondatrabontás során a szövegben levő link szétszakad, így a maradék elem nem
+kapja meg a megfelelő címkét. Mivel a mondat első szavát csak akkor tekintjük
+potenciális entitásnak, ha főnév, a szófajmeghatározás hibája folytán előfordul,
+hogy átsiklunk egy mondatkezdő néven. Például a \textit{Hél visszaengedte volna}
+mondatban a \textit{Hél} szót igeként azonosította a HunDisambig, így nem
+tekintettük tulajdonnév-jelöltnek. E problémákra esetleg megoldást jelenthet az
+alkalmazott eszközök teljesítményének javítása, vagy más eszközök használata.
+
+Tipikus jelenség a magyarban, amikor egy név összetétel eleme lesz,
+pl.~\textit{Bizánc-ellenes}. Ilyen esetekben a köznév az összetétel alaptagja,
+vagyis az határozza meg a referenciát. A referenciaváltozást természetesen a
+címkézés változásának kell követnie, ami viszont nem, vagy nehezen kezelhető
+automatikusan, mivel nagyjából bármilyen köznév kapcsolódhat névhez. A helyzetet
+bonyolítja az is, hogy a mozaikszavakhoz, rövidítésekhez, valamint nem ejtett
+magánhangzóra végződő idegen nevekhez is kötőjellel kapcsoljuk a toldalékokat,
+amelynek a felszíni szerkezete nagyon hasonlít az összetételekéhez. Ennek a
+problémának a megoldása még további vizsgálatokat követel.
+
+A felsoroltak mellett természetesen implementációs hibákra is fény derült, amelyek azonban összességében 
+csak néhány esetben okoztak rossz címkézést. Ezeket a jövőben javítani fogjuk. 
+%a \ref{tab:fp_fn}.~táblázat 5-6 eleméért felelősek. Ezeket a korpusz  % TODO na ez de hülyén hangzik 
+%második verziójára javítani fogjuk.
+
+%A felsorolt problémás esetek ellenére összességében azt mondhatjuk, hogy a magyar
+%Wikipédia korpusz annotálási pontossága közelíti a gold standard korpuszokét.
 
@@
 
-A korpuszokat Creative Commons Attribution-Sharealike 3.0 licensz alatt publikáltuk, vagyis ugyanolyan feltételekkel adjuk tovább, ahogy a Wikipédiából letöltöttük.
+A korpuszokat Creative Commons Attribution-Sharealike 3.0 licensz alatt publikáljuk, vagyis ugyanolyan feltételekkel adjuk tovább, ahogy a Wikipédiából letöltöttük.
 Szabadon elérhetőek a \texttt{http://hlt.sztaki.hu} oldalon keresztül, valamint a META-SHARE tárhelyről (http://www.meta-net.eu/).
-(A META-SHARE célja egy nyílt rendszer létrehozása, amely lehetővé teszi a nyelvi erőforrások megosztását; létrehozója a META-NET, az Európai Bizottság által alapított kiválósági hálózat.)
+A META-SHARE egy nyílt rendszer, amely lehetővé teszi a nyelvi erőforrások megosztását.
+Létrehozója a META-NET, az Európai Bizottság által alapított nyelvtechnológiai hálózat.
 
-A fájlok ún.~\emph{multitag} formátumban vannak, amelyben a tartalmas sorok tabulátorral vannak elválasztva.
+A fájlok ún.~\emph{multitag} formátumban vannak, amelyben a tartalmas sorokat tabulátor választja el.
 Az első oszlop tartalmazza magukat a szövegszavakat, az egyes oszlopokban pedig a különböző szintű annotációk találhatók.
@@
 A névcímkéken kívül minden token mellett szerepel a töve és a hozzá tartozó teljes morfológiai elemzése KR-kódokkal.
-Két további oszlopban közöljük, hogy sima szöveg vagy Wikipédia link-e az adott token, és ha utóbbi, akkor melyik szócikkre utal.
+Két további oszlopban közöljük, hogy sima szöveg vagy kereszthivatkozás-e az adott token, és ha utóbbi, akkor melyik szócikkre utal.
 
@@
 
-Az automatikusan generált korpuszok nyilvánvaló előnyei mellett azonban megvan az a hátrányuk, hogy nem használhatók gold standard adatként.
-Kiértékelésünkben megmutatjuk, hogy a létrehozott silver standard adathalmaz viszont kiválóan használható a tulajdonnév-felismerés teljesítményének növelésére több módon is.
+Kiértékelésünkben megmutatjuk, hogy a létrehozott magyar nyelvű korpusz kiválóan használható a tulajdonnév-felismerés teljesítményének növelésére több módon is.
 
 A kiértékeléshez a Hunner \cite{Varga} tulajdonnév-felismerő rendszert használtuk.
-A csak az egyes korpuszokra jellemző jegyeket (pl.~főnévi csoportok jelölése, Wikipédia linkek) kidobtuk, hogy növeljük a korpuszok összehasonlíthatóságát.
+A csak az egyes korpuszokra jellemző jegyeket (pl.~főnévi csoportok jelölése, Wikipédia-linkek) kidobtuk, hogy növeljük a korpuszok összehasonlíthatóságát.
 Így a következő jegykészlettel dolgoztunk: mondatkezdő és -vég pozíciók, szóalakon alapuló jegyek, morfológiai információ és listajegyek.
 
-A kiértékeléshez a sztenderd CoNLL módszert alkalmaztuk, vagyis az annotációt csak akkor vettük helyesnek, ha a kezdő- és végpozíció is stimmelt, és a rendszer által kibocsátott címke megegyezett a gold standard címkével.
+Az eredmények kiszámításához a sztenderd CoNLL-módszert alkalmaztuk, vagyis az annotációt csak akkor vettük helyesnek, ha a kezdő- és végpozíció is stimmelt, és a rendszer által kibocsátott címke megegyezett a gold standard címkével.
 Ezen alapulva a szokásos pontosságot, fedést és F-mértéket számoltuk.
@@
 Viszont amikor megvizsgáltuk az egyes korpuszok telítettségét a nevek szempontjából, arra jutottunk, hogy a gold standard adathalmazzal való összevetéskor inkább sűrítjük a szöveget, vagyis kivesszük azokat a mondatokat, amelyekben nincs név.
-A \ref{tab:size_hu}.~táblázat mutatja a magyar korpuszokra vonatkozó számszerű adatokat, melyekből jól látható, hogy a Wikipédiából generált korpusz telítettsége meglehetősen alacsony, a nevet nem tartalmazó mondatok kiszűrése után viszont hasonlóak lettek a számok.
-Mivel a Wikipédia az entitások kincsesbányájaként van számon tartva, ez az adat kissé meglepő lehet.
+A \ref{tab:size_hu}.~táblázat mutatja a magyar korpuszokra vonatkozó számszerű adatokat, melyekből jól látható, hogy a Wikipédiából generált korpusz telítettsége meglehetősen alacsony.
+%A nevet nem tartalmazó mondatok kiszűrése után viszont hasonlóak lettek a számok. Mivel a Wikipédia az entitások kincsesbányájaként van számon tartva, ez az adat kissé meglepő lehet. 
 A szövegnek ez a hígsága valószínűleg annak köszönhető, hogy a módszerünk szigorú, vagyis inkább minden olyan mondatot eltávolítottunk, amelyben nem lehetett beazonosítani a nevet, minthogy rosszul annotált nevek maradjanak benne.
@@
 \toprule
- & \textbf{huwiki} & \textbf{huwiki filter} & \textbf{Szeged NER} \\
+ & \textbf{huwiki} & \textbf{sűrített huwiki} & \textbf{Szeged NER} \\
 \midrule
@@
 \end{center}
-\caption{A magyar korpuszok mérete és telítettsége.}
+\caption{A magyar Wikipédia és a Szeged NER korpusz mérete és telítettsége.}
 \label{tab:size_hu}
@@
 
-Mivel a névkategóriák leképezésénél a Szeged NER korpusz címkekészletét használtuk, ezért adta magát, hogy a magyar korpusz kiértékeléséhez is ugyanezt alkalmazzuk.
+Mivel a névkategóriák leképezésénél a Szeged NER korpusz címkekészletét használtuk, ezért adta magát, hogy a korpusz kiértékeléséhez is ugyanezt alkalmazzuk.
 Többek által (pl.
@@
 Különböző típusú szövegek esetén a különbség 20-30\% is lehet.
-A helyzet a mi esetünkben is nagyon hasonló (lásd a \ref{tab:huresults}.~táblázatot az eredményekért): a Wikipédián tanított rendszer teljesítménye közel sem olyan jó a gold standard korpusz kiértékelő halmazán mérve, mint a saját kiértékelő halmazán.
+A helyzet a mi esetünkben is nagyon hasonló (lásd az \ref{tab:huresults}.~táblázatot az eredményekért): a Wikipédián tanított rendszer teljesítménye közel sem olyan jó a gold standard korpusz kiértékelő halmazán mérve, mint a saját kiértékelő halmazán.
 
@@
 \toprule 
-\bf tanítás & \bf teszt & \bf pontosság & \bf fedés & \bf F-mérték \\ 
+\bf tanítás & \bf teszt & \bf Pontosság (\%) & \bf Fedés (\%) & \bf F-mérték (\%) \\ 
 \midrule
@@
 
-A Wikipédia korpuszt további módokon is használhatjuk a tulajdonnév-felismerés teljesítményének növelése érdekében.
+Az általunk épített korpuszt további módokon is használhatjuk a tulajdonnév-felismerés teljesítményének növelése érdekében.
 Egy kézenfekvő megoldás nagyméretű névlisták kinyerése a Wikipédiából, és azok hozzáadása gazetteer listaként a tanításhoz.
@@
 
-%na ezt még majd át kell írni a végleges eredmények ismeretében. A Misc-es fejtegetést lehet, hogy inkább a hibaelemzés részbe kéne átrakni. 
-Ha összehasonlítjuk a magyar és az angol eredményeket, azt látjuk, hogy a magyar korpuszokon szignifikánsan jobb eredményeket értünk el. Ennek több oka is lehet. 
-Az egyik, hogy a magyar nyelv esetében a \texttt{Misc} kategóriába sokkal kevesebb dolog tartozik, vagyis kevésbé heterogén, így kevesebb hibát lehet elkövetni az automatikus annotálás során.
-És ahogy a magyar korpuszon végzett kézi hibaelemzés eredménye is mutatja, tényleg sikerült elkerülni az esetek nagy részében a félrecímkézéseket.
-A másik ok az lehet, hogy a magyar esetében használt silver és gold standard adathalmazok egyaránt sokkal tisztábbak, a NE-annotálás pontosabb, mint az angol erőforrások esetében.
-Mivel az automatikusan épített magyar korpuszunk közelíti a gold standard minőséget, mondhatjuk, hogy olyan erőforrást sikerült létrehoznunk, amely használható tulajdonnév-felismerő rendszerek tanító és kiértékelő korpuszaként, akár önállóan, akár kiegészítve a gold standard korpuszokat.
+A kiértékelés legfontosabb eredményének a saját teszthalmazon elért 89,76\%-os F-mértéket tartjuk.
+A kézi hibaelemzés tanulságaival együtt ez arról tanúskodik, hogy az általunk épített korpusz akár önálló gold standard adathalmazként, akár kiegészítő erőforrásként jól használható automatikus névfelismerő rendszerek építéséhez.
 
@@
 
-Cikkünkben egy olyan új módszert mutattunk be, amellyel magyar és angol nyelvű, automatikusan tulajdonnév-annotált korpuszokat hoztunk létre a Wikipédiából.
+Cikkünkben egy új módszert mutattunk be, amellyel létrehoztunk egy magyar nyelvű, automatikusan tulajdonnév-annotált korpuszt a Wikipédiából.
 Az eddig alkalmazottakkal ellentétben a mi metódusunk egy leképezést valósít meg a DBpedia ontológiai osztályairól a hagyományos címkekészletekre.
@@
 Módszerünk nyilvánvaló előnyei, hogy nagyban csökkenti az annotálás költségeit, valamint hogy sokkal nagyobb adathalmazokat állíthatunk elő általa, mint kézi annotációval.
-Egy másik előnye, hogy bármely Wikipédiával rendelkező nyelvre alkalmazható, így kevés erőforrással rendelkező nyelvekre is előállíthatunk silver standard korpuszokat.
+Egy másik előnye, hogy bármely Wikipédiával rendelkező nyelvre alkalmazható, így kevés erőforrással rendelkező nyelvekre is előállíthatunk a gold standard minőséget közelítő korpuszokat.
 A létrehozott korpuszok a továbbiakban számos módon alkalmazhatók a tulajdonnév-felismerő rendszerek hatékonyságának növelésére.
@@
 Tudomásunk szerint ez az első magyar nyelvű automatikusan előállított tulajdonnév-annotált korpusz.
-Az angol erőforrások tekintetében is hasonló a helyzet: jelenleg a Semantically Annotated Snapshot of English Wikipedia mellett az itt publikált korpusz az egyetlen szabadon felhasználható tulajdonnév-annotált korpusz.
+Az angol erőforrások tekintetében is hasonló a helyzet: tudomásunk szerint a Semantically Annotated Snapshot of English Wikipedia \cite{Atserias:08} mellett az itt publikált korpusz az egyetlen szabadon felhasználható tulajdonnév-annotált korpusz.
 
@@
 \section*{Köszönetnyilvánítás}
-A fejlesztés az OTKA 82333 számú projektjén belül valósult meg.
-A fejlesztést támogatta továbbá a CESAR projekt (No.
-271022).
-A szerzők ezúton fejezik ki köszönetüket Zséder Attilának a Wikipédia szövegek feldolgozásában végzett munkájáért, és Kornai Andrásnak támogatásáért.
-%vagy mi
+A fejlesztés az OTKA 82333.~számú projektjén belül valósult meg.
+A fejlesztést támogatta továbbá a CESAR projekt (No.~271022).
+A szerzők ezúton fejezik ki köszönetüket Zséder Attilának a Wikipédia-szövegek feldolgozásában végzett munkájáért, és Kornai Andrásnak támogatásáért.
 

### 2012-11-30 paper/MSZNY2013/cikk.tex
Én is hozzászóltam. 	modified:   cikk.tex

@@
 
-Először, a Wikipédia nem minden szócikke szól megnevezett entitásokról: köznevek, dátumok és egyéb, nem tulajdonnévi elemek is
-saját oldalt kaptak. Annak érdekében, hogy az algoritmus ne
+Először, a Wikipédia nem minden szócikke szól megnevezett entitásokról:
+egyes köznevek, dátumok és egyéb, nem tulajdonnévi elemek is
+kaptak saját oldalt. Annak érdekében, hogy az algoritmus ne
 kezelje ezeket \texttt{Unk} típusú entitásként, majd dobja el a mondatokat,

### 2012-11-30 paper/MSZNY2013/cikk.tex
Final version? 	modified:   cikk.tex

@@
 felcímkézett, angol nyelvű tulajdonnév-listát kaptunk. Utolsó lépésként
-meghatározzuk ezen oldalak magyar és egyéb nyelvű megfelelőit, és ezeket
-is felvesszük a listára. Az idegen nyelvű oldalak meghagyásának oka, hogy
+meghatároztuk ezen oldalak magyar és egyéb nyelvű megfelelőit, és ezeket
+is felvettük a listára. Az idegen nyelvű oldalak meghagyásának oka, hogy
 a magyar Wikipédia időnként ezekre is hivatkozik.
@@
 írók műveit felsoroló oldalakon fordul elő. Ezen műveknek nincs saját szócikkük,
-ezért címkézésük nehéz.
-Mivel egy műcímben bármilyen nyelvi elem előfordulhat, az általunk alkalmazott szűrők együttese sem képes kiszűrni ezeket.
-Megoldást jelenthetne az, ha ezeket a kizárólag műcímeket felsoroló oldalakat külön kezelnénk, és egy komplex rendszert építenénk a feldolgozásukra.
-Mivel ez egy külön nyelvfeldolgozási feladat, jelen fejlesztésen belül nem vállalkozunk rá.
-%Ezért a jövőben az ilyen oldalakat kihagyjuk a korpuszból.
-% TODO: így ok?
+ezért címkézésük nehéz. Mivel egy műcímben bármilyen nyelvi elem előfordulhat,
+az általunk alkalmazott szűrők együttese sem képes kiszűrni ezeket. Megoldást
+jelenthetne az, ha ezeket a kizárólag műcímeket felsoroló oldalakat külön kezelnénk,
+és egy komplex rendszert építenénk a feldolgozásukra. Mivel ez egy külön
+nyelvfeldolgozási feladat, jelen fejlesztésen belül nem vállalkozunk rá;
+a jövőben az ilyen oldalakat kihagyjuk a korpuszból.
 

### 2013-01-03 paper/MSZNY2013/mszny_slides.tex
slide-ok elkezdve

--- 2012-07-02 paper/jeju_presentation.tex
@@
 
-\usepackage[english]{babel}
+\usepackage[magyar]{babel}
 \usepackage[utf8x]{inputenc}
@@
 \title % (optional, use only with long paper titles)
-{Automatically generated NE tagged corpora for English and Hungarian}
+{Automatikus korpuszépítés tulajdonnév-felismerés céljára}
 
@@
 \author % (optional, use only with lots of authors)
-{Eszter Simon\inst{1}, Dávid Márk Nemeskey\inst{2}}
+{Nemeskey Dávid Márk\inst{1}, Simon Eszter\inst{2}}
 
-\institute{\inst{1} Research Institute for Linguistics, Hungarian Academy of Sciences \\
-\inst{2} Computer and Automation Research Institute, Hungarian Academy of Sciences} % (optional, but mostly needed)
+\institute{\inst{1} MTA SZTAKI \\
+\inst{2} MTA Nyelvtudományi Intézet} % (optional, but mostly needed)
 % - Use the \inst command only if there are several affiliations.
@@
 \date % (optional)
-{2012.07.12. \\ ACL NEWS2012 workshop}
+{2013.01.07. \\ MSZNY 2013}
 
@@
    \begin{frame}
-       \frametitle{Outline}
+       \frametitle{Az előadás vázlata}
        \tableofcontents[currentsection]
@@
 
-\section{Introduction}
+\section{Bevezetés}
 
@@
 
-%\section{Wikipedia and NER}
+\section{Wikipédia és tulajdonnév-felismerés}
 
@@
 
-\section{Creating the English Corpus}
+\section{Korpuszépítés}
 
@@
 
-\section{Creating the Hungarian Corpus}
+\section{Problémás esetek, hibaelemzés}
 
@@
 
-\section{Data Description}
+\section{A korpuszok leírása}
 
@@
 
-\section{Evaluation}
+\section{Kiértékelés}
 
@@
 
-\section{Conclusions}
+\section{Összegzés}
 \begin{frame}{Conclusions}
@@
 
-{\huge Thank you for your attention!}
+{\huge Köszönjük a figyelmet!}
 
@@
 
-emails: \\ {\tt simon.eszter@nytud.mta.hu \\ nemeskey.david@sztaki.mta.hu} \\
+e-mail: \\ {\tt nemeskey.david@sztaki.mta.hu \\ simon.eszter@nytud.mta.hu} \\
 
@@
 
-the corpora are freely downloadable from {\tt http://hlt.sztaki.hu/resources/}
+URL: \\ {\tt http://hlt.sztaki.hu/resources/}
 

### 2013-01-03 paper/MSZNY2013/mszny_slides.tex
elso 3 slide

@@
 
-\begin{frame}{Datasets}
+\begin{frame}{A feladat és a probléma}
 
@@
 
-The standard datasets are
-\smallskip
-\begin{itemize}
-\item highly domain-specific
-\item restricted in size
-\end{itemize}
+A feladat: a szövegben felismerni és előre definiált osztályok valamelyikébe besorolni a neveket
 
@@
 
-To get a bigger training corpus:
-\smallskip
+A probléma:
 \begin{itemize}
-\item merging datasets \nyil problems of combining different tagsets
-\item manual annotation \nyil not cost-effective
+\item a sztenderd adathalmazok korlátozott méretűek és témaspecifikusak 
+\item robusztus rendszerek építéséhez nagyméretű és heterogén tulajdonnév-annotált korpuszokra van szükség
+\item a kézi annotálás rendkívül költséges és erőforrás-igényes
 \end{itemize}
@@
 
-\begin{frame}{Reducing the Annotation Cost}
+\begin{frame}{A kihívás}
 
-\bigskip
+az annotációs költség csökkentése \nyil automatikusan annotált korpuszok előállítása
 
-Reducing the annotation cost:
-\smallskip
 \begin{itemize}
-\item to generate the resource automatically 
-\item to use collaboratively constructed resources (e.g.~Wikipedia, DBpedia)
+\item meglévő korpuszok egyesítése \nyil az annotációs sémák és címkekészletek különbözősége 
+\item rendelkezésre álló korpuszon tanított tagger futtatása új szövegen \nyil a domének közötti különbség teljesítménycsökkenést okoz
+\item közösségi tartalmak felhasználása \nyil leképezés a sztenderd címkekészletekre
 \end{itemize}
 
-\bigskip
+\smallskip
 
 \begin{nix}
-Combining these approaches: \\ automatically generating NE tagged corpora from Wikipedia
+A fenti megközelítések kombinációja: \\ automatikus eszközökkel tulajdonnév-annotált korpuszok építése a Wikipédiából
 \end{nix}
@@
 
-\begin{frame}{Wikipedia and NER}
+\begin{frame}{Wikipédia és NER}
+
+\bigskip
 
 \begin{nix}
-Wikipedia is one of the largest available collections of entities
+mivel a Wikipédia címszavainak nagy része tulajdonnév, kiválóan használható tulajdonnév-felismeréshez
 \end{nix}
 
-\smallskip
+\bigskip
 
-Improving the accuracy of NER with WP:
+Alkalmazási módjai:
 \smallskip
 \begin{itemize}
-\item extracting gazetteers containing person names, locations and organizations 
-\item extracting the knowledge embodied in WP and incorporating it in NER learning as features
-\item automatic creation of training data from WP
+\item nagyméretű névlisták előállítása 
+\item a Wikipédia kategóriacímkéinek kinyerése és beépítése NER rendszerekbe
+\item tanító korpusz előállítása a Wikipédia szövegbázisából
 \begin{itemize}
-\item by using the category structure of WP
-\item by using the category head nouns and the opening sentences of articles
+\item az inherens kategóriastruktúra felhasználásával
+\item a szócikkek első mondatának felhasználásával
 \end{itemize} 

### 2013-01-03 paper/MSZNY2013/mszny_slides.tex
Data availability slide has been translated.

@@
 
-\begin{frame}{Data Availability}
+\begin{frame}{A korpusz elérhetősége}
 \bigskip
-The corpora are
+A korpuszok
 \begin{itemize}
-\vitem available under CC-BY-SA 3.0 Unported License
-\vitem freely downloadable from \url{http://hlt.sztaki.hu/resources/} 
-\vitem also distributed through META-SHARE (from August)
+\vitem a CC-BY-SA 3.0 Unported licensz alatt kerültek publikálásra
+\vitem szabadon letölthetőek a \url{http://hlt.sztaki.hu/resources/} honlapról
+\vitem szintén elérhetőek a META-SHARE-en keresztül. Utóbbi
   \begin{itemize}
-  \vitem an open, distributed facility for exchanging and sharing resources
-  \vitem one of the lines of action of META-NET, a Network of Excellence funded
-        by the European Commission
+  \vitem egy nyílt rendszer, mely lehetővé teszi a nyelvi erőforrások megosztását
+  \vitem létrehozója a META-NET, az Európai Bizottság által alapított nyelvtechnológiai hálózat
   \vitem \url{http://www.meta-net.eu/meta-share}

### 2013-01-03 paper/MSZNY2013/mszny_slides.tex
I'm going home.

@@
 
-\subsection*{Motivation}
-\begin{frame}{Motivation}
+\subsection*{Motiváció}
+\begin{frame}{Motiváció}
   \bigskip
@@
   \begin{itemize}
-  \item Our goal is to create a large NE annotated corpus, automatically
-        generated from WP articles.
+  \item Célunk egy NE korpusz készítése automatikusan, a Wikipédia cikkei alapján
+% TODO: WP két tulajdonsága
   \smallskip
-  \vitem Our method is similar to Nothman et al. (2008)
-  \vitem The process consists of four steps:
+  \vitem Módszerünk hasonló Nothman et al. (2008)-hoz
+  \vitem A korpuszépítés a következő lépésekből áll:
     \begin{enumerate}
-    \vitem Classify WP articles into entity classes.
-    \vitem \textit{Parse WP and split articles into sentences.}
-    \vitem Label named entities in the text.
-    \vitem Select the sentences for inclusion in the corpus.
+    \vitem a Wikipédia-cikkeket entitásosztályokba soroljuk;
+    \vitem a cikkeket mondatokra bontjuk;
+    \vitem felcímkézzük a tulajdonneveket a szövegben;
+    \vitem kiszűrjük a rossz minőségű mondatokat.
     \end{enumerate}
@@
 
-\subsection{Articles to Entities}
-\begin{frame}{Articles to Entities}
+\subsection{Cikkek, mint entitások}
+\begin{frame}{Cikkek, mint entitások}
   \bigskip

### 2013-01-04 paper/MSZNY2013/mszny_slides.tex
Two pages translated.

@@
 
-\AtBeginSection[]
-{
-   \begin{frame}
-       \frametitle{Az előadás vázlata}
-       \tableofcontents[currentsection]
-   \end{frame}
-}
+%\AtBeginSection[]
+%{
+%   \begin{frame}
+%       \frametitle{Az előadás vázlata}
+%       \tableofcontents[currentsection]
+%   \end{frame}
+%}
 
@@
 
-\subsection*{Motiváció}
-\begin{frame}{Motiváció}
-  \bigskip
+\subsection*{A módszer}
 
+% TODO: bigskip?
+\begin{frame}{Korpuszépítés}
   \begin{itemize}
-  \item Célunk egy NE korpusz készítése automatikusan, a Wikipédia cikkei alapján
-% TODO: WP két tulajdonsága
-  \smallskip
-  \vitem Módszerünk hasonló Nothman et al. (2008)-hoz
-  \vitem A korpuszépítés a következő lépésekből áll:
+  \vitem A Wikipédia két feltételezett tulajdonsága:
+    \begin{enumerate}
+    \vitem a szócikkek nagy része entitásokról szól
+    \vitem a kereszthivatkozások azonosítják a tulajdonneveket a szövegben
+    \end{enumerate}
+  \vitem Módszerünk hasonló Nothman et al. (2008)-hoz:
     \begin{enumerate}
@@
   \bigskip
-  Our method is based on thinking of WP articles as NEs.
+  Cikkek entitásosztályokba sorolása: % TODO what?
   \smallskip
   \begin{itemize}
-  \vitem Usual approach: semi-supervised learning (infoboxes, first sentence)
-  \vitem Our approach: create entity list from DBpedia
+  \vitem Szokásos megoldás: felügyelt tanulás (infoboxok, első mondat)
+  \vitem A mi megközelítésünk: entitás lista a DBpedia alapján
     \begin{itemize}
-    \vitem OWL ontology with 320 classes
-    \vitem We take the most specific class
-    \vitem Manual mapping to CoNLL categories that exploits inheritence
+    \vitem OWL ontológia kb. 320 osztállyal: kézileg leképeztük CoNLL (Szeged) kategóriákra
+    \vitem Minden entitást a legspecifikusabb osztály szerinti kategóriába raktunk
     \end{itemize}
-  \item Decreased coverage, increased precision
+  \item Csökkent fedés, jobb pontosság
   \end{itemize}
@@
 
-\subsection{Named Entity Tagging}
-\begin{frame}{Named Entity Tagging}
+\subsection{Tulajdonnevek címkézése}
+\begin{frame}{Tulajdonnevek címkézése}
   \bigskip
-  The entity list from DBpedia is used for NE tagging.
-  \smallskip
   \begin{itemize}

### 2013-01-04 paper/MSZNY2013/mszny_slides.tex
I don't know.

@@
   \begin{itemize}
-  \vitem Entity identification:
+  \vitem Tulajdonnevek felismerése:
     \begin{itemize}
@@
   \vitem NE tagging: from the DBpedia--CoNLL category mapping
-  \vitem Exceptions:
+  \vitem Kivételek:
     \begin{itemize}
-    \vitem Not entities: common nouns, date expressions, etc.
-    \vitem Unmarked entities: link to already identified entities
-    \vitem Special cases: derived words, first word in a sentence, etc.
+    \vitem Cikkek, amik nem entitások: nagybetűvel kell kezdődjön a szócikk
+    \vitem Jelöletlen entitások: megpróbáljuk őket korábbi tulajdonnevekhez kötni
+    \vitem Speciális esetek: tulajdonnév-köznév összetételek, mondatkezdő főnevek, stb.
     \end{itemize}
@@
 
-\subsection{Sentence Filtering}
-\begin{frame}{Sentence Filtering}
+\subsection{Mondatok szűrése}
+\begin{frame}{Mondatok szűrése}
   \bigskip
-  We only include fully tagged sentences in the corpus.
+  Csak a teljesen felcímkézett mondatokat tartjuk meg. Eldobjuk:
   \smallskip
   \begin{itemize}
-  \vitem Sentences that contain unknown links / uppercase words
-  \vitem To filter incomplete sentences:
+  \vitem azokat a mondatokat, amiben címkézetlen hivatkozások vagy nagybetűs szavak vannak
+  \vitem a mondattöredékeket:
     \begin{itemize}
-    \vitem Sentences without an EOS punctuation mark
-    \vitem Sentences without a predicate (verb)
+    \vitem amikben nincs mondatvégi írásjel
+    \vitem állítmány nélküli mondatok %TODO: tényleg?
     \end{itemize}

### 2013-01-04 paper/MSZNY2013/mszny_slides.tex
The method part has been translated.
plusz egy-ket slide

@@
 
-\begin{frame}{Our Approach}
+\begin{frame}{Az általunk alkalmazott módszer}
 
 \begin{nix}
-Mapping the DBpedia ontology classes to standard NE tags, and assign them to WP entities
+A DBpedia ontológiai osztályait képeztük le a CoNLL-névosztályokra, majd ezeket Wikipédia-entitásokhoz kötöttük
 \end{nix}
@@
 
-Our method is largely language-independent \nyil sufficiently large corpora for less resourced languages can be built
+Nyelvfüggetlen módszer \nyil minden Wikipédia-nyelvre építhető automatikusan névannotált korpusz
 
@@
 
-\begin{nix}
-This is the first automatically annotated corpus for Hungarian 
-\end{nix}
+\begin{description}
+\item[Magyar:] ez az első automatikusan névannotált korpusz
+\item[Angol:] a \textit{Semantically Annotated Snapshot of English Wikipedia} mellett ez az egyetlen
+\end{description}
 
@@
     \end{enumerate}
-  \vitem Módszerünk hasonló Nothman et al. (2008)-hoz:
+  \vitem Módszerünk hasonló Nothman et al.~(2008)-hoz:
     \begin{enumerate}
@@
 
-\subsection{Cikkek, mint entitások}
-\begin{frame}{Cikkek, mint entitások}
+\subsection{Cikkek mint entitások}
+\begin{frame}{Cikkek mint entitások}
   \bigskip
@@
     \begin{itemize}
-    \vitem amikben nincs mondatvégi írásjel
-    \vitem állítmány nélküli mondatok %TODO: tényleg?
+    \vitem amikben nincs mondatvégi írásjel %és amik nem nagybetűvel kezdődnek
+    \vitem állítmány nélküli mondatok %TODO: tényleg? SE: ezeket végül nem dobtuk el, csak leírtuk, hogy el lehet
     \end{itemize}
@@
 
-\begin{frame}{Data Format}
+\begin{frame}{Multitag formátum}
 
-\begin{tabular}{llllll}
+\begin{tabular}{lllllll}
 \textbf{token} & \textbf{text/link} & \textbf{WP link} & \textbf{POS} & \textbf{lemma} & \textbf{NE} \\
-This & text & 0 &    DT &   This & O \\
-is &   text & 0 &    VBZ &  be &   O \\
-a &    text & 0 &    DT &   a  &    O\\
-list & text & 0 &    NN &   list  & O\\
-of &   text & 0 &    IN &   of &     O\\
-films  & B-link & films &  NNS &  film & O\\
-by &   text & 0 &    IN &   by &     O\\
-year & text & 0 &    NN &   year & O\\
-produced & text & 0 &    VBN &  produce & O\\
-in &   text & 0 &    IN &   in &     O\\
-the &  text & 0 &    DT &   the &   O\\
-country &text & 0 &    NN &   country &O\\
-of &   text & 0 &    IN &   of &     O\\
-South &  B-link & South Korea &  NNP &  South  &B-LOC\\
-Korea &  I-link & South Korea &  NNP &  Korea   & I-LOC\\
+Neumann & B-link & Neumann János & NOUN & neumann & B-PER & B-PER \\
+János & I-link & Neumann János & NOUN & jános & I-PER & I-PER \\
+( & text & 0 & PUNCT & ( & O & O \\
+1903–1957 & text & 0 & NUM & 1903–1957 & O & O \\
+) & text & 0 & PUNCT & ) & O & O \\
+matematikai & text & 0 & ADJ & matematikai & O & O \\
+szemszögből & text & 0 & NOUN<CAS<ELA>> & szemszög & O & O \\
+közelíti & text & 0 & VERB<DEF> & közelít & O & O \\
+meg & text & 0 & PREV & meg & O & O \\
+a & text & 0 & ART & a & O & O \\
+kérdést & text & 0 & NOUN<CAS<ACC>> & kérdés & O & O \\
 \end{tabular}
+
+%\begin{tabular}{llllll}
+%\textbf{token} & \textbf{text/link} & \textbf{WP link} & \textbf{POS} & \textbf{lemma} & \textbf{NE} \\
+%This & text & 0 &    DT &   This & O \\
+%is &   text & 0 &    VBZ &  be &   O \\
+%a &    text & 0 &    DT &   a  &    O\\
+%list & text & 0 &    NN &   list  & O\\
+%of &   text & 0 &    IN &   of &     O\\
+%films  & B-link & films &  NNS &  film & O\\
+%by &   text & 0 &    IN &   by &     O\\
+%year & text & 0 &    NN &   year & O\\
+%produced & text & 0 &    VBN &  produce & O\\
+%in &   text & 0 &    IN &   in &     O\\
+%the &  text & 0 &    DT &   the &   O\\
+%country &text & 0 &    NN &   country &O\\
+%of &   text & 0 &    IN &   of &     O\\
+%South &  B-link & South Korea &  NNP &  South  &B-LOC\\
+%Korea &  I-link & South Korea &  NNP &  Korea   & I-LOC\\
+%\end{tabular}
 

### 2013-01-04 paper/MSZNY2013/mszny_slides.tex
Something.
The corpus description has been moved to the end.

@@
   \begin{itemize}
-  \vitem Tulajdonnevek felismerése:
+  \vitem Tulajdonnevek felismerése szövegben:
     \begin{itemize}
-    \vitem Look up linked pages in the entity list
-    \vitem References to the current page, if in the list
+    \vitem Kereszthivatkozások, ha a hivatkozott cikk szerepel az entitáslistában
+    \vitem Az aktuális cikk címe, ha szerepel a listában
     \end{itemize}
-  \vitem NE tagging: from the DBpedia--CoNLL category mapping
+  \vitem Címkézés: a DBpedia--Szeged leképezés alapján
   \vitem Kivételek:
@@
 %TODO: the "Hungarian Corpus" phrase must appear in the frame title 
+\begin{frame}{Nyelvfüggetlenség}
+\end{frame}
+
 \begin{frame}{Differences from English}
@@
   \end{Examplee}
-\end{frame}
-
-\section{A korpuszok leírása}
-
-\begin{frame}{Multitag formátum}
-
-\begin{tabular}{lllllll}
-\textbf{token} & \textbf{text/link} & \textbf{WP link} & \textbf{POS} & \textbf{lemma} & \textbf{NE} \\
-Neumann & B-link & Neumann János & NOUN & neumann & B-PER & B-PER \\
-János & I-link & Neumann János & NOUN & jános & I-PER & I-PER \\
-( & text & 0 & PUNCT & ( & O & O \\
-1903–1957 & text & 0 & NUM & 1903–1957 & O & O \\
-) & text & 0 & PUNCT & ) & O & O \\
-matematikai & text & 0 & ADJ & matematikai & O & O \\
-szemszögből & text & 0 & NOUN<CAS<ELA>> & szemszög & O & O \\
-közelíti & text & 0 & VERB<DEF> & közelít & O & O \\
-meg & text & 0 & PREV & meg & O & O \\
-a & text & 0 & ART & a & O & O \\
-kérdést & text & 0 & NOUN<CAS<ACC>> & kérdés & O & O \\
-\end{tabular}
-
-%\begin{tabular}{llllll}
-%\textbf{token} & \textbf{text/link} & \textbf{WP link} & \textbf{POS} & \textbf{lemma} & \textbf{NE} \\
-%This & text & 0 &    DT &   This & O \\
-%is &   text & 0 &    VBZ &  be &   O \\
-%a &    text & 0 &    DT &   a  &    O\\
-%list & text & 0 &    NN &   list  & O\\
-%of &   text & 0 &    IN &   of &     O\\
-%films  & B-link & films &  NNS &  film & O\\
-%by &   text & 0 &    IN &   by &     O\\
-%year & text & 0 &    NN &   year & O\\
-%produced & text & 0 &    VBN &  produce & O\\
-%in &   text & 0 &    IN &   in &     O\\
-%the &  text & 0 &    DT &   the &   O\\
-%country &text & 0 &    NN &   country &O\\
-%of &   text & 0 &    IN &   of &     O\\
-%South &  B-link & South Korea &  NNP &  South  &B-LOC\\
-%Korea &  I-link & South Korea &  NNP &  Korea   & I-LOC\\
-%\end{tabular}
-
-\end{frame}
-
-\begin{frame}{A korpusz elérhetősége}
-\bigskip
-A korpuszok
-\begin{itemize}
-\vitem a CC-BY-SA 3.0 Unported licensz alatt kerültek publikálásra
-\vitem szabadon letölthetőek a \url{http://hlt.sztaki.hu/resources/} honlapról
-\vitem szintén elérhetőek a META-SHARE-en keresztül. Utóbbi
-  \begin{itemize}
-  \vitem egy nyílt rendszer, mely lehetővé teszi a nyelvi erőforrások megosztását
-  \vitem létrehozója a META-NET, az Európai Bizottság által alapított nyelvtechnológiai hálózat
-  \vitem \url{http://www.meta-net.eu/meta-share}
-  \end{itemize}
-\end{itemize}
-
 \end{frame}
@@
 
+\section{A korpuszok leírása}
+
+\begin{frame}{Multitag formátum}
+
+% TODO: nem fér ki minden oszlop
+\begin{tabular}{lllllll}
+\textbf{token} & \textbf{text/link} & \textbf{WP link} & \textbf{POS} & \textbf{lemma} & \textbf{NE} \\
+Neumann & B-link & Neumann János & NOUN & neumann & B-PER & B-PER \\
+János & I-link & Neumann János & NOUN & jános & I-PER & I-PER \\
+( & text & 0 & PUNCT & ( & O & O \\
+1903–1957 & text & 0 & NUM & 1903–1957 & O & O \\
+) & text & 0 & PUNCT & ) & O & O \\
+matematikai & text & 0 & ADJ & matematikai & O & O \\
+szemszögből & text & 0 & NOUN<CAS<ELA{>}> & szemszög & O & O \\
+közelíti & text & 0 & VERB<DEF> & közelít & O & O \\
+meg & text & 0 & PREV & meg & O & O \\
+a & text & 0 & ART & a & O & O \\
+kérdést & text & 0 & NOUN<CAS<ACC{>}> & kérdés & O & O \\
+\end{tabular}
+
+%\begin{tabular}{llllll}
+%\textbf{token} & \textbf{text/link} & \textbf{WP link} & \textbf{POS} & \textbf{lemma} & \textbf{NE} \\
+%This & text & 0 &    DT &   This & O \\
+%is &   text & 0 &    VBZ &  be &   O \\
+%a &    text & 0 &    DT &   a  &    O\\
+%list & text & 0 &    NN &   list  & O\\
+%of &   text & 0 &    IN &   of &     O\\
+%films  & B-link & films &  NNS &  film & O\\
+%by &   text & 0 &    IN &   by &     O\\
+%year & text & 0 &    NN &   year & O\\
+%produced & text & 0 &    VBN &  produce & O\\
+%in &   text & 0 &    IN &   in &     O\\
+%the &  text & 0 &    DT &   the &   O\\
+%country &text & 0 &    NN &   country &O\\
+%of &   text & 0 &    IN &   of &     O\\
+%South &  B-link & South Korea &  NNP &  South  &B-LOC\\
+%Korea &  I-link & South Korea &  NNP &  Korea   & I-LOC\\
+%\end{tabular}
+
+\end{frame}
+
+\begin{frame}{A korpusz elérhetősége}
+\bigskip
+A korpuszok
+\begin{itemize}
+\vitem a CC-BY-SA 3.0 Unported licensz alatt kerültek publikálásra
+\vitem szabadon letölthetőek a \url{http://hlt.sztaki.hu/resources/} honlapról
+\vitem szintén elérhetőek a META-SHARE-en keresztül. Utóbbi
+  \begin{itemize}
+  \vitem egy nyílt rendszer, mely lehetővé teszi a nyelvi erőforrások megosztását
+  \vitem létrehozója a META-NET, az Európai Bizottság által alapított nyelvtechnológiai hálózat
+  \vitem \url{http://www.meta-net.eu/meta-share}
+  \end{itemize}
+\end{itemize}
+
+\end{frame}
+
 \section{Összegzés}

### 2013-01-04 paper/MSZNY2013/mszny_slides.tex
3 slides assigned to ppl.
Hahaha!

@@
 
-\section{Problémás esetek, hibaelemzés}
-
-%\begin{frame}{The Hungarian Corpus}
-%  The algorithm is, in theory, language-agnostic. We tested this property by
-%  applying it to a medium-sized language, Hungarian. \\
-%  Difference from bigger, well-resourced languages:
-%  \begin{itemize}
-%  \vitem E % TODO: ide mit?
-%  \end{itemize}
-%\end{frame}
-
-%TODO: the "Hungarian Corpus" phrase must appear in the frame title 
-\begin{frame}{Nyelvfüggetlenség}
-\end{frame}
-
-\begin{frame}{Differences from English}
+\begin{frame}{Nyelvfüggető elemek}
   \bigskip
   \begin{itemize}
-  \vitem Language processing tools
+  \vitem Nyelvi eszközök
     \begin{itemize}
-    \vitem Inflectional morphology: different set of tools were needed 
-    \vitem We developed a flexible framework that enables the plugging of any NLP tool
-    % TODO: what kind of linguistic tools are needed: table?
+    \vitem kidolgoztunk egy keretrendszert, amibe bármilyen eszközt be lehet kötni
+    \vitem tokenizáló, lemmatizáló, szófajmeghatározó
     \end{itemize}
-  \vitem Wikipedia size and coverage
+  \vitem Wikipédia
     \begin{itemize}
-    \vitem Hungarian is medium-size language: \nagytilde 200k articles vs.~\nagytilde 4M English pages
-    \vitem Cross-references may point to English pages
-    \vitem Templates are in Hungarian: per-language template configuration
+    \vitem Sablonok nyelvfüggőek
+    \vitem a magyar egy közepes méretű nyelv: \nagytilde 200k cikk a kb.~\nagytilde 4M angollal szemben
+    \end{itemize}
+  \vitem DBpedia
+    \begin{itemize}
+    \vitem nagyobb nyelvekre honosított fejezetek --- magyarra nincs
+    \vitem csak az angol Wikipédiában is szereplő magyar tulajdonneveket tartalmazza
+    \vitem távlati terv a lista kiegészítése gépi tanulással  % TODO "Future work"
     \end{itemize}
@@
 
-\begin{frame}{Differences from English -- cont.}
-  \bigskip
-  \begin{itemize}
-  \vitem DBpedia coverage
-    \smallskip
-    \begin{itemize}
-    \vitem Internationalized chapters for big languages only
-    \vitem "Hungarian" DBpedia list contains English articles % Nem egyértelmű
-      \begin{itemize}
-      \vitem Hungarian equivalents must be added to the list % Added: lásd előző oldal 2/2
-      \vitem Hungarian entities without an English page are missing:
-             the NER language model might not be accurate
-      \vitem Future work: extend the list with the missing Hungarian entities
-      \end{itemize}
-    \end{itemize}
-  \end{itemize}
-\end{frame}
+\section{Problémás esetek, hibaelemzés}
 
+% Eszter
+% TODO a következő fejezetbe
 \begin{frame}{Differences from English -- cont.}

### 2013-01-04 paper/MSZNY2013/mszny_slides.tex
nagyjabol kesz

--- 2013-01-04 paper/MSZNY2013/mszny_slides.tex
@@
 \usepackage{bm}  % For bold math
+
+\usepackage{booktabs}
 
@@
 
+\begin{frame}{Kézi ellenőrzés}
+
+\bigskip
+
+A teljes korpusz 0,001\%-át kézzel annotáltuk, majd összehasonlítottuk a gépi és kézi annotálás eredményét.
+
+\bigskip
+
+\begin{tabular}{lcccc}
+\toprule
+& P (\%) &  R (\%) & F$_{\beta=1}$ (\%) & NE (\#) \\
+\midrule
+LOC     &   98.72 &  95.65 &  97.16 & 161 \\
+MISC    &   95.24 &  76.92 &  85.11 & 26 \\
+ORG     &   89.66 &  89.66 &  89.66 & 29 \\
+PER     &   88.30 &  89.25 &  88.77 & 93 \\
+\midrule
+Összesítve &   94.33 &  91.59 &  92.94 & 309 \\
+\bottomrule
+\end{tabular}
+
+\end{frame}
+
+\begin{frame}{Igazságmátrix}
+
+\bigskip
+
+\begin{tabular}{l@{\hspace{0.5em}}|@{\hspace{0.5em}}cccc}
+\toprule
+Auto$\downarrow$ / Gold$\rightarrow$ & PER & ORG & LOC & MISC \\
+\midrule
+PER & 83 & 1 & & 2 \\
+ORG &  & 26 & 1 & 1 \\
+LOC &  & 1 & 154 & \\
+MISC &  &  & 1 & 20 \\
+\bottomrule
+\end{tabular}
+
+\bigskip
+
+A típustévesztés okai:
+
+\begin{itemize}
+\item a DBpedia típusinformációja helytelen: \\ $[\mbox{MTA}]_{LOC}$
+\item a Wikipédia rossz kereszthivatkozásai: \\ $[\mbox{Walt Disney}]_{PER}$ Co.
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}{További hibák}
+
+\bigskip
+
+\begin{tabular}{l@{\hspace{0.5em}}cccc}
+\toprule
+	& PER & ORG & LOC & MISC \\
+\midrule
+Álpozitív & 1 &  0  &  1  &  0 \\
+Álnegatív & 3 &  0  &  5  &  4 \\
+Részlegesen felismert nevek & 7 & 1 & 0 & 0 \\
+\bottomrule
+\end{tabular}
+
+\bigskip
+
+\begin{itemize}
+\item Álpozitív: hibásan névként felismert szavak: $[\mbox{Bizánc-ellenes}]_{LOC}$
+\item Álnegatív: fel nem ismert nevek: $[\mbox{ókori Róma}]_O$
+\item Részlegesen felismert nevek: pontatlan entitáshatárok: $[\mbox{Szent István király}]_{PER}$
+\end{itemize}
+
+\end{frame}
+
 %\begin{frame}{The Hungarian Corpus}
@@
 
+% Dávid
 \begin{frame}{Differences from English}
@@
 
+% Dávid
 \begin{frame}{Differences from English -- cont.}
@@
 
+% Eszter
+% TODO a következő fejezetbe
 \begin{frame}{Differences from English -- cont.}
@@
 
-\begin{frame}{Silver Standard Corpora}
+%\begin{frame}{Silver standard korpuszok}
+
+%\bigskip
+
+%A silver standard korpuszok
+
+%\smallskip
+
+%\begin{itemize}
+%\vitem for less resourced languages they can serve as training corpora in lieu of gold standard datasets
+%\vitem they can serve as training sets for domains differing from newswire
+%\vitem they can be sources of huge entity lists
+%\vitem they can be sources of feature extraction
+%\end{itemize}
+
+%\end{frame}
+
+\begin{frame}{Kiértékelés}
 
@@
 
-Silver standard corpora can be useful for improving NER in several ways:
-
-\smallskip
-
 \begin{itemize}
-\vitem for less resourced languages they can serve as training corpora in lieu of gold standard datasets
-\vitem they can serve as training sets for domains differing from newswire
-\vitem they can be sources of huge entity lists
-\vitem they can be sources of feature extraction
+\item A kiértékeléshez a \texttt{hunner} taggert használtuk az alábbi jegyekkel:
+\begin{itemize}
+\vitem mondatkezdő- és vég pozíciók
+\vitem szóalakon alapuló jegyek
+\vitem morfológiai információ
+\vitem listajegyek
 \end{itemize}
-
-\end{frame}
-
-\begin{frame}{Evaluation}
-
-\bigskip
-
-A maxent NE tagger was used with the feature set consisting of
-
-\smallskip
-
-\begin{itemize}
-\vitem gazetteer features
-\vitem sentence start and end position
-\vitem Boolean-valued orthographic properties of the word form
-\vitem string-valued properties of the word form
-\vitem morphological information
+\item Az eredmények kiszámításához a sztenderd CoNLL-módszert alkalmaztuk
 \end{itemize}
@@
 
-The CoNLL standard method was used for evaluation.
+\end{frame}
+
+\begin{frame}{Korpuszméret és telítettség}
+
+\bigskip
+
+\begin{tabular}{lccc}
+\toprule
+ & \textbf{huwiki} & \textbf{sűrített huwiki} & \textbf{Szeged NER} \\
+\midrule
+token & 19.108.027 & 3.512.249  & 225.963 \\
+NE & 456.281 & 456.281  & 25.896 \\
+telítettség (\%) & 2,38 & 12,99 & 11,46 \\
+\bottomrule
+\end{tabular}
+
+\bigskip
+
+\begin{nix}
+szigorú módszer \nyil kevés, de pontos NE \nyil híg korpusz
+\end{nix}
+
+\bigskip
+
+A kiértékeléshez eltávolítottuk a NE-t nem tartalmazó mondatokat.
 
@@
 
-\begin{frame}{Corpus Size and NE Density}
-
-English:
-
-\begin{center}
-\begin{tabular}{llll}
-\hline  & \bf enwiki & \bf enwiki filtered & \bf CoNLL \\ \hline
-token & 60,520,819 & 21,718,854 & 302,811 \\
-NE & 3,169,863 & 3,169,863 & 50,758 \\
-NE density & 5.23\% & 14.59\% & 16.76\% \\ \hline
-\end{tabular}
-\end{center}
-
-%\smallskip
-
-Hungarian:
-
-\begin{center}
-\begin{tabular}{llll}
-\hline  & \bf huwiki & \bf huwiki filtered  & \bf Szeged NER \\ \hline
-token & 19,108,027 & 3,512,249  & 225,963 \\
-NE & 456,281 & 456,281  & 25,896 \\
-NE density & 2.38\% & 12.99\%  & 11.46\% \\ \hline 
-\end{tabular}
-\end{center}
-
-Quite low NE density \nyil sentences without NEs were filtered out from the evaluation datasets
-
-\end{frame}
-
-\begin{frame}{Experiments}
+\begin{frame}{Kísérletek}
 
 \begin{nix}
-Our experiments were performed with a sample of 3.5 million tokens
+a 3,5 millió tokenes korpuszt 90-10\% arányban tanító és kiértékelő halmazra osztottuk
 \end{nix}
@@
 \begin{itemize}
-\item \textit{CoNLL} train -- \textit{CoNLL} test
-\item \textit{enwiki} train -- \textit{enwiki} test
-\item \textit{enwiki} train -- \textit{CoNLL} test
-\item \textit{CoNLL} train with \textit{wikilists} (gazetteer lists extracted from WP) -- \textit{CoNLL} test
-\item \textit{CoNLL} train with \textit{wikitags} (labeling the CoNLL datasets by the model trained on WP, and giving these labels as extra features) -- \textit{CoNLL} test
+\item Szeged NER tanítókorpusz -- Szeged NER tesztkorpusz
+\item huwiki tanítóhalmaz -- huwiki teszthalmaz
+\item huwiki tanítóhalmaz -- Szeged NER tesztkorpusz
+\item Szeged NER tanítókorpusz, a Wikipédiából kinyert névlistákkal -- Szeged NER tesztkorpusz
+\item Szeged NER tanítókorpusz, a Wikipédia korpuszon tanított modell által felcímkézve, és a címkék extra jegyként hozzáadva -- Szeged NER tesztkorpusz
 \end{itemize}
@@
 
-\begin{frame}{English Results}
+\begin{frame}{Eredmények}
 
 \begin{center}
-\begin{tabular}{lllll}
-\hline \bf Train & \bf Test & \bf P (\%) & \bf R (\%) & \bf F (\%) \\ \hline
-CoNLL & CoNLL & 85.13 & 85.13 & 85.13 \\
-enwiki & enwiki & 72.46 & 73.33 &  72.89 \\
-enwiki & CoNLL & 56.55 & 49.77 & 52.94 \\
-CoNLL with wikilists & CoNLL & 86.33 & 86.35 & 86.34 \\
-CoNLL with wikitags & CoNLL & 85.88 & 85.94 & 85.91 \\
-\hline
+\begin{tabular}{lcccc}
+\toprule 
+\bf tanítás & \bf teszt & \bf P (\%) & \bf R (\%) & \bf F (\%) \\ 
+\midrule
+Szeged & Szeged & 94,50 & 94,35 & 94,43 \\
+huwiki & huwiki & 90,64 & 88,91 &  \textbf{89,76}  \\
+huwiki & Szeged & 63,08 & 70,46 & 66,57  \\
+Szeged\_wikilisták & Szeged & 95,48 & 95,48 & 95,48  \\
+Szeged\_wikitag & Szeged & 95,38 & 94,92 & 95,15 \\
+\bottomrule
 \end{tabular}
@@
 
-\begin{itemize}
-\item training and testing across different corpora decreases F-measure
-\item the NE tagger trained on WP does not achieve as high performance tested against CoNLL test set as one trained on its own train set
-\item gazetteer lists extracted from WP and training with extra features given by the model trained on WP improve F-measure
-\end{itemize}
+\bigskip
 
-\end{frame}
-
-\begin{frame}{Hungarian Results}
-
-\begin{center}
-\begin{tabular}{lllll}
-\hline \bf Train & \bf Test & \bf P (\%) & \bf R (\%) & \bf F (\%) \\ \hline
-Szeged  & Szeged  & 94.50 & 94.35 & 94.43 \\
-huwiki & huwiki & 90.64 & 88.91 &  \textbf{89.76} \\
-huwiki & Szeged  & 63.08 & 70.46 & 66.57 \\
-Szeged  with wikilists & Szeged  & 95.48 & 95.48 & 95.48 \\
-Szeged  with wikitags & Szeged  & 95.38 & 94.92 & 95.15 \\
-\hline
-\end{tabular}
-\end{center}
-
-\begin{itemize}
-\vitem F-measures for Hungarian are higher 
-\begin{itemize}
-\vitem the MISC category is more consistent
-\vitem the Szeged NER corpus is highly accurately tagged 
-\end{itemize}
-\vitem the Hungarian corpus can serve as a training corpus to build NE taggers for non-newswire domains
-\end{itemize}
+A saját teszthalmazon elért eredmény arról tanúskodik, hogy a korpusz akár önálló gold standard adathalmazként is használható.
 
@@
 % TODO: nem fér ki minden oszlop
-\begin{tabular}{lllllll}
+%SE: kicsit összébb húztam, de így se fér ki. vagy ki kéne szedni belőle a linkes oszlopokat, vagy valahogy kicsinyíteni az egészet
+\begin{tabular}{@{}l@{ }l@{ }l@{ }l@{ }l@{ }l@{ }l}
 \textbf{token} & \textbf{text/link} & \textbf{WP link} & \textbf{POS} & \textbf{lemma} & \textbf{NE} \\
@@
 \section{Összegzés}
-\begin{frame}{Conclusions}
+\begin{frame}{Összegzés}
   \begin{itemize}
-  \vitem We presented freely available NE tagged corpora for English and Hungarian, fully automatically generated from WP
-  \vitem We applied a new approach: mapping DBpedia ontology classes to standard CoNLL NE tags, and assigning them to WP entities
-  \vitem The huge amount of WP articles opens the possibility of building large enough corpora for less resourced languages such as Hungarian
-  \vitem Silver standard corpora can improve NER accuracy in more ways
-  \vitem The Hungarian corpus can serve as a training corpus to build NE taggers for non-newswire domains
+  \vitem módszerünk nagyban csökkenti az annotációs költségeket 
+  \vitem bármely Wikipédia-nyelvre alkalmazható \nyil kevés erőforrással rendelkező nyelvekre is előállítható tulajdonnév-annotált korpusz
+  \vitem a létrehozott korpuszok több módon is felhasználhatóak a NER rendszerek hatékonyságának növelésére
+  \vitem mivel korpuszunk közelíti a gold standard minőséget, akár önálló tanítóhalmazként is használható a sajtó stílustól eltérő szövegekhez
+  \vitem az általunk előállított korpuszok szabadon felhasználhatóak
+  \vitem a módszerben benne rejlik a lehetőség finomabb tulajdonnév-hierarchiák támogatására is
   \end{itemize}

### 2013-01-04 paper/MSZNY2013/mszny_slides.tex

@@
 
+\begin{frame}{Nyelvfüggető elemek}
+  \bigskip
+  \begin{itemize}
+  \vitem Nyelvi eszközök
+    \begin{itemize}
+    \vitem kidolgoztunk egy keretrendszert, amibe bármilyen eszközt be lehet kötni
+    \vitem tokenizáló, lemmatizáló, szófajmeghatározó
+    \end{itemize}
+  \vitem Wikipédia
+    \begin{itemize}
+    \vitem Sablonok nyelvfüggőek
+    \vitem a magyar egy közepes méretű nyelv: \nagytilde 200k cikk a kb.~\nagytilde 4M angollal szemben
+    \end{itemize}
+  \vitem DBpedia
+    \begin{itemize}
+    \vitem nagyobb nyelvekre honosított fejezetek --- magyarra nincs
+    \vitem csak az angol Wikipédiában is szereplő magyar tulajdonneveket tartalmazza
+    \vitem távlati terv a lista kiegészítése gépi tanulással  % TODO "Future work"
+    \end{itemize}
+  \end{itemize}
+\end{frame}
+
 \section{Problémás esetek, hibaelemzés}
@@
 %\end{frame}
-
-%TODO: the "Hungarian Corpus" phrase must appear in the frame title 
-\begin{frame}{Nyelvfüggetlenség}
-\end{frame}
-
-% Dávid
-\begin{frame}{Differences from English}
-  \bigskip
-  \begin{itemize}
-  \vitem Language processing tools
-    \begin{itemize}
-    \vitem Inflectional morphology: different set of tools were needed 
-    \vitem We developed a flexible framework that enables the plugging of any NLP tool
-    % TODO: what kind of linguistic tools are needed: table?
-    \end{itemize}
-  \vitem Wikipedia size and coverage
-    \begin{itemize}
-    \vitem Hungarian is medium-size language: \nagytilde 200k articles vs.~\nagytilde 4M English pages
-    \vitem Cross-references may point to English pages
-    \vitem Templates are in Hungarian: per-language template configuration
-    \end{itemize}
-  \end{itemize}
-\end{frame}
-
-% Dávid
-\begin{frame}{Differences from English -- cont.}
-  \bigskip
-  \begin{itemize}
-  \vitem DBpedia coverage
-    \smallskip
-    \begin{itemize}
-    \vitem Internationalized chapters for big languages only
-    \vitem "Hungarian" DBpedia list contains English articles % Nem egyértelmű
-      \begin{itemize}
-      \vitem Hungarian equivalents must be added to the list % Added: lásd előző oldal 2/2
-      \vitem Hungarian entities without an English page are missing:
-             the NER language model might not be accurate
-      \vitem Future work: extend the list with the missing Hungarian entities
-      \end{itemize}
-    \end{itemize}
-  \end{itemize}
-\end{frame}
-
-% Eszter
-% TODO a következő fejezetbe
-\begin{frame}{Differences from English -- cont.}
-  \bigskip
-  \begin{itemize}
-  \vitem NE conventions
-    \smallskip
-    \begin{itemize}
-    \vitem Szeged NER guideline: similar to CoNLL
-    \vitem Common nouns in Hungarian: names of languages, religions, months, etc. \nyil more consistent MISC category
-    \vitem NEs can be subject of compounding \nyil the joint common noun can modify the original sense and category of the NE:
-    \end{itemize}
-  \end{itemize}
-  \bigskip
-  \begin{Examplee}
-  \small{WorldCom {\normalfont(\texttt{ORG})} \nyil WorldCom-botrány ['WorldCom scandal'] {\normalfont(\texttt{O})}}
-  \end{Examplee}
-\end{frame}
 

### 2013-01-06 paper/MSZNY2013/mszny_slides.tex
Hmph.
Fixed some stuff, added a few TODOs...

@@
 
-% TODO: kell outline?
 \begin{frame}{}
@@
 
-az annotációs költség csökkentése \nyil automatikusan annotált korpuszok előállítása
+Az annotációs költség csökkentése \nyil automatikusan annotált korpuszok előállítása
 
@@
 \begin{nix}
-mivel a Wikipédia címszavainak nagy része tulajdonnév, kiválóan használható tulajdonnév-felismeréshez
+Mivel a Wikipédia címszavainak nagy része tulajdonnév, kiválóan használható tulajdonnév-felismeréshez
 \end{nix}
@@
 
-Alkalmazási módjai:
+Alkalmazási módjai:  % TODO: eddigi alkalmazások?
 \smallskip
@@
 \begin{frame}{Az általunk alkalmazott módszer}
-
+\bigskip
 \begin{nix}
@@
 
-% TODO: bigskip?
 \begin{frame}{Korpuszépítés}
+  \bigskip
   \begin{itemize}
@@
   \bigskip
-  Cikkek entitásosztályokba sorolása: % TODO what?
+  Cikkek entitásosztályokba sorolása:
   \smallskip
@@
   \vitem Szokásos megoldás: felügyelt tanulás (infoboxok, első mondat)
+    \begin{itemize}
+    \vitem klasszifikációs hibák
+    \end{itemize}
   \vitem A mi megközelítésünk: entitás lista a DBpedia alapján
     \begin{itemize}
+    \vitem szemantikus tudásbázis Wikipédia entitásokról
     \vitem OWL ontológia kb. 320 osztállyal: kézileg leképeztük CoNLL (Szeged) kategóriákra
-    \vitem Minden entitást a legspecifikusabb osztály szerinti kategóriába raktunk
+    \vitem minden entitást a legspecifikusabb osztály szerinti kategóriába raktunk
+    \vitem csökkent fedés, jobb pontosság
     \end{itemize}
-  \item Csökkent fedés, jobb pontosság
   \end{itemize}
@@
   Csak a teljesen felcímkézett mondatokat tartjuk meg. Eldobjuk:
-  \smallskip
+%  \smallskip
   \begin{itemize}
@@
     \begin{itemize}
-    \vitem amikben nincs mondatvégi írásjel %és amik nem nagybetűvel kezdődnek
-    \vitem állítmány nélküli mondatok %TODO: tényleg? SE: ezeket végül nem dobtuk el, csak leírtuk, hogy el lehet
+    \vitem amik nem nagybetűvel kezdődnek
+    \vitem amikben nincs mondatvégi írásjel
+    \vitem befejezett ige nélküli mondatok % Végül nem dobtuk el
     \end{itemize}
@@
 
-\begin{frame}{Nyelvfüggető elemek}
-  \bigskip
+\begin{frame}{Nyelvfüggő elemek}
   \begin{itemize}
@@
     \begin{itemize}
+    \vitem tokenizáló, lemmatizáló, morfológiai egyértelműsítő
     \vitem kidolgoztunk egy keretrendszert, amibe bármilyen eszközt be lehet kötni
-    \vitem tokenizáló, lemmatizáló, szófajmeghatározó
     \end{itemize}
@@
     \vitem csak az angol Wikipédiában is szereplő magyar tulajdonneveket tartalmazza
-    \vitem távlati terv a lista kiegészítése gépi tanulással  % TODO "Future work"
+%    \vitem távlati terv a lista kiegészítése gépi tanulással
     \end{itemize}
@@
 
+\begin{center}
 \begin{tabular}{lcccc}
@@
 \end{tabular} 
+\end{center}
 
@@
 
+\begin{center}
 \begin{tabular}{l@{\hspace{0.5em}}|@{\hspace{0.5em}}cccc}
@@
 \end{tabular}
+\end{center}
 
@@
 
-%\begin{frame}{The Hungarian Corpus}
-%  The algorithm is, in theory, language-agnostic. We tested this property by
-%  applying it to a medium-sized language, Hungarian. \\
-%  Difference from bigger, well-resourced languages:
-%  \begin{itemize}
-%  \vitem E % TODO: ide mit?
-%  \end{itemize}
-%\end{frame}
-
 \section{Kiértékelés}
-
-%\begin{frame}{Silver standard korpuszok}
-
-%\bigskip
-
-%A silver standard korpuszok
-
-%\smallskip
-
-%\begin{itemize}
-%\vitem for less resourced languages they can serve as training corpora in lieu of gold standard datasets
-%\vitem they can serve as training sets for domains differing from newswire
-%\vitem they can be sources of huge entity lists
-%\vitem they can be sources of feature extraction
-%\end{itemize}
-
-%\end{frame}
 
@@
 \bigskip
+A Szeged NER korpusz 
+% TODO: valami!!!!!
 
@@
 
+% Sűrített = csak NE-s mondatok?
 \begin{tabular}{lccc}
@@
 \item huwiki tanítóhalmaz -- Szeged NER tesztkorpusz
-\item Szeged NER tanítókorpusz, a Wikipédiából kinyert névlistákkal -- Szeged NER tesztkorpusz
+\item Szeged NER tanítókorpusz, a Wikipédiából kinyert névlistákkal -- Szeged NER tesztkorpusz % DBpediából, nem?
 \item Szeged NER tanítókorpusz, a Wikipédia korpuszon tanított modell által felcímkézve, és a címkék extra jegyként hozzáadva -- Szeged NER tesztkorpusz
@@
 
-% TODO: nem fér ki minden oszlop
-%SE: kicsit összébb húztam, de így se fér ki. vagy ki kéne szedni belőle a linkes oszlopokat, vagy valahogy kicsinyíteni az egészet
+\bigskip
+
+{\footnotesize
 \begin{tabular}{@{}l@{ }l@{ }l@{ }l@{ }l@{ }l@{ }l}
 \textbf{token} & \textbf{text/link} & \textbf{WP link} & \textbf{POS} & \textbf{lemma} & \textbf{NE} \\
-Neumann & B-link & Neumann János & NOUN & neumann & B-PER & B-PER \\
-János & I-link & Neumann János & NOUN & jános & I-PER & I-PER \\
-( & text & 0 & PUNCT & ( & O & O \\
-1903–1957 & text & 0 & NUM & 1903–1957 & O & O \\
-) & text & 0 & PUNCT & ) & O & O \\
-matematikai & text & 0 & ADJ & matematikai & O & O \\
-szemszögből & text & 0 & NOUN<CAS<ELA{>}> & szemszög & O & O \\
-közelíti & text & 0 & VERB<DEF> & közelít & O & O \\
-meg & text & 0 & PREV & meg & O & O \\
-a & text & 0 & ART & a & O & O \\
-kérdést & text & 0 & NOUN<CAS<ACC{>}> & kérdés & O & O \\
+Neumann & B-link & Neumann János & NOUN & neumann & B-PER \\
+János & I-link & Neumann János & NOUN & jános & I-PER \\
+( & text & 0 & PUNCT & ( & O \\
+1903–1957 & text & 0 & NUM & 1903–1957 & O \\
+) & text & 0 & PUNCT & ) & O \\
+matematikai & text & 0 & ADJ & matematikai & O \\
+szemszögből & text & 0 & NOUN<CAS<ELA{>}> & szemszög & O \\
+közelíti & text & 0 & VERB<DEF> & közelít & O \\
+meg & text & 0 & PREV & meg & O \\
+a & text & 0 & ART & a & O \\
+kérdést & text & 0 & NOUN<CAS<ACC{>}> & kérdés & O \\
 \end{tabular}
+}
 
@@
   \begin{itemize}
+  \vitem létrehozója a META-NET, az Európai Bizottság által alapított nyelvtechnológiai hálózat
   \vitem egy nyílt rendszer, mely lehetővé teszi a nyelvi erőforrások megosztását
-  \vitem létrehozója a META-NET, az Európai Bizottság által alapított nyelvtechnológiai hálózat
   \vitem \url{http://www.meta-net.eu/meta-share}
@@
   \vitem az általunk előállított korpuszok szabadon felhasználhatóak
-  \vitem a módszerben benne rejlik a lehetőség finomabb tulajdonnév-hierarchiák támogatására is
+%  \vitem a módszerben benne rejlik a lehetőség finomabb tulajdonnév-hierarchiák támogatására is
   \end{itemize}

### 2013-01-06 paper/MSZNY2013/mszny_slides.tex
Removed some pages, added some...

@@
 
-\begin{frame}{A feladat és a probléma}
+\begin{frame}{A tulajdonnév-felismerés (Named Entity Recognition)}
 
@@
 
-\begin{frame}{Wikipédia és NER}
-
-\bigskip
-
-\begin{nix}
-Mivel a Wikipédia címszavainak nagy része tulajdonnév, kiválóan használható tulajdonnév-felismeréshez
-\end{nix}
-
-\bigskip
-
-Alkalmazási módjai:  % TODO: eddigi alkalmazások?
-\smallskip
-  \begin{itemize}
-  \item nagyméretű névlisták előállítása 
-  \item a Wikipédia kategóriacímkéinek kinyerése és beépítése NER rendszerekbe
-  \item tanító korpusz előállítása a Wikipédia szövegbázisából
-    \begin{itemize}
-    \item az inherens kategóriastruktúra felhasználásával
-    \item a szócikkek első mondatának felhasználásával
-    \end{itemize} 
-  \end{itemize}
-
-\end{frame}
+%\begin{frame}{Wikipédia és NER}
+%
+%\bigskip
+%
+%\begin{nix}
+%Mivel a Wikipédia címszavainak nagy része tulajdonnév, kiválóan használható tulajdonnév-felismeréshez
+%\end{nix}
+%
+%\bigskip
+%
+%Alkalmazási módjai:  % TODO: eddigi alkalmazások?
+%\smallskip
+%  \begin{itemize}
+%  \item nagyméretű névlisták előállítása 
+%  \item a Wikipédia kategóriacímkéinek kinyerése és beépítése NER rendszerekbe
+%  \item tanító korpusz előállítása a Wikipédia szövegbázisából
+%    \begin{itemize}
+%    \item az inherens kategóriastruktúra felhasználásával
+%    \item a szócikkek első mondatának felhasználásával
+%    \end{itemize} 
+%  \end{itemize}
+%
+%\end{frame}
 
@@
 \begin{nix}
-A DBpedia ontológiai osztályait képeztük le a CoNLL-névosztályokra, majd ezeket Wikipédia-entitásokhoz kötöttük
+A Wikipédia entitásait felcímkéztük CoNLL-névosztályokkal, majd a Wikipédia azon mondataiból,
+melyekben ezek az entitások szerepelnek, NER korpuszt építettünk.
 \end{nix}
@@
   \vitem Címkézés: a DBpedia--Szeged leképezés alapján
-  \vitem Kivételek:
-    \begin{itemize}
-    \vitem Cikkek, amik nem entitások: nagybetűvel kell kezdődjön a szócikk
-    \vitem Jelöletlen entitások: megpróbáljuk őket korábbi tulajdonnevekhez kötni
-    \vitem Speciális esetek: tulajdonnév-köznév összetételek, mondatkezdő főnevek, stb.
-    \end{itemize}
   \end{itemize}
@@
   Csak a teljesen felcímkézett mondatokat tartjuk meg. Eldobjuk:
-%  \smallskip
   \begin{itemize}
-  \vitem azokat a mondatokat, amiben címkézetlen hivatkozások vagy nagybetűs szavak vannak
+  \item azokat a mondatokat, amiben címkézetlen hivatkozások vagy nagybetűs szavak vannak
   \vitem a mondattöredékeket:
@@
     \vitem befejezett ige nélküli mondatok % Végül nem dobtuk el
+    \end{itemize}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Kivételek}
+  \bigskip
+  \begin{itemize}
+  \vitem Cikkek, amik nem entitások
+    \begin{itemize}
+    \vitem köznevek (\textit{kocsi}), dátumok (\textit{2012})
+    \vitem megoldás: nagybetűvel kell kezdődjön a hivatkozás
+    \end{itemize}
+  \vitem Jelöletlen entitások:
+    \begin{itemize}
+    \vitem csak az első előfordulást linkelik
+    \vitem megoldás: nagybetűs szavakat próbáljuk már felismert tulajdonnevekhez kötni
+    \end{itemize}
+  \vitem Speciális esetek:
+    \begin{itemize}
+    \vitem tulajdonnév-köznév összetételek
+    \vitem mondatkezdő főnevek
+    \vitem ...
     \end{itemize}
@@
 
-\begin{frame}{Igazságmátrix}
+\begin{frame}{Típustévesztések igazságmátrix}
 
@@
 \begin{itemize}
+\item a Wikipédia rossz kereszthivatkozásai: \\ $[\mbox{Walt Disney}]_{PER}$ Co.
 \item a DBpedia típusinformációja helytelen: \\ $[\mbox{MTA}]_{LOC}$
-\item a Wikipédia rossz kereszthivatkozásai: \\ $[\mbox{Walt Disney}]_{PER}$ Co.
 \end{itemize}
@@
 
-\begin{frame}{További hibák}
+\begin{frame}{A tulajdonnév-felismerés hibái}
 
@@
 \bigskip
-A Szeged NER korpusz 
-% TODO: valami!!!!!
+Statisztikai tanulók eredményességét mértük az automatikusan előállított (\textbf{huwiki}),
+és a Szeged NER korpuszon.
 
@@
 \begin{nix}
-a 3,5 millió tokenes korpuszt 90-10\% arányban tanító és kiértékelő halmazra osztottuk
+A korpuszokat 90-10\% arányban tanító és kiértékelő halmazra osztottuk
 \end{nix}
@@
 \item huwiki tanítóhalmaz -- Szeged NER tesztkorpusz
-\item Szeged NER tanítókorpusz, a Wikipédiából kinyert névlistákkal -- Szeged NER tesztkorpusz % DBpediából, nem?
-\item Szeged NER tanítókorpusz, a Wikipédia korpuszon tanított modell által felcímkézve, és a címkék extra jegyként hozzáadva -- Szeged NER tesztkorpusz
 \end{itemize}
@@
 huwiki & Szeged & 63,08 & 70,46 & 66,57  \\
-Szeged\_wikilisták & Szeged & 95,48 & 95,48 & 95,48  \\
-Szeged\_wikitag & Szeged & 95,38 & 94,92 & 95,15 \\
 \bottomrule
@@
 A saját teszthalmazon elért eredmény arról tanúskodik, hogy a korpusz akár önálló gold standard adathalmazként is használható.
+
+\end{frame}
+
+\begin{frame}{A korpusz egyéb felhasználásai}
+
+\bigskip
+
+A Szeged NER korpuszon futó tagger jegykészletét kiegészítjük
+
+\begin{itemize}
+\item a DBpediából kinyert névlistákkal
+\item a címkével, amit adott szóra a huwiki korpuszon tanított tagger jósol
+\end{itemize}
+
+\begin{center}
+\begin{tabular}{lcccc}
+\toprule 
+\bf tanítás & \bf teszt & \bf P (\%) & \bf R (\%) & \bf F (\%) \\ 
+\midrule
+Szeged & Szeged & 94,50 & 94,35 & 94,43 \\
+Szeged\_wikilisták & Szeged & 95,48 & 95,48 & \textbf{95,48} \\
+Szeged\_wikitag & Szeged & 95,38 & 94,92 & 95,15 \\
+\bottomrule
+\end{tabular}
+\end{center}
 
@@
 
-\begin{frame}{Multitag formátum}
-
-\bigskip
-
-{\footnotesize
-\begin{tabular}{@{}l@{ }l@{ }l@{ }l@{ }l@{ }l@{ }l}
-\textbf{token} & \textbf{text/link} & \textbf{WP link} & \textbf{POS} & \textbf{lemma} & \textbf{NE} \\
-Neumann & B-link & Neumann János & NOUN & neumann & B-PER \\
-János & I-link & Neumann János & NOUN & jános & I-PER \\
-( & text & 0 & PUNCT & ( & O \\
-1903–1957 & text & 0 & NUM & 1903–1957 & O \\
-) & text & 0 & PUNCT & ) & O \\
-matematikai & text & 0 & ADJ & matematikai & O \\
-szemszögből & text & 0 & NOUN<CAS<ELA{>}> & szemszög & O \\
-közelíti & text & 0 & VERB<DEF> & közelít & O \\
-meg & text & 0 & PREV & meg & O \\
-a & text & 0 & ART & a & O \\
-kérdést & text & 0 & NOUN<CAS<ACC{>}> & kérdés & O \\
-\end{tabular}
-}
-
-%\begin{tabular}{llllll}
-%\textbf{token} & \textbf{text/link} & \textbf{WP link} & \textbf{POS} & \textbf{lemma} & \textbf{NE} \\
-%This & text & 0 &    DT &   This & O \\
-%is &   text & 0 &    VBZ &  be &   O \\
-%a &    text & 0 &    DT &   a  &    O\\
-%list & text & 0 &    NN &   list  & O\\
-%of &   text & 0 &    IN &   of &     O\\
-%films  & B-link & films &  NNS &  film & O\\
-%by &   text & 0 &    IN &   by &     O\\
-%year & text & 0 &    NN &   year & O\\
-%produced & text & 0 &    VBN &  produce & O\\
-%in &   text & 0 &    IN &   in &     O\\
-%the &  text & 0 &    DT &   the &   O\\
-%country &text & 0 &    NN &   country &O\\
-%of &   text & 0 &    IN &   of &     O\\
-%South &  B-link & South Korea &  NNP &  South  &B-LOC\\
-%Korea &  I-link & South Korea &  NNP &  Korea   & I-LOC\\
-%\end{tabular}
-
-\end{frame}
-
 \begin{frame}{A korpusz elérhetősége}
@@
 \begin{itemize}
+\vitem multitag formátumúak
 \vitem a CC-BY-SA 3.0 Unported licensz alatt kerültek publikálásra
@@
 
+\begin{frame}{Multitag formátum}
+
+\bigskip
+
+{\footnotesize
+\begin{tabular}{@{}l@{ }l@{ }l@{ }l@{ }l@{ }l@{ }l}
+\textbf{token} & \textbf{text/link} & \textbf{WP link} & \textbf{POS} & \textbf{lemma} & \textbf{NE} \\
+Neumann & B-link & Neumann János & NOUN & neumann & B-PER \\
+János & I-link & Neumann János & NOUN & jános & I-PER \\
+( & text & 0 & PUNCT & ( & O \\
+1903–1957 & text & 0 & NUM & 1903–1957 & O \\
+) & text & 0 & PUNCT & ) & O \\
+matematikai & text & 0 & ADJ & matematikai & O \\
+szemszögből & text & 0 & NOUN<CAS<ELA{>}> & szemszög & O \\
+közelíti & text & 0 & VERB<DEF> & közelít & O \\
+meg & text & 0 & PREV & meg & O \\
+a & text & 0 & ART & a & O \\
+kérdést & text & 0 & NOUN<CAS<ACC{>}> & kérdés & O \\
+\end{tabular}
+}
+
+\end{frame}
+
 \end{document}

### 2013-01-07 paper/MSZNY2013/mszny_slides.tex
Deleted a few lines.
Final version?

@@
 \begin{frame}{A kihívás}
-
-Az annotációs költség csökkentése \nyil automatikusan annotált korpuszok előállítása
+\bigskip
+Az annotációs költség csökkentése:
 
 \begin{itemize}
-\item meglévő korpuszok egyesítése \nyil az annotációs sémák és címkekészletek különbözősége 
-\item rendelkezésre álló korpuszon tanított tagger futtatása új szövegen \nyil a domének közötti különbség teljesítménycsökkenést okoz
-\item közösségi tartalmak felhasználása \nyil leképezés a sztenderd címkekészletekre
+\item crowdsourcing
+\item automatikusan annotált korpuszok előállítása
 \end{itemize}
@@
     \vitem OWL ontológia kb. 320 osztállyal: kézileg leképeztük CoNLL (Szeged) kategóriákra
-    \vitem minden entitást a legspecifikusabb osztály szerinti kategóriába raktunk
     \vitem csökkent fedés, jobb pontosság
@@
     \vitem Kereszthivatkozások, ha a hivatkozott cikk szerepel az entitáslistában
-    \vitem Az aktuális cikk címe, ha szerepel a listában
     \end{itemize}
@@
 
-\begin{frame}{Típustévesztések igazságmátrix}
+\begin{frame}{Típustévesztések igazságmátrixa}
 
@@
 \begin{frame}{Összegzés}
+  \bigskip
   \begin{itemize}
@@
   \vitem bármely Wikipédia-nyelvre alkalmazható \nyil kevés erőforrással rendelkező nyelvekre is előállítható tulajdonnév-annotált korpusz
+  \vitem mivel korpuszunk közelíti a gold standard minőséget, akár önálló tanítóhalmazként is használható a sajtó stílustól eltérő szövegekhez
   \vitem a létrehozott korpuszok több módon is felhasználhatóak a NER rendszerek hatékonyságának növelésére
-  \vitem mivel korpuszunk közelíti a gold standard minőséget, akár önálló tanítóhalmazként is használható a sajtó stílustól eltérő szövegekhez
-  \vitem az általunk előállított korpuszok szabadon felhasználhatóak
+  \vitem az általunk előállított korpuszok szabadon hozzáférhető
 %  \vitem a módszerben benne rejlik a lehetőség finomabb tulajdonnév-hierarchiák támogatására is
